### 2018-09-03 sample-sigconf.tex
initial commit

\documentclass[sigconf]{acmart}

\settopmatter{printacmref=false} % Removes citation information below abstract
\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in first column
\pagestyle{plain} % removes running headers

\usepackage{booktabs} % For formal tables

% Copyright
\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
%\acmDOI{10.475/123_4}

% ISBN
%\acmISBN{123-4567-24-567/08/06}

%Conference
%\acmConference[Short Name N/A]{Long Name N/A}{Date N/A}{Location N/A}
%\acmYear{N/A}
%\copyrightyear{2018}

%\acmArticle{N/A}
%\acmPrice{N/A}

% These commands are optional
%\acmBooktitle{Transactions of the ACM Woodstock conference}
%\editor{Jennifer B. Sartor}
%\editor{Theo D'Hondt}
%\editor{Wolfgang De Meuter}

\begin{document}
\title{A Covert Channel Detector for Machine Learning Models}
%\titlenote{Produces the permission block, and
  %copyright information}
%\subtitle{Extended Abstract}
%\subtitlenote{The full version of the author's guide is available as
  %\texttt{acmart.pdf} document}

\author{John Stein}
%\authornote{Dr.~Trovato insisted his name be first.}
%\orcid{1234-5678-9012}
\affiliation{%
  \institution{Indiana University}
  \streetaddress{107 S Indiana Ave}
  \city{Bloomington}
  \state{Indiana}
  \postcode{47405}
}
\email{jodstein@iu.edu}

%\author{David Crandall, PhD}
%%\authornote{The secretary disavows any knowledge of this author's actions.}
%\affiliation{%
%  \institution{Indiana University}
%  \streetaddress{107 S Indiana Ave}
%  \city{Bloomington}
%  \state{Indiana}
%  \postcode{47405}
%}
%\email{djcran@indiana.edu}

% The default list of authors is too long for headers.
%\renewcommand{\shortauthors}{B. Trovato et al.}

\begin{abstract}
It has been shown that malicious model training code can augment sensitive training data with specially-crafted samples to enable later exfiltration of the sensitive data by an attacker who has black-box access to the completed model, but no direct access to the sensitive training data. \cite{DBLP:journals/corr/abs-1709-07886}  This is known as a Covert Channel for machine learning models.  It has also been shown an attacker can infer whether a specific sample was included in the training set for a black-box model if the attacker has knowledge of the model's structure or can generate models with similar structure. \cite{DBLP:journals/corr/ShokriSS16}  This is known as Membership Inference on machine learning models.  We demonstrate that a variation of the Membership Inference `attack' can be used to detect whether a given machine learning model may have a covert channel.  We review some performance factors of the detector, and discuss how performance is directly related to model capacity and sparsity of the sensitive and maliciously-crafted data.
\end{abstract}

\keywords{machine learning, models, covert channels, detector}

\maketitle

\input{samplebody-conf}

\bibliographystyle{ACM-Reference-Format}
\bibliography{sample-bibliography}

\end{document}

### 2018-10-12 main.tex
Rev1 - Removed unneeded ACM files
Initial ShareLaTeX Import

--- 2018-09-03 sample-sigconf.tex
@@
-\documentclass[sigconf]{acmart}
+\documentclass{article}
+\usepackage[utf8]{inputenc}
 
-\settopmatter{printacmref=false} % Removes citation information below abstract
-\renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in first column
-\pagestyle{plain} % removes running headers
+\title{QE}
+\author{John Stein}
+\date{October 2018}
 
-\usepackage{booktabs} % For formal tables
-
-% Copyright
-\setcopyright{none}
-%\setcopyright{acmcopyright}
-%\setcopyright{acmlicensed}
-%\setcopyright{rightsretained}
-%\setcopyright{usgov}
-%\setcopyright{usgovmixed}
-%\setcopyright{cagov}
-%\setcopyright{cagovmixed}
-
-% DOI
-%\acmDOI{10.475/123_4}
-
-% ISBN
-%\acmISBN{123-4567-24-567/08/06}
-
-%Conference
-%\acmConference[Short Name N/A]{Long Name N/A}{Date N/A}{Location N/A}
-%\acmYear{N/A}
-%\copyrightyear{2018}
-
-%\acmArticle{N/A}
-%\acmPrice{N/A}
-
-% These commands are optional
-%\acmBooktitle{Transactions of the ACM Woodstock conference}
-%\editor{Jennifer B. Sartor}
-%\editor{Theo D'Hondt}
-%\editor{Wolfgang De Meuter}
+\usepackage{natbib}
+\usepackage{graphicx}
 
 \begin{document}
-\title{A Covert Channel Detector for Machine Learning Models}
-%\titlenote{Produces the permission block, and
-  %copyright information}
-%\subtitle{Extended Abstract}
-%\subtitlenote{The full version of the author's guide is available as
-  %\texttt{acmart.pdf} document}
-
-\author{John Stein}
-%\authornote{Dr.~Trovato insisted his name be first.}
-%\orcid{1234-5678-9012}
-\affiliation{%
-  \institution{Indiana University}
-  \streetaddress{107 S Indiana Ave}
-  \city{Bloomington}
-  \state{Indiana}
-  \postcode{47405}
-}
-\email{jodstein@iu.edu}
-
-%\author{David Crandall, PhD}
-%%\authornote{The secretary disavows any knowledge of this author's actions.}
-%\affiliation{%
-%  \institution{Indiana University}
-%  \streetaddress{107 S Indiana Ave}
-%  \city{Bloomington}
-%  \state{Indiana}
-%  \postcode{47405}
-%}
-%\email{djcran@indiana.edu}
-
-% The default list of authors is too long for headers.
-%\renewcommand{\shortauthors}{B. Trovato et al.}
-
-\begin{abstract}
-It has been shown that malicious model training code can augment sensitive training data with specially-crafted samples to enable later exfiltration of the sensitive data by an attacker who has black-box access to the completed model, but no direct access to the sensitive training data.
-\cite{DBLP:journals/corr/abs-1709-07886}  This is known as a Covert Channel for machine learning models.
-It has also been shown an attacker can infer whether a specific sample was included in the training set for a black-box model if the attacker has knowledge of the model's structure or can generate models with similar structure.
-\cite{DBLP:journals/corr/ShokriSS16}  This is known as Membership Inference on machine learning models.
-We demonstrate that a variation of the Membership Inference `attack' can be used to detect whether a given machine learning model may have a covert channel.
-We review some performance factors of the detector, and discuss how performance is directly related to model capacity and sparsity of the sensitive and maliciously-crafted data.
-\end{abstract}
-
-\keywords{machine learning, models, covert channels, detector}
 
@@
 
-\input{samplebody-conf}
+\section{Introduction}
+There is a theory which states that if ever anyone discovers exactly what the Universe is for and why it is here, it will instantly disappear and be replaced by something even more bizarre and inexplicable.
+There is another theory which states that this has already happened.
 
-\bibliographystyle{ACM-Reference-Format}
-\bibliography{sample-bibliography}
+\begin{figure}[h!]
+\centering
+\includegraphics[scale=1.7]{universe}
+\caption{The Universe}
+\label{fig:universe}
+\end{figure}
 
+\section{Conclusion}
+``I always thought something was fundamentally wrong with the universe'' \citep{adams1995hitchhiker}
+
+\bibliographystyle{plain}
+\bibliography{references}
 \end{document}

### 2018-10-14 main.tex
181014-1749_sync

@@
-\documentclass{article}
+\documentclass[11pt]{article}
 \usepackage[utf8]{inputenc}
 
-\title{QE}
-\author{John Stein}
-\date{October 2018}
+\title{Qualifying Examination Paper}
+\author{John Stein (jodstein@iu.edu)}
+\date{November 2018}
 
@@
 \usepackage{graphicx}
+\usepackage{setspace}\singlespacing
+\usepackage[left=1in, right=1in, top=1in, bottom=1in]{geometry}
+\usepackage{appendix}
 
@@
 
+\begin{titlepage}
+    % https://stackoverflow.com/questions/3141702/vertically-centering-a-title-page
+    \null
+    \nointerlineskip
+    \vfill
+    \let\snewpage \newpage
+    \let\newpage \relax
     \maketitle
+    \let \newpage \snewpage
+    \vfill
+    \thispagestyle{empty}
+\end{titlepage}
 
-\section{Introduction}
-There is a theory which states that if ever anyone discovers exactly what the Universe is for and why it is here, it will instantly disappear and be replaced by something even more bizarre and inexplicable.
-There is another theory which states that this has already happened.
+\tableofcontents
+\thispagestyle{empty}
 
-\begin{figure}[h!]
-\centering
-\includegraphics[scale=1.7]{universe}
-\caption{The Universe}
-\label{fig:universe}
-\end{figure}
+%%%%%%%%%%%%%%%%% Paper %%%%%%%%%%%%%%%%%%%%%%
+\newpage
+\setcounter{page}{1}
+\section{Generative Adversarial Networks (GANs)}
 
-\section{Conclusion}
-``I always thought something was fundamentally wrong with the universe'' \citep{adams1995hitchhiker}
+\subsection{Introduction}
+The Generative Adversarial Nets (GAN) framework was first introduced by Goodfellow et al.
+in 2014 \cite{NIPS2014_5423}.
+Deep learning approaches, although highly successful in discrimination or classification tasks, had enjoyed limited success in generative tasks.
+The GAN framework represents an innovative approach for using existing deep learning principles to address generative tasks.
+The GAN framework involves the simultaneous training of both a generative model $G$ with parameters $\theta_g$ and a discriminative model $D$ with parameters $\theta_d$.
+Given a target data space $T$ with distribution $p_t$, $G$'s task is to learn a mapping from an input $z$ to a data space $G(z, \theta_g)$ with distribution $p_g$; while $D$'s task is take input $x$ belonging to either $T$ or $G(z)$ and learn to classify $x$ as belonging either to $p_t$ or $p_g$.
+The adversarial component of the task is that while $D$ is adjusting $\theta_d$ to classify $x$ as belonging to $p_t$ or $p_g$, $G$ is simultaneously adjusting $\theta_g$ to maximize $D$'s error by trying to capture $p_t$ with $p_g$.
+In this way, $G$ learns how to generate data that is hard for an ever-improving $D$ to distinguish.
+Importantly, labeled data is not required for either $G$ or $D$ because the loss function used by $G$ is defined in terms of $D(G(z)) \in \{0, 1\}$ (ie.
+did $G$ fool $D$?) and the labels for $x$ are $\{p_t, p_g\}$ (ie.
+from target or generated data?).
+Thus, the GAN framework presents a promising approach for generating data that approximates a target distribution in an unsupervised manner.
 
+\subsection{GAN Advancements}
+The GAN framework has resulted in significant advancement within many Computer Vision related areas of study, and has inspired many GAN variations that represent improvements or specialized implementations.
+
+In the original GAN, $G$ learns to generate data that estimates $p_t$, but cannot target specific sub-distributions within $p_t$, such as class labels or modes within $p_t$.
+Shortly after the original GAN was introduced, Mirza and Osindero introduced an extension called Conditional GAN \cite{mirza2014conditional} which addresses this need.
+By essentially adding a conditioning input layer to both models $G$ and $D$, one can present the conditioning data $y$ alongside the normal data ($z$ for $G$ and $x \in G(z) \cup T$ for $D$).
+$G$ will therefore learn a mapping given $z$ and $y$ that will result in making classification difficult for $D$ given $x$ and $y$.
+This advancement seems obvious in retrospect, but warrants contemplation.
+While $G$ represents a mapping that captures the distribution of the target data space $T$ in general, the input $z$ can be thought of as the diversity component of the input and the input $y$ can be thought of as the specification component of the input.
+Since the $y$ specification value is consistent within modes, a smaller portion of the overall networks for $G$ and $D$ can focus on the generation and discrimination, respectively, of the modal contribution whereas the remaining portions of the networks can focus on the unspecified contributions from $z$.
+Of course, Conditional GAN requires that conditioning data (eg.
+labels) is available for all target samples $x \in X$.
+
+\subsection{Remaining Challenges and Opportunities}
+
+\begin{itemize}
+    \item Introduction to GANs \cite{NIPS2014_5423} \\ %seminal
+    \item How they work\\
+    \item Why they are novel/interesting\\
+    \item Benefits, Security (Defense, Offense)\\
+
+    \item GAN types: % https://github.com/nashory/gans-awesome-applications
+    \begin{itemize}
+        \item DCGAN \cite{DBLP:journals/corr/RadfordMC15}: (replace pooling with conv, use batchnorm everywhere, remove fully connected, relu for gen, leaky relu for discrim) %seminal
+        \item Hash
+        \item Attention
+        \item Conditional \cite{mirza2014conditional}\\
+        \item InfoGAN https://arxiv.org/abs/1606.03657 (unsupervised, repr disentanglement)
+        \item Wasserstein GAN (WGAN) https://arxiv.org/abs/1701.07875 (improves learning stability and mode collapse)
+        \item BEGAN: Boundary Equilibrium Generative Adversarial Networks
+    \end{itemize}
+
+    \item Uses: 
+    \begin{itemize}
+        \item Image Translation \& 3D
+        \begin{itemize}
+            \item Cross-View Image Synthesis Using Conditional GANs \cite{regmi2018crossview}
+            \item Learning a Probabilistic Latent Space of Object Shapes via 3D Generative-Adversarial Modeling \cite{NIPS2016_6096}
+        \end{itemize}
+
+        \item Domain Transfer
+        \begin{itemize}
+            \item Domain-Adversarial Training of Neural Networks \cite{ganin2016domain}
+            \item Image-to-Image Translation with Conditional GANs \cite{isola2017image}
+        \end{itemize}
+
+        \item Image Recovery 
+        \begin{itemize}
+            \item inpainting: Context Encoders - Feature Learning by Inpainting\_Pathak \cite{pathakCVPR16context}
+            \item resolution: Photo-Realistic Single Image Super-Resolution Using a Generative Adversarial Network \cite{ledig2017photo}
+            \item honorable mentions (face resolution: "Finding Tiny Faces in the Wild with Generative Adversarial Network" \cite{bai2018finding}, "Generative Adversarial Networks recover features in astrophysical images of galaxies beyond the deconvolution limit" \cite{schawinski2017generative})
+        \end{itemize}
+
+        \item Image Synthesis (including from text "Generative Adversarial Text to Image Synthesis" \cite{reed2016generative}, semi-supervised learning "Improved Techniques for Training GANs" \cite{salimans2016improved})
+
+        \item Pattern recognition and segmentation (Semantic Segmentation using Adversarial Networks \cite{Luc2016SemanticSU}
+
+        \item Adversarial/Security
+        \begin{itemize}
+            \item Adversarial Examples Generation and Defense Based on Generative Adversarial Network \cite{xia2016adversarial}
+            \item Recycle-GAN \cite{bansal2018recyclegan}
+            \item H.m.: deepfakes \cite{deepfakes-faceswap}
+            \item My paper
+        \end{itemize}
+
+        \item Video
+        \begin{itemize}
+        \item Deep multi-scale video prediction beyond mean square error \cite{Mathieu2015DeepMV}
+        \item Video-to-Video Synthesis \cite{wang2018videotovideo}
+            \item MoCoGAN \cite{tulyakov2017mocogan}
+        \end{itemize}
+
+        \item Task-Oriented (robotics)
+
+    \end{itemize}
+
+    \item Areas of Advancement:
+    \begin{itemize}
+        \item use cases
+        \item mode collapse
+        \item unsupervised and semi-supervised learning ability
+        \item learning stability
+        \item regularization
+        \item vanishing gradient
+        \item loss types
+        \item structures
+        \item video, sequencing
+    \end{itemize}
+
+    \item Other Potentials:
+    \begin{itemize}
+        \item FaceID-GAN: Learning a Symmetry Three-Player GAN for Identity-Preserving Face Synthesis
+        \item GAGAN: Geometry-Aware Generative Adversarial Networks
+        \item MoCoGAN: Decomposing Motion and Content for Video Generation
+        \item DA-GAN: Instance-Level Image Translation by Deep Attention Generative Adversarial Networks (unsupervised)
+        \item SeGAN: Segmenting and Generating the Invisible \cite{ehsani2017segan}
+        \item From Source to Target and Back: Symmetric Bi-Directional Adaptive GAN 
+        \item Learning Pose Specific Representations by Predicting Different Views
+        \item Cross-Modal Deep Variational Hand Pose Estimation
+        \item Disentangled Person Image Generation
+    \end{itemize}
+
+\end{itemize}
+
+%%%%%%%%%%%%%%%%% BIB %%%%%%%%%%%%%%%%%%%%
+\newpage
 \bibliographystyle{plain}
 \bibliography{references}
+\addcontentsline{toc}{section}{References}
+
+%%%%%%%%%%%%%%%%% Appendices %%%%%%%%%%%%%%%%%%%%
+\newpage
+\appendixtitleon
+\begin{appendices}
+
+\begin{center}
+    \section{Examination Committee Questions}
+\end{center}
+
+\subsection{Dr. Apu Kapadia}
+
+\subsubsection{GANs (2.5-3 pages excluding citations)}
+Researchers have recently made much progress in the area of adversarial machine learning using "generative adversarial networks." Identify 8-12 papers in this area specific to computer vision, and synthesize and summarize the various research contributions in this area; draw connections between the various papers and make sure the writeup does NOT look like a string of summaries.
+Instead group and summarize categories and discuss how they relate to each other.
+Make sure you critique the various works when you discuss them, discussing how these works address the weaknesses of the other works you discuss.
+
+Have a clearly marked section at the end (about 0.75 page) that talks about research challenges and directions forward from your perspective, i.e., what all remains to be done after your discussion of the various papers (including your own published work)?
+
+\subsubsection{Adversarial machine learning (1.5-2 pages excluding citations)}
+Model this section on the previous section, but expand the focus beyond computer vision and talk about 5-8 papers NOT related to computer vision and how machine learning algorithms can be 'tricked' by adversaries.
+Discuss how machine learning algorithms in the context of security can be made robust against adversaries trying to escape classification, for example.
+
+\subsection{Dr. David Crandall}
+
+\subsection{Dr. Donald Williamson}
+
+\end{appendices}
+
 \end{document}

### 2018-10-16 main.tex
181016-0830

@@
 
+(TBD) LAPGAN
+
+As the appreciation and use of GAN-based approaches increased, it was observed that the structure design of the $G$ and $D$ models deserved reconsideration.
+Radford, Metz, and Chintala \cite{DBLP:journals/corr/RadfordMC15} evaluated various model architectures and recommended a set of guidelines and constraints for $G$ and $D$ architectures that appeared to perform well in terms of stability, convergence, resolution, and depth of representation.
+These guidelines describe a Deep Convolutional GAN (DCGAN) and are summarized as follows: (a) use convolution and fractional convolution layers with striding to achieve adaptive down-sampling ($D$) or up-sampling ($G$), respectively, rather than fixed/deterministic methods (i.e.
+max pooling), (b) use batch normalization between convolution layers to mitigate collapse or failure due to poor initialization or vanishing gradients, (c) avoid fully-connected layers, with the exception of the input to $G$ and the output of $D$, to help earlier convergence, and (d) use Rectifier Linear Unit (ReLU) activation for input and hidden layers in $G$, tanh for the output layer in $G$, and Leaky ReLU for all layers in $D$.
+In addition to performance improvements, the authors demonstrate that $D$ is able to learn semantically-relevant features that are both explainable (via latent-space walk based validation) and usable (via feature-based supervised classification).
+In fact, the extracted features from $D$ were so informative that they were able to discover associations between the input noise vector $Z$ and some features, and even modify the generated images (such as face rotation) using simple interpolation along the correct axis in $Z$-space.
+The strong semantic learning power offered by DCGAN made it a strong candidate for enabling full unsupervised learning tasks such that the features learned by $D$ can be later explained and applied for re-use.
+
+%In Machine Learning, and especially Computer Vision, domain variance between training and test data has been a long standing challenge.
+The domain of a data set can be explained as components of the distribution that are invariant to class label, but unique to the data set as a whole, in contrast to other similar data sets.
+Ganin et al.
+recognized that an adversarial approach known as a Domain Adversarial Neural Network (DANN) could be used to help improve Domain Adaption (DA).
+Their adversarial approach essentially minimizes the loss with respect to the basic classification task, and maximizes the loss with respect to an adversarial model trying to distinguish between the training domain and a test domain.
+
+Wtihin the theme of semantic learning and generation, several significant contributions were made during 2016 and 2017.
+Pathak, Krahenbuhl, Donahue, Darrell, and Efros successfully used an adversarial approach to inpaint (fill in missing portions of) images with large extents of missing pixel data.
+Their architecture was based on an encoder-decoder (TBD ref) combined with GAN.
+In their approach, the encoder encodes both whole and masked (portions missing) images and $D$ tries to discriminate whole vs.
+masked, while $G$ generates the completed images, conditioned on the context.
+...
+
+ serves as $G$, attempting to reconstruct the context encoding from images with missing portions and the decoder feeds $D$ which tries to distinguish between reconstructed images and complete images.
+Interestingly, and in contrast with \cite{mirza2014conditional}, the authors achieved better results when $G$ was conditioned
+
 \subsection{Remaining Challenges and Opportunities}
@@
         \begin{itemize}
-            \item Domain-Adversarial Training of Neural Networks \cite{ganin2016domain}
+            %\item Domain-Adversarial Training of Neural Networks \cite{ganin2016domain} - ADVERSARIAL BUT NOT GAN
             \item Image-to-Image Translation with Conditional GANs \cite{isola2017image}

### 2018-10-18 main.tex
181018-0913 sync

@@
 Wtihin the theme of semantic learning and generation, several significant contributions were made during 2016 and 2017.
-Pathak, Krahenbuhl, Donahue, Darrell, and Efros successfully used an adversarial approach to inpaint (fill in missing portions of) images with large extents of missing pixel data.
+Pathak, Krahenbuhl, Donahue, Darrell, and Efros successfully used an adversarial approach to inpaint (fill in missing portions of) images with large extents of missing pixel data \cite{pathakCVPR16context}.
 Their architecture was based on an encoder-decoder (TBD ref) combined with GAN.
-In their approach, the encoder encodes both whole and masked (portions missing) images and $D$ tries to discriminate whole vs.
-masked, while $G$ generates the completed images, conditioned on the context.
-...
-
- serves as $G$, attempting to reconstruct the context encoding from images with missing portions and the decoder feeds $D$ which tries to distinguish between reconstructed images and complete images.
+In their approach, the encoder-decoder learns a semantic encoding based on the non-missing portions and attempts to reproduce a whole image based on that semantic encoding.
+The `reconstruction loss' is defined in terms of closesness ($L2$) of the filled area to the removed area.
+The encoder-decoder is regarded as $G$, and a new Discriminator $D$ is trained to predict whether its input is a reconstructed output from $G$ or an original image.
 Interestingly, and in contrast with \cite{mirza2014conditional}, the authors achieved better results when $G$ was conditioned
@@
 \subsection{Remaining Challenges and Opportunities}
+
+semantics discovery (semantics of security)
 

### 2018-10-19 main.tex
181019-1611 sync

@@
 Wtihin the theme of semantic learning and generation, several significant contributions were made during 2016 and 2017.
-Pathak, Krahenbuhl, Donahue, Darrell, and Efros successfully used an adversarial approach to inpaint (fill in missing portions of) images with large extents of missing pixel data \cite{pathakCVPR16context}.
-Their architecture was based on an encoder-decoder (TBD ref) combined with GAN.
+Pathak, Krahenbuhl, Donahue, Darrell, and Efros used an adversarial approach to inpaint (fill in missing portions of) images with large extents of missing pixel data \cite{pathakCVPR16context}.
+Their architecture is based on an autoencoder \cite{hinton2006reducing} combined with GAN.
 In their approach, the encoder-decoder learns a semantic encoding based on the non-missing portions and attempts to reproduce a whole image based on that semantic encoding.
-The `reconstruction loss' is defined in terms of closesness ($L2$) of the filled area to the removed area.
+The `reconstruction loss' is defined in terms of distance ($L2$) from the filled area to the removed area.
 The encoder-decoder is regarded as $G$, and a new Discriminator $D$ is trained to predict whether its input is a reconstructed output from $G$ or an original image.
-Interestingly, and in contrast with \cite{mirza2014conditional}, the authors achieved better results when $G$ was conditioned
+Interestingly, and in contrast with \cite{mirza2014conditional}, the authors achieved better results when $G$, but not $D$, was conditioned with context information \footnote{The paper defines context information as the missing pixels which, if true, undermines the objective (any statement of success should be based on withholding the missing pixels from the architecture).
+I have requested clarification on this from the author (17 Oct 2018).}.
+
+Luc, Couprie,Chintala and Verbeek applied a straightforward GAN approach to the task of semantic segmentation on images \cite{Luc2016SemanticSU}.
+Their approach uses a Convolutional Neural Network (CNN) based generative model $G$ to learn a class label for each pixel in the image based on a ground-truth labeling (supervised), thereby creating a segmentation map of the input image.
+The discriminative model $D$ takes a segmentation map and tries to classify it as being a ground truth map or a generated map.
+While $D$'s objective function tries to maximize classification accuracy of the input segmentation map as being either ground truth or generated, $G$'s objective function tries to maximize pixel class labeling accuracy with respect to ground truth while also minimizing accuracy.
+
+Reed et al. applied a GAN approach to the task of generating realistic images from text-based descriptions \cite{reed2016generative}.
 

### 2018-10-23 main.tex
181023-0732 sync

@@
 
-\subsection{Introduction}
+%\subsection{Introduction}
 The Generative Adversarial Nets (GAN) framework was first introduced by Goodfellow et al.
 in 2014 \cite{NIPS2014_5423}.
-Deep learning approaches, although highly successful in discrimination or classification tasks, had enjoyed limited success in generative tasks.
-The GAN framework represents an innovative approach for using existing deep learning principles to address generative tasks.
 The GAN framework involves the simultaneous training of both a generative model $G$ with parameters $\theta_g$ and a discriminative model $D$ with parameters $\theta_d$.
-Given a target data space $T$ with distribution $p_t$, $G$'s task is to learn a mapping from an input $z$ to a data space $G(z, \theta_g)$ with distribution $p_g$; while $D$'s task is take input $x$ belonging to either $T$ or $G(z)$ and learn to classify $x$ as belonging either to $p_t$ or $p_g$.
-The adversarial component of the task is that while $D$ is adjusting $\theta_d$ to classify $x$ as belonging to $p_t$ or $p_g$, $G$ is simultaneously adjusting $\theta_g$ to maximize $D$'s error by trying to capture $p_t$ with $p_g$.
+Given a target data space $T$ with distribution $p_t$, $G$'s task is to learn a mapping from an noisy input $z$ to the data space $p_t$ ($G:Z \rightarrow G(z) \sim p_g \approx p_t$); while $D$'s task is to discriminate whether some input $x$ originated from either $T$ or $G(z)$ ($D: X \rightarrow \{0, 1\}$).
+Since $G$ would like to generate samples that are representative of $T$, and therefore make $D$'s task more difficult by capturing $p_t$ with $p_g$, $D$ includes a loss term that is based on whether $D$ successfully discriminated inputs as $T$ or $G(z)$ and back-propagates $D$'s success as error to $\theta_g$ appropriately.
 In this way, $G$ learns how to generate data that is hard for an ever-improving $D$ to distinguish.
-Importantly, labeled data is not required for either $G$ or $D$ because the loss function used by $G$ is defined in terms of $D(G(z)) \in \{0, 1\}$ (ie.
-did $G$ fool $D$?) and the labels for $x$ are $\{p_t, p_g\}$ (ie.
+Importantly, labeled data is not required for either $G$ or $D$ because the loss function used by $G$ is defined in terms of $D(x)$ and $(1-D(G(z))) \in \{0, 1\}$ (ie.
+did $G$ fool $D$?) and the labels for $x$ essentially map to $\{p_t, p_g\}$ (ie.
 from target or generated data?).
@@
 
-\subsection{GAN Advancements}
+\subsection{GANs in Computer Vision}
+
 The GAN framework has resulted in significant advancement within many Computer Vision related areas of study, and has inspired many GAN variations that represent improvements or specialized implementations.
+
+\subsubsection{GAN Implementation Advancements}
 
@@
 Shortly after the original GAN was introduced, Mirza and Osindero introduced an extension called Conditional GAN \cite{mirza2014conditional} which addresses this need.
-By essentially adding a conditioning input layer to both models $G$ and $D$, one can present the conditioning data $y$ alongside the normal data ($z$ for $G$ and $x \in G(z) \cup T$ for $D$).
-$G$ will therefore learn a mapping given $z$ and $y$ that will result in making classification difficult for $D$ given $x$ and $y$.
+By adding a conditioning input layer to both models $G$ and $D$, one can present the conditioning data $y$ alongside the normal data.
+$G$ will learn a mapping from $z$ to $X$ given $y$ that will result in making classification of $X$ given $y$ difficult for $D$.
 This advancement seems obvious in retrospect, but warrants contemplation.
-While $G$ represents a mapping that captures the distribution of the target data space $T$ in general, the input $z$ can be thought of as the diversity component of the input and the input $y$ can be thought of as the specification component of the input.
+While $G$ represents a mapping that captures the distribution of the target data space $T$ in general, the noisy input $z$ can be thought of as the diversity component of the input and the input $y$ can be thought of as the specification component of the input.
 Since the $y$ specification value is consistent within modes, a smaller portion of the overall networks for $G$ and $D$ can focus on the generation and discrimination, respectively, of the modal contribution whereas the remaining portions of the networks can focus on the unspecified contributions from $z$.
@@
 
-(TBD) LAPGAN
+%(TBD) LAPGAN
 
@@
 In fact, the extracted features from $D$ were so informative that they were able to discover associations between the input noise vector $Z$ and some features, and even modify the generated images (such as face rotation) using simple interpolation along the correct axis in $Z$-space.
-The strong semantic learning power offered by DCGAN made it a strong candidate for enabling full unsupervised learning tasks such that the features learned by $D$ can be later explained and applied for re-use.
+
+%The strong semantic learning power offered by DCGAN made it a strong candidate for enabling full unsupervised learning tasks such that the features learned by $D$ can be later explained and applied for re-use.
 
@@
 
-Wtihin the theme of semantic learning and generation, several significant contributions were made during 2016 and 2017.
+\subsubsection{GAN in Semantic Learning and Generation}
+Within the theme of semantic learning and generation, several significant contributions were made during 2016 and 2017.
 Pathak, Krahenbuhl, Donahue, Darrell, and Efros used an adversarial approach to inpaint (fill in missing portions of) images with large extents of missing pixel data \cite{pathakCVPR16context}.
-Their architecture is based on an autoencoder \cite{hinton2006reducing} combined with GAN.
+Their architecture is based on an auto-encoder \cite{hinton2006reducing} combined with GAN.
 In their approach, the encoder-decoder learns a semantic encoding based on the non-missing portions and attempts to reproduce a whole image based on that semantic encoding.
 The `reconstruction loss' is defined in terms of distance ($L2$) from the filled area to the removed area.
-The encoder-decoder is regarded as $G$, and a new Discriminator $D$ is trained to predict whether its input is a reconstructed output from $G$ or an original image.
+The encoder-decoder is regarded as $G$, and a new discriminator $D$ is trained to predict whether its input is a reconstructed output from $G$ or an original image.
 Interestingly, and in contrast with \cite{mirza2014conditional}, the authors achieved better results when $G$, but not $D$, was conditioned with context information \footnote{The paper defines context information as the missing pixels which, if true, undermines the objective (any statement of success should be based on withholding the missing pixels from the architecture).
@@
 
-Luc, Couprie,Chintala and Verbeek applied a straightforward GAN approach to the task of semantic segmentation on images \cite{Luc2016SemanticSU}.
+Wu et al.
+took a somewhat similar approach and extended a variational autoencoder \cite{larsen2015autoencoding} within a 3-dimensional GAN architecture (3D-VAE-GAN) which learns to generate 3D images from 2D examples \cite{NIPS2016_6096}.
+
+Luc, Couprie, Chintala and Verbeek applied a straightforward GAN approach to the task of semantic segmentation on images \cite{Luc2016SemanticSU}.
 Their approach uses a Convolutional Neural Network (CNN) based generative model $G$ to learn a class label for each pixel in the image based on a ground-truth labeling (supervised), thereby creating a segmentation map of the input image.
 The discriminative model $D$ takes a segmentation map and tries to classify it as being a ground truth map or a generated map.
-While $D$'s objective function tries to maximize classification accuracy of the input segmentation map as being either ground truth or generated, $G$'s objective function tries to maximize pixel class labeling accuracy with respect to ground truth while also minimizing accuracy.
+%While $D$'s objective function tries to maximize classification accuracy of the input segmentation map as being either ground truth or generated, $G$'s objective function tries to maximize pixel class labeling accuracy with respect to ground truth while also minimizing accuracy.
 
-Reed et al. applied a GAN approach to the task of generating realistic images from text-based descriptions \cite{reed2016generative}.
+Ehsani, Mottaghi, and Farhadi used a GAN approach to address the related dual purpose task of semantically segmenting objects within an image and completing those objects that were occluded by other objects in the image \cite{ehsani2017segan}.
 
-\subsection{Remaining Challenges and Opportunities}
+Reed et al.
+applied a GAN approach to the dual-task of (1) generating realistic images (2) from text-based descriptions \cite{reed2016generative}.
+Essentially, they create $G$ by concatenating the input noise with an encoding of the text description and feed the result into a feed-forward CNN using fractional-striding\footnote{The paper claims this to be a de-convolutional network, but I presume it to mean a CNN using fractional striding to achieve up-sampling.}.
+They create $D$ by downsampling with CNN layers until a 4x4 representation is achieved, concatenate the text encoding onto each of the 16 frames, and continue convolutions to achieve the binary discriminator output.
+The learning objectives are consistent with \cite{NIPS2014_5423}.
+However, since there are two sources of error (error associated with producing fake-looking images and error associated with producing real-looking images that do not match the description), their approach addresses the second source of error by augmenting the labeled data with real images having non-matching descriptions, which $D$ must learn to recognize.
 
-semantics discovery (semantics of security)
+\subsubsection{Image Translation}
 
-\begin{itemize}
-    \item Introduction to GANs \cite{NIPS2014_5423} \\ %seminal
-    \item How they work\\
-    \item Why they are novel/interesting\\
-    \item Benefits, Security (Defense, Offense)\\
+Another rich area of research that has greatly benefited from the GAN architecture is image translation.
+Generally, image translation deals with the learning of a meaningful semantic representation of an image in one domain, and then generating an image with the same semantic representation from a different domain.
+Isola, Zhu, Zhou, and Efros designed a solution that appears to generalize well across various domain pairs for this task \cite{isola2017image}.
+Their architecture follows a U-Net auto-encoder architecture \cite{ronneberger2015u} with DCGAN constraints \cite{salimans2016improved}.
+Further, similar to \cite{pathakCVPR16context}, they constrain $D$'s task to focusing on high-frequency semantic accuracy using learned image patch losses that are fed from skipped connections from encoder layers (PatchGAN), while leveraging an $L1$ term in $G$'s loss to ensure low-frequency accuracy.
+A later paper from Liu, Breuel, and Kautz also displays impressive results \cite{NIPS2017_6672} in image translation.
 
-    \item GAN types: % https://github.com/nashory/gans-awesome-applications
-    \begin{itemize}
-        \item DCGAN \cite{DBLP:journals/corr/RadfordMC15}: (replace pooling with conv, use batchnorm everywhere, remove fully connected, relu for gen, leaky relu for discrim) %seminal
-        \item Hash
-        \item Attention
-        \item Conditional \cite{mirza2014conditional}\\
-        \item InfoGAN https://arxiv.org/abs/1606.03657 (unsupervised, repr disentanglement)
-        \item Wasserstein GAN (WGAN) https://arxiv.org/abs/1701.07875 (improves learning stability and mode collapse)
-        \item BEGAN: Boundary Equilibrium Generative Adversarial Networks
-    \end{itemize}
+\subsubsection{Video Implementations}
 
-    \item Uses: 
-    \begin{itemize}
-        \item Image Translation \& 3D
-        \begin{itemize}
-            \item Cross-View Image Synthesis Using Conditional GANs \cite{regmi2018crossview}
-            \item Learning a Probabilistic Latent Space of Object Shapes via 3D Generative-Adversarial Modeling \cite{NIPS2016_6096}
-        \end{itemize}
+A logical step forwards from image-based applications is to apply GAN to video.
+Mathieu, Couprie, and LeCun first adopted a GAN approach to a video prediction task \cite{Mathieu2015DeepMV}.
+They designed a scaled generator $G_k$ that convolves input frames at scale $k$ and predicts a future frame at that scale $Y_k$.
+The overall generator $G$ then generates the final predicted frame $Y$ at full resolution using a Laplacian pyramid \cite{denton2015deep} built on each $G_k$.
+The discriminator $D$ attempts to classify the last frame(s) in a frame sequence as being either real or generated by $G$ (the first frames in the input sequence are always real), and thus provides the usable adversarial loss to $G$.
 
-        \item Domain Transfer
-        \begin{itemize}
-            %\item Domain-Adversarial Training of Neural Networks \cite{ganin2016domain} - ADVERSARIAL BUT NOT GAN
-            \item Image-to-Image Translation with Conditional GANs \cite{isola2017image}
-        \end{itemize}
+Tulyakov, Liu, Yang, and Kautz introduce a method for decomposing video into content (the objects in the video) and motion (the movement of those objects) components, each having its own generator and discriminator ($G_C, G_M, D_C, D_M$) \cite{tulyakov2017mocogan}.
+$D_C$ attempts to detect whether a frame is sampled from a real video clip or a generated one, while $D_M$ attempts to detect whether a video clip (series of frames) is sampled from a real or generated clip.
+$G_C$ (a CNN) takes motion vectors from $G_M$ (a Recurrent Neural Network), and it's objective includes loss terms from both $D_C$ (a CNN) and $D_M$ (a spatio-temporal CNN), while $G_M$'s objective includes a loss term from $D_M$ only.
+In this way, the noise input to $G_C$ can control which objects are in the video, while the noise input to $G_M$ can control the objects' motion.
+Another notable application involves segmentation-map-to-video synthesis \cite{wang2018videotovideo}.
 
-        \item Image Recovery 
-        \begin{itemize}
-            \item inpainting: Context Encoders - Feature Learning by Inpainting\_Pathak \cite{pathakCVPR16context}
-            \item resolution: Photo-Realistic Single Image Super-Resolution Using a Generative Adversarial Network \cite{ledig2017photo}
-            \item honorable mentions (face resolution: "Finding Tiny Faces in the Wild with Generative Adversarial Network" \cite{bai2018finding}, "Generative Adversarial Networks recover features in astrophysical images of galaxies beyond the deconvolution limit" \cite{schawinski2017generative})
-        \end{itemize}
+Bansal et al.
+tackle the task of video-to-video translation and achieve fairly convincing results with a hierarchical GAN approach (RecycleGAN) \cite{bansal2018recyclegan}.
+Their approach does not require matched video pairs (between domains), but instead introduces loss terms related to reconstruction (accuracy in translating from domain A to domain B and back to domain A again), next frame prediction, and straightforward generative and aversarial losses.
 
-        \item Image Synthesis (including from text "Generative Adversarial Text to Image Synthesis" \cite{reed2016generative}, semi-supervised learning "Improved Techniques for Training GANs" \cite{salimans2016improved})
+\subsubsection{Implications to Security}
 
-        \item Pattern recognition and segmentation (Semantic Segmentation using Adversarial Networks \cite{Luc2016SemanticSU}
+Like any powerful technology, there are significant security implications associated with GANs, both with respect to how they are implemented and how they are used.
+Indeed, most of the papers involving GAN implementations for computer vision applications rely on the ability of generated image/video to fool a human observer as an important qualitative measure of success.
+The use of deep learning as a tool for generating convincing, non-authentic video was put in the public eye when a toolset developed by deepfake \cite{deepfakes-faceswap} was used maliciously to produce pornographic video content with public personalities, and later by celebrity Jordan Peele as a public service announcement, warning of the dangers of the technology \cite{peeleDeepFakes}.
 
-        \item Adversarial/Security
-        \begin{itemize}
-            \item Adversarial Examples Generation and Defense Based on Generative Adversarial Network \cite{xia2016adversarial}
-            \item Recycle-GAN \cite{bansal2018recyclegan}
-            \item H.m.: deepfakes \cite{deepfakes-faceswap}
-            \item My paper
-        \end{itemize}
+Although this paper has only discussed the use of GANs to achieve a singular objective, learning and/or generative models can be juxtaposed by two owners with opposing goals.
+Szegedy et al.
+introduced this as a potential issue, showing how minimally-perturbed examples can be generated which result in miss-classification (with high confidence) by a classification model.
+Several follow-up studies have explored this phenomena further \cite{carlini2016towards, kos2017adversarial, moosavi2017universal, athalye2017synthesizing}, noting potentially serious consequences for applications such as robotics, autonomous vehicles, and military.
 
-        \item Video
-        \begin{itemize}
-        \item Deep multi-scale video prediction beyond mean square error \cite{Mathieu2015DeepMV}
-        \item Video-to-Video Synthesis \cite{wang2018videotovideo}
-            \item MoCoGAN \cite{tulyakov2017mocogan}
-        \end{itemize}
+%Lastly, Shokri et al.
+developed an adversarial methodology for determining whether a given sample was specifically included in the training set (which, in many cases, may be considered private) for a given classification model \cite{shokri2017membership}.
+Song et al.
+developed a methodology
 
-        \item Task-Oriented (robotics)
+\subsubsection{Remaining Challenges and Opportunities}
 
-    \end{itemize}
+As discussed, the GAN approach has enabled major improvements to several research areas including semi-supervised and unsupervised learning, generative models, the learning of complex loss functions, data synthesis for other learning tasks, image and domain learning and translation, and semantic learning and generation.
+In general, the GAN approach provides a relatively simple and tractable way of isolating deep learning objectives and sources of error in such a way that they can be addressed individually and researchers can make headway toward a complex objective, as is apparent in the Conditional GAN, in-painting, MoCoGAN, and text-to-image generation, among others.
+Broadly stated, as deep learning models adopt more complex architectures and greater capacity, there is an increasing need to provide more meaningful and comprehensive constraint to the model's objective.
+GANs provide a tractable methodology for learning sets of constraints which can be as complex as the main learning objective, or even more-so.
 
-    \item Areas of Advancement:
-    \begin{itemize}
-        \item use cases
-        \item mode collapse
-        \item unsupervised and semi-supervised learning ability
-        \item learning stability
-        \item regularization
-        \item vanishing gradient
-        \item loss types
-        \item structures
-        \item video, sequencing
-    \end{itemize}
+To this end, I predict that a likely trend in GAN research is the emergence of more complex GAN architectures.
+Subdividing complex goals into multiple generators and complex constraints into multiple discriminators is likely to lead to elegant solutions to tasks previously considered to be too complex for more conventional singular deep learning models.
+Likewise, as GAN architectures grow in complexity, the overall objective function (combination of loss terms, heuristics, and discriminator outputs) for a complex learning task may become a focus of future work.
+Whereas, for example, current research is learning semantic representations in videos that involve specific objects and their spatial movement across several frames, deeper semantic learning of \textit{who, what, where, when, and why} information may demand these more complex hierarchical structures.
 
-    \item Other Potentials:
-    \begin{itemize}
-        \item FaceID-GAN: Learning a Symmetry Three-Player GAN for Identity-Preserving Face Synthesis
-        \item GAGAN: Geometry-Aware Generative Adversarial Networks
-        \item MoCoGAN: Decomposing Motion and Content for Video Generation
-        \item DA-GAN: Instance-Level Image Translation by Deep Attention Generative Adversarial Networks (unsupervised)
-        \item SeGAN: Segmenting and Generating the Invisible \cite{ehsani2017segan}
-        \item From Source to Target and Back: Symmetric Bi-Directional Adaptive GAN 
-        \item Learning Pose Specific Representations by Predicting Different Views
-        \item Cross-Modal Deep Variational Hand Pose Estimation
-        \item Disentangled Person Image Generation
-    \end{itemize}
+Further, as research yields models and approaches that efficiently and effectively solve moderately complex tasks, it is reasonable to expect growing a demand for standardized, modular, and extensible model interfaces (eg.
+Model Based Engineering ecosystem) for use among research and commercial entities to solve more complex tasks and continue the evolution of artificial intelligence.
 
-\end{itemize}
+In the area of security, GAN approaches pose both a threat (fake video, adversarial examples) as well as a benefit (eg.
+efficient fuzzing \cite{hu2018ganfuzz}) to security, privacy, and safety.
+Specifically, the ability to generate adversarial examples potentially applies to many systems in which the input to that system is corruptible/mutable by a malicious actor.
+Unfortunately, deep learning approaches to build robustness against these attacks into the system are only advancing the adversarial game one move forward.
+A more effective (though admittedly not immune) approach may involve research into understanding and increasing the robustness of the manifold (hidden layers) of deep learning models.
+Some objectives of research in this area may be to level-load the sensitivity of the manifold to small changes or to explore the robustness of the manifold in terms of spacial, temporal, or modal variation.
+
+\subsection{Adversarial Machine Learning}
+
+This section will explore the broader topic of Adversarial Machine Learning (ML) and it's role in security and privacy.
+Back in 2008, Barreno, Nelson, Joseph, and Tygar discussed some security implications of ML and proposed a taxonomy to identify and analyze different attacks \cite{barreno2010security}.
+In the proposed taxonomy, an attack was to be identified in terms of \textit{Influence}, \textit{Security Violation}, and \textit{Specificity}.
+\textit{Influence} is similar to the concept of attack surface, and refers to whether the attacker (a) can control training data to influence the model (causative) or (b) must exploit an existing model during test/classification (exploratory).
+\textit{Security Violation} refers to whether the attacker's goal is to result in misclassification of inputs (integrity compromise) or in general performance degradation (model availability compromise).
+Lastly, \textit{Specificity} refers to whether the attack is targeted at a particular instance (eg.
+a single input) or a wide class of instances (one or multiple modes).
+
+More recently, research has shown that legitimate attacks could be performed to expose potentially private training data or reveal enough information about model structure and parametrics to effectively learn sensitive/proprietary latent semantic representation through white- and black-box methods, which is discussed further in the following sections.
+
+\subsubsection{Malicious Perturbations}
+
+When ML methods are used as a tool in an adversarial context like security, robustness of the model to adversarial examples becomes a primary design objective.
+Spam filtering and intrusion detection systems (IDS) are two classic examples wherein the design objective is to accurate classify messages or traffic as benign or malicious, while the attacker's objective is to design his/her malicious content to be classified as benign.
+%This scenario depicts an \textit{exploratory}, \textit{integrity}, and \textit{specific} attack.
+
+Spam filtering has a relatively rich history from an adversarial learning perspective.
+Word-based Bayesian spam filters date back to 1998 \cite{sahami1998bayesian}, and the first `Bayes vs.
+Bayes' attack surfaced (in the public eye) in 2004 \cite{graham2004beat}.
+
+\subsubsection{Training Data Confidentiality}
+
+Shokri, Stronati, Song, and Shmatikov have demonstrated a methodology to determine whether a specific sample was included within the training set for a target model \cite{shokri2017membership}.
+Given a specific data sample $x_?$ and a target model $M_t$, their paper shows how to create multiple `shadow' models with different (real or synthesized) data from that used to train $M_t$, obtain prediction responses from each of the shadow models that can be labeled as `in' or `out' depending on whether each sample was part of that shadow models training set, and then use the labeled response data to then train an attack model $M_a$.
+$M_a$ achieves good performance in predicting whether black-box responses from $M_t$ are associated with training input (`in') or test input (`out').
+The paper addresses the problem of simply assuming that higher-confidence responses corresponds to `in' data by the following methodology: (a) start running randomized data samples through $m_t$, (b) use a hill-climbing algorithm to search the sample space until a high-confidence response is obtained, (c) repeat a and b until asufficiently large set of data is obtained, then (d) sample from the obtained samples and split into training and test sets for each shadow model.
+In this way, the distributions for `in' and `out' synthesized data are the same from a confidence perspective.
+
+Similarly, Song, Ristenpart, and Shmatikov demonstrated several white- and black-box approaches to build covert channels into ML models for the purpose of extracting sensitive/proprietary training data.
+In the attack scenario, the model provider and the training data owner may be different entities.
+The data owner may not wish to reveal the training data to the model developer, but instead submit the data to a secure ML service broker that then trains and delivers the working model back to the data owner.
+White-box methods generally include methods for steganographically encoding training data into model parametrics that would then be available for extraction by an attacker with white-box access to the completed model.
+More interestingly, a `capacity abuse' attack was described that addresses the black-box scenario.
+Essentially, the learning algorithm automatically augments the training data with chosen (by the attacker) data and labels that data with an encoding of the private training data.
+The model learns to classify the original and malicious training data according to their labels, and then the attacker, having balack-box access to the model, can present the known malicious data to the model and the responses will be an encoding of the private training data.
+I hypothesize that performance of the model for both it's benign purpose as well as it's covert purpose depends on the malicious data being sparse.
+The intuition is that since sparse data is statistically distant from the private training data, it will not have adverse affects on classification; and since each sparse sample is distant from other sparse samples, they can result in higher accuracy predictions.
+As a follow-up independent study, I demonstrated, under ideal conditions, that the Membership Inference \cite{shokri2017membership} methodology could be used to detect whether a given model was trained on a sparse distribution or not, thereby serving as a potential indicator of covert channel(s) \cite{stein2018covert}.
 

### 2018-10-24 main.tex
181024-1943 sync

@@
 The GAN framework involves the simultaneous training of both a generative model $G$ with parameters $\theta_g$ and a discriminative model $D$ with parameters $\theta_d$.
-Given a target data space $T$ with distribution $p_t$, $G$'s task is to learn a mapping from an noisy input $z$ to the data space $p_t$ ($G:Z \rightarrow G(z) \sim p_g \approx p_t$); while $D$'s task is to discriminate whether some input $x$ originated from either $T$ or $G(z)$ ($D: X \rightarrow \{0, 1\}$).
-Since $G$ would like to generate samples that are representative of $T$, and therefore make $D$'s task more difficult by capturing $p_t$ with $p_g$, $D$ includes a loss term that is based on whether $D$ successfully discriminated inputs as $T$ or $G(z)$ and back-propagates $D$'s success as error to $\theta_g$ appropriately.
+Given a target data space $T$ with distribution $p_t$, $G$'s task is to learn a mapping from a noisy input $z$ to the data space $T$ ($G:Z \rightarrow G(z) \sim p_g \approx p_t$); while $D$'s task is to discriminate whether some input $x$ originated from either $T$ or $G(z)$ ($D: X \rightarrow \{0, 1\}$).
+Since $G$ would like to generate samples that are representative of $T$, and therefore make $D$'s task more difficult by capturing $p_t$ with $p_g$, $G$ includes a loss term that is based on whether $D$ successfully discriminated inputs as $T$ or $G(z)$ and back-propagates $D$'s success as error to $\theta_g$ appropriately.
 In this way, $G$ learns how to generate data that is hard for an ever-improving $D$ to distinguish.
@@
 This advancement seems obvious in retrospect, but warrants contemplation.
-While $G$ represents a mapping that captures the distribution of the target data space $T$ in general, the noisy input $z$ can be thought of as the diversity component of the input and the input $y$ can be thought of as the specification component of the input.
-Since the $y$ specification value is consistent within modes, a smaller portion of the overall networks for $G$ and $D$ can focus on the generation and discrimination, respectively, of the modal contribution whereas the remaining portions of the networks can focus on the unspecified contributions from $z$.
+While $G$ represents a mapping that captures the distribution of the target data space $T$ in general, the noisy input $z$ can be regarded as the diversity component of the input and $y$ can regarded as the specification component of the input.
+Since $y$ is constant within a given mode, a smaller portion of the overall networks for $G$ and $D$ can focus on the generation and discrimination, respectively, of the modal contribution whereas the remaining portions of the networks can focus on the unspecified contributions from $z$.
 Of course, Conditional GAN requires that conditioning data (eg.
-labels) is available for all target samples $x \in X$.
+labels) are available for all target samples $x \in X$.
 
-%(TBD) LAPGAN
-
-As the appreciation and use of GAN-based approaches increased, it was observed that the structure design of the $G$ and $D$ models deserved reconsideration.
 Radford, Metz, and Chintala \cite{DBLP:journals/corr/RadfordMC15} evaluated various model architectures and recommended a set of guidelines and constraints for $G$ and $D$ architectures that appeared to perform well in terms of stability, convergence, resolution, and depth of representation.
@@
 In addition to performance improvements, the authors demonstrate that $D$ is able to learn semantically-relevant features that are both explainable (via latent-space walk based validation) and usable (via feature-based supervised classification).
-In fact, the extracted features from $D$ were so informative that they were able to discover associations between the input noise vector $Z$ and some features, and even modify the generated images (such as face rotation) using simple interpolation along the correct axis in $Z$-space.
-
-%The strong semantic learning power offered by DCGAN made it a strong candidate for enabling full unsupervised learning tasks such that the features learned by $D$ can be later explained and applied for re-use.
-
-%In Machine Learning, and especially Computer Vision, domain variance between training and test data has been a long standing challenge.
-The domain of a data set can be explained as components of the distribution that are invariant to class label, but unique to the data set as a whole, in contrast to other similar data sets.
-Ganin et al.
-recognized that an adversarial approach known as a Domain Adversarial Neural Network (DANN) could be used to help improve Domain Adaption (DA).
-Their adversarial approach essentially minimizes the loss with respect to the basic classification task, and maximizes the loss with respect to an adversarial model trying to distinguish between the training domain and a test domain.
+In one case they were able to discover a linear vector in $Z$-space that could control rotation of faces.
 
 \subsubsection{GAN in Semantic Learning and Generation}
-Within the theme of semantic learning and generation, several significant contributions were made during 2016 and 2017.
-Pathak, Krahenbuhl, Donahue, Darrell, and Efros used an adversarial approach to inpaint (fill in missing portions of) images with large extents of missing pixel data \cite{pathakCVPR16context}.
+
+Within the theme of semantic learning and generation, Pathak, Krahenbuhl, Donahue, Darrell, and Efros used an adversarial approach to inpaint (fill in missing portions of) images with large extents of missing pixel data \cite{pathakCVPR16context}.
 Their architecture is based on an auto-encoder \cite{hinton2006reducing} combined with GAN.
 In their approach, the encoder-decoder learns a semantic encoding based on the non-missing portions and attempts to reproduce a whole image based on that semantic encoding.
-The `reconstruction loss' is defined in terms of distance ($L2$) from the filled area to the removed area.
+The `reconstruction loss' is defined in terms of distance ($L_2$) from the filled area to the removed area.
 The encoder-decoder is regarded as $G$, and a new discriminator $D$ is trained to predict whether its input is a reconstructed output from $G$ or an original image.
@@
 
-Wu et al.
+%Wu et al.
 took a somewhat similar approach and extended a variational autoencoder \cite{larsen2015autoencoding} within a 3-dimensional GAN architecture (3D-VAE-GAN) which learns to generate 3D images from 2D examples \cite{NIPS2016_6096}.
@@
 Their approach uses a Convolutional Neural Network (CNN) based generative model $G$ to learn a class label for each pixel in the image based on a ground-truth labeling (supervised), thereby creating a segmentation map of the input image.
-The discriminative model $D$ takes a segmentation map and tries to classify it as being a ground truth map or a generated map.
+$D$ then takes a segmentation map and tries to classify it as being a ground truth map or a generated map.
 %While $D$'s objective function tries to maximize classification accuracy of the input segmentation map as being either ground truth or generated, $G$'s objective function tries to maximize pixel class labeling accuracy with respect to ground truth while also minimizing accuracy.
 
-Ehsani, Mottaghi, and Farhadi used a GAN approach to address the related dual purpose task of semantically segmenting objects within an image and completing those objects that were occluded by other objects in the image \cite{ehsani2017segan}.
+%Ehsani, Mottaghi, and Farhadi used a GAN approach to address the related dual purpose task of semantically segmenting objects within an image and completing those objects that were occluded by other objects in the image \cite{ehsani2017segan}.
 
@@
 Essentially, they create $G$ by concatenating the input noise with an encoding of the text description and feed the result into a feed-forward CNN using fractional-striding\footnote{The paper claims this to be a de-convolutional network, but I presume it to mean a CNN using fractional striding to achieve up-sampling.}.
-They create $D$ by downsampling with CNN layers until a 4x4 representation is achieved, concatenate the text encoding onto each of the 16 frames, and continue convolutions to achieve the binary discriminator output.
+$D$ then down-samples with CNN layers until a 4x4 representation is achieved, concatenates the text encoding onto each of the 16 frames, and continues convolutions to achieve the binary discriminator output.
 The learning objectives are consistent with \cite{NIPS2014_5423}.
@@
 Their architecture follows a U-Net auto-encoder architecture \cite{ronneberger2015u} with DCGAN constraints \cite{salimans2016improved}.
-Further, similar to \cite{pathakCVPR16context}, they constrain $D$'s task to focusing on high-frequency semantic accuracy using learned image patch losses that are fed from skipped connections from encoder layers (PatchGAN), while leveraging an $L1$ term in $G$'s loss to ensure low-frequency accuracy.
+Further, similar to \cite{pathakCVPR16context}, they constrain $D$'s task to focusing on high-frequency semantic accuracy using learned image patch losses that are fed from skipped connections from encoder layers (PatchGAN), while leveraging an $L_1$ term in $G$'s loss to ensure low-frequency accuracy.
 A later paper from Liu, Breuel, and Kautz also displays impressive results \cite{NIPS2017_6672} in image translation.
@@
 
-Bansal et al.
+%Bansal et al.
 tackle the task of video-to-video translation and achieve fairly convincing results with a hierarchical GAN approach (RecycleGAN) \cite{bansal2018recyclegan}.
@@
 Indeed, most of the papers involving GAN implementations for computer vision applications rely on the ability of generated image/video to fool a human observer as an important qualitative measure of success.
-The use of deep learning as a tool for generating convincing, non-authentic video was put in the public eye when a toolset developed by deepfake \cite{deepfakes-faceswap} was used maliciously to produce pornographic video content with public personalities, and later by celebrity Jordan Peele as a public service announcement, warning of the dangers of the technology \cite{peeleDeepFakes}.
+The use of deep learning as a tool for generating convincing, non-authentic video was put in the public eye when a tool set developed by deepfake \cite{deepfakes-faceswap} was used maliciously to produce pornographic video content with public personalities, and later by celebrity Jordan Peele as a public service announcement, warning of the dangers of misuse of the technology \cite{peeleDeepFakes}.
 
@@
 This section will explore the broader topic of Adversarial Machine Learning (ML) and it's role in security and privacy.
-Back in 2008, Barreno, Nelson, Joseph, and Tygar discussed some security implications of ML and proposed a taxonomy to identify and analyze different attacks \cite{barreno2010security}.
+In 2008, Barreno, Nelson, Joseph, and Tygar discussed some security implications of ML and proposed a taxonomy to identify and analyze different attacks \cite{barreno2010security}.
 In the proposed taxonomy, an attack was to be identified in terms of \textit{Influence}, \textit{Security Violation}, and \textit{Specificity}.
@@
 When ML methods are used as a tool in an adversarial context like security, robustness of the model to adversarial examples becomes a primary design objective.
-Spam filtering and intrusion detection systems (IDS) are two classic examples wherein the design objective is to accurate classify messages or traffic as benign or malicious, while the attacker's objective is to design his/her malicious content to be classified as benign.
+Intrusion detection systems (IDS) and Spam filters are two classic examples wherein the design objective is to accurate classify messages or traffic as benign or malicious, while the attacker's objective is to design his/her malicious content to be classified as benign (i.e.
+evade detection).
 %This scenario depicts an \textit{exploratory}, \textit{integrity}, and \textit{specific} attack.
 
-Spam filtering has a relatively rich history from an adversarial learning perspective.
-Word-based Bayesian spam filters date back to 1998 \cite{sahami1998bayesian}, and the first `Bayes vs.
+Russu, Demontis, Biggio, Fumera, and Roli explore a formalization of the classification evasion problem and some principled reasoning toward the design of robustness in the context of linear models and kernels \cite{russu2016secure}.
+Essentially, they formalize the attack paradigm as trying to transform a malicious input $x$ into a still-malicious input $x^\prime$ such that $x^\prime$ will be classified as benign and the distance (norm) between $x^\prime$ and $x$ is minimized.
+They discuss different paradigms for sparse attacks, where the attacker would like to modify as few features as possible (eg.
+Spam), and dense attacks, where the attacker can make minimal modifications to a large number of features (such as image perturbations).
+The paper provides proof, under the assumption that an attacker will try to minimize their $L_1$ distance (from $x^\prime$ to $x$) for sparse attacks and $L_2$ distance for dense attacks, that classifiers should use $L_{\infty}$ regularization and $L_2$ regularization on the parameters, respectively.
+They extend the proof to non-linear kernels, resulting in recommendations of Laplacian kernels for sparse attacks, and Radial Basis Function for dense.
+They did not extend their work to highly complex adversarial distance functions (like those from GANs), but it nevertheless provides a principled basis for the nature of the adversarial learning task.
+
+Spam filtering is a very real example of this problem, and has a relatively rich history from an adversarial learning perspective.
+Word-based Bayesian Spam filters date back to 1998 \cite{sahami1998bayesian}, and the first `Bayes vs.
 Bayes' attack surfaced (in the public eye) in 2004 \cite{graham2004beat}.
+These early examples demonstrated the concept, but were not human-spoofable.
+Generating adversarial perturbations that (a) retain the same general Spam content, (b) are not detectable by humans, and (c) result in classification evasion, has proved to be a difficult challenge.
+Recently, Iyyer, Wieting, Gimpel, and Zettlemoyer developed an approach known as Syntactically Controlled Paraphrase Networks \cite{DBLP:journals/corr/abs-1804-06059}.
+First, they implemented a parse generator which takes, as input, a sentence $s_1$, a full parse $p_1$ of $s_1$ (obtained using \cite{manning2014stanford}), and a target template $t_2$, and provides a target parse $p_2$.
+Next, they implemented a paraphrase generator which takes, as input, $s_1$ and $p_2$, and generates a paraphrase sentence $s_2$.
+They show that randomly generated paraphrase samples (not adversarially learned) result in a relatively high number of miss-classifications, and also that augmenting those classification models at train-time with these syntactically-diverse adversarial examples results in better performance in the face of new adversarial examples.
 
@@
 Shokri, Stronati, Song, and Shmatikov have demonstrated a methodology to determine whether a specific sample was included within the training set for a target model \cite{shokri2017membership}.
-Given a specific data sample $x_?$ and a target model $M_t$, their paper shows how to create multiple `shadow' models with different (real or synthesized) data from that used to train $M_t$, obtain prediction responses from each of the shadow models that can be labeled as `in' or `out' depending on whether each sample was part of that shadow models training set, and then use the labeled response data to then train an attack model $M_a$.
-$M_a$ achieves good performance in predicting whether black-box responses from $M_t$ are associated with training input (`in') or test input (`out').
-The paper addresses the problem of simply assuming that higher-confidence responses corresponds to `in' data by the following methodology: (a) start running randomized data samples through $m_t$, (b) use a hill-climbing algorithm to search the sample space until a high-confidence response is obtained, (c) repeat a and b until asufficiently large set of data is obtained, then (d) sample from the obtained samples and split into training and test sets for each shadow model.
-In this way, the distributions for `in' and `out' synthesized data are the same from a confidence perspective.
+Given a specific data sample $x_?$ and a target model $M_t$, their paper shows how to create multiple `shadow' models with different (real or synthesized) data from that used to train $M_t$, obtain prediction responses from each of the shadow models that can be labeled as `in' or `out' depending on whether each sample was part of that shadow models training set, and then use the labeled response data to train an attack model $M_a$.
+After training, $M_a$ achieves good performance in predicting whether black-box responses from $M_t$ are associated with training input (`in') or test input (`out').
+The paper addresses the issue of bias from confidence in the responses by synthesizing the data and sampling from the subset of that data that results in similar confidence distributions from $M_t$.
 
 Similarly, Song, Ristenpart, and Shmatikov demonstrated several white- and black-box approaches to build covert channels into ML models for the purpose of extracting sensitive/proprietary training data.
-In the attack scenario, the model provider and the training data owner may be different entities.
-The data owner may not wish to reveal the training data to the model developer, but instead submit the data to a secure ML service broker that then trains and delivers the working model back to the data owner.
-White-box methods generally include methods for steganographically encoding training data into model parametrics that would then be available for extraction by an attacker with white-box access to the completed model.
-More interestingly, a `capacity abuse' attack was described that addresses the black-box scenario.
-Essentially, the learning algorithm automatically augments the training data with chosen (by the attacker) data and labels that data with an encoding of the private training data.
-The model learns to classify the original and malicious training data according to their labels, and then the attacker, having balack-box access to the model, can present the known malicious data to the model and the responses will be an encoding of the private training data.
-I hypothesize that performance of the model for both it's benign purpose as well as it's covert purpose depends on the malicious data being sparse.
-The intuition is that since sparse data is statistically distant from the private training data, it will not have adverse affects on classification; and since each sparse sample is distant from other sparse samples, they can result in higher accuracy predictions.
-As a follow-up independent study, I demonstrated, under ideal conditions, that the Membership Inference \cite{shokri2017membership} methodology could be used to detect whether a given model was trained on a sparse distribution or not, thereby serving as a potential indicator of covert channel(s) \cite{stein2018covert}.
+In the attack scenario, the data owner may not wish to reveal the training data to the model developer, but rather have the model trained and delivered by a secure third part.
+White-box methods were presented but, more interestingly, a `capacity abuse' attack was described that addresses the black-box scenario.
+Essentially, the learning algorithm automatically augments the training data with chosen (by the attacker) data that is labeled with an encoding of the private training data.
+The model learns on both the original and malicious training data, and then the attacker, having black-box access to the model, can present the known malicious data to the model and receive the encoded private data via the responses.
+I hypothesized that performance of the model for both its benign purpose as well as its covert purpose depends on the malicious data being sparse.
+The intuition is that since sparse data is distant  from the private training data (in the feature space), it will not have adverse affects on classification; and since each sparse sample is distant from other sparse samples, they are mutually separable resulting in higher accuracy predictions.
+As a follow-up independent study, I demonstrated under ideal conditions that the Membership Inference \cite{shokri2017membership} methodology could be used to detect whether a given model was trained on a sparse distribution or not, thereby serving as a potential indicator of covert channel(s) \cite{stein2018covert}.
+
+\subsubsection{Remaining Challenges and Opportunities}
+
+ML robustness in the face of adversarial examples is likely to continue being an open problem for some time.
+The challenge is not a transient implementation issue that will be worked out as ML technology matures, but rather, like generalization, a core learning issue at the heart of the gap between ML and true intelligence.
+In this vein, I feel that a deeper understanding of complex models, specifically deep learning networks, including properties, topographies, and structures, and their effects in adversarial scenarios.
+For example, characterizing parameter evolution in the manifold during training on benign data sets vs data sets with adversarial augmentation may reveal indicators that could be measured directly on future models (in the absence of adversarial data).
+Further, the formalized approach discussed in \cite{russu2016secure} could be extended to more complex networks.
+Since knowledge of the adversary's perturbation minimization objective helped inform provable design guidelines for the classifier, perhaps a finer-grained approach to adversarial learning (eg.
+segmentation of large generative and classification networks with complex tasks into matched ensembles of smaller networks with simpler tasks) will help increase model robustness to adversarial examples.
 

### 2018-11-05 main.tex
181105-0723 sync

@@
 \usepackage{appendix}
+\usepackage{amsmath}
+\usepackage{tikz}
+\usetikzlibrary{positioning, shapes, arrows}
+\usepackage[ruled]{algorithm2e}
 
@@
 
+\newpage
+\section{Privacy Prediction}
+
+Two privacy prediction tasks have been identified: (1) predict whether each pixel in a given image is part of a private object, and (2) predict whether or not a given image has private content\footnote{The tasks are addressed in reverse order from how they were presented in \ref{app:crandall} because, in the approach described herein, the solution for (2) can be largely derived from the solution for (1).}.
+For both tasks, assume that $N$ images are sampled, each with $M$ pixels having $H$ channels defined as $I : {\lbrack0,256\rbrack}^{M \times H} \rightarrow \{i_{mh} | \forall m \in \lbrack0,M), \forall h \in \lbrack0,H) \}$.
+Also, the following forms of evidence are available for each image $I_n$ sampled from $I$:
+
+\begin{table}[h]
+\centering
+\caption{Evidence}
+{\renewcommand{\arraystretch}{2}
+\begin{tabular}{p{7cm}|l}
+    \textbf{Evidence} & \textbf{Formalized as} \\
+    %\hline \\
+    Prediction vector of a categorical scene type $S$ (eg. natural, man-made, bathroom, etc.)
+        & $\{p(S_n = s \mid I_n) \mid \forall s \in dom(S) \}$\\
+    %\hline \\
+    Structured meta-data $D$ (GPS coordinates, timestamp, etc.)
+        & $D_n$\\
+    %\hline \\
+    Posterior probability of image if a pixel lies on some categorical object type $O$
+        & $\{p(I_n \mid i_{nm}\textrm{ on }o) \mid \forall m \in \lbrack0,M), \forall o \in dom(O) \}$\\
+    %\hline \\
+    A set $t$, where $dom(t)=\lbrack0,M)$, of coordinates for human-selected pixels that are part of private objects
+        & $t_n \subseteq \lbrack0,M)$
+\end{tabular}}
+\end{table}
+
+Let's define some assumptions.
+
+\paragraph{Evidence Assumptions:}
+\begin{enumerate}
+    \item The distribution for $I$ is equivalent to the distribution upon which the scene estimator providing $p(S_n = s \mid I_n)$ was trained.
+    \item The distribution for $I$ is equivalent to the distribution upon which the object estimator providing $p(I_n \mid i_{nm} \textrm{ on } o)$ was trained.
+    \item Both tasks are semi-supervised, such that the only form of labeled data is $t_n$\footnote{The semi-supervised assumption was adopted because human point-and-click labeling of private image regions for both training and updating of models is thought to be a realistic scenario.}.
+Note that this assumption implies that $t_n$ will not be available at test time.
+\end{enumerate}
+
+In the following sections, the $n$ subscript will usually be omitted for simplicity when considering a single image.
+
+\subsection{Task 1: Which Pixels Represent Private Data?}
+
+The objective for this task can be stated as estimating $\{\forall m \in \lbrack0,M), P(C_{P_{m}} = 1 \mid I, \textrm{ evidence}) \}$ where $C_{P_m}$ is the binary class label for whether a pixel $i_m$ is part of a benign object/region (0) or part of a private object/region (1).
+This set or probabilities can be regarded as a `privacy map' of $I$ for use in sensorship, for example.
+
+\subsubsection{Probabalistic Model Formulation}
+
+A simplistic approach might involve using the evidence $p(I \mid i_{m} \textrm{ on } o)$ directly but, since it is noisy evidence, false positives are likely.
+One would like to incorporate any and all useful evidence into the determination, and also reduce noise to achieve congruency for pixel-on-object predictions within object areas to reduce false positives.
+Therefore, a probabilistic model is needed.
+
+When defining a probabilistic model, it is important to define the central assumptions behind what the model is attempting to capture in terms of privacy.
+For this approach, the following assumptions are used: (a) certain objects, or object details, are private, and (b) whether an image is private is determined by whether it contains private objects or object details.
+Notice that certain contextual privacy scenarios, such as an image of a car known to belong to `Bob' being parked outside of a liquor store during work hours, for example, may not necessarily be marked private, even though `Bob' might argue that it is.
+
+Under these assumptions, we can model the objective task as a marginal (over all object types) of the joint probability that $i_m$ belongs to a certain object type $o$ and that object type $o$ is a private type.
+More formally, $\forall m \in \lbrack0,M)$:
+
+\begin{align}\label{eqn:task1.0}
+    p(C_{P_{m}} = 1 \mid I, \textrm{ evidence}) = \sum_{o \in O} \Big( p(i_m \textrm{ on } o, o \textrm{ is private} \mid I, \textrm{ evidence}) \Big)
+    %p(C_I = 1 \mid \textrm{evidence}) = 1 - \prod_{k \in K} \Big( 1 - p(\textrm{image region for $o_k$ is private} \mid \textrm{evidence}) \Big)
+\end{align}
+
+\subsubsection{Basis of the Model}
+
+To form the basis of the probabilistic model, a belief network which juxtaposes our evidence and other random variables into a reasonable set of dependency assumptions can be created.
+The dependency assumptions are defined as follows.
+Firstly, an image ($I$) is the result of, and therefore depends on, the true object-to-pixel mapping ($i_m$ on $o_k$) that is captured by the camera.
+Secondly, the scene type ($S$) prediction is based on, and therefore depends on, the image ($I$).
+Thirdly, whether or not the object $o_k$ is private is dependent on $t$, but since $t$ is not available at test time $o_k$'s privacy will be regarded as independent from any observable variable.
+Lastly, we'll assume for now that image metadata ($D$) in its raw form doesn't meaningfully contribute to the privacy determination problem.
+GPS location and time stamps may be useful for influencing the scene probabilities or accounting for lighting differences in the image for object recognition, but we do not have direct influence over those predictors.
+The resulting belief network is shown in Figure \ref{fig:belief}.
+
+%Thirdly, whether or not a given pixel is marked as being private ($m \in t$) is going to depend on (a) which object type the pixel is on ($i_m$ on $o_k$), and (b) whether that object type is private ($o_k$ is private).
+However, since the number of pixels marked per private object is small (likely $\leq 1$), Bayesian inference at the pixel level involving $p(m \in t)$ will result in a weak predictor of $p(o_k \textrm{ is private})$.
+A better approach is to use the knowledge that is given in the problem statement about how $t$ is generated, and perform inference at the object level.
+Therefore, we will represent the event that there exists at least one pixel that is marked as private as follows $(\exists i_l \textrm{ on } o_k \textrm{ s.t.
+} l \in t)$.
+This event is dependent on whether or not the object type is private, but we assume independence from whether or not a specific pixel is marked as private.
+This approach is discussed more in the following sections.
+
+\begin{figure}[ht]
+\centering
+    \begin{tikzpicture}[
+        every node/.style={shape=ellipse, draw}
+    ]
+    %NODES
+    %\node (mt) {$m \in t$};
+    \node (io) {$i_{m}\textrm{ on }o_k$};
+    \node (i)  [right = of io] {$I$};
+    \node (s) [right = of i] {$S$};
+    %\node (t) [below = of po, text width=2cm, align=center] {$i_m \textrm{ on } o_k, \exists i_l \textrm{ on } o_k$ s.t. $l \in t$};
+    %\node (op) [below = of io] {$\exists i_l \textrm{ on } o_k \textrm{ s.t. } l \in t$};
+    \node (or) [below = of i] {$o_k$ is private};
+    \node (d) [right = of s] {$D$};
+
+    %PATHS
+    \path   (io) edge [->] (i);
+    \path   (i) edge [->] (s);
+    %\path   (po) edge [->] (t);
+    %\path   (po) edge [->] node [left, draw=none] {$f_t$} (t);
+    %\path   (or) edge [->] (op);
+    \end{tikzpicture}
+    \caption{Belief Network}\label{fig:belief}
+    %\medskip
+    %\small
+\end{figure}
+
+Now we are ready to perform some Bayesian inference to obtain $p(i_m \textrm{ on } o, o \textrm{ is private} \mid I, \textrm{ evidence})$.
+
+\begin{align}
+    & p(i_m \textrm{ on } o_k, o_k \textrm{ is priv} \mid I, \textrm{ evidence})     && = p(i_m \textrm{ on } o_k, o_k \textrm{ is priv} \mid I, S, D) \notag\\
+    & \quad && = \frac
+        {p(S, I, i_m \textrm{ on } o_k, o_k \textrm{ is priv}, D)}
+        {p(S, I, D)} \notag\\
+    & \quad && = \frac
+        {p(S \mid I) p(I \mid i_m \textrm{ on } o_k) p(i_m \textrm{ on } o_k) p(o_k \textrm{ is priv}) p(D)}
+        {p(S \mid I) p(I) p(D)} \notag\\
+    & \quad && = \frac
+        {p(I \mid i_m \textrm{ on } o_k) p(i_m \textrm{ on } o_k) p(o_k \textrm{ is priv})}
+        {p(I)}\\
+    & p(i_m \textrm{ on } o_k) 
+        && = \sum\limits_{n \in \lbrack0,N)} \Big( p(I_n \mid i_m \textrm{ on } o_k) p(i_m \textrm{ on } o_k) \Big) \label{eqn:joint}\\
+    & p(I)
+        && = \sum\limits_{o_k \in O} \Big( p(I \mid i_m \textrm{ on } o_k) p(i_m \textrm{ on } o_k) \Big)%
+\end{align}%
+
+\subsubsection{Algorithm}
+
+We will consider all the factors in Equation \ref{eqn:joint}.
+$p(I \mid i_m \textrm{ on } o_k)$ is provided as noisy evidence, which has a high potential for resulting in a falsely high number of contiguous object regions.
+Aside from the affect of identifying only small portions of private objects, this will also adversely affect $p(o_k \textrm{ is priv})$ by driving up the number of unmarked private objects and potentially mislabeling those that are marked.
+Efficient belief propagation over pixels arranged as Markov Random Fields (MRF) has been shown to produce good results in scenarios where labels are unlikely to change between two neighboring pixels except in border regions \cite{felzenszwalb2006efficient}.
+We use a variant of this approach to help regularize the noisy $p(I \mid i_m \textrm{ on } o_k)$ evidence to achieve $p_R(I \mid i_m \textrm{ on } o_k)$, as shown in Algorithm \ref{alg:regObjMap}.
+
+\begin{algorithm}[h!]\label{alg:regObjMap}
+\caption{Regularize the Object Map}
+\KwData{
+    $I: \lbrack 0,256 \rbrack^{M \times H} \rightarrow \{i_{mh} \mid \forall m \in \lbrack 0, M ), \forall h \in \lbrack 0, H)\}$
+}
+\KwResult{$p_R(I \mid i_m \textrm{ on } o_k)$}
+
+$T \leftarrow$ small number // number of iterations for message updates\;
+$d \leftarrow$ small positive constant // penalty for labeling variation\;
+$D^{M \times K} \leftarrow -log(p(I \mid i_m \textrm{ on } o))$ /* get object probability map from noisy evidence and convert values to negative log */\;
+MSG$^t \leftarrow 0^{M \times \lvert m\textrm{'s neighbors} \rvert \times K}$\;
+MSG$^{t-1}$\;
+res $\leftarrow$ null$^M$\;
+
+\SetKwFunction{h}{h}
+\SetKwProg{Def}{def}{:}{}
+\Def{\h{x: coord}}{
+    res $\leftarrow 0^K$\;
+    \For{$k \leftarrow 0$ \KwTo $K$}{
+        \For{$q \in$ Neighbors($x$)}{
+            $res_k \leftarrow res_k +$ MSG$^{t-1}_{q,x,k}$\;
+        }
+        $res_k \leftarrow res_k + D_{x,k}$\;
+    }
+    return $res$
+}
+
+\SetKwFunction{Nei}{Neighbors}
+\Def{\Nei{x}}{return 8 pixel coordinates around x // could be 4\;}
+
+\For{$t \leftarrow 0$ \KwTo $T$}{
+    // Update messages $T$ times\;
+    MSG$^{t-1} \leftarrow$ MSG$^t$\;
+    \For{$m \leftarrow 0$ \KwTo $M$}{
+        // for each pixel coordinate\;
+        $h_m \leftarrow$ \h{$m$}\;
+        $h_{m_{min}} \leftarrow \min_{k} (h_m + d)$\;
+        $Q \leftarrow$ Neighbors($m$)\;
+        \For{$q \in Q$}{
+            $h_q \leftarrow$ \h{$q$}\;
+            \For{$k \leftarrow 0$ \KwTo $K$}{
+                MSG$^t_{m,q,k} \leftarrow \min (h_{q_{k}}, h_{m_{min}})$
+            }
+        }
+    }
+}
+\For{$m \leftarrow 0$ \KwTo $M$}{
+    res$_m \leftarrow$ $\operatorname{arg\,min}_k$ \h{$m$}\;
+}
+return res\;
+\end{algorithm}
+
+There are three factors in Equation \ref{eqn:joint} which are not supplied by the evidence or known, and need to be found.
+$p(o_k \textrm{ is priv})$ is the prior distribution for a given object type being private.
+This distribution will need to be learned through an enumeration of object types and instances in all the images and tallying which types, and how many, have been marked private in our training data.
+
+\begin{algorithm}
+\KwData{this text}
+\KwResult{this result}
+\caption{hey there}
+\end{algorithm}
+
+\subsection{Task 2: Does the Image Contain Private Data?}
+
+The objective for this task can be stated as estimating $p(C_I = 1 \mid I)$ where $C_I$ is the binary class label for whether the image is benign (0) or private (1).
+
+\subsubsection{Probabilistic Model Formulation}
+
+Assuming there are $K$ objects in an image, if even one object in the image is private, then the image contains private content.
+A simplistic approach might involve using the evidence $p(I_n \mid i_{nm} \textrm{ on } o)$ directly but, since it is noisy evidence, false positives are likely.
+One would like to incorporate any and all useful evidence into the determination, and also reduce noise to achieve congruency for pixel-on-object predictions within object areas to reduce false positives.
+Therefore, a probabilistic model is needed.
+
+When defining a probabilistic model, it is important to define the central assumptions behind what the model is attempting to capture in terms of privacy.
+For this approach, the following assumptions are used: (a) certain objects, or object details, are private, and (b) whether an image is private is determined by whether it contains private objects or object details.
+Notice that certain contextual privacy scenarios, such as an image of a car known to belong to `Bob' being parked outside of a liquor store during work hours, for example, may not necessarily be marked private, even though `Bob' might argue that it is.
+
+Under these assumptions, we can informally model the objective task as a set of Bernoulli trials over all objects in the image:
+
+\begin{align}\label{eqn:task2.0}
+    p(C_I = 1 \mid \textrm{evidence}) = 1 - \prod_{k \in K} \Big( 1 - p(\textrm{image region for $o_k$ is private} \mid \textrm{evidence}) \Big)
+\end{align}
+
+where $p(\textrm{image region for $o_k$ is private})$ is the marginal distribution of $o_k$ being object type $o$ and $o$ being a private type, as follows...
+
+\begin{align}\label{eqn:task2.1}
+    p(\textrm{image region for $o_k$ is private}) = \sum_{o \in O} p(o_k = o, o\textrm{ is private} \mid \textrm{evidence})    
+\end{align}
+
+\subsubsection{Basis of the Model}
+
+Now we are ready to infer the marginal probability distribution for $p(\textrm{image region for $o_k$ is private}) = \sum_{o \in O} p(o_k = o, o\textrm{ is private} \mid \textrm{evidence})$.
+First, solve for $p(o_k = o, o\textrm{ is private} \mid \textrm{evidence})$.
+
+\begin{align*}
+    p(o_k = o, o\textrm{ is private} \mid \textrm{evidence}) =
+    %p((i_m \textrm{ on } o_k, \exists i_l \textrm{ on } o_k \textrm{ s.t.
+} l \in t) \mid o_k \textrm{ is private}, i_m \textrm{ on } o_k) p(o_k \textrm{ is private}) p(i_m \textrm{ on } o_k) p(I \mid i_m \textrm{ on } o_k) p(S \mid I) p(D)
+    %p(I, S, D, i_{m} \textrm{ on } o,o \textrm{ is private}, t, Z_h) = & p(t \mid i_m \textrm{ on o}, o \textrm{ is private}, Z_h) \\
+        %& p(i_m \textrm{ on o} \mid S, I) p(S \mid I) p(I) p(Z_h) p(o \textrm{ is private}) p(D)\\
+\end{align*}
+
+\subsubsection{Algorithm}
+
+1. get joint prob map $p(i_{nm} \textrm{ on o}, S_n \mid I_n, t_n)$\\
+2. use MRF (Felzenszwalb) to reduce noise \& increase congruence \cite{felzenszwalb2006efficient}.  Record max $o$ for segmentation.\\
+3.
+use watershed segmentation to label objects (type \& instance) \cite{87344}.
+During algorithm, calculate average probability for each object instance (add all pixel probabilities and keep a queue pop counter for final division).
+Also build dataset of object types \& instances and if $t_n=1$ is encountered, mark that instance private.
+This will be used to build a distribution for $p(o \textrm{ is private}$\\
+4. use \#3 to get $p(o_k = o)$\\
+
 %%%%%%%%%%%%%%%%% BIB %%%%%%%%%%%%%%%%%%%%
@@
 
-\subsection{Dr. David Crandall}
+\subsection{Dr. David Crandall}\label{app:crandall}
 

### 2018-11-09 main.tex
181109-0918 sync

@@
 \usepackage[ruled]{algorithm2e}
+\usepackage{amssymb}
 
@@
         {p(I \mid i_m \textrm{ on } o_k) p(i_m \textrm{ on } o_k) p(o_k \textrm{ is priv})}
-        {p(I)}\\
+        {p(I)} \label{eqn:joint}\\
+    & \quad && = p(i_m \textrm{ on } o_k \mid I) p(o_k \textrm{ is priv}) \label{eqn:indep}\\
     & p(i_m \textrm{ on } o_k) 
-        && = \sum\limits_{n \in \lbrack0,N)} \Big( p(I_n \mid i_m \textrm{ on } o_k) p(i_m \textrm{ on } o_k) \Big) \label{eqn:joint}\\
+        && = \textrm{?}\\
     & p(I)
@@
 
-We will consider all the factors in Equation \ref{eqn:joint}.
-$p(I \mid i_m \textrm{ on } o_k)$ is provided as noisy evidence, which has a high potential for resulting in a falsely high number of contiguous object regions.
+Equation \ref{eqn:indep} illustrates independence of the events that a specific pixel belongs to a specific object and that object being a private object type.
+However, the prior distribution $p(i_m \textrm{ on } o_k \ I)$ is still unknown.
+We will consider the factors in Equation \ref{eqn:joint} to estimate it.
+Estimation of $p(i_m \textrm{ on } o_k \ I)$ can be regarded as a constrained Maximum a Priori (MAP) estimation follows:
+
+\begin{gather}
+	\theta : \Big(p(i_m \textrm{ on } o_0), p(i_m \textrm{ on } o_1),
+		\dots, p(i_m \textrm{ on } o_K)\Big) \notag\\
+	L(\theta) = \prod_{n=0}^{N} \sum_{k=0}^{K} p(I_n \mid i_{nm} 
+		\textrm{ on } o_k) \theta_k \notag\\
+	LL(\theta) = \sum_{n=0}^{n=N} \log \Big( \sum_{k=0}^{k=K} 
+		p(I_n \mid i_{nm} \textrm{ on } o_k) \theta_k \Big) \label{eqn:loglik}\\
+	\sum_{k=0}^{k=K} \theta_k - 1 = 0 \label{eqn:constr}\\
+	\mathcal{L}(\theta, \lambda) = \sum_{n=0}^{n=N} \log 
+		\Big( \sum_{k=0}^{k=K} p(I_n \mid i_{nm} \textrm{ on } o_k) \theta_k \Big) + \lambda - \lambda \sum_{k=0}^{k=K} \theta_k \label{eqn:lagr}\\
+	\mathbf{F}^{K \times 1}: \mathbf{F}_k = \frac{\partial \mathcal{L}}{\partial \theta_k} = 
+		\sum_{n=0}^{N} \frac
+			{p(I_n \mid i_{nm \textrm{ on } o_k})}
+			{\sum_{j=0}^{K} p(I_n \mid i_{nm \textrm{ on } o_j}) \theta_j} = 0;
+		\frac{\partial \mathcal{L}}{\partial \lambda} = 1 - \sum_{k=0}^{K} 
+			\theta_k = 0 \label{eqn:lagr1der}\\
+	\mathbf{J}^{K \times K}: \mathbf{J}_{i,j} = \frac{\partial}{\partial \theta_j} \left( \frac
+		{\partial \mathcal{L}}{\partial \theta_i} \right) = 
+		\sum_{n=0}^{N} \frac
+			{-p(I_n \mid i_{nm \textrm{ on } o_i})p(I_n \mid i_{nm \textrm{ on } o_j})}
+			{(\sum_{l=0}^{K} p(I_n \mid i_{nm \textrm{ on } o_l}) \theta_l)^2} \label{eqn:lagr2der};
+		\frac{\partial}{\partial \lambda} \left( \frac{\partial \mathcal{L}}{\partial \lambda} \right) = 0
+\end{gather}
+
+Equation \ref{eqn:loglik} represents the Log Likelihood of the observed data to be maximized over $\theta$, and Equation \ref{eqn:constr} represents the constraint that prior probabilities for all object types sum to 1.
+A Lagrangian (Equation \ref{eqn:lagr}) is constructed to define the boundary on the Log Likelihood along which the constraint is satisfied.
+The roots of the vector of first derivatives for this boundary (Equation \ref{eqn:lagr1der}) will contain the prior distribution that maximizes the posterior distribution.
+Equation \ref{eqn:lagr1der} is difficult to solve in closed form.
+Thus, we will use the Newton-Raphson root finding algorithm to estimate the roots in quadratic time.
+This algorithm also requires the second derivative of the Lagrangian, a Jacobian matrix, Equation \ref{eqn:lagr2der}.
+
+\begin{algorithm}[h]
+\caption{Newton-Raphson Root Finding Algorithm}
+	\KwData{
+		$X \leftarrow p(I_n \mid i_{nm} \textrm{ on } o_k)$; $\nabla_{\theta} \mathcal{L}$
+	}
+	\KwResult{$\theta_{root}: \{\theta_{0_{root}}, \theta_{1_{root}},
+		\dots, \theta_{K_{root}}\}$}
+	$\vec{\theta} \leftarrow 0^{M \times K}$\;
+	$\alpha \leftarrow$ small constant; $T \leftarrow$ small constant\;
+	\For{$m \leftarrow 0$ \KwTo $M$}{
+		\For{$n \leftarrow 0$ \KwTo $N$}{
+			$\vec{\theta_{m}} \leftarrow \vec{\theta_{m}} + 
+				\frac{\vec{p(I_n \mid i_{nm} \textrm{ on } o)}}{N \times \sum_{k=0}^{K} p(I_n \mid i_{nm} \textrm{ on } o_k)}$
+				// initial guess for $\theta_{m}$ assuming $I$ is uniform\;
+		}
+		Build $\mathbf{F}$ // complexity $O(NK)$\;
+		Build $\mathbf{J}^{-1}$ // complexity $O(NK+K^3)$\;
+		$t \leftarrow 0$\;
+		\While{$\mathbf{F}(\theta_{m}) > \alpha$ and $t < T$}{
+			$\vec{\theta_m} \leftarrow \vec{\theta_m} - 
+				\mathbf{J}(\vec{\theta_m})^{-1} \mathbf{F}(\vec{\theta_m})$
+				// complexity $O(NK)$\;
+			$t \leftarrow t + 1$\;
+		}
+	$\vec{\theta_m} \leftarrow \vec{\theta_m} / sum_{k=0}^{K} \theta_{mk}$
+		// satisfy constraint (i.e. solve for $\lambda$)\;
+	}
+\end{algorithm}
+
+Now that the prior $p(i_m \textrm{ on } o_k)$ has been estimated, we can calculate $p(i_m \textrm{ on } o_k \mid I)$ as follows:
+
+$$p(i_m \textrm{ on } o_k \mid I) = 
+	\frac{p(I \mid i_m \textrm{ on } o_k) p(i_m \textrm{ on } o_k)}{\sum_{j=0}^{K} p(I \mid i_m \textrm{ on } o_j) p(i_m \textrm{ on } o_j)}$$
+
+However, $p(i_m \textrm{ on } o_k \mid I)$ is still based on noisy evidence, which has a high potential for resulting in a falsely high number of contiguous object regions.
 Aside from the affect of identifying only small portions of private objects, this will also adversely affect $p(o_k \textrm{ is priv})$ by driving up the number of unmarked private objects and potentially mislabeling those that are marked.
 Efficient belief propagation over pixels arranged as Markov Random Fields (MRF) has been shown to produce good results in scenarios where labels are unlikely to change between two neighboring pixels except in border regions \cite{felzenszwalb2006efficient}.
-We use a variant of this approach to help regularize the noisy $p(I \mid i_m \textrm{ on } o_k)$ evidence to achieve $p_R(I \mid i_m \textrm{ on } o_k)$, as shown in Algorithm \ref{alg:regObjMap}.
+We use a variant of this approach to help regularize the noisy $p(i_m \textrm{ on } o_k \mid I)$ evidence to achieve $p_R(i_m \textrm{ on } o_k \mid I)$, as shown in Algorithm \ref{alg:regObjMap}.
 
@@
 }
-\KwResult{$p_R(I \mid i_m \textrm{ on } o_k)$}
+\KwResult{$p_R(i_m \textrm{ on } o_k \mid I)$}
 
@@
 $d \leftarrow$ small positive constant // penalty for labeling variation\;
-$D^{M \times K} \leftarrow -log(p(I \mid i_m \textrm{ on } o))$ /* get object probability map from noisy evidence and convert values to negative log */\;
+$D^{M \times K} \leftarrow -log(p(i_m \textrm{ on } o \mid I))$ /* get object probability map from noisy evidence and convert values to negative log */\;
 MSG$^t \leftarrow 0^{M \times \lvert m\textrm{'s neighbors} \rvert \times K}$\;
 MSG$^{t-1}$\;
-res $\leftarrow$ null$^M$\;
+PMAP $\leftarrow$ null$^{M \times K}$\;
 
@@
 \For{$m \leftarrow 0$ \KwTo $M$}{
-    res$_m \leftarrow$ $\operatorname{arg\,min}_k$ \h{$m$}\;
+    $p_R(i_m \textrm{ on } o \mid I) \leftarrow \frac
+	    {\vec{log^{-1}(-\textrm{\h{m}})}}
+	    {\sum_{k=0}^{K}log^{-1}(-\textrm{\h{m}})}$
+	    // conv. to probability and ensure normalized\;
 }
@@
 
-There are three factors in Equation \ref{eqn:joint} which are not supplied by the evidence or known, and need to be found.
-$p(o_k \textrm{ is priv})$ is the prior distribution for a given object type being private.
-This distribution will need to be learned through an enumeration of object types and instances in all the images and tallying which types, and how many, have been marked private in our training data.
+The final task is to estimate $p(o_k \textrm{ is priv})$.
+To achieve this, we need to enumerate all the object types and instances in the image set.
+One we have enumerated all the instances, we tally all those instances on which a human-marked pixel falls to build a prior distribution for $p(o_k \textrm{ is priv})$.
 
-\begin{algorithm}
-\KwData{this text}
-\KwResult{this result}
-\caption{hey there}
+\begin{algorithm}[h!]\label{alg:objEnum}
+	\KwData{$p_R(i_{nm} \textrm{ on } o_k \mid I_n) \forall n \in \lbrack0,N), \forall k \in \lbrack0,K), \forall m \in \lbrack0,M)$}
+	\KwResult{$E: \{(Obj, Prob, Priv)_0, \dots, (Obj, P_{type}, Priv)_L\}$ where Obj is the set of identified objects, Prob is the probability associated with the object type labeling, Priv is True if marked, False otherwise, and L is total number of objects}
+	\SetKwProg{Def}{def}{:}{}
+	\SetKwFunction{N}{Neighbors}
+	\Def{\N{x: coord}}{returns previously marked neighbors to $x$}
+	%$Q_{unseen}, Q_{cur}$ // declare empty queues\;
+	\For{$m \leftarrow 0$ \KwTo $M$}{
+		nbs $\leftarrow$ \N{m}\;
+		\uIf{}{}
+		\uElse{}
+		%$Q_{unseen} \leftarrow I_n$\;
+		%$Q_{cur}.\textrm{clear}$\;
+		%$Q_{cur}.\textrm{push}(Q_{unseen}.\textrm{pop}())$
+	}
+\caption{Object Enumeration}
 \end{algorithm}

### 2018-11-10 main.tex
181109-2039 first draft of Q2 complete

@@
 \usepackage{amsmath}
+\DeclareMathOperator*{\argmax}{arg\,max}
 \usepackage{tikz}
@@
 
-In the following sections, the $n$ subscript will usually be omitted for simplicity when considering a single image.
+In the following sections, the $n$ subscript will often be omitted for simplicity when considering a single image.
 
@@
 
-A simplistic approach might involve using the evidence $p(I \mid i_{m} \textrm{ on } o)$ directly but, since it is noisy evidence, false positives are likely.
 One would like to incorporate any and all useful evidence into the determination, and also reduce noise to achieve congruency for pixel-on-object predictions within object areas to reduce false positives.
@@
 
-Under these assumptions, we can model the objective task as a marginal (over all object types) of the joint probability that $i_m$ belongs to a certain object type $o$ and that object type $o$ is a private type.
+Under these assumptions, we can model the objective task as a marginal (over all $K$ object types) of the joint probability that $i_m$ belongs to a certain object type $o_k$ and that object type $o_k$ is a private type.
 More formally, $\forall m \in \lbrack0,M)$:
@@
 \begin{align}\label{eqn:task1.0}
-    p(C_{P_{m}} = 1 \mid I, \textrm{ evidence}) = \sum_{o \in O} \Big( p(i_m \textrm{ on } o, o \textrm{ is private} \mid I, \textrm{ evidence}) \Big)
-    %p(C_I = 1 \mid \textrm{evidence}) = 1 - \prod_{k \in K} \Big( 1 - p(\textrm{image region for $o_k$ is private} \mid \textrm{evidence}) \Big)
+    p(C_{P_{m}} = 1 \mid I, \textrm{ evidence}) = \sum_{k = 0}^{K} \Big( p(i_m \textrm{ on } o_k, o_k \textrm{ is private} \mid I, \textrm{ evidence}) \Big)
 \end{align}
@@
 
-To form the basis of the probabilistic model, a belief network which juxtaposes our evidence and other random variables into a reasonable set of dependency assumptions can be created.
+To form the basis of the probabilistic model, a belief network which juxtaposes our evidence and other random variables into the believed set of dependency assumptions can be created.
 The dependency assumptions are defined as follows.
@@
 The resulting belief network is shown in Figure \ref{fig:belief}.
-
-%Thirdly, whether or not a given pixel is marked as being private ($m \in t$) is going to depend on (a) which object type the pixel is on ($i_m$ on $o_k$), and (b) whether that object type is private ($o_k$ is private).
-However, since the number of pixels marked per private object is small (likely $\leq 1$), Bayesian inference at the pixel level involving $p(m \in t)$ will result in a weak predictor of $p(o_k \textrm{ is private})$.
-A better approach is to use the knowledge that is given in the problem statement about how $t$ is generated, and perform inference at the object level.
-Therefore, we will represent the event that there exists at least one pixel that is marked as private as follows $(\exists i_l \textrm{ on } o_k \textrm{ s.t.
-} l \in t)$.
-This event is dependent on whether or not the object type is private, but we assume independence from whether or not a specific pixel is marked as private.
-This approach is discussed more in the following sections.
 
@@
 \begin{align}
-    & p(i_m \textrm{ on } o_k, o_k \textrm{ is priv} \mid I, \textrm{ evidence})     && = p(i_m \textrm{ on } o_k, o_k \textrm{ is priv} \mid I, S, D) \notag\\
+    & p(i_m \textrm{ on } o, o \textrm{ is priv} \mid I, \textrm{ evidence})     && = p(i_m \textrm{ on } o, o \textrm{ is priv} \mid I, S, D) \notag\\
     & \quad && = \frac
-        {p(S, I, i_m \textrm{ on } o_k, o_k \textrm{ is priv}, D)}
+        {p(S, I, i_m \textrm{ on } o, o \textrm{ is priv}, D)}
         {p(S, I, D)} \notag\\
     & \quad && = \frac
-        {p(S \mid I) p(I \mid i_m \textrm{ on } o_k) p(i_m \textrm{ on } o_k) p(o_k \textrm{ is priv}) p(D)}
+        {p(S \mid I) p(I \mid i_m \textrm{ on } o) p(i_m \textrm{ on } o) p(o \textrm{ is priv}) p(D)}
         {p(S \mid I) p(I) p(D)} \notag\\
     & \quad && = \frac
-        {p(I \mid i_m \textrm{ on } o_k) p(i_m \textrm{ on } o_k) p(o_k \textrm{ is priv})}
+        {p(I \mid i_m \textrm{ on } o) p(i_m \textrm{ on } o) p(o \textrm{ is priv})}
         {p(I)} \label{eqn:joint}\\
-    & \quad && = p(i_m \textrm{ on } o_k \mid I) p(o_k \textrm{ is priv}) \label{eqn:indep}\\
-    & p(i_m \textrm{ on } o_k) 
+    & \quad && = p(i_m \textrm{ on } o \mid I) p(o \textrm{ is priv}) \label{eqn:indep}\\
+    & p(i_m \textrm{ on } o) 
         && = \textrm{?}\\
     & p(I)
-        && = \sum\limits_{o_k \in O} \Big( p(I \mid i_m \textrm{ on } o_k) p(i_m \textrm{ on } o_k) \Big)%
+        && = \sum_{k = 0}^{K} \Big( p(I \mid i_m \textrm{ on } o_k) p(i_m \textrm{ on } o_k) \Big)%
 \end{align}%
 
-\subsubsection{Algorithm}
+\subsubsection{Algorithm}\label{sec:t1alg}
+
+Time complexity for all algorithms are given in their respective captions.
+$N$ is number of images, $M$ is number of pixels per image, $K$ is number of object types, $L_n$ is number of objects in image $I_n$, $Q$ is the number of neighbors to pixels, and $T$ is number of iterations (for root finding).
 
 Equation \ref{eqn:indep} illustrates independence of the events that a specific pixel belongs to a specific object and that object being a private object type.
-However, the prior distribution $p(i_m \textrm{ on } o_k \ I)$ is still unknown.
+However, the prior distribution $p(i_m \textrm{ on } o \mid I)$ is still unknown.
 We will consider the factors in Equation \ref{eqn:joint} to estimate it.
-Estimation of $p(i_m \textrm{ on } o_k \ I)$ can be regarded as a constrained Maximum a Priori (MAP) estimation follows:
+Estimation of $p(i_m \textrm{ on } o \mid I)$ can be regarded as a constrained Maximum a Priori (MAP) estimation follows:
 
@@
 	\theta : \Big(p(i_m \textrm{ on } o_0), p(i_m \textrm{ on } o_1),
-		\dots, p(i_m \textrm{ on } o_K)\Big) \notag\\
+		\dots, p(i_m \textrm{ on } o_K)\Big) \\
 	L(\theta) = \prod_{n=0}^{N} \sum_{k=0}^{K} p(I_n \mid i_{nm} 
-		\textrm{ on } o_k) \theta_k \notag\\
-	LL(\theta) = \sum_{n=0}^{n=N} \log \Big( \sum_{k=0}^{k=K} 
+		\textrm{ on } o_k) \theta_k \\
+	LL(\theta) = \sum_{n=0}^{N} \log \Big( \sum_{k=0}^{K} 
 		p(I_n \mid i_{nm} \textrm{ on } o_k) \theta_k \Big) \label{eqn:loglik}\\
-	\sum_{k=0}^{k=K} \theta_k - 1 = 0 \label{eqn:constr}\\
-	\mathcal{L}(\theta, \lambda) = \sum_{n=0}^{n=N} \log 
-		\Big( \sum_{k=0}^{k=K} p(I_n \mid i_{nm} \textrm{ on } o_k) \theta_k \Big) + \lambda - \lambda \sum_{k=0}^{k=K} \theta_k \label{eqn:lagr}\\
+	\sum_{k=0}^{K} \theta_k - 1 = 0 \label{eqn:constr}\\
+	\mathcal{L}(\theta, \lambda) = \sum_{n=0}^{N} \log 
+		\Big( \sum_{k=0}^{K} p(I_n \mid i_{nm} \textrm{ on } o_k) \theta_k \Big) + \lambda - \lambda \sum_{k=0}^{K} \theta_k \label{eqn:lagr}\\
 	\mathbf{F}^{K \times 1}: \mathbf{F}_k = \frac{\partial \mathcal{L}}{\partial \theta_k} = 
@@
 Equation \ref{eqn:lagr1der} is difficult to solve in closed form.
-Thus, we will use the Newton-Raphson root finding algorithm to estimate the roots in quadratic time.
-This algorithm also requires the second derivative of the Lagrangian, a Jacobian matrix, Equation \ref{eqn:lagr2der}.
+Thus, we will use the Newton-Raphson root finding algorithm to estimate the roots in quadratic time, Algorithm \ref{alg:newtraph}.
+This algorithm also requires the second derivative of the Lagrangian, a Jacobian matrix, expressed in Equation \ref{eqn:lagr2der}.
 
-\begin{algorithm}[h]
-\caption{Newton-Raphson Root Finding Algorithm}
+\begin{algorithm}[h!]\label{alg:newtraph}
+\caption{Newton-Raphson Root Finding Algorithm, $O(M(NK + K^3 + TNK))$}
 	\KwData{
-		$X \leftarrow p(I_n \mid i_{nm} \textrm{ on } o_k)$; $\nabla_{\theta} \mathcal{L}$
-	}
-	\KwResult{$\theta_{root}: \{\theta_{0_{root}}, \theta_{1_{root}},
-		\dots, \theta_{K_{root}}\}$}
-	$\vec{\theta} \leftarrow 0^{M \times K}$\;
+		$X \leftarrow p(I_n \mid i_{nm} \textrm{ on } o_k)$}
+	\KwResult{$\theta^{M \times K}: \{\vec{\theta}_{0}^K, \dots, \vec{\theta}_{M}^K\}$}
+	$\theta \leftarrow 0^{M \times K}$\;
 	$\alpha \leftarrow$ small constant; $T \leftarrow$ small constant\;
@@
 		}
-		Build $\mathbf{F}$ // complexity $O(NK)$\;
-		Build $\mathbf{J}^{-1}$ // complexity $O(NK+K^3)$\;
+		Build $\mathbf{F}$ // Equation \ref{eqn:lagr1der} complexity $O(NK)$\;
+		Build $\mathbf{J}^{-1}$ // Inverse matrix of Equation \ref{eqn:lagr2der} complexity $O(NK+K^3)$\;
 		$t \leftarrow 0$\;
@@
 
-Now that the prior $p(i_m \textrm{ on } o_k)$ has been estimated, we can calculate $p(i_m \textrm{ on } o_k \mid I)$ as follows:
+Now that the prior $p(i_m \textrm{ on } o)$ has been estimated, we can calculate $p(i_m \textrm{ on } o \mid I)$ as follows:
 
-$$p(i_m \textrm{ on } o_k \mid I) = 
-	\frac{p(I \mid i_m \textrm{ on } o_k) p(i_m \textrm{ on } o_k)}{\sum_{j=0}^{K} p(I \mid i_m \textrm{ on } o_j) p(i_m \textrm{ on } o_j)}$$
+\begin{align}
+    p(i_m \textrm{ on } o \mid I) = 
+	    \frac{p(I \mid i_m \textrm{ on } o) p(i_m \textrm{ on } o)}{\sum_{j=0}^{K} p(I \mid i_m \textrm{ on } o_j) p(i_m \textrm{ on } o_j)}    
+\end{align}
 
-However, $p(i_m \textrm{ on } o_k \mid I)$ is still based on noisy evidence, which has a high potential for resulting in a falsely high number of contiguous object regions.
+However, $p(i_m \textrm{ on } o \mid I)$ is still based on noisy evidence, which has a high potential for resulting in a falsely high number of contiguous object regions.
 Aside from the affect of identifying only small portions of private objects, this will also adversely affect $p(o_k \textrm{ is priv})$ by driving up the number of unmarked private objects and potentially mislabeling those that are marked.
 Efficient belief propagation over pixels arranged as Markov Random Fields (MRF) has been shown to produce good results in scenarios where labels are unlikely to change between two neighboring pixels except in border regions \cite{felzenszwalb2006efficient}.
-We use a variant of this approach to help regularize the noisy $p(i_m \textrm{ on } o_k \mid I)$ evidence to achieve $p_R(i_m \textrm{ on } o_k \mid I)$, as shown in Algorithm \ref{alg:regObjMap}.
+We use a variant of this approach to help regularize the noisy $p(i_m \textrm{ on } o \mid I)$ evidence to achieve $p_R(i_m \textrm{ on } o \mid I)$, as shown in Algorithm \ref{alg:regObjMap}.
 
 \begin{algorithm}[h!]\label{alg:regObjMap}
-\caption{Regularize the Object Map}
+\caption{Regularize the Object Map, $O(NTMQ^2K)$}
 \KwData{
-    $I: \lbrack 0,256 \rbrack^{M \times H} \rightarrow \{i_{mh} \mid \forall m \in \lbrack 0, M ), \forall h \in \lbrack 0, H)\}$
-}
-\KwResult{$p_R(i_m \textrm{ on } o_k \mid I)$}
-
+    $I: \lbrack 0,256 \rbrack^{M \times H} \rightarrow \{i_{mh} \mid \forall m \in \lbrack 0, M ), \forall h \in \lbrack 0, H)\}$}
+\KwResult{$p_R(i_m \textrm{ on } o \mid I)$}
 $T \leftarrow$ small number // number of iterations for message updates\;
@@
 PMAP $\leftarrow$ null$^{M \times K}$\;
-
 \SetKwFunction{h}{h}
@@
 }
-
 \SetKwFunction{Nei}{Neighbors}
 \Def{\Nei{x}}{return 8 pixel coordinates around x // could be 4\;}
-
 \For{$t \leftarrow 0$ \KwTo $T$}{
@@
 
-The final task is to estimate $p(o_k \textrm{ is priv})$.
-To achieve this, we need to enumerate all the object types and instances in the image set.
-One we have enumerated all the instances, we tally all those instances on which a human-marked pixel falls to build a prior distribution for $p(o_k \textrm{ is priv})$.
+The final task is to estimate $p(o \textrm{ is priv})$.
+To achieve this, we need to enumerate all the object instances, their types, their respective probabilities, and whether they are private.
+To achieve this, we us a variant of the Two-Pass Connected Component Labeling algorithm \cite{hoshen1976percolation}, described in Algorithm \ref{alg:objEnum}.
+It is important to note that although objects are enumerated according to the high object type probabilities for each pixel, probability summations for all objects types are preserved at each pixel.
+This will be important later.
 
 \begin{algorithm}[h!]\label{alg:objEnum}
-	\KwData{$p_R(i_{nm} \textrm{ on } o_k \mid I_n) \forall n \in \lbrack0,N), \forall k \in \lbrack0,K), \forall m \in \lbrack0,M)$}
-	\KwResult{$E: \{(Obj, Prob, Priv)_0, \dots, (Obj, P_{type}, Priv)_L\}$ where Obj is the set of identified objects, Prob is the probability associated with the object type labeling, Priv is True if marked, False otherwise, and L is total number of objects}
+\caption{Object Enumeration, $O(NMQ + L)$}
+	\KwData{$P^{N \times M \times K} \leftarrow p_R(i_{nm} \textrm{ on } o_k \mid I_n) \forall n \in \lbrack 0,N), \forall m \in \lbrack0,M), \forall k \in \lbrack0,K)$}
+	\KwResult{$\Big( E^{N \times L_n}, Y^{N \times L_n} \Big)$ s.t.
+$E: (((\textrm{typ, } \vec{\textrm{psum}} \textrm{, cnt})_{00}, \dots, (\textrm{typ, } \vec{\textrm{psum}} \textrm{, cnt})_{0L_{0}}),$ \\ $((\textrm{typ, } \vec{\textrm{psum}} \textrm{, cnt})_{0N}, \dots, (\textrm{typ, } \vec{\textrm{psum}} \textrm{, cnt})_{NL_{N}})); Y: \{True, False\}^{N \times L_n}$}
+	/* where typ is the object type label, psum is the sum of probabilities for that type within the object region, cnt is the number of pixels in the object region, the training label $Y$ is True if object region is marked, False otherwise, and $L_n$ is total number of objects within image $I_n$ */\;
 	\SetKwProg{Def}{def}{:}{}
     \SetKwFunction{N}{Neighbors}
-	\Def{\N{x: coord}}{returns previously marked neighbors to $x$}
-	%$Q_{unseen}, Q_{cur}$ // declare empty queues\;
+    \Def{\N{x: coord}}{returns Queue of previously marked neighbors to $x$ in ascending order}
+	\For{$n \leftarrow 0$ \KwTo $N$}{
+    	$\vec{\lambda} \leftarrow \infty^{M}$\;
+    	$\lambda_{cur} \leftarrow -1$\;
+    	$\vec{\Pi} \leftarrow ()$ // each element points to its parent label\;
     	\For{$m \leftarrow 0$ \KwTo $M$}{
-		nbs $\leftarrow$ \N{m}\;
-		\uIf{}{}
-		\uElse{}
-		%$Q_{unseen} \leftarrow I_n$\;
-		%$Q_{cur}.\textrm{clear}$\;
-		%$Q_{cur}.\textrm{push}(Q_{unseen}.\textrm{pop}())$
+    		$Q \leftarrow$ \N{m}\;
+    		\While{$Q \neq \emptyset$}{
+    		    $b \leftarrow Q$.pop()\;
+    		    $k \leftarrow \argmax_k P_{nm}$\;
+    		    \uIf{$k == (\argmax_k P_{nb})$}{
+    		        \uIf{$\lambda_b > \lambda_m$}{
+    		            $\pi_b \leftarrow \lambda_m$ // record $\lambda_m$ as parent of $\lambda_b$\;
     		        }
-\caption{Object Enumeration}
+    		        \uElse{
+    		            $\lambda_m \leftarrow \lambda_b$ // adopt neighbor's label\;
+    		            $E_{n\lambda_{m}}$.psum $+= P_{nm}$\;
+    		            $E_{n\lambda_{m}}$.cnt $+= 1$\;
+    		            \uIf{$m \in t_n$}{$Y_{n\lambda_{m}} \leftarrow True$ // if this is training data}
+    		        }
+    		    }
+    		}
+    		\uIf{$\lambda_m == \infty$}{
+    		    $\lambda_{cur} \leftarrow \lambda_{cur} + 1$\;
+    		    k $\leftarrow \argmax_k P_{nm}$\;
+    		    $\lambda_m \leftarrow \lambda_{cur}$\;
+    		    $E_n$.append(typ, $P_{nm}$, 1)\;
+    		    $Y_n$.append($m \in t_n$) // if this is training data\;
+    		    $\Pi$.append($-1$)\;
+    	    }
+    	}
+    	\For{$l \leftarrow (\left| E_n \right| -1)$ \KwTo $0$}{
+    	    \uIf{$\pi_l \geq 0$}{
+    	        // collapse children into parents\;
+    	        $E_{n\pi_{l}} \leftarrow (E_{n\pi_{l}}\textrm{.type}, E_{n\pi_{l}}\textrm{.psum} + E_{nl}\textrm{.psum}, E_{n\pi_{l}}\textrm{.cnt} + E_{nl}\textrm{.cnt})$\;
+    	        $Y_{n\pi_{l}} \leftarrow Y_{n\pi_{l}} \textrm{ OR } Y_{nl}$\;
+    	        $E_n$.remove($l$); $Y_n$.remove($l$)\;
+    	    }
+    	}
+	}
+	return $E$\;
 \end{algorithm}
+
+Given the enumerated object data, we estimate $p(o \textrm{ is priv})$ as described in Algorithm \ref{alg:private}.
+Object privacy is tallied according to the type that was chosen as most probable for the image region as opposed to soft-attribution of privacy across the type probabilities.
+This is because the other types may well have resulted in very different object regions.
+
+\begin{algorithm}[h!]\label{alg:private}
+	\KwData{$E^{N}$ // the enumerated object vectors}
+	\KwResult{$P^K$ // probability distribution $p(o \textrm{ is priv})$}
+	$C: ((t_0, p_0), \dots, (t_K, p_K)) \leftarrow (0,0)^K$ // total and private counts for each object type\;
+	\For{$n \leftarrow 0$ \KwTo $N$}{
+	    \For{$l \leftarrow 0$ \KwTo $\left| E_n \right|$}{
+	        $k \leftarrow E_{nl}$.typ\;
+	        $C_{k}.t += 1$\;
+	        \uIf{$E_{nl}$.priv}{$C_{k}.p += 1$}
+	    }
+	}
+	\For{$k \leftarrow 0$ \KwTo $K$}{
+	    $P_k \leftarrow C_k.p / C_k.t$\;
+	}
+	return $P$\;
+\caption{Object Privacy Prior Estimate, $O(NL  K)$}
+\end{algorithm}
+
+Finally, all the information has been obtained to answer Task 1.  For any given pixel $i_{nm}$, we can estimate the probability of whether it lies on a private object as follows:
+
+\begin{align}
+    p(C_{P_{nm}} = 1 \mid I, \textrm{ evidence}) = \sum_{k=0}^{K} p(i_{nm} \textrm{ on } o_k \mid I_n) P_k
+\end{align}
 
@@
 
-The objective for this task can be stated as estimating $p(C_I = 1 \mid I)$ where $C_I$ is the binary class label for whether the image is benign (0) or private (1).
+The objective for this task can be stated as estimating $p(C_I = 1 \mid I, \textrm{ evidence})$ where $C_I$ is the binary class label for whether the image is benign (0) or private (1).
+The model formulation, basis, and algorithms for Task 1 are completely reused for Task 2.
+A simple extension to the knowledge we've built in Task 1 is applied in order to complete Task 2, as follows.
 
-\subsubsection{Probabilistic Model Formulation}
-
-Assuming there are $K$ objects in an image, if even one object in the image is private, then the image contains private content.
-A simplistic approach might involve using the evidence $p(I_n \mid i_{nm} \textrm{ on } o)$ directly but, since it is noisy evidence, false positives are likely.
-One would like to incorporate any and all useful evidence into the determination, and also reduce noise to achieve congruency for pixel-on-object predictions within object areas to reduce false positives.
-Therefore, a probabilistic model is needed.
-
-When defining a probabilistic model, it is important to define the central assumptions behind what the model is attempting to capture in terms of privacy.
-For this approach, the following assumptions are used: (a) certain objects, or object details, are private, and (b) whether an image is private is determined by whether it contains private objects or object details.
-Notice that certain contextual privacy scenarios, such as an image of a car known to belong to `Bob' being parked outside of a liquor store during work hours, for example, may not necessarily be marked private, even though `Bob' might argue that it is.
-
-Under these assumptions, we can informally model the objective task as a set of Bernoulli trials over all objects in the image:
+Assuming there are $L_n$ objects in an image $I_n$, if even one object in the image is private, then the image contains private content.
+We can model the objective task over a Binomial distribution.
+Specifically, choose $L_n$ successes given $L_n$ trials (objects) with the probability of success being $p_s = (1-p(o_l \textrm{ is priv} \mid I, \textrm{ evidence}))$.
+The resulting probability is the complement of our objective probability.
 
 \begin{align}\label{eqn:task2.0}
-    p(C_I = 1 \mid \textrm{evidence}) = 1 - \prod_{k \in K} \Big( 1 - p(\textrm{image region for $o_k$ is private} \mid \textrm{evidence}) \Big)
+    p(C_{P_{nm}} = 1 \mid I, \textrm{ evidence}) & = 1 - \prod_{l = 0}^{L_n} \left( 1 - p(\textrm{image region $E_{nl}$ 
+        is private} \mid \textrm{evidence}) \right)\\
+    & = 1 - \prod_{l = 0}^{L_n} \left( 1 - \sum_{k=0}^{K} (E_{nl}\textrm{.psum}_k / E_{nl}\textrm{.cnt}) p(o_k \textrm{ is priv}) \right) \label{eqn:objpriv}
 \end{align}
 
-where $p(\textrm{image region for $o_k$ is private})$ is the marginal distribution of $o_k$ being object type $o$ and $o$ being a private type, as follows...
-
-\begin{align}\label{eqn:task2.1}
-    p(\textrm{image region for $o_k$ is private}) = \sum_{o \in O} p(o_k = o, o\textrm{ is private} \mid \textrm{evidence})    
-\end{align}
-
-\subsubsection{Basis of the Model}
-
-Now we are ready to infer the marginal probability distribution for $p(\textrm{image region for $o_k$ is private}) = \sum_{o \in O} p(o_k = o, o\textrm{ is private} \mid \textrm{evidence})$.
-First, solve for $p(o_k = o, o\textrm{ is private} \mid \textrm{evidence})$.
-
-\begin{align*}
-    p(o_k = o, o\textrm{ is private} \mid \textrm{evidence}) =
-    %p((i_m \textrm{ on } o_k, \exists i_l \textrm{ on } o_k \textrm{ s.t.
-} l \in t) \mid o_k \textrm{ is private}, i_m \textrm{ on } o_k) p(o_k \textrm{ is private}) p(i_m \textrm{ on } o_k) p(I \mid i_m \textrm{ on } o_k) p(S \mid I) p(D)
-    %p(I, S, D, i_{m} \textrm{ on } o,o \textrm{ is private}, t, Z_h) = & p(t \mid i_m \textrm{ on o}, o \textrm{ is private}, Z_h) \\
-        %& p(i_m \textrm{ on o} \mid S, I) p(S \mid I) p(I) p(Z_h) p(o \textrm{ is private}) p(D)\\
-\end{align*}
-
-\subsubsection{Algorithm}
-
-1. get joint prob map $p(i_{nm} \textrm{ on o}, S_n \mid I_n, t_n)$\\
-2. use MRF (Felzenszwalb) to reduce noise \& increase congruence \cite{felzenszwalb2006efficient}.  Record max $o$ for segmentation.\\
-3.
-use watershed segmentation to label objects (type \& instance) \cite{87344}.
-During algorithm, calculate average probability for each object instance (add all pixel probabilities and keep a queue pop counter for final division).
-Also build dataset of object types \& instances and if $t_n=1$ is encountered, mark that instance private.
-This will be used to build a distribution for $p(o \textrm{ is private}$\\
-4. use \#3 to get $p(o_k = o)$\\
+Notice in Equation \ref{eqn:objpriv} that the probability of an object region being of some type $k$ is represented as the average of the $k$-probabilities across all pixels in that $l$ object region.
+As mentioned in section \ref{sec:t1alg}, the actual object regions would likely have been different for different values of $k$, but capturing average probabilities for each type across all pixels in the dominant region is presumed (by the author) to be a reasonable compromise.
+The independent joint probabilities for object classification and class privacy are marginalized over all object types to achieve the probability of the object region being private.
 

### 2018-11-14 main.tex
181114-0632 sync

@@
 
-\paragraph{Evidence Assumptions:}
 \begin{enumerate}
-    \item The distribution for $I$ is equivalent to the distribution upon which the scene estimator providing $p(S_n = s \mid I_n)$ was trained.
-    \item The distribution for $I$ is equivalent to the distribution upon which the object estimator providing $p(I_n \mid i_{nm} \textrm{ on } o)$ was trained.
     \item Both tasks are semi-supervised, such that the only form of labeled data is $t_n$\footnote{The semi-supervised assumption was adopted because human point-and-click labeling of private image regions for both training and updating of models is thought to be a realistic scenario.}.
@@
 
-The objective for this task can be stated as estimating $\{\forall m \in \lbrack0,M), P(C_{P_{m}} = 1 \mid I, \textrm{ evidence}) \}$ where $C_{P_m}$ is the binary class label for whether a pixel $i_m$ is part of a benign object/region (0) or part of a private object/region (1).
+The objective for this task can be stated as estimating $\{\forall m \in \lbrack0,M), p(C_{P_{m}} = 1 \mid I, \textrm{ evidence}) \}$ where $C_{P_m}$ is the binary class label for whether a pixel $i_m$ is part of a benign object/region (0) or part of a private object/region (1).
 This set or probabilities can be regarded as a `privacy map' of $I$ for use in sensorship, for example.
@@
 Secondly, the scene type ($S$) prediction is based on, and therefore depends on, the image ($I$).
-Thirdly, whether or not the object $o_k$ is private is dependent on $t$, but since $t$ is not available at test time $o_k$'s privacy will be regarded as independent from any observable variable.
-Lastly, we'll assume for now that image metadata ($D$) in its raw form doesn't meaningfully contribute to the privacy determination problem.
+Lastly, we'll assume for now that image meta-data ($D$) in its raw form doesn't meaningfully contribute to the privacy determination problem.
 GPS location and time stamps may be useful for influencing the scene probabilities or accounting for lighting differences in the image for object recognition, but we do not have direct influence over those predictors.
@@
 \centering
-    \begin{tikzpicture}[
-        every node/.style={shape=ellipse, draw}
-    ]
+    \begin{tikzpicture}[every node/.style={shape=ellipse, draw}]
     %NODES
-    %\node (mt) {$m \in t$};
     \node (io) {$i_{m}\textrm{ on }o_k$};
@@
     \node (s) [right = of i] {$S$};
-    %\node (t) [below = of po, text width=2cm, align=center] {$i_m \textrm{ on } o_k, \exists i_l \textrm{ on } o_k$ s.t. $l \in t$};
-    %\node (op) [below = of io] {$\exists i_l \textrm{ on } o_k \textrm{ s.t. } l \in t$};
     \node (or) [below = of i] {$o_k$ is private};
     \node (d) [right = of s] {$D$};
-
     %PATHS
@@
     \path   (i) edge [->] (s);
-    %\path   (po) edge [->] (t);
-    %\path   (po) edge [->] node [left, draw=none] {$f_t$} (t);
-    %\path   (or) edge [->] (op);
     \end{tikzpicture}
@@
 
-Now we are ready to perform some Bayesian inference to obtain $p(i_m \textrm{ on } o, o \textrm{ is private} \mid I, \textrm{ evidence})$.
+We can use Bayes rule and the chain rule to obtain the conditional joint probability distribution $p(i_m \textrm{ on } o, o \textrm{ is private} \mid I, \textrm{ evidence})$.
 
@@
 $N$ is number of images, $M$ is number of pixels per image, $K$ is number of object types, $L_n$ is number of objects in image $I_n$, $Q$ is the number of neighbors to pixels, and $T$ is number of iterations (for root finding).
+$T$ and $Q$ are presumed to be small.
 
-Equation \ref{eqn:indep} illustrates independence of the events that a specific pixel belongs to a specific object and that object being a private object type.
+Equation \ref{eqn:indep} in the previous section illustrates independence of the events that a specific pixel belongs to a specific object and that object being a private object type.
 However, the prior distribution $p(i_m \textrm{ on } o \mid I)$ is still unknown.
-We will consider the factors in Equation \ref{eqn:joint} to estimate it.
 Estimation of $p(i_m \textrm{ on } o \mid I)$ can be regarded as a constrained Maximum a Priori (MAP) estimation follows:
@@
 		p(I_n \mid i_{nm} \textrm{ on } o_k) \theta_k \Big) \label{eqn:loglik}\\
-	\sum_{k=0}^{K} \theta_k - 1 = 0 \label{eqn:constr}\\
+	1 - \sum_{k=0}^{K} \theta_k = 0 \label{eqn:constr}\\
 	\mathcal{L}(\theta, \lambda) = \sum_{n=0}^{N} \log 
-		\Big( \sum_{k=0}^{K} p(I_n \mid i_{nm} \textrm{ on } o_k) \theta_k \Big) + \lambda - \lambda \sum_{k=0}^{K} \theta_k \label{eqn:lagr}\\
+		\Big( \sum_{k=0}^{K} p(I_n \mid i_{nm} \textrm{ on } o_k) \theta_k \Big) - \lambda + \lambda \sum_{k=0}^{K} \theta_k \label{eqn:lagr}\\
 	\mathbf{F}^{K \times 1}: \mathbf{F}_k = \frac{\partial \mathcal{L}}{\partial \theta_k} = 
@@
 			{p(I_n \mid i_{nm \textrm{ on } o_k})}
-			{\sum_{j=0}^{K} p(I_n \mid i_{nm \textrm{ on } o_j}) \theta_j} = 0;
+			{\sum_{j=0}^{K} p(I_n \mid i_{nm \textrm{ on } o_j}) \theta_j} 
+			    + \lambda = 0;
 		\frac{\partial \mathcal{L}}{\partial \lambda} = 1 - \sum_{k=0}^{K} 
@@
 		\sum_{n=0}^{N} \frac
-			{-p(I_n \mid i_{nm \textrm{ on } o_i})p(I_n \mid i_{nm \textrm{ on } o_j})}
+			{-p(I_n \mid i_{nm \textrm{ on } o_i})p(I_n \mid i_{nm 
+			    \textrm{ on } o_j})}
 			{(\sum_{l=0}^{K} p(I_n \mid i_{nm \textrm{ on } o_l}) \theta_l)^2} \label{eqn:lagr2der};
-		\frac{\partial}{\partial \lambda} \left( \frac{\partial \mathcal{L}}{\partial \lambda} \right) = 0
+		%\frac{\partial}{\partial \lambda} \left( \frac{\partial \mathcal{L}}{\partial \lambda} \right) = 0
 \end{gather}
@@
 Equation \ref{eqn:lagr1der} is difficult to solve in closed form.
-Thus, we will use the Newton-Raphson root finding algorithm to estimate the roots in quadratic time, Algorithm \ref{alg:newtraph}.
+Thus, we will use the Newton-Raphson root finding algorithm to estimate the roots in $O(M(NK + K^3 + TNK))$ time, Algorithm \ref{alg:newtraph}.
 This algorithm also requires the second derivative of the Lagrangian, a Jacobian matrix, expressed in Equation \ref{eqn:lagr2der}.
 
-\begin{algorithm}[h!]\label{alg:newtraph}
+\begin{algorithm}[htp!]\label{alg:newtraph}
 \caption{Newton-Raphson Root Finding Algorithm, $O(M(NK + K^3 + TNK))$}
@@
 
-However, $p(i_m \textrm{ on } o \mid I)$ is still based on noisy evidence, which has a high potential for resulting in a falsely high number of contiguous object regions.
+However, $p(i_m \textrm{ on } o \mid I)$ is still based on noisy evidence $p(I \mid i_m \textrm{ on } o)$, which has a high potential for resulting in a falsely high number of contiguous object regions.
 Aside from the affect of identifying only small portions of private objects, this will also adversely affect $p(o_k \textrm{ is priv})$ by driving up the number of unmarked private objects and potentially mislabeling those that are marked.
@@
 
-\begin{algorithm}[h!]\label{alg:regObjMap}
+\begin{algorithm}[htp!]\label{alg:regObjMap}
 \caption{Regularize the Object Map, $O(NTMQ^2K)$}
 \KwData{
-    $I: \lbrack 0,256 \rbrack^{M \times H} \rightarrow \{i_{mh} \mid \forall m \in \lbrack 0, M ), \forall h \in \lbrack 0, H)\}$}
+    $I: \{i_{mh} \mid \forall m \in \lbrack 0, M ), \forall h \in \lbrack 0, H)\}$}
 \KwResult{$p_R(i_m \textrm{ on } o \mid I)$}
@@
 To achieve this, we us a variant of the Two-Pass Connected Component Labeling algorithm \cite{hoshen1976percolation}, described in Algorithm \ref{alg:objEnum}.
-It is important to note that although objects are enumerated according to the high object type probabilities for each pixel, probability summations for all objects types are preserved at each pixel.
+It is important to note that although objects are enumerated according to the highest object type probabilities for each pixel, probability summations for all objects types are preserved at each pixel.
 This will be important later.
 
-\begin{algorithm}[h!]\label{alg:objEnum}
+\begin{algorithm}[htp!]\label{alg:objEnum}
 \caption{Object Enumeration, $O(NMQ + L)$}
@@
 
-\begin{algorithm}[h!]\label{alg:private}
+\begin{algorithm}[htp!]\label{alg:private}
 	\KwData{$E^{N}$ // the enumerated object vectors}
@@
 \begin{align}
-    p(C_{P_{nm}} = 1 \mid I, \textrm{ evidence}) = \sum_{k=0}^{K} p(i_{nm} \textrm{ on } o_k \mid I_n) P_k
+    p(C_{P_{nm}} = 1 \mid I, \textrm{ evidence}) = \sum_{k=0}^{K} p_R(i_{nm} \textrm{ on } o_k \mid I_n) P_k
 \end{align}
+
+In summary, we can estimate $p(i_m \textrm{ on } o)$ once in $O(M(NK + K^3 + TNK))$ time, regularize $p(i_m \textrm{ on } o \mid I)$ for all images in $O(NTMQ^2K)$ time, enumerate the objects in all images in $O(NMQ + L_n)$  time, estimate $p(o \textrm{ is priv})$ once in $O(N L_n K)$ time, and finally estimate the pixel privacy probability for all images in $O(N M K)$ time.
+Since $L_n < M$, total running time for estimating all pixels for all images is $O(MK^3 + NMTKQ^2)$ where $T$ and $Q$ are presumed to be small.
 
@@
 \begin{align}\label{eqn:task2.0}
-    p(C_{P_{nm}} = 1 \mid I, \textrm{ evidence}) & = 1 - \prod_{l = 0}^{L_n} \left( 1 - p(\textrm{image region $E_{nl}$ 
-        is private} \mid \textrm{evidence}) \right)\\
+    p(C_{I_{n}} = 1 \mid I_n, \textrm{ evidence}) & = 1 - \prod_{l = 0}^{L_n} \left( 1 - p(\textrm{image region $E_{nl}$ 
+        is private} \mid I_n, \textrm{evidence}) \right)\\
+    & = 1 - \prod_{l = 0}^{L_n} \left( 1 - \sum_{k=0}^{K} (E_{nl}\textrm{ on } o_k) p(o_k \textrm{ is priv}) \right)\\
     & = 1 - \prod_{l = 0}^{L_n} \left( 1 - \sum_{k=0}^{K} (E_{nl}\textrm{.psum}_k / E_{nl}\textrm{.cnt}) p(o_k \textrm{ is priv}) \right) \label{eqn:objpriv}
@@
 
-Notice in Equation \ref{eqn:objpriv} that the probability of an object region being of some type $k$ is represented as the average of the $k$-probabilities across all pixels in that $l$ object region.
-As mentioned in section \ref{sec:t1alg}, the actual object regions would likely have been different for different values of $k$, but capturing average probabilities for each type across all pixels in the dominant region is presumed (by the author) to be a reasonable compromise.
+Notice in Equation \ref{eqn:objpriv} that the probability of an object region being of some type $k$ is represented as the arithmetical average of the $k$-probabilities across all pixels in that $l$ object region.
+This implies that individual pixel probabilities are identically distributed within a given region, which is likely not true since each is the result of belief propagation over a Markov Random Field.
+However, as mentioned in section \ref{sec:t1alg}, the actual object regions would likely have been different for different values of $k$, so use of the arithmetical mean for each type across all pixels in the dominant region is presumed (by the author) to be a reasonable compromise.
 The independent joint probabilities for object classification and class privacy are marginalized over all object types to achieve the probability of the object region being private.
+Then, the final probability of the image being private is the complement of the probability over the Binomial distribution that zero object regions are private.
+This solution relies on the entirety of Task 1, except the final calculation of pixel privacy, and then calculates the probability over the Binomial distribution in $O(N L_n K)$ time, for the same total running time of $O(MK^3 + NMTKQ^2)$ where $T$ and $Q$ are presumed to be small.
 
@@
 
+\textit{Important note.}  A qualifying exam tests your technical
+knowledge and abilities, but also your potential to conduct
+independent, high-quality research using  the highest ethical standards.
+In answering the question below, you may consult existing
+published work, but you may not discuss the question with people other than
+members of the advisory committee.  (The sole exception is that you
+may seek non-technical help with English from the writing center on
+campus, but then you must submit both your original draft as well as
+the final draft after incorporating feedback from the editor or
+writing center.)  The University's academic integrity policy will
+govern this exam. In particular, all material you submit
+must be your own work which you
+personally and independently completed. If any portion of your
+submitted material is not your own work, you must make this explicitly
+clear using proper citations. For instance, all sentences in your
+report must be your own unless you place the relevant sentences in
+quotation marks and give a citation to the source.  \textit{Note that
+  descriptions and discussions of existing work are to be in your own
+  words based on your understanding of the work; it is \textbf{not}
+  acceptable to copy or paraphrase even a small part of another work
+  unless it is placed inside quotation marks and properly cited.}
+{In accordance with University policy, consequences of academic
+  integrity violations  include failure on this exam and expulsion
+  from the University.}
+
+\vspace{-12pt}
+\begin{enumerate}
+\item
+Many photos contain sensitive information, especially images
+  taken by smartphone or home cameras. A potential solution to
+  this problem is to develop automatic algorithms that analyze an
+  image's content and context to estimate if it contains private
+  content, and to censor the image if so.  Such an  algorithm
+  may have access to a variety of information about the image, but
+  this evidence is likely to be noisy. For instance, for a given
+  image $I$, consider four classes of potential noisy evidence:
+
+\begin{enumerate}
+\item The output of automatic computer vision algorithms that estimate
+  properties of the scene, like whether it is taken indoors or
+  outdoors, what type of room it is taken in (office, bathroom, etc.),
+  whether it is a natural (e.g.\ forest) or manmade (e.g.\ city) scene,
+  etc. Since image analysis is a hard problem,  these classifiers
+  give probabilistic outputs instead of hard labels,
+  i.e. for the natural versus manmade class, they might give two
+probabilities, $P(\textsf{natural}|I)$ and $P(\textsf{manmade}|I)$.
+
+\item The output of automatic object recognition algorithms that
+  classify each pixel $p \in I$ according to whether or not it
+  contains an object of interest.  Again, since these recognition
+algorithms are imperfect, the output
+is typically provided as a probability distribution $P(I |
+  p \mbox{ on object type } o)$, the probability of the observed image
+given that object $o$ includes pixel $p$.
+
+\item Non-visual metadata like GPS coordinates and timestamps.
+
+\item Explicit feedback from a user, in a form that requires very
+  little human effort. For instance, users might click on one or more
+  points in the image that lie within sensitive objects, yielding a
+  (very small) set of image coordinates ${\cal P} \subseteq I$.
+\end{enumerate}
+
+After analyzing an image, we would want our automated analysis to
+produce two outputs. The first is an estimate of the probability
+that $I$
+contains at least one
+privacy-sensitive object or situation, $P(I \mbox{ is private} |I)$. The
+second is an estimate for each pixel $p$ of the probability that
+$p$ lies within a private object or image region (and thus should be
+blurred or obscured), $P(p \mbox{ is private} | I)$. Since privacy is
+unlikely to change dramatically from one pixel to the  next, this
+latter probability should encourage smoothness of the resulting
+``privacy map'' by incorporating some prior probability on adjacent
+pixels (e.g.\ that two neighboring pixels are both private or both
+not private with some high probability, and have different privacy
+values with some low probability).
+
+In a roughly 5-page paper, describe how you would formulate and solve
+each of the above two problems. Make sure to: (1) propose a probabilistic
+model for solving the problem, using the evidence from (a) through (d)
+above, (2) derive the model from first
+principles, or at least argue for why it is well-principled, and (3)
+propose an inference algorithm that would solve each problem (using
+the probabilistic model).  For (1), make sure to precisely define any
+variables you use.
+For (2), make sure to explicitly state and defend any assumptions that
+your model makes.
+ For (3), give pseudocode and an intuitive
+explanation of how the algorithm works. The algorithm does not need to
+be fast but should be no worse than polynomial time.  You can assume
+that the computer vision features ((a) and (b) above) are already
+implemented.
+
+In preparing this solution, the following two papers may be helpful:
+
+\begin{itemize}
+\item  Lawrence R. Rabiner. ``A  tutorial on hidden Markov models and
+selected applications in speech recognition.'' \textit{Proceedings of
+the IEEE,} 1989, pages 257-286.
+\item Pedro F. Felzenszwalb and Daniel P. Huttenlocher. ``Efficient
+Belief Propagation for Early Vision.''
+\textit{International Journal of Computer Vision}, Vol. 70, No. 1, October 2006.
+\end{itemize}
+
+\end{enumerate}
+
 \subsection{Dr. Donald Williamson}

### 2018-11-18 main.tex
181118-0921 ready to submit

@@
 The GAN framework involves the simultaneous training of both a generative model $G$ with parameters $\theta_g$ and a discriminative model $D$ with parameters $\theta_d$.
-Given a target data space $T$ with distribution $p_t$, $G$'s task is to learn a mapping from a noisy input $z$ to the data space $T$ ($G:Z \rightarrow G(z) \sim p_g \approx p_t$); while $D$'s task is to discriminate whether some input $x$ originated from either $T$ or $G(z)$ ($D: X \rightarrow \{0, 1\}$).
+Given a target data space $T$ with distribution $p_t$, $G$'s task is to learn a mapping from a noisy input $z$ to a data space that estimates the target data space $T$ ($G:Z \rightarrow G(z) \sim p_g \approx p_t$); while $D$'s task is to discriminate whether some input $x$ originated from either $T$ or $G(z)$ ($D: X \rightarrow \{0, 1\}$).
 Since $G$ would like to generate samples that are representative of $T$, and therefore make $D$'s task more difficult by capturing $p_t$ with $p_g$, $G$ includes a loss term that is based on whether $D$ successfully discriminated inputs as $T$ or $G(z)$ and back-propagates $D$'s success as error to $\theta_g$ appropriately.
 In this way, $G$ learns how to generate data that is hard for an ever-improving $D$ to distinguish.
-Importantly, labeled data is not required for either $G$ or $D$ because the loss function used by $G$ is defined in terms of $D(x)$ and $(1-D(G(z))) \in \{0, 1\}$ (ie.
-did $G$ fool $D$?) and the labels for $x$ essentially map to $\{p_t, p_g\}$ (ie.
-from target or generated data?).
+Importantly, `target' or `non-target' (or $\{p_t,p_{g}\}$, respectively) are the only labels (of $x$) required by either $G$ or $D$ because the loss function used by $G$ is defined in terms of $D(t)$ and $(1-D(G(z)))$ (ie.
+$G$ tries to minimize while $D$ tries to maximize, or vise versa).
 Thus, the GAN framework presents a promising approach for generating data that approximates a target distribution in an unsupervised manner.
@@
 Of course, Conditional GAN requires that conditioning data (eg.
-labels) are available for all target samples $x \in X$.
+labels) are available for all target samples $t \in T$.
 
@@
 In addition to performance improvements, the authors demonstrate that $D$ is able to learn semantically-relevant features that are both explainable (via latent-space walk based validation) and usable (via feature-based supervised classification).
-In one case they were able to discover a linear vector in $Z$-space that could control rotation of faces.
+In one example case they were able to discover a linear vector in $Z$-space that could control rotation of faces.
 
@@
 In their approach, the encoder-decoder learns a semantic encoding based on the non-missing portions and attempts to reproduce a whole image based on that semantic encoding.
-The `reconstruction loss' is defined in terms of distance ($L_2$) from the filled area to the removed area.
 The encoder-decoder is regarded as $G$, and a new discriminator $D$ is trained to predict whether its input is a reconstructed output from $G$ or an original image.
+The `reconstruction loss' for $G$ is defined in terms of distance ($L_2$) from the filled area to the removed area and the adversrial loss comes from $D$.
 Interestingly, and in contrast with \cite{mirza2014conditional}, the authors achieved better results when $G$, but not $D$, was conditioned with context information \footnote{The paper defines context information as the missing pixels which, if true, undermines the objective (any statement of success should be based on withholding the missing pixels from the architecture).
-I have requested clarification on this from the author (17 Oct 2018).}.
+I have requested clarification on this from the author (17 Oct 2018) and have not received a response as of 18 Nov 2018.}.
 
@@
 Their approach uses a Convolutional Neural Network (CNN) based generative model $G$ to learn a class label for each pixel in the image based on a ground-truth labeling (supervised), thereby creating a segmentation map of the input image.
-$D$ then takes a segmentation map and tries to classify it as being a ground truth map or a generated map.
+$D$ then takes a segmentation map and tries to classify it as being a ground truth map or a generated map, and passes its loss back to $G$.
 %While $D$'s objective function tries to maximize classification accuracy of the input segmentation map as being either ground truth or generated, $G$'s objective function tries to maximize pixel class labeling accuracy with respect to ground truth while also minimizing accuracy.
@@
 Essentially, they create $G$ by concatenating the input noise with an encoding of the text description and feed the result into a feed-forward CNN using fractional-striding\footnote{The paper claims this to be a de-convolutional network, but I presume it to mean a CNN using fractional striding to achieve up-sampling.}.
+The architecture is a variant of the Conditional GAN.
 $D$ then down-samples with CNN layers until a 4x4 representation is achieved, concatenates the text encoding onto each of the 16 frames, and continues convolutions to achieve the binary discriminator output.
@@
 
-A logical step forwards from image-based applications is to apply GAN to video.
+A logical progression from image-based applications is to apply GAN to video.
 Mathieu, Couprie, and LeCun first adopted a GAN approach to a video prediction task \cite{Mathieu2015DeepMV}.
@@
 The overall generator $G$ then generates the final predicted frame $Y$ at full resolution using a Laplacian pyramid \cite{denton2015deep} built on each $G_k$.
-The discriminator $D$ attempts to classify the last frame(s) in a frame sequence as being either real or generated by $G$ (the first frames in the input sequence are always real), and thus provides the usable adversarial loss to $G$.
+The discriminator $D$ attempts to classify the last frame(s) in a sequence as being either real or generated by $G$ (the first frames in the input sequence are always real), and thus provides the usable adversarial loss to $G$.
 
-Tulyakov, Liu, Yang, and Kautz introduce a method for decomposing video into content (the objects in the video) and motion (the movement of those objects) components, each having its own generator and discriminator ($G_C, G_M, D_C, D_M$) \cite{tulyakov2017mocogan}.
+Tulyakov, Liu, Yang, and Kautz introduce MoCoGAN, a method for decomposing video into content (the objects in the video) and motion (the movement of those objects) components, each having its own generator and discriminator ($G_C, G_M, D_C, D_M$) \cite{tulyakov2017mocogan}.
 $D_C$ attempts to detect whether a frame is sampled from a real video clip or a generated one, while $D_M$ attempts to detect whether a video clip (series of frames) is sampled from a real or generated clip.
@@
 
-As discussed, the GAN approach has enabled major improvements to several research areas including semi-supervised and unsupervised learning, generative models, the learning of complex loss functions, data synthesis for other learning tasks, image and domain learning and translation, and semantic learning and generation.
+As discussed, the GAN approach has enabled major improvements to several research areas including enablement of semi-supervised and unsupervised learning, improved performance of generative models, the learning of highly complex loss functions, data synthesis for enabling/improving other learning tasks, image and domain learning and translation, and semantic learning and generation.
 In general, the GAN approach provides a relatively simple and tractable way of isolating deep learning objectives and sources of error in such a way that they can be addressed individually and researchers can make headway toward a complex objective, as is apparent in the Conditional GAN, in-painting, MoCoGAN, and text-to-image generation, among others.
-Broadly stated, as deep learning models adopt more complex architectures and greater capacity, there is an increasing need to provide more meaningful and comprehensive constraint to the model's objective.
-GANs provide a tractable methodology for learning sets of constraints which can be as complex as the main learning objective, or even more-so.
+Broadly stated, as deep learning models adopt more complex architectures and greater capacity, there is an increasing need to provide more meaningful and comprehensive constraint(s) to the model's objective.
+GANs provide a tractable methodology for learning sets of constraints (losses) which can be as complex as the main learning objective, or even more-so.
 
@@
 
-Further, as research yields models and approaches that efficiently and effectively solve moderately complex tasks, it is reasonable to expect growing a demand for standardized, modular, and extensible model interfaces (eg.
+Further, as research yields models and approaches that efficiently and effectively solve moderately complex tasks, it is reasonable to expect an increase in demand for standardized, modular, and extensible model interfaces (eg.
 Model Based Engineering ecosystem) for use among research and commercial entities to solve more complex tasks and continue the evolution of artificial intelligence.
 
-In the area of security, GAN approaches pose both a threat (fake video, adversarial examples) as well as a benefit (eg.
+In the area of security, GAN approaches pose both a threat (eg.
+fake video, adversarial examples) as well as a benefit (eg.
 efficient fuzzing \cite{hu2018ganfuzz}) to security, privacy, and safety.
@@
 When ML methods are used as a tool in an adversarial context like security, robustness of the model to adversarial examples becomes a primary design objective.
-Intrusion detection systems (IDS) and Spam filters are two classic examples wherein the design objective is to accurate classify messages or traffic as benign or malicious, while the attacker's objective is to design his/her malicious content to be classified as benign (i.e.
+Intrusion detection systems (IDS) and Spam filters are two classic examples wherein the design objective is to accurate classify messages or traffic as benign or malicious in the presence of an attacker who's objective is to design his/her malicious content to be classified as benign (i.e.
 evade detection).
@@
 Word-based Bayesian Spam filters date back to 1998 \cite{sahami1998bayesian}, and the first `Bayes vs.
-Bayes' attack surfaced (in the public eye) in 2004 \cite{graham2004beat}.
+Bayes' publicly-demonstrated attack was presented in 2004 \cite{graham2004beat}.
 These early examples demonstrated the concept, but were not human-spoofable.
@@
 Next, they implemented a paraphrase generator which takes, as input, $s_1$ and $p_2$, and generates a paraphrase sentence $s_2$.
-They show that randomly generated paraphrase samples (not adversarially learned) result in a relatively high number of miss-classifications, and also that augmenting those classification models at train-time with these syntactically-diverse adversarial examples results in better performance in the face of new adversarial examples.
+They show that randomly generated paraphrase samples (not adversarially learned) result in a relatively high number of miss-classifications using a common classifier trained on common data sets, and also that augmenting the training data with these syntactically-diverse adversarial examples results in better performance in the face of new adversarial examples.
 
@@
 
+The next two papers address the problem of confidentiality of training data used to train models.
+An andividual or company may develop and train a model based on data they own and wish to protect, such as financial, medical, social, or other information, while making the model available to those who are not authorized to see the training data.
 Shokri, Stronati, Song, and Shmatikov have demonstrated a methodology to determine whether a specific sample was included within the training set for a target model \cite{shokri2017membership}.
-Given a specific data sample $x_?$ and a target model $M_t$, their paper shows how to create multiple `shadow' models with different (real or synthesized) data from that used to train $M_t$, obtain prediction responses from each of the shadow models that can be labeled as `in' or `out' depending on whether each sample was part of that shadow models training set, and then use the labeled response data to train an attack model $M_a$.
+Given a specific data sample $x_?$ and a target model $M_t$, their paper shows how to (a) create multiple `shadow' models with different (real or synthesized) data from that used to train $M_t$, (b) obtain prediction responses from each of the shadow models that can be labeled as `in' or `out' depending on whether each sample was part of that shadow models training set, and then (c) use the labeled response data to train an attack model $M_a$.
 After training, $M_a$ achieves good performance in predicting whether black-box responses from $M_t$ are associated with training input (`in') or test input (`out').
-The paper addresses the issue of bias from confidence in the responses by synthesizing the data and sampling from the subset of that data that results in similar confidence distributions from $M_t$.
+The paper addresses the issue of bias from confidence in the responses by synthesizing the data and sampling from only the subset of that synthesized data which results in similar confidence distributions from $M_t$.
 
 Similarly, Song, Ristenpart, and Shmatikov demonstrated several white- and black-box approaches to build covert channels into ML models for the purpose of extracting sensitive/proprietary training data.
-In the attack scenario, the data owner may not wish to reveal the training data to the model developer, but rather have the model trained and delivered by a secure third part.
+In the attack scenario, the data owner may not wish to reveal the training data to the model developer, but rather have the model trained and delivered by a secure third party.
 White-box methods were presented but, more interestingly, a `capacity abuse' attack was described that addresses the black-box scenario.
-Essentially, the learning algorithm automatically augments the training data with chosen (by the attacker) data that is labeled with an encoding of the private training data.
-The model learns on both the original and malicious training data, and then the attacker, having black-box access to the model, can present the known malicious data to the model and receive the encoded private data via the responses.
-I hypothesized that performance of the model for both its benign purpose as well as its covert purpose depends on the malicious data being sparse.
-The intuition is that since sparse data is distant  from the private training data (in the feature space), it will not have adverse affects on classification; and since each sparse sample is distant from other sparse samples, they are mutually separable resulting in higher accuracy predictions.
-As a follow-up independent study, I demonstrated under ideal conditions that the Membership Inference \cite{shokri2017membership} methodology could be used to detect whether a given model was trained on a sparse distribution or not, thereby serving as a potential indicator of covert channel(s) \cite{stein2018covert}.
+Essentially, the malicious learning algorithm automatically augments the training data with attacker-chosen data (think seeded number generator) that is then labeled with an encoding of the private training data just prior to training.
+The model learns on both the original and malicious training data.
+Later on, the attacker, having black-box access to the model, can present the known malicious data to the model and receive the encoded private data via the responses.
+Stein built upon this work, and hypothesized that performance of the model for both its benign purpose as well as its covert purpose depends on the malicious data being sparse.
+\cite{stein2018covert} The intuition is that since sparse data is distant  from the private training data (in the feature space), it will not have adverse affects on classification; and since each sparse sample is distant from other sparse samples, they are mutually separable resulting in higher accuracy training data reconstruction.
+Stein demonstrated under ideal conditions that the Membership Inference \cite{shokri2017membership} methodology could be used to detect whether a given model was trained on a sparse distribution or not, thereby serving as a potential indicator of covert channel(s).
 
@@
 ML robustness in the face of adversarial examples is likely to continue being an open problem for some time.
-The challenge is not a transient implementation issue that will be worked out as ML technology matures, but rather, like generalization, a core learning issue at the heart of the gap between ML and true intelligence.
-In this vein, I feel that a deeper understanding of complex models, specifically deep learning networks, including properties, topographies, and structures, and their effects in adversarial scenarios.
+The challenge is not a transient implementation issue that will be worked out as ML technology matures, but rather, like bias vs variance, a core learning issue at the heart of the gap between ML and true intelligence.
+In this vein, I feel that a deeper understanding of complex models, specifically deep learning networks, including properties, topographies, and structures, and their effects in adversarial scenarios, is a promising area of study.
 For example, characterizing parameter evolution in the manifold during training on benign data sets vs data sets with adversarial augmentation may reveal indicators that could be measured directly on future models (in the absence of adversarial data).
@@
 Since knowledge of the adversary's perturbation minimization objective helped inform provable design guidelines for the classifier, perhaps a finer-grained approach to adversarial learning (eg.
-segmentation of large generative and classification networks with complex tasks into matched ensembles of smaller networks with simpler tasks) will help increase model robustness to adversarial examples.
+segmentation of large generative and classification networks with complex tasks into matched ensembles of smaller networks with simpler tasks) will lead to discovery of better weakness causality and eventually prescriptions for model robustness to specific adversarial goals.
 
 \newpage
-\section{Privacy Prediction}
+\section{Privacy Prediction}\label{sec:privpred}
 
-Two privacy prediction tasks have been identified: (1) predict whether each pixel in a given image is part of a private object, and (2) predict whether or not a given image has private content\footnote{The tasks are addressed in reverse order from how they were presented in \ref{app:crandall} because, in the approach described herein, the solution for (2) can be largely derived from the solution for (1).}.
+Two privacy prediction tasks have been identified: (1) predict whether each pixel in a given image is part of a private object, and (2) predict whether or not a given image has private content\footnote{The tasks are addressed in reverse order from how they were presented in \ref{app:questions} because, in the approach described herein, the solution for (2) can be largely derived from the solution for (1).}.
 For both tasks, assume that $N$ images are sampled, each with $M$ pixels having $H$ channels defined as $I : {\lbrack0,256\rbrack}^{M \times H} \rightarrow \{i_{mh} | \forall m \in \lbrack0,M), \forall h \in \lbrack0,H) \}$.
@@
 
-Let's define some assumptions.
-
-\begin{enumerate}
-    \item Both tasks are semi-supervised, such that the only form of labeled data is $t_n$\footnote{The semi-supervised assumption was adopted because human point-and-click labeling of private image regions for both training and updating of models is thought to be a realistic scenario.}.
-Note that this assumption implies that $t_n$ will not be available at test time.
-\end{enumerate}
-
 In the following sections, the $n$ subscript will often be omitted for simplicity when considering a single image.
@@
 When defining a probabilistic model, it is important to define the central assumptions behind what the model is attempting to capture in terms of privacy.
-For this approach, the following assumptions are used: (a) certain objects, or object details, are private, and (b) whether an image is private is determined by whether it contains private objects or object details.
+For this approach, the following assumptions are used:
+
+\begin{enumerate}
+    \item Both tasks are semi-supervised, such that the only form of labeled data is $t_n$\footnote{The semi-supervised assumption was adopted because human point-and-click labeling of private image regions for both training and updating of models is thought to be a realistic scenario.}.
+Note that this assumption implies that $t_n$ will not be available at test time.
+    \item Certain objects, or object details, are private.
+    \item Whether an image is private is determined by whether it contains private objects or object details.
+\end{enumerate}
+
 Notice that certain contextual privacy scenarios, such as an image of a car known to belong to `Bob' being parked outside of a liquor store during work hours, for example, may not necessarily be marked private, even though `Bob' might argue that it is.
@@
 
-To form the basis of the probabilistic model, a belief network which juxtaposes our evidence and other random variables into the believed set of dependency assumptions can be created.
+To form the basis of the probabilistic model, a belief network which juxtaposes our evidence and other random variables into the believed set of dependency assumptions is created.
 The dependency assumptions are defined as follows.
@@
 
-Time complexity for all algorithms are given in their respective captions.
+All pseudo code can be found in Appendix \ref{app:algs}.
+Time complexity for all algorithms is given in their respective captions.
 $N$ is number of images, $M$ is number of pixels per image, $K$ is number of object types, $L_n$ is number of objects in image $I_n$, $Q$ is the number of neighbors to pixels, and $T$ is number of iterations (for root finding).
@@
 This algorithm also requires the second derivative of the Lagrangian, a Jacobian matrix, expressed in Equation \ref{eqn:lagr2der}.
+Essentially, the algorithm uses the hyperplane formed by the gradient at any coordinate to estimate the root as the point at which the plane passes through the axes, and repeats the process using the estimated root until a sufficiently close approximation is reached.
 
-\begin{algorithm}[htp!]\label{alg:newtraph}
+Now that the prior $p(i_m \textrm{ on } o)$ has been estimated, we can use Bayes rule to calculate $p(i_m \textrm{ on } o \mid I)$ as follows:
+
+\begin{align}
+    p(i_m \textrm{ on } o \mid I) = 
+	    \frac{p(I \mid i_m \textrm{ on } o) p(i_m \textrm{ on } o)}{\sum_{j=0}^{K} p(I \mid i_m \textrm{ on } o_j) p(i_m \textrm{ on } o_j)}    
+\end{align}
+
+However, $p(i_m \textrm{ on } o \mid I)$ is still based on noisy evidence $p(I \mid i_m \textrm{ on } o)$, which has a high potential for resulting in a falsely high number of contiguous object regions.
+Aside from the affect of identifying only small portions of private objects, this will also adversely affect $p(o_k \textrm{ is priv})$ by driving up the number of unmarked private objects and potentially mislabeling those that are marked.
+Efficient belief propagation over pixels arranged as Markov Random Fields (MRF) has been shown to produce good results in scenarios where labels are unlikely to change between two neighboring pixels except in border regions \cite{felzenszwalb2006efficient}.
+We use a variant of this approach to help regularize the noisy $p(i_m \textrm{ on } o \mid I)$ evidence to achieve $p_R(i_m \textrm{ on } o \mid I)$, as shown in Algorithm \ref{alg:regObjMap}.
+Essentially, the algorithm iteratively updates messages between pixels which serve to penalize/regularize any two neighboring pixels having different object type probabilities.
+Multiple iterations are needed to ensure influences can propagate sufficiently across potentially large object regions.
+
+The final task is to estimate $p(o \textrm{ is priv})$.
+To achieve this, we need to enumerate all the object instances, their types, their respective probabilities, and whether they are private.
+We us a variant of the Two-Pass Connected Component Labeling algorithm \cite{hoshen1976percolation}, described in Algorithm \ref{alg:objEnum}, which labels all the contiguous object regions and collects their associated privacy labels and statistics.
+It is important to note that although objects are enumerated according to the highest object type probabilities for each pixel, probability summations for all object types are preserved at each pixel because we'll need to marginalize privacy over all object types.
+
+Given the enumerated object data, we estimate $p(o \textrm{ is priv})$ as described in Algorithm \ref{alg:private}.
+Object privacy is tallied according to the type that was chosen as most probable for the image region as opposed to soft-attribution of privacy across the type probabilities.
+This is because the other types may well have resulted in very different object regions, on the order of $K^{L_n}$ different mappings.
+
+As an aside, there is an opportunity to modify the original belief network such that $p(o \textrm{ is priv})$ is dependent on the scene $S$.
+Specifically, if the object types are not sufficiently granular to result in good performance of $p(o \textrm{ is priv})$, the scene information may serve as a valuable conditioning variable.
+An example might be if `text' is a lowest-level object type, then the appearance of `text' in a bathroom scene (prescription label?) might have very different privacy implications than `text' in an outdoor scene (road sign?).
+To incorporate this dependency, Equation \ref{eqn:indep} becomes Equation \ref{eqn:indep-scene} and Algorithm \ref{alg:private} would tally privacy markings conditioned on scene type and output a probability map of dimension $K \times \left| S \right|$.
+This approach, as described, is essentially intended to discover sub-classes of object types which may or may not be needed, as opposed to modeling a believed causality between scene and object privacy, which is why the dependency was omitted in the original belief network.
+
+\begin{align}
+    p(i_m \textrm{ on } o, o \textrm{ is priv} \mid I, \textrm{ evidence}) = p(i_m \textrm{ on } o \mid I) p(o \textrm{ is priv} \mid S) \label{eqn:indep-scene}
+\end{align}
+
+Finally, all the information has been obtained to answer Task 1.  For any given pixel $i_{nm}$, we can estimate the probability of whether it lies on a private object as follows:
+
+\begin{align}
+    p(C_{P_{nm}} = 1 \mid I, \textrm{ evidence}) = \sum_{k=0}^{K} p_R(i_{nm} \textrm{ on } o_k \mid I_n) P_k
+\end{align}
+
+In summary, we can estimate $p(i_m \textrm{ on } o)$ once in $O(M(NK + K^3 + TNK))$ time, regularize $p(i_m \textrm{ on } o \mid I)$ for all images in $O(NTMQ^2K)$ time, enumerate the objects in all images in $O(NMQ + L_n)$  time, estimate $p(o \textrm{ is priv})$ once in $O(N L_n K)$ time, and finally estimate the pixel privacy probability for all images in $O(N M K)$ time.
+Since $L_n < M$, total running time for estimating all pixels for all images is $O(MK^3 + NMTKQ^2)$ where $T$ and $Q$ are presumed to be small.
+
+\subsection{Task 2: Does the Image Contain Private Data?}
+
+The objective for this task can be stated as estimating $p(C_I = 1 \mid I, \textrm{ evidence})$ where $C_I$ is the binary class label for whether the image is benign (0) or private (1).
+The model formulation, basis, and algorithms for Task 1 are completely reused for Task 2.
+A simple extension to the knowledge we've built in Task 1 is applied in order to complete Task 2, as follows.
+
+Assuming there are $L_n$ objects in an image $I_n$, if even one object in the image is private, then the image contains private content.
+We can model the objective task over a Binomial distribution.
+Specifically, choose $L_n$ successes given $L_n$ trials (objects) with the probability of success being $p_s = (1-p(E_l \textrm{ is priv} \mid I, \textrm{ evidence}))$, where $E_l$ is the $l^\textrm{th}$ object region.
+The resulting probability is the complement of our objective probability.
+
+\begin{align}\label{eqn:task2.0}
+    p(C_{I_{n}} = 1 \mid I_n, \textrm{ evidence}) & = 1 - \prod_{l = 0}^{L_n} \left( 1 - p(\textrm{image region $E_{nl}$ 
+        is private} \mid I_n, \textrm{evidence}) \right)\\
+    & = 1 - \prod_{l = 0}^{L_n} \left( 1 - \sum_{k=0}^{K} (E_{nl}\textrm{ on } o_k \mid I_n) p(o_k \textrm{ is priv}) \right)\\
+    & = 1 - \prod_{l = 0}^{L_n} \left( 1 - \sum_{k=0}^{K} (E_{nl}\textrm{.psum}_k / E_{nl}\textrm{.cnt}) p(o_k \textrm{ is priv}) \right) \label{eqn:objpriv}
+\end{align}
+
+Notice in Equation \ref{eqn:objpriv} that the probability of an object region being of some type $k$ given the image is represented as the arithmetical average of the $k$-probabilities across all pixels in that $l^\textrm{th}$ object region.
+This implies that individual pixel probability distributions are identically distributed within a given region, which is not true since each is the result of belief propagation over a Markov Random Field.
+However, as mentioned in section \ref{sec:t1alg}, the actual object regions would likely have been different for different values of $k$, so use of the arithmetical mean for each type across all pixels in the dominant region is presumed (by the author) to be a reasonable compromise.
+The independent joint probabilities for object classification and class privacy are marginalized over all object types to achieve the probability of the object region being private.
+Then, the final probability of the image being private is the complement of the probability over the Binomial distribution that zero object regions are private.
+This solution relies on the entirety of Task 1, except the final calculation of pixel privacy, and then calculates the probability over the Binomial distribution in $O(N L_n K)$ time, for the same total running time of $O(MK^3 + NMTKQ^2)$ where $T$ and $Q$ are presumed to be small.
+
+%%%%%%%%%%%%%%%%% BIB %%%%%%%%%%%%%%%%%%%%
+\newpage
+\addcontentsline{toc}{section}{References}
+\bibliographystyle{plain}
+\bibliography{references}
+
+%%%%%%%%%%%%%%%%% Appendices %%%%%%%%%%%%%%%%%%%%
+
+\appendixtitleon
+\begin{appendices}
+
+%%%%%%%%%% Appendix A %%%%%%%%%%
+
+\clearpage
+\begin{center}
+    \section{Algorithms from Section \ref{sec:privpred}\label{app:algs}}
+\end{center}
+
+\begin{algorithm}[!h]\label{alg:newtraph}
+\tiny
 \caption{Newton-Raphson Root Finding Algorithm, $O(M(NK + K^3 + TNK))$}
@@
 
-Now that the prior $p(i_m \textrm{ on } o)$ has been estimated, we can calculate $p(i_m \textrm{ on } o \mid I)$ as follows:
-
-\begin{align}
-    p(i_m \textrm{ on } o \mid I) = 
-	    \frac{p(I \mid i_m \textrm{ on } o) p(i_m \textrm{ on } o)}{\sum_{j=0}^{K} p(I \mid i_m \textrm{ on } o_j) p(i_m \textrm{ on } o_j)}    
-\end{align}
-
-However, $p(i_m \textrm{ on } o \mid I)$ is still based on noisy evidence $p(I \mid i_m \textrm{ on } o)$, which has a high potential for resulting in a falsely high number of contiguous object regions.
-Aside from the affect of identifying only small portions of private objects, this will also adversely affect $p(o_k \textrm{ is priv})$ by driving up the number of unmarked private objects and potentially mislabeling those that are marked.
-Efficient belief propagation over pixels arranged as Markov Random Fields (MRF) has been shown to produce good results in scenarios where labels are unlikely to change between two neighboring pixels except in border regions \cite{felzenszwalb2006efficient}.
-We use a variant of this approach to help regularize the noisy $p(i_m \textrm{ on } o \mid I)$ evidence to achieve $p_R(i_m \textrm{ on } o \mid I)$, as shown in Algorithm \ref{alg:regObjMap}.
-
-\begin{algorithm}[htp!]\label{alg:regObjMap}
+\begin{algorithm}[hp!]\label{alg:regObjMap}
+\tiny
 \caption{Regularize the Object Map, $O(NTMQ^2K)$}
@@
 
-The final task is to estimate $p(o \textrm{ is priv})$.
-To achieve this, we need to enumerate all the object instances, their types, their respective probabilities, and whether they are private.
-To achieve this, we us a variant of the Two-Pass Connected Component Labeling algorithm \cite{hoshen1976percolation}, described in Algorithm \ref{alg:objEnum}.
-It is important to note that although objects are enumerated according to the highest object type probabilities for each pixel, probability summations for all objects types are preserved at each pixel.
-This will be important later.
-
-\begin{algorithm}[htp!]\label{alg:objEnum}
+\begin{algorithm}[hp!]\label{alg:objEnum}
+\tiny
 \caption{Object Enumeration, $O(NMQ + L)$}
@@
 
-Given the enumerated object data, we estimate $p(o \textrm{ is priv})$ as described in Algorithm \ref{alg:private}.
-Object privacy is tallied according to the type that was chosen as most probable for the image region as opposed to soft-attribution of privacy across the type probabilities.
-This is because the other types may well have resulted in very different object regions.
-
-\begin{algorithm}[htp!]\label{alg:private}
+\begin{algorithm}[hp!]\label{alg:private}
+\tiny
 	\KwData{$E^{N}$ // the enumerated object vectors}
@@
 
-Finally, all the information has been obtained to answer Task 1.  For any given pixel $i_{nm}$, we can estimate the probability of whether it lies on a private object as follows:
+%%%%%%%%% Appendix B %%%%%%%%%%%
 
-\begin{align}
-    p(C_{P_{nm}} = 1 \mid I, \textrm{ evidence}) = \sum_{k=0}^{K} p_R(i_{nm} \textrm{ on } o_k \mid I_n) P_k
-\end{align}
-
-In summary, we can estimate $p(i_m \textrm{ on } o)$ once in $O(M(NK + K^3 + TNK))$ time, regularize $p(i_m \textrm{ on } o \mid I)$ for all images in $O(NTMQ^2K)$ time, enumerate the objects in all images in $O(NMQ + L_n)$  time, estimate $p(o \textrm{ is priv})$ once in $O(N L_n K)$ time, and finally estimate the pixel privacy probability for all images in $O(N M K)$ time.
-Since $L_n < M$, total running time for estimating all pixels for all images is $O(MK^3 + NMTKQ^2)$ where $T$ and $Q$ are presumed to be small.
-
-\subsection{Task 2: Does the Image Contain Private Data?}
-
-The objective for this task can be stated as estimating $p(C_I = 1 \mid I, \textrm{ evidence})$ where $C_I$ is the binary class label for whether the image is benign (0) or private (1).
-The model formulation, basis, and algorithms for Task 1 are completely reused for Task 2.
-A simple extension to the knowledge we've built in Task 1 is applied in order to complete Task 2, as follows.
-
-Assuming there are $L_n$ objects in an image $I_n$, if even one object in the image is private, then the image contains private content.
-We can model the objective task over a Binomial distribution.
-Specifically, choose $L_n$ successes given $L_n$ trials (objects) with the probability of success being $p_s = (1-p(o_l \textrm{ is priv} \mid I, \textrm{ evidence}))$.
-The resulting probability is the complement of our objective probability.
-
-\begin{align}\label{eqn:task2.0}
-    p(C_{I_{n}} = 1 \mid I_n, \textrm{ evidence}) & = 1 - \prod_{l = 0}^{L_n} \left( 1 - p(\textrm{image region $E_{nl}$ 
-        is private} \mid I_n, \textrm{evidence}) \right)\\
-    & = 1 - \prod_{l = 0}^{L_n} \left( 1 - \sum_{k=0}^{K} (E_{nl}\textrm{ on } o_k) p(o_k \textrm{ is priv}) \right)\\
-    & = 1 - \prod_{l = 0}^{L_n} \left( 1 - \sum_{k=0}^{K} (E_{nl}\textrm{.psum}_k / E_{nl}\textrm{.cnt}) p(o_k \textrm{ is priv}) \right) \label{eqn:objpriv}
-\end{align}
-
-Notice in Equation \ref{eqn:objpriv} that the probability of an object region being of some type $k$ is represented as the arithmetical average of the $k$-probabilities across all pixels in that $l$ object region.
-This implies that individual pixel probabilities are identically distributed within a given region, which is likely not true since each is the result of belief propagation over a Markov Random Field.
-However, as mentioned in section \ref{sec:t1alg}, the actual object regions would likely have been different for different values of $k$, so use of the arithmetical mean for each type across all pixels in the dominant region is presumed (by the author) to be a reasonable compromise.
-The independent joint probabilities for object classification and class privacy are marginalized over all object types to achieve the probability of the object region being private.
-Then, the final probability of the image being private is the complement of the probability over the Binomial distribution that zero object regions are private.
-This solution relies on the entirety of Task 1, except the final calculation of pixel privacy, and then calculates the probability over the Binomial distribution in $O(N L_n K)$ time, for the same total running time of $O(MK^3 + NMTKQ^2)$ where $T$ and $Q$ are presumed to be small.
-
-%%%%%%%%%%%%%%%%% BIB %%%%%%%%%%%%%%%%%%%%
 \newpage
-\bibliographystyle{plain}
-\bibliography{references}
-\addcontentsline{toc}{section}{References}
-
-%%%%%%%%%%%%%%%%% Appendices %%%%%%%%%%%%%%%%%%%%
-\newpage
-\appendixtitleon
-\begin{appendices}
-
 \begin{center}
@@
 
-\subsection{Dr. David Crandall}\label{app:crandall}
+\subsection{Dr. David Crandall}\label{app:questions}
 
@@
 
-\subsection{Dr. Donald Williamson}
+%\subsection{Dr. Donald Williamson}
 

### 2020-05-03 paper.tex
Initial Overleaf Import

\documentclass[11pt,letterpaper]{article}
\usepackage[letterpaper, margin=1in]{geometry}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[english]{babel}
\usepackage{natbib,latexsym}
\bibliographystyle{plainnat}
\usepackage{graphicx}
\usepackage{hyperref}
\hypersetup{
    colorlinks=true,
    linkcolor=blue,
    filecolor=magenta,      
    urlcolor=cyan,
}
\title{Triplet Extraction from sentences}

\author{Sumith Baddam, Deepthi Raghu, Third Author, ... \\
  Affiliation
  {\tt \{srbaddam, draghu, email3, ...\}@iu.edu} \\}

\date{\today}

\begin{document}
\maketitle

\begin{abstract}
  In this paper we present an approach for extracting triplets from sentences using Json-NLP semantic relationship extractor and dependency parser. Triplet extraction is the processing of extracting subject-verb-object relationships from multiple sentences. This paper presents various traditional approaches previously used by others and evaluates our approach. We further compares our results with the results from traditional approaches. We would also discuss the evaluation procedure we followed and the challenges we faced in this project.
\end{abstract}

\section{Introduction}

Triplet Extraction is a crucial cog in the field of Natural Language Processing (NLP) and linguistics. It’s widely used for tasks such as Question Answering Systems, Machine Translation, Entity Extraction, Event Extraction, Named Entity Linking, Co-reference Resolution, Relation Extraction, etc.\\

A triplet represents a couple of entities and a relation between them. For example, (Damir Cavar, teach, NLP course) is a triple in which ‘Damir Cavar’ and ‘NLP course’ are the related entities, and the relation between them is ‘teach’. In this paper, we will focus on the extraction of these types of triples from a given text.\\

The main scope of this project is the triple extraction from sentences. Three relation types are explored.

\begin{itemize}
    \item (subject, verb, object)
    \item (event, occurs after, event)
    \item (event, preposition, proposition object)
\end{itemize}

This problem requires understanding large corpora, which is an increasingly popular problem these days. In this report, we will discuss and explain some Natural Language Processing (NLP) techniques that we used to extract the subject, verb and object of a particular sentence and the different approaches used in particular by our team.

\section{Previous Work}
Our research about the previous work for this task exposed us to two different approaches. Traditional triplet extraction and open triplets extraction. In Traditional Extraction, the relations to be extracted are pre-defined. In Open Extraction, the relations are not pre-defined. The system is free to extract any relations it comes across while going through the text data. We also came across different methods from some of the research papers we read. Although many of these approaches involved building a neural model, we mainly focused on approaching the problem of triple extraction using a linguistic method (rule based extraction).

\subsection{Semantic Relationships: Get Structured Knowledge from Unstructured Text}
Let's have a look at text snippet, “Food Tutorials are better when directed by Wes Anderson. Bruce Lee's biopic, 'Little Dragon' to be directed by Shekhar Kapur.” \\

In the first sentence, we have two entities (“Food Tutorials” and “Wes Anderson”). These entities are related by the term “Directed”. Hence, (Wes Anderson, directed, Food Tutorials) is a triple.Similarly, we can extract relations from the other sentences as well, (Shekhar Kapur, directed, Little dragon). It turns out that we can get structured information based on the syntactic structure and grammar of the text, as illustrated in the example above.

\subsection{Different Approaches to Triplet Extraction}
In the previous section, we managed to easily extract triples from a few sentences. However, in the real world, the data size is huge and manual extraction of structured information is not feasible. Therefore, automating this information extraction becomes important.\\

There are multiple approaches to perform information extraction automatically. Let’s understand them one-by-one:
\begin{itemize}
    \item \textbf{Rule-based Approach:} We define a set of rules for the syntax and other grammatical properties of a natural language and then use these rules to extract information from text. 
    \item \textbf{Supervised:} Let’s say we have a sentence S. It has two entities E1 and E2. Now, the supervised machine learning model has to detect whether there is any relation (R) between E1 and E2. So, in a supervised approach, the task of relation extraction turns into the task of relation detection. The only drawback of this approach is that it needs a lot of labeled data to train a model.
    \item \textbf{Semi-supervised:} When we don’t have enough labeled data, we can use a set of seed examples (triples) to formulate high-precision patterns that can be used to extract more relations from the text.

\end{itemize}

\section{Approach}
For this task, we used NLTK, JSON-NLP and requests framework. We extract structured information from unstructured data (text data in our case). We’ve already seen that the information in text lies in the form of relations between different entities. \subsection{Subtree Matching for triplet extraction}
Our methodology is similar to subtree matching for relation extraction.\\

The simple rule-based methods work well for information extraction tasks. However, they have a few drawbacks and shortcomings. We have to be extremely creative to come up with new rules to capture different patterns. It is difficult to build patterns that generalize well across different sentences. To enhance the rule-based methods for relation/information extraction, we should try to understand the dependency structure of the sentences at hand. We use JSON-NLP API for getting the dependency structure of the sentence. From the output of the API, we try to find interesting relationships. \\

For example, let's consider the sentence to be, “Tableau was recently acquired by Salesforce.” Output from the API is <Image>\\

If you look at the entities in the sentence – Tableau and Salesforce – they are related by the term ‘acquired’. So, the pattern I can extract from this sentence is either “Salesforce acquired Tableau” or “X acquired Y”.\\

Now consider another statement, “Careem, a ride-hailing major in the middle east, was acquired by Uber.” Output from the API is <Image>\\

From the graph, all we have to check is which dependency paths are common between multiple sentences. This method is known as Subtree matching. We will just consider the common dependency paths and extract the entities and the relation (acquired) between them. Hence, the relations extracted from these sentences are:
\begin{itemize}
    \item Salesforce acquired Tableau
    \item Uber acquired Careem
\end{itemize}

After extracting a basic subject - verb - object extracted from a single sentence by the above approach, we had to handle compounds such as adjectives and modifiers. Also, we had to handle sentences with multiple subject/verb/object occurrences, define a particular output structure and handle reading multiple files containing multiple sentences. In addition to this, event ordering also had to be included. The various experiments done to achieve these results have been described in the next section.

\section{Experiments}
To bring in some creativity, we developed two different rule-based methods to handle the triple extraction. One of our approaches was to extend the base code that extracts a single subject/verb/object and write rules to identify compounds such as adjectives and modifiers from the dependency tree.

\noindent
The rules are specifically based on the 'tokenlist' and the 'dependency' parts of the JSON from the NLP-API. In this approach, we take as input, a list of files containing multiple sentences line by line. The output will be a list of JSONs corresponding to each sentence in the input files. \\

\noindent
For example, if the input is a file containing two sentences:

"Mary marinated the chicken."

"John cooked the chicken."

\noindent \newline
Then the output will be:

\{'sentence': 1, 'event': 1, 'subject': 'Mary', 'verb': 'marinate', 'object': 'chicken'\}

\{'sentence': 2, 'event': 2, 'subject': 'John', 'verb': 'cook', 'object': 'chicken'\}

\noindent \newline
Explanation of output:
\begin{itemize}
    \item The sentence key in the output JSON provides a reference to the sentence number in the input sentence file.

    \item The event key in the output JSON indicates the event ordering i.e, the ordering of the predicates.

    \item The subject, verb and object keys indicate the triples that have been extracted.
\end{itemize}
The rules were extended as and when new scenarios and variations of the triples have to be handled.
Below are some example sentences from file15.txt in webanno to show the different variations handled:
\begin{enumerate}
    \item
    \textbf{Sentence:} Prepare the cake mixes according to the package directions. \newline
    \textbf{Variation in triple:} This sentence has a compound object 'cake mixes'.
    \newline
    \textbf{Extracted output:}

    \{'sentence': 1, 'event': 1, 'subject': '<imperative-chef>', 'verb': 'prepare', 'object': 'cake mix'\}
    \newline
    \textbf{Inference from output:} 
    An important point to note is that in sentences that do not have a subject, we replace the subject with "<imperative-chef>". In this particular example, we can see that the compound object "cake mix" has been extracted successfully.

    \item
    \textbf{Sentence:} Stir to combine, then add the milk, cream and butter.
    \newline
    \textbf{Variation in triple:} This sentence has a multiple objects - milk, cream, butter
    \newline
    \textbf{Extracted output:}

    \{'sentence': 1, 'event': 1, 'subject': '<imperative-chef>', 'verb': 'add', 'object': 'milk'\}

    \{'sentence': 1, 'event': 1, 'subject': '<imperative-chef>', 'verb': 'add', 'object': 'cream'\}

    \{'sentence': 1, 'event': 1, 'subject': '<imperative-chef>', 'verb': 'add', 'object': 'butter'\}

    \textbf{Inference from output:} 
    Whenever there are multiple subjects/verbs/objects, our approach will naively produce all combinations of triples. In this case, the three objects, milk, cream and butter have been identified and put into three different triples. Since all the extracted triples are from the same sentence, the sentence key remains unchanged.

    \item
    \textbf{Sentence:} Cover and cook until cauliflower begins to soften, about 4 minutes.
    \newline
    \textbf{Variation in triple:} This sentence has a multiple verbs - cover, cook and it has no object
    \newline
    \textbf{Extracted output:}

    \{'sentence': 1, 'event': 1, 'subject': 'cauliflower', 'verb': 'cover', 'object': '<None>'\}

    \{'sentence': 1, 'event': 2, 'subject': 'cauliflower', 'verb': 'cook', 'object': '<None>'\}

    \textbf{Inference from output:} 
    Whenever there are multiple subjects/verbs/objects, our approach will naively produce all combinations of triples. In this case, the two verbs cover and cook have been identified and put into two different triples. Also, since this sentence does not have an object, we replace the object with '<None>'. Another point to note here is that since there are two verbs in this sentence, the event changes accordingly, capturing the order in which the actions occurred.

\end{enumerate}

\noindent
In this manner, multiple files containing multiple sentences line by line can be processed to produce an output JSON.

\noindent \newline
\textbf{Steps to run the code for this experiment:}
\begin{itemize}
    \item Run the python code \textbf{extract\_triples\_deepthi.py} using the command:

    \textbf{python extract\_triples\_deepthi.py file1.txt file2.txt ...}

    where, file1.txt, file2.txt .. are the input files

    \item After the execution of code, the result is generated in the file \textbf{Result\_JSON.txt}
\end{itemize}

\noindent
In order to standardize the outputs of all the experiments, we decided to utilize the graph.py code to give a graph object as the output. To incorporate this output format, we modified the existing \textbf{extract\_triples\_deepthi.py} into \textbf{extract\_triples\_graph\_deepthi.py} and constructed two relations:

\begin{itemize}
    \item Triples (Subject, Verb, Object)
    \item Event relation (Event2 occurs\_after Event1)
\end{itemize}

\noindent
The graph.py enabled us to produce two types of results:
\begin{enumerate}
    \item A graph object represented as a JSON 
    \item A line by line relation extraction
\end{enumerate}
\noindent
We produced both these outputs. Below is an example:
\noindent \newline
\textbf{Sentences:}

Prepare the cake mixes according to the package directions.

Stir to combine, then add the milk, cream and butter.

\noindent 
\textbf{Extracted output1 : A graph object represented as a JSON } 
\newline

\includegraphics[width=9cm, height=9cm]{graph_obj1.png}
\includegraphics[width=9cm, height=9cm]{graph_obj2.png}

\noindent
\textbf{Extracted output2 : A line by line relation extraction }

\begin{flushleft}
\{'source': '<imperative-chef>', 'relation': 'prepare', 'target': 'cake mix'\}

\{'source': '<imperative-chef>', 'relation': 'add', 'target': 'milk'\}

\{'source': '<imperative-chef>', 'relation': 'add', 'target': 'cream'\}

\{'source': '<imperative-chef>', 'relation': 'add', 'target': 'butter'\}

\{'source': '2\_<imperative-chef>\_add\_milk', 'relation': 'occurs\_after', 'target': '1\_<imperative-chef>\_prepare\_cake mix'\}

\{'source': '3\_<imperative-chef>\_add\_cream', 'relation': 'occurs\_after', 'target': '2\_<imperative-chef>\_add\_milk'\}

\{'source': '4\_<imperative-chef>\_add\_butter', 'relation': 'occurs\_after', 'target': '3\_<imperative-chef>\_add\_cream'\}

\end{flushleft}
\noindent 
\textbf{Inference from output1:}

In the graph object, Subjects, Objects and Events are constructed as nodes. An edge is drawn in two cases:
\begin{enumerate}
    \item Edge over a verb, connecting a subject and an object to capture triple relations
    \item Edge over "occur\_after", connecting two events to capture event relations
\end{enumerate}
\noindent 
\textbf{Inference from output2:}

This output gives a line by line relation extraction. It extracts two relations:

\begin{itemize}
    \item Triple relations like \{'source': '<imperative-chef>', 'relation': 'add', 'target': 'milk'\}
    \item Event relations like \{'source': '3\_<imperative-chef>\_add\_cream', 'relation': 'occurs\_after', 'target': '2\_<imperative-chef>\_add\_milk'\}

\end{itemize}
\noindent \newline
Also, in this experiment, the event ID has been generated by concatinating the verb ID from its corresponding graph node with with subject, verb and object as "verbID\_subject\_verb\_object to get a unique event ID.

\noindent \newline
\textbf{Steps to run the code for this experiment:}
\begin{itemize}
    \item Run the python code \textbf{extract\_triples\_graph\_deepthi.py} using the command:

    \textbf{python extract\_triples\_graph\_deepthi.py file1.txt file2.txt ...}

    where, file1.txt, file2.txt .. are the input files

    \item After execution of the code, the graph object (output1) is generated in \textbf{output\_relations\_graph.json} and the line by line relation extraction (output2) is generated in \textbf{output\_relations\_triples.txt}
\end{itemize}
\section{Experiments}

Our experiments...

\section{Evaluation}

This is how we evaluated our work, presenting the evaluation scores in comparison to previous projects...

\section{Discussion}

The discussion of the issues and...

\subsection{Conclusion}

Some conclusion...

% the bibliography
\bibliography{main}

\end{document}

### 2020-05-05 paper.tex
Stein finished his experiments section with examples.

@@
 
-\author{Sumith Baddam, Deepthi Raghu, Third Author, ... \\
+\author{Sumith Baddam, Deepthi Raghu, John Stein, ... \\
   Affiliation
-  {\tt \{srbaddam, draghu, email3, ...\}@iu.edu} \\}
+  {\tt \{srbaddam, draghu, jodstein, ...\}@iu.edu} \\}
 
@@
 \begin{itemize}
-    \item (subject, verb, object)
-    \item (event, occurs after, event)
-    \item (event, preposition, proposition object)
+    \item Verb Relation: (subject, verb, object)
+    \item Temporal Relation: (event, occurs after, event)
+    \item Verb-Preposition Relation: (event, preposition, proposition object)
 \end{itemize}
@@
 This problem requires understanding large corpora, which is an increasingly popular problem these days.
-In this report, we will discuss and explain some Natural Language Processing (NLP) techniques that we used to extract the subject, verb and object of a particular sentence and the different approaches used in particular by our team.
+In this report, we will discuss and explain some Natural Language Processing (NLP) techniques that we used to extract these relations from sentences and the different approaches used in particular by our team.
 
@@
 We also came across different methods from some of the research papers we read.
-Although many of these approaches involved building a neural model, we mainly focused on approaching the problem of triple extraction using a linguistic method (rule based extraction).
+Although many of these approaches involved building a neural model, we mainly focused on approaching the problem of triple extraction using a linguistic method (rule based extraction) because no labeled training data were available.
 
@@
 
-\section{Experiments}
-To bring in some creativity, we developed two different rule-based methods to handle the triple extraction.
-One of our approaches was to extend the base code that extracts a single subject/verb/object and write rules to identify compounds such as adjectives and modifiers from the dependency tree.
+\section{Experiments (draghu)}
+This section discusses the experiments and implementation done by Deepthi Raghu.
 
-\noindent
+\noindent \newline
+One of the approaches for rule-based triple extraction was to perform dependency parsing by writing rules to identify compounds such as adjectives and modifiers from the dependency tree.
 The rules are specifically based on the 'tokenlist' and the 'dependency' parts of the JSON from the NLP-API.
-In this approach, we take as input, a list of files containing multiple sentences line by line.
-The output will be a list of JSONs corresponding to each sentence in the input files.
-\\
 
+\subsection{Method 1}
+In this approach, the input is a list of files containing multiple sentences line by line. The output will be a list of JSONs corresponding to each sentence in the input files. \\
 \noindent
@@
     \textbf{Inference from output:} 
-    An important point to note is that in sentences that do not have a subject, we replace the subject with "<imperative-chef>".
-In this particular example, we can see that the compound object "cake mix" has been extracted successfully.
+    An important point to note is that in sentences that do not have a subject, the subject is replaced with "<imperative-chef>".
+In this particular example, the output shows that the compound object "cake mix" has been extracted successfully.
 
@@
     \textbf{Inference from output:} 
-    Whenever there are multiple subjects/verbs/objects, our approach will naively produce all combinations of triples.
+    Whenever there are multiple subjects/verbs/objects, this approach will naively produce all combinations of triples.
 In this case, the three objects, milk, cream and butter have been identified and put into three different triples.
@@
     \textbf{Inference from output:} 
-    Whenever there are multiple subjects/verbs/objects, our approach will naively produce all combinations of triples.
+    Whenever there are multiple subjects/verbs/objects, this approach will naively produce all combinations of triples.
 In this case, the two verbs cover and cook have been identified and put into two different triples.
-Also, since this sentence does not have an object, we replace the object with '<None>'.
+Also, since this sentence does not have an object, the object is replaced with '<None>'.
 Another point to note here is that since there are two verbs in this sentence, the event changes accordingly, capturing the order in which the actions occurred.
@@
 \noindent
-In this manner, multiple files containing multiple sentences line by line can be processed to produce an output JSON.
+In this manner, multiple files containing multiple sentences line by line is processed to produce an output JSON.
 
@@
 \begin{itemize}
-    \item Run the python code \textbf{extract\_triples\_deepthi.py} using the command:
+    \item Run the python code \textbf{extract\_triples\_draghu.py} using the command:
 
-    \textbf{python extract\_triples\_deepthi.py file1.txt file2.txt ...}
+    \textbf{python extract\_triples\_draghu.py file1.txt file2.txt ...}
 
@@
 
-    \item After the execution of code, the result is generated in the file \textbf{Result\_JSON.txt}
+    \item After the execution of code, the result is generated in the file \textbf{method1\_relations\_draghu.txt}
 \end{itemize}
 
+\subsection{Method 2}
 \noindent
-In order to standardize the outputs of all the experiments, we decided to utilize the graph.py code to give a graph object as the output.
-To incorporate this output format, we modified the existing \textbf{extract\_triples\_deepthi.py} into \textbf{extract\_triples\_graph\_deepthi.py} and constructed two relations:
+In order to standardize the outputs of all the experiments, the graph.py code was utilized to give a graph object as the output.
+To incorporate this output format, the existing code \textbf{extract\_triples\_draghu.py} was modified to \textbf{extract\_triples\_graph\_draghu.py} and two relations were constructed:
 
@@
 \begin{enumerate}
-    \item A graph object represented as a JSON 
+    \item A graph object with nodes and edges represented as a JSON 
     \item A line by line relation extraction
@@
 \noindent
-We produced both these outputs. Below is an example:
+Below is an example:
 \noindent \newline
@@
 
-\includegraphics[width=9cm, height=9cm]{graph_obj1.png}
-\includegraphics[width=9cm, height=9cm]{graph_obj2.png}
+\includegraphics[width=9cm, height=7.75cm]{graph_obj1.png}
+\includegraphics[width=9cm, height=7.75cm]{graph_obj2.png}
 
@@
 \noindent \newline
-Also, in this experiment, the event ID has been generated by concatinating the verb ID from its corresponding graph node with with subject, verb and object as "verbID\_subject\_verb\_object to get a unique event ID.
+Also, in this experiment, the event ID has been generated by concatenating the verb ID from its corresponding graph node with with subject, verb and object as "verbID\_subject\_verb\_object to get a unique event ID.
 
@@
 \begin{itemize}
-    \item Run the python code \textbf{extract\_triples\_graph\_deepthi.py} using the command:
+    \item Run the python code \textbf{extract\_triples\_graph\_draghu.py} using the command:
 
-    \textbf{python extract\_triples\_graph\_deepthi.py file1.txt file2.txt ...}
+    \textbf{python extract\_triples\_graph\_draghu.py file1.txt file2.txt ...}
 
@@
 
-    \item After execution of the code, the graph object (output1) is generated in \textbf{output\_relations\_graph.json} and the line by line relation extraction (output2) is generated in \textbf{output\_relations\_triples.txt}
+    \item After execution of the code, the graph object (output1) is generated in
+
+    \textbf{method2\_relations\_graph\_draghu.json} and the line by line relation extraction (output2) is generated in \textbf{method2\_relations\_triples\_draghu.txt}
 \end{itemize}
-\section{Experiments}
+\newpage
 
-Our experiments...
+\section{Experiments (jodstein)}\label{jodstein_experiments}
+
+This section discusses the evolution of approaches and experiments undertaken by John Stein.
+
+\subsection{A brief discussion of Neural approaches}
+The initial approach considered was to build a neural network, or set of neural networks, trained to extract various relation types from a sample text.
+The primary advantage of a neural network approach is that they can be arbitrarily complex (number of hidden layers, number of nodes per hidden layer, any architecture) and so are able to learn very rich (semantically-speaking) representations of inputs and produce outputs (extractions, classifications) that would not be achievable through lexical approaches.
+However, a neural approach implies several pre-requisites and intermediate tasks:
+
+\begin{itemize}
+
+    \item Labeled Data - Neural networks must have access to labeled data so that errors can be identified and their tensor values can be correspondingly adjusted.
+The data available to the relation extraction group was not labeled, and was relatively small (in comparison to many other neural network datasets).
+These constraints would have resulted in having to annotate relations for the entire dataset, creation of negative (false) relations to aid in the training process, and placing limitations on the number of relation types that could be extracted.
+
+    \item Architecture - The architecture plays a central role to a network's performance against the assigned task.
+For example, recurrent networks have the ability to carry information from previous inputs and network state into subsequent inputs, which is particularly well suited for linear sequential input types.
+Language is a little more complicated than that, so more complicated structures such as Long Short-Term Memory (LSTM) \cite{article}, Gated Recurrent Units (GRU) \cite{DBLP:journals/corr/ChoMGBSB14}, or Attention \cite{bahdanau2014neural} are likely to perform better.
+
+    \item Input and Data Representation - It is best practice when using neural models for NLP tasks to represent the input text as a feature vector vice raw text.
+This is because the words, characters, and character encodings aren't defined in a way that is at all relevant to the semantic meaning of the text.
+Multiple approaches for encodings have been used with success, including some popular examples such as Word2Vec \cite{mikolov2015computing}, FastText \cite{DBLP:journals/corr/JoulinGBDJM16}, and Bidirectional Encoder Representations from Transformers (BERT) \cite{DBLP:journals/corr/abs-1810-04805}.
+
+    \item External Data - Features derived from annotations (parts of speech, named entities, dependencies, constituency, etc.) may provide valuable in the automated learning task, so an encoded representation would need to be learned.
+Further, external data sources such as ontologies, knowledge graphs, or databases may help a great deal in performance of the neural model, if they can be leveraged.
+FrameNET, in particular, contains very powerful information about semantic concepts, how they are most often presented in text, and what their argument structure looks like.
+Use of FrameNET, or any of these external data, in a neural model might be accomplished in a way similar to word embeddings mentioned above but, instead of training the embeddings to predict co-occurrence of other words, the embeddings would be trained to classify the conceptual frame being represented by the sentence or word span.
+
+\end{itemize}
+
+\subsection{The Implemented Approach}
+
+This section discusses the approach and associated experiments undertaken by John Stein.
+Code for these experiments can be found in \texttt{ML4NLP/group3/} as \texttt{extract\_jodstein\_v0.py}, or previous committed versions as \texttt{extract\_jodstein.py}.
+File can be run from the ML4NLP/group3 folder in the console using \texttt{python extract\_jodstein\_v0.py <path-for-file(s)>}.
+
+\begin{enumerate}
+
+    \item Constituency Parsing - The first experiments were based on a constituency parse of the sentence.
+The thinking behind using constituency parsing was that, for complex sentences, an entire phrase or clause may be playing the role of semantic subject, verb, or object in the context of a given sentence.
+The theory behind following the constituency parse was that a recursive algorithm could inspect clauses and phrases for arbitrarily-complex sentences, and could build some context representation from the top-level on down into the recursion so as to enable more powerful entity and relation extraction at the deepest levels.
+This is similar to the notion of tail recursion where an environment is continuously built up through the recursion process and when actual extraction occurs, it happens in the presence of a well-structured trace through the constituency parsing, ignoring sentence elements that are not a part of that trace.
+This approach turned out to be fraught with complicating factors, including (a) constituency labels don't themselves reveal even the lexical subject/object parts of a sentence, much less the semantic parts, and (b) the representation of context down through the recursion remained a task-too-far without the benefit of embeddings or semantically-encoded represetnations.
+This work can be viewed under commit \texttt{27bfd29394075086c777625976caadb525404748}, but was never complete to the point it would run.
+
+    \item Dependency Parsing - The next approach taken, based on feedback from the professor, was a Dependency-based approach.
+This approach is much simpler because the subjects and objects of verbs are already labeled/annotated.
+However, the successful extraction of relations is inextricably tied to the quality of the dependency annotations.
+Initial dependency-based success simply extracted verb-relations using the verb token as the relation, dependent tokens labeled with \texttt{nsubj} that were governed by that verb as the subjects, and dependent tokens labeled with \texttt{dobj} that were goverened by the verb as the objects.
+The next success was to successfully trace any subject or object token to include any other token with which it shared a compound chunk.
+Further, many verbs have multiple subjects and/or multiple objects via conjunctions.
+To handle this, the algorithm permutates over every combination of subject and noun around the verb that serves as the governer for both.
+Hence, all verbs resulted in $ \mid subjects \mid \times \mid objects \mid $ verb-relations.
+Show below is an example sentence with extracted relations:
+
+        \begin{verbatim}
+            "This vegan salad makes a healthy and filling meal, providing 4 of
+            your 5-a-day."
+
+            Extracted: (vegan salad,  make,  meal)
+        \end{verbatim}
+
+    This approach produced around 100 relations from the Corpus, and can be viewed/run under commit \texttt{86aea17093c51cf51fcb3669de0750fc1f413766}.
+
+    \item Temporal Relations and Dummy Values - The next step taken was to include the notion of event timelines.
+This was at the specific request of Group 4, who wanted to demonstrate some event-based knowledge representation in their graph work.
+To capture event timelines, the notion of an event must first be established.
+Essentially, the approach was to create an event node/entity for each verb-relation and embed the Edge.id from the corresponding verb-relation into the event node's label.
+Next, the algorithm keeps a short term memory containing the set of events corresponding to the last verb encountered.
+Finally, for every new verb encountered, the algorithm creates \texttt{occurs\_after} temporal-relations between the events associated with the current verb and those in short term memory.
+The short term memory is cleared between documents since there is no temporal relationship between entirely separate text sources.
+Also, the group noticed that any verbs that didn't have both a noun subject and a direct object were not being captured.
+Since the corpora consisted of recipes, many of the statements were imperative without a subject.
+To solve this, the group decided to create a dummy entity \texttt{<imperative-chef>} to serve as the subject for imperative statements, and a dummy entity \texttt{<None>} to serve as the direct object when one wasn't present.
+Show below is an example sentence with extracted relations:
+
+    \begin{verbatim}
+        "With the mixer on low speed, add 1/2 cup of the flour at a time, 
+            waiting for each addition to be fully incorporated before 
+            adding more."
+
+        Extracted: (<imperative-chef>,  add,  cup)
+        Extracted: (<imperative-chef> ,  wait ,  <None>)
+        Extracted: (5009_<imperative-chef>_wait_<None>,  occurs_after,  
+            5005_<imperative-chef>_add_cup)
+        Extracted: (<imperative-chef>,  be,  <None>)
+        Extracted: (5012_<imperative-chef>_be_<None>,  occurs_after,  
+            5009_<imperative-chef>_wait_<None>)
+        Extracted: (<imperative-chef> ,  add ,  more)
+        Extracted: (5014_<imperative-chef>_add_more,  occurs_after,  
+            5012_<imperative-chef>_be_<None>)
+    \end{verbatim}
+
+    The addition of Temporal relations and other dummy values resulted in around 300 relations from the Corpus, and can be viewed/run under commit \texttt{3ea35b27eb3aa289fbc5a0f5bdaf06e053178131}.
+
+    \item Prepositional Relations - The next step taken was to include prepositional relations.
+This task was somewhat straightforward to implement, having become familiar with how to navigate the dependency tree.
+Essentially, after defining verb-relations, the algorithm gathers all tokens (prepositions) that are governed by that same verb under a \texttt{prep} dependency type and iterates on each.
+For each preposition, a lit of tokens dependent on that preposition under a \texttt{pobj} dependency type (with compounds) are then collected as objects for that preposition.
+Finally, a set of relations are created with the events corresponding to the verb as the sources, the prepositions as the relation labels, and the propositions' objects as the relation targets, permutated over all combinations as before.
+Show below is an example sentence with extracted relations:
+
+    \begin{verbatim}
+        "Make bread crumbs by placing the bread in the work bowl of a food 
+            processor and processing until finely ground."
+
+        Extracted: (<imperative-chef>,  make,  <None>)
+        Extracted: (4968_<imperative-chef>_make_<None>,  occurs_after,  
+            4966_<imperative-chef>_remain_<None>)
+        Extracted: (<imperative-chef>,  place,  bread)
+        Extracted: (4970_<imperative-chef>_place_bread,  occurs_after,  
+            4968_<imperative-chef>_make_<None>)
+        Extracted: (4970_<imperative-chef>_place_bread,  in,  work bowl)
+        Extracted: (4970_<imperative-chef>_place_bread,  until,  grind)
+        Extracted: (<imperative-chef>,  process,  <None>)
+        Extracted: (4974_<imperative-chef>_process_<None>,  occurs_after, 
+            4970_<imperative-chef>_place_bread)
+    \end{verbatim}
+
+    This approach resulted in around 5,000 relations from the Corpus, and can be viewed/run under commit \texttt{f636830a9751e43732a27ede19d0ac02b64f45a1}.
+
+    \item Dependency with Verb-as-Entity - The final significant change to the relation extractor was to regard verb tokens as entities, rather than as relations.
+As stated above in the Temporal Relations step, the current method for linking a verb-relation to an event was to include the verb-relation identification number in the event label.
+That hack may serve the immediate purpose, but is counter to the purist aspects of knowledge management.
+An improved approach is recognize that each verb is an entity in its own right, representing an activity, action, or being, and doesn't get its meaning from any one event instance.
+So, after creating an entity for each verb encountered (the idea of the verb) as well as an event for each verb encountered (the instance of a verb being involved in a specific event), it is possible to add the relation \texttt{involves\_verb} which ties the verb to the event.
+In this way, the event also becomes the center for all subjects, objects, and prepositional statements.
+Afterall, these other entities are only linked to each other through the event in which they all had a role.
+This change was implemented late in the project, but is not planned for processing by the graph subgroup due to the fundamental change in the way even the most basic relations are captured.
+Show below is an example sentence with extracted relations:
+
+    \begin{verbatim}
+        "Sprinkle the cream of tartar over the egg whites and continue to 
+            beat until they hold a soft peak."
+
+        Extracted: (34_Sprinkle, involves_verb, Sprinkle) 
+        Extracted: (34_Sprinkle, has_object, cream)
+        Extracted: (37_continue, involves_verb, continue)
+        Extracted: (37_continue, occurs_after, 34_Sprinkle)
+        Extracted: (39_beat, involves_verb, beat)
+        Extracted: (39_beat, occurs_after, 37_continue)
+        Extracted: (41_hold, involves_verb, hold)
+        Extracted: (41_hold, has_subject, -PRON-)
+        Extracted: (41_hold, has_object, peak)
+        Extracted: (41_hold, occurs_after, 39_beat)
+    \end{verbatim}
+
+    This approach resulted in around 6,100 relations from the Corpus, and can be viewed/run under commit \texttt{c002c099852dce47468a7d366e614f4514e9f146} as \texttt{extractor\_jodstein\_v1.py} (the previous approach continues to be maintained as \texttt{extractor\_jodstein\_v0.py}).
+
+\end{enumerate}
 
@@
 
+Describe manual evaluation methodology: (1) look at sentence, (2) observe extracted triplets, (3) observe annotations in context of how our rule-based approach is programmed.
+Walk through several examples.
+
 \section{Discussion}
 
-The discussion of the issues and...
+\subsection{Knowledge Graph vs Ontology}\label{ont_vs_kg}
+
+Current popular opinion \cite{science_where_2018, schrader_whats_2020} has the definitions for ontology and knowledge graph loosely differentiated.
+The key difference, it would appear, has ontologies capturing schema, while knowledge graphs capture facts.
+Stated another way, an ontology may be the tool of choice to capture or define how a universe of relevant entities (classes, attributes, relations, rules, etc.) are related to one another, while a knowledge graph may be the tool of choice to capture factual knowledge in a structured way as it is encountered in the real world.
+The difference here is not clear in all cases.
+Take a simple example: "The police officer ordered a custard-filled donut."  Is "police officer" an instance of "person::employee::law\_enforcement", or is it a concept in its own right.
+A common approach is to treat "police officer" as an instance of an existing entity until such time that it becomes necessary/desirable to reason about "police officer" in relation to other entities.
+From an ontological perspective, the schema continues to be defined at lower levels as more granular schemas and relations emerge from real data.
+From a knowledge graph perspective, there must be an expectation that the knowledge graph is re-defined when new schemas emerge and subsequently used to better organize and reason over the facts captured in real data.
+
+\subsection{Everything as Entities}
+
+One of the challenges encountered during the progression of various approaches to triple extraction relates to pre-suppositions about what the resulting relation set might contain, and how one might want to reason over those relations.
+More specifically, the initial expectation among many in the class was that the relations extracted would always be of the form (subject, verb, object).
+While it is true that this was the initial approach (easiest to accomplish), this expectation resulted in some implementation decisions on the part of group 3 and the API group which may constrain the versatility and power of the knowledge graph representation.
+For example, as discussed in Section \ref{jodstein_experiments} above, the initial choice to extract verbs as the relation versus as an independent entity constrained the resulting knowledge graph in several ways:
+
+\begin{enumerate}
+    \item Could not define temporal relations between verbs
+    \item Could not define preposition relations between verbs and preposition objects
+    \item Could not reason about verbs as concepts (identifying all the objects of 'chop')
+    \item Could not truly link verb-relations to the concept of an event (using graph methodology)
+\end{enumerate}
+
+Although some of these constraints were overcome, the means by which they were overcome involve, in some cases, implicit knowledge about how the relations were constructed and how to interpret them.
+In other words, the reasoning engine relies on implicit knowledge about the relations it is reasoning over (e.g.
+the verb-relation id embedded in event labels or knowledge that verb-relation ids are representative of their temporal placement in the text).
+Through this exercise, a couple of important lessons emerged which build on the concepts discussed in Section \ref{ont_vs_kg}.
+Firstly, a pattern emerges wherein as the desire to learn and reason over more abstract and semantically-rich concepts and notions increases, so does the necessity to represent these higher-level concepts as entities.
+This is necessary because in order to capture more subtle nuances between texts, the reasoning engine must operate at higher levels of semantic meaning and/or abstraction, and so relations must be drawn at those levels to enable that reasoning.
+Secondly, a design requirement for the reasoning engine should be that it can reason over arbitrary relation types.
+Obviously the queries for the engine will likely be specific to the set of relation types that exist in the knowledge graph, but the engine logic should not be.
 

### 2020-05-06 paper.tex
jodstein added to conclusion section

@@
 Our research about the previous work for this task exposed us to two different approaches.
-Traditional triplet extraction and open triplets extraction.
+Traditional triplet extraction and Open triplets extraction.
 In Traditional Extraction, the relations to be extracted are pre-defined.
@@
 
-\section{Approach}
+\section{Experiments (srbaddam)}
 For this task, we used NLTK, JSON-NLP and requests framework.
@@
 We’ve already seen that the information in text lies in the form of relations between different entities.
-\subsection{Subtree Matching for triplet extraction}
-Our methodology is similar to subtree matching for relation extraction.\\
+\subsection{Sub-tree Matching for triplet extraction}
+Our methodology is similar to sub-tree matching for relation extraction.\\
 
@@
 
-For example, let's consider the sentence to be, “Tableau was recently acquired by Salesforce.” Output from the API is <Image>\\
+For example, let's consider the sentence to be, “Tableau was recently acquired by Salesforce.” Output from the API is, \\
+
+\includegraphics[width=9cm, height=7.75cm]{image1.PNG}
 
@@
 
-Now consider another statement, “Careem, a ride-hailing major in the middle east, was acquired by Uber.” Output from the API is <Image>\\
+Now consider another statement, “Careem, a ride-hailing major in the middle east, was acquired by Uber.” Output from the API is,\\
+
+\includegraphics[width=15cm, height=9cm]{image2.PNG}
 
@@
 
-    \item Dependency with Verb-as-Entity - The final significant change to the relation extractor was to regard verb tokens as entities, rather than as relations.
+    \item Verb-as-Entity - The final significant change to the relation extractor was to regard verb tokens as entities, rather than as relations.
 As stated above in the Temporal Relations step, the current method for linking a verb-relation to an event was to include the verb-relation identification number in the event label.
@@
 
-\section{Evaluation}
+\section{Evaluation (rakoduru)}
 
@@
 
-Some conclusion...
+This project produced several approaches to relation extraction.
+Constraints on the data size and the lack of relation labels for training purposes prevented the use of supervised learning methods for neural approaches.
+Still, several challenges surfaced during the project.
+Some of these challenges pertained to the approach taken (John's constituency parsing approach), while other challenges pertained to implementation and code - getting the extractor to execute the designer's intent.
+The most pronounced challenge, arguably, was the coordination that was required between multiple teams toward an overall pipeline.
+The individuals within the Relation Extraction group found it difficult at times to keep a consistent format, as new design requirements often spawned impacts to the extractor output.
+Further, the interface between the Relation Extraction Group and the API group became a point of contention on more than one occasion.
+Partially, this was due to the open extraction approach that was taken.
+In other words, a pre-defined set of relation types was not established a priori, but rather each member was given the liberty to experiment with different relation types as inspiration dictated.
+
+Despite these challenges, the Relation Extraction group successfully extracted many relations from the corpus, and gained valuable experience and lessons learned to be used on future data science and language processing projects.
 

### 2020-05-08 paper.tex
Final commit for Group 3

@@
 
-\author{Sumith Baddam, Deepthi Raghu, John Stein, ... \\
+\author{Sumith Baddam, Deepthi Raghu, John Stein, Manisha Suresh Kumar, Rasmitha Choudary Koduru \\
   Affiliation
-  {\tt \{srbaddam, draghu, jodstein, ...\}@iu.edu} \\}
+  {\tt \{srbaddam, draghu, jodstein, msureshk, rakoduru\}@iu.edu} \\}
 
@@
 
-\section{Previous Work}
+\section{Previous Work (srbaddam)}
 Our research about the previous work for this task exposed us to two different approaches.
@@
 
-\noindent \newline
+\noindent
 One of the approaches for rule-based triple extraction was to perform dependency parsing by writing rules to identify compounds such as adjectives and modifiers from the dependency tree.
@@
 \subsection{Method 1}
-In this approach, the input is a list of files containing multiple sentences line by line. The output will be a list of JSONs corresponding to each sentence in the input files. \\
-\noindent
+In this approach, the input is a list of files containing multiple sentences line by line.
+The output will be a list of JSONs corresponding to each sentence in the input files.
 For example, if the input is a file containing two sentences:
@@
 \newpage
+\section{Experiments (msureshk)}
+
+Adding on to the previous experiment, this is Manisha Suresh's approach to extracting prepositional relation.
+Before we talk more about the implementations, the below are the 'carry forwards' from the previous experiment:
+\begin{itemize}
+    \item \textbf{The sequential parsing of token list:}
+
+    The token list i indexed from 0 - (number of words in input text - 1) and we sequentially parse through each word to slot them into objects, subjects, verb or prepositions and also identify relations.
+
+    \item \textbf{Sentences without subjects, have subject tagged as “<imperative-chef>":}
+
+    Often, sentences may not have a subject. below is one such example from our food corpus:
+    "Prepare the cake mixes according to the package directions."
+
+    So, we use "<imperative-chef>" as a placeholder while also being relevant to the context.
+    Extracted relation : \{'source': 'imperative-chef', 'relation': 'prepare', 'target': 'cake mix'\}
+
+    \item \textbf{Sentences without objects, have object tagged as ’<None>’:}
+
+    Similar to “<imperative-chef>", we use ’<None>’ to indicate the absence of an object.
+
+    "Cook until cauliflower begins to soften, about 4 minutes."
+
+    Extracted Relation : {'source': 'cauliflower', 'relation': 'cook', 'target': '<None>'}
+
+    \item \textbf{Utilizing graph.py to give a graph object as the output:}
+
+    \begin{itemize}
+        \item Triples (Subject, Verb, Object)
+        \item Event relation (Event2 occurs\_after Event1)
+        \item Prepositional Relations (Event, preposition, object)
+    \end{itemize}
+     \item \textbf{Generation of Event ID:}
+
+      Event ID is created by concatenating the ID of that node with its subject and object
+
+      An Example of one such event would be: '1\_imperative-chef\_prepare\_cake mix'
+\end{itemize}
+This section has two parts to it, both of which are handled in \textbf{'relations\_msureshk.py'}:
+\begin{itemize}
+    \item To take care of some unrelated triplets that get generated as was mentioned in the previous experiment.
+    \item To capture Prepositional Relations to add on to the triples.
+\end{itemize}
+Next, we shall discuss this in detail.
+\subsection{Part 1 : Coherent Triple Extraction}
+With the previous experiments, we successfully extracted all possible triples.
+Concretely, all the verbs, objects and subjects were captured and then a mechanism where from these three lists, all possible combinations and permutations of triples was implemented.
+Though this would contain the correct triples, it would also contain some incoherent ones. Below is one such example:
+\newpage
+\textbf{Sentence : }
+
+"Cook the fish 5 to 6 minutes and gently break up the haddock into chunks."\\
+
+\textbf{Extracted Triples : }
+
+\{'source': 'imperative-chef', 'relation': 'cook', 'target': 'fish'\}
+
+\{'source': 'imperative-chef', 'relation': 'cook', 'target': 'haddock'\}
+
+\{'source': 'imperative-chef', 'relation': 'break', 'target': 'fish'\}
+
+\{'source': 'imperative-chef', 'relation': 'break', 'target': 'haddock'\}\\
+
+To tackle this, we extract triples as we go instead of 'collecting' them for later.
+This helps with capturing prepositional relations as well.
+We assume that when prepositions or a punctuation is encountered, the subject-verb-object triple captured till then will be one triple.
+Then the verb is discarded to allow a new triple to form.
+So, working with the previous sentence, the new triples extracted are:\\
+
+\textbf{\{'source': 'imperative-chef', 'relation': 'cook', 'target': 'fish'\}}
+
+\textbf{\{'source': 'imperative-chef', 'relation': 'break', 'target': 'haddock'\}}\\
+
+Although this did away with the extra triples, it also would miss out on triples in certain kinds of sentences.
+Recall that we discard verbs as and when a triple is extracted.
+To implement this, at a time, we can work with one verb only.
+So if a sentence has consecutive verbs, only the latest verb would be captured and any possible triples with the missed out verbs will also be missed out.
+Were we using the approach adopted in Deepthi's experiment, none of the triples would be missed out.
+To understand this better here's a comparison between the two approaches.\\
+
+\textbf{Sentence:}\\
+
+"Cover and cook until cauliflower begins to soften."\\
+
+\textbf{Previous Extraction:}\\
+
+\{'source': 'cauliflower', 'relation': 'cover', 'target': '<None>'\}
+
+\{'source': 'cauliflower', 'relation': 'cook', 'target': '<None>'\}\\
+
+\textbf{Current Extraction:}\\
+
+\textbf{\{'source': 'cauliflower', 'relation': 'cook', 'target': '<None>'\}}\\
+
+Note how the verb 'cover' wasn't captured as it was immediately succeeded without prepositions or punctuations by another verb.
+In short, this mechanism works well for parts of sentences with single verbs and for ones with more than one, only the triple with the latest verb will be captured.\\
+
+\section{Part 2 : Prepositional Relations}
+
+Keeping up with the same logic, we can now extract prepositional relations. Lets consider the following sentence:
+
+\textbf{"Prepare the cake mixes according to the package directions."}
+
+\begin{itemize}
+    \item Here the preposition is "according to". Using our triple extraction mechanism, we first get the triple as such:
+
+    \textbf{\{'source': 'imperative-chef', 'relation': 'prepare', 'target': 'cake mix'\}}
+
+    \item This is now captured as an even as such: \textbf{'1\_imperative-chef\_prepare\_cake mix'}\\
+
+\end{itemize}
+
+Now, when we encounter our preposition ("according to"), we retain it separately and continue with our traversal to capture the compound object 'package directions'.
+When the next preposition or punctuation is encountered, we have a prepositional relation at hand from the previous iteration.
+For this example, the period encountered at the end is when we have our relation:\\
+
+\textbf{\{'source': '1\_imperative-chef\_prepare\_cake mix', 'relation': 'according to', 'target': 'package direction'\}}\\
+
+Together, the two relations extracted for the example sentence captures the essence of the sentence. Here are more examples capturing prepositional relations
+\begin{itemize}
+    \item Example 1
+
+    \textbf{Sentence : }
+
+    "Add panko and Parmesan to first bowl and mix with your hands."
+
+    \textbf{Relations : }
+
+    \{'source': 'imperative-chef', 'relation': 'add', 'target': 'panko'\}
+
+    \{'source': 'imperative-chef', 'relation': 'add', 'target': 'Parmesan'\}
+
+    \{'source': '30\_imperative-chef\_add\_Parmesan', \textbf{'relation': 'to'}, 'target': 'first bowl'\}
+
+    \{'source': 'imperative-chef', 'relation': 'mix', 'target': 'first bowl'\}
+
+    \{'source': '32\_imperative-chef\_mix\_first bowl', \textbf{'relation': 'with'}, 'target': 'hand'\}
+
+    \item Example 2
+
+    \textbf{Sentence:}
+
+    "Arrange the pita quarters on a parchment-lined baking sheet."
+
+     \textbf{Relations : }
+
+     \{'source': 'imperative-chef', 'relation': 'arrange', 'target': 'pita quarter'\}
+
+    \{'source': '15\_imperative-chef\_arrange\_pita quarter', \textbf{'relation': 'on'}, 'target': 'lined sheet'\}
+
+\end{itemize}
+\textbf{Running 'relations\_msureshk.py' :}\\
+
+To run the code, execute relations\_msureshk.py and as arguments pass the file names in sequence. Example : \textbf{relations\_msureshk.py test1.txt test2.txt}\\
+
+\textbf{relations\_graph\_msureshk.json} captures the graph object, R in JSON format and \textbf{relations\_triples\_msureshk.txt} contains the relations extracted for each line in the input file.
+Note that one of each o these two files will be generated for all input files (In this case, 'test1.txt' and 'test2.txt').
 
@@
 
-This is how we evaluated our work, presenting the evaluation scores in comparison to previous projects...
+This section describes the evaluation method followed for the work presented.
+Section 8.1 describes the evaluation method.
+Section 8.2 evaluates performance comparing to existing Open IE extractor.
+Section 8.3 compares few examples of triples in the existing and linguistic approaches.
+Section 8.4 describes the limitations of the evaluation.
 
-Describe manual evaluation methodology: (1) look at sentence, (2) observe extracted triplets, (3) observe annotations in context of how our rule-based approach is programmed.
-Walk through several examples.
+\subsection{Evaluation Method}
+
+We chose to calculate the precision, recall, F-measure and accuracy by manually comparing the machine annotations against the gold standard.
+We have arrived to a gold standard by manually extracting triples from the food corpus shared by the corpus team.
+We have considered ford corpus file having 100 lines of text for the ease of evaluation.
+(1)
+
+Precision, Recall, F-measure and accuracy are traditional evaluation methods in information retrieval.~\cite{precision} 
+For this task we defined precision as the fraction of returned extractions that are correct, and recall as the fraction of correct extractions in the total corpus.
+For each extracted triplet, we manually check whether it is correct or not.
+Following are the expressions used for Precision, Recall, F-measure and accuracy
+    \begin{verbatim}
+        precision: true positives (TP)/true positives (TP) + false positives (FP)
+        recall : true positives(TP) /true positives (TP) + false negatives (FN)
+        Accuracy = (TN + TP)/(TN+TP+FN+FP) 
+        F1 = 2* precision*recall/precision+recall
+    \end{verbatim}
+\begin{enumerate}
+    \item True Positive (TP) : A true positive test result is one that detects the right triple  when the triple is present in ground truth.
+    \item True Negative (TN) : A true negative test result is one that does not detect the triple when the triple is not present in the ground truth.
+    \item False Positive (FP) : A false positive test result is one that detects the triple when the triple is not present in ground truth.
+    \item False Negative (FN) : A false negative test result is one that does not detect the triple when the triple is present in ground truth.
+\end{enumerate}
+
+We have also compared the triples extracted by Open IE with the gold standard and calculated the respective Accuracy, recall, precision score and compared both the results.
+
+\subsection{Performance}
+
+We started with analyzing the existing approaches to measure the performance.
+For neural approaches of extracting triples many follow supervised techniques to verify if the extracted triplet is positive or negative ~\cite{classification} but the most primarily taken approach is the ground-truth based approach.
+We have compared each triple manually with the ground truth to classify into positive and negative classes.
+Therefore we have restricted our evaluation to one approach and compare it with the existing approach.
+~\cite{paul_groth,shi_yuan}
+
+Stanford Open IE ~\cite{openIE_extractor} follows a linguistic approach to extract triples which is closer to the aim of the paper.
+So we chose to compare the results with Open IE.
+Open information extraction (open IE) refers to the extraction of relation tuples, typically binary relations, from plain text, such as (Mark Zuckerberg; founded; Facebook).
+The central difference from other information extraction is that the schema for these relations does not need to be specified in advance; typically the relation name is just the text linking two arguments.\newline
+
+To extract triples from open IE we have used a python open IE module and run the script with the input file.
+We then compared the triples of open IE manually with the ground truth.
+Here are the respective scores obtained.
+\newline
+
+\includegraphics[width=9cm, height=2.5cm]{score.png}\newline
+
+\noindent Due to the complexity and time constraints we have restricted the task of comparing every triple and classifying to positive and negative to one experiment.
+The code file to generate the open IE triples and data files for classification of triples are all uploaded on to github (1).
+
+\subsection{Analysis}
+
+When you compare the accuracy of both the approaches it is clear that the rule based approach followed in the paper yields better accuracy than Open IE since we have improved the rules like adding imperatives for the sentences where the subject is missing as discussed in the Experiments section.\newline
+
+Here are some of the observations comparing manual extraction, open IE and the approaches discussed.
+\begin{itemize}
+\item Example 1
+\begin{verbatim}
+"When sauce is almost done, cook pasta in a large pot of boiling salted water, 
+stirring occasionally, until al dente; drain pasta."
+
+Manually Extracted: 
+{'subject': 'chef', 'verb': 'cook', 'object': 'pasta'}
+{'subject': chef, 'verb': 'stir', 'object': 'pasta'}
+Open IE extracted: 
+{'subject': 'sauce', 'relation': 'is', 'object': 'almost done'}
+{'subject': 'sauce', 'relation': 'is', 'object': 'When done'}
+{'subject': 'sauce', 'relation': 'is', 'object': 'When almost done'}
+{'subject': 'cook pasta', 'relation': 'is in', 'object': 'large pot boiling'}
+{'subject': 'sauce', 'relation': 'is', 'object': 'done'}
+Extracted: 
+{'sentence': 26, 'event': 27, 'subject': 'sauce', 'verb': 'be', 'object': 'cook'}
+{'sentence': 26, 'event': 27, 'subject': 'sauce', 'verb': 'be', 'object': 'al'}
+\end{verbatim}
+
+Inference: Since the subject is already present our approach does not add an imperative subject but when you look at it manually adding imperative subject makes much more sense.
+
+\end{itemize}
+
+\subsection{Limitations}
+
+The evaluation measures can be improved by constantly verifying the ground truth and also by implementing more set of rules and extending the discussed experiments.
+First challenge is to verify if the ground truth is accurate since the evaluation is on the basis of ground truth.
+Current approach to calculate precision and recall is based on true positives and negatives.
+Since these are also compared and calculated manually there might be an error rate in the final values.
+\newline
+
+\noindent There might be cases of correct triples not being detected manually but are generated by the code.
+\begin{verbatim}
+"Arrange in a single layer, then roast in upper third of oven until 
+green beans begin to blister, about 15 minutes."
+
+Manually Extracted: 
+{'subject': 'chef', 'verb': 'arranges', 'object': 'green beans'}
+{'subject': 'chef', 'verb': 'roasts', 'object': 'green beans'}
+Open IE extracted: 
+{'subject': 'upper third', 'relation': 'in roast is', 'object': 'about 15 minutes'}
+{'subject': 'green beans', 'relation': 'begin to','object': 'blister'}
+{'subject': 'sauce', 'relation': 'is', 'object': 'When almost done'}
+{'subject': 'beans', 'relation': 'begin to', 'object': 'blister'}
+{'subject': 'roast', 'relation': 'is in', 'object': 'upper third of oven'}
+Extracted: 
+{'sentence': 34, 'event': 33, 'subject': 'bean', 'verb': 'arrange', 'object': '<None>'}
+\end{verbatim}
+
+Inference : When you look at the above text {'subject': 'green beans', 'relation': 'begin to', 'object': 'blister'} is considered to be a valid relation which was overlooked manually and approach followed in Experiment 4 was unable to extract the right relation.
+\newline
+
+\noindent Example of a case where the JSON-NLP API fails to provide the right dependency parse tree.
+\begin{verbatim}
+"1 x 15 oz can cut wax beans."
+
+Manually Extracted: {'subject': 'none', 'verb': 'none', 'object': 'can cut beans'}
+Open IE extracted: {'subject': 'oz', 'relation': 'can cut', 'object': 'wax beans'}
+Extracted: 
+{'sentence': 12, 'event': 13, 'subject': 'oz', 'verb': 'cut', 'object': 'wax bean'}
+\end{verbatim}
+
+Inference : The API assumes that 'can cut' is a verb where as can cut beans is a single entity and can here refers to the container.
+
+Also The evaluation of sequence of event is currently not considered there is a confusion for the approach followed for sequencing.
+The current approach is to add a counter to every sentence and increment it.
+This approach is agreed up on the basis if requirements of group 4 but we think that current approach does not make sense because event  sequence is formed on the basis of a recipe in this context.
+\newline
+For Example in the text sentence "Let all the ingredients sit and marinate for at least 30 minutes." depends on previous sequences where 'all the ingredients' are defined but according to the present approach we are providing a static counter number sequence to reduce the complexity of the task.
+\newline
+\newpage
+\begin{itemize}
+    \item Improvements and Further Evaluation :
+    \begin{enumerate}
+        \item Evaluate the results after integrating with the API team.
+        \item Improve the correctness of the ground truth.
+        \item Evaluation of triplets with event sequence.
+        \item Extend the evaluation all the experiments discussed.
+        \item Expand the ground truth data set to the entire food corpus.
+    \end{enumerate}
+\end{itemize}
+
+\noindent (1) All the files and code can be viewed from\newline
+https://github.com/dcavar/ML4NLP/commit/d0669d59e5096939c7a7c027ebc5b10d63539b29. 
+Details about the files are mentioned in the README.md file.
 
@@
 More specifically, the initial expectation among many in the class was that the relations extracted would always be of the form (subject, verb, object).
-While it is true that this was the initial approach (easiest to accomplish), this expectation resulted in some implementation decisions on the part of group 3 and the API group which may constrain the versatility and power of the knowledge graph representation.
+While it is true that this was the initial approach (easiest to accomplish), this expectation resulted in some implementation decisions on the part of the Relation Extraction group and the API group which may constrain the versatility and power of the knowledge graph representation.
 For example, as discussed in Section \ref{jodstein_experiments} above, the initial choice to extract verbs as the relation versus as an independent entity constrained the resulting knowledge graph in several ways:

### 2020-08-28 B547A1/text/attack_tree.tex
moved non-repo papers from overleaf into this repo

\section{Attack Tree}
\newlist{myEnumerate}{enumerate}{9}
\setlist[myEnumerate,1]{label*=\arabic*.}
\setlist[myEnumerate,2]{label*=\arabic*.}
\setlist[myEnumerate,3]{label*=\arabic*.}
\setlist[myEnumerate,4]{label*=\arabic*.}
\setlist[myEnumerate,5]{label*=\arabic*.}
\setlist[myEnumerate,6]{label*=\arabic*.}
\setlist[myEnumerate,7]{label*=\arabic*.}
\setlist[myEnumerate,8]{label*=\arabic*.}
\setlist[myEnumerate,9]{label*=\arabic*.}

\begin{myEnumerate}

\item Compromise the software/firmware update
  \begin{myEnumerate}[label*=\arabic*.]
  \item Gather knowledge (and)
    \begin{myEnumerate}[label*=\arabic*.]
    \item Find the IP address or URL of the webserver (or)
      \begin{myEnumerate}[label*=\arabic*.]
      \item Google the website of the target USB
      \item Collect the static contents of the web site (e.g. CSS, javascript,
        images, \dots).
      \end{myEnumerate}
    \item Get the same model as the target USB
      \begin{myEnumerate}[label*=\arabic*.]
      \item Try to see if it is easy to feed in any malware/keylogger
      \end{myEnumerate}
    \item Find the place (IP range) where he/she uses the target USB
      \begin{myEnumerate}[label*=\arabic*.]
      \item Use port scanner in the same subnet
      \item Tap his/her network cable
      \item Find the public key of the target USB
      \end{myEnumerate}
    \end{myEnumerate}
  \item Build a fake web site which looks like the real one (and)
    \begin{myEnumerate}[label*=\arabic*.]
    \item Prepare the MITM attack
      \begin{myEnumerate}[label*=\arabic*.]
      \item Do the reverse engineering to build a malicious app to feed into the
        target USB
      \end{myEnumerate}
    \end{myEnumerate}
  \item Gain an access (and)
    \begin{myEnumerate}[label*=\arabic*.]
    \item Use the DNS Spoofing trick to redirect the victims access of app
      update to the fake website
      \begin{myEnumerate}[label*=\arabic*.]
      \item Wait until the target USB tries to access the website to get it updated (or)
        \begin{myEnumerate}[label*=\arabic*.]
        \item Intercept the connection to the real website (or)
          \begin{myEnumerate}[label*=\arabic*.]
          \item Push the malicious app/firmware (and)
            \begin{myEnumerate}[label*=\arabic*.]
            \item Collect all the passwords of the target USB
            \end{myEnumerate}
          \end{myEnumerate}
        \end{myEnumerate}
      \item Mail a fake notification of the new release of the app with the URL of
        the fake website
        \begin{myEnumerate}[label*=\arabic*.]
        \item Intercept the connection to the real website (or)
          \begin{myEnumerate}[label*=\arabic*.]
          \item Push the malicious app/firmware (and)
            \begin{myEnumerate}[label*=\arabic*.]
            \item Collect all the passwords of the target USB
            \end{myEnumerate}
          \end{myEnumerate}
        \end{myEnumerate}
      \end{myEnumerate}
    \item Use the IP Spoofing trick to intercept the victims access of app
      update to the fake website
      \begin{myEnumerate}[label*=\arabic*.]
      \item Wait until the target USB tries to access the website to get it updated (or)
        \begin{myEnumerate}[label*=\arabic*.]
        \item Intercept the connection to the real website (or)
          \begin{myEnumerate}[label*=\arabic*.]
          \item Push the malicious app/firmware (and)
            \begin{myEnumerate}[label*=\arabic*.]
            \item Collect all the passwords of the target USB
            \end{myEnumerate}
          \end{myEnumerate}
        \end{myEnumerate}
      \item Mail a fake notification of the new release of the app with the URL of
        the fake website
        \begin{myEnumerate}[label*=\arabic*.]
        \item Intercept the connection to the real website (or)
          \begin{myEnumerate}[label*=\arabic*.]
          \item Push the malicious app/firmware (and)
            \begin{myEnumerate}[label*=\arabic*.]
            \item Collect all the passwords of the target USB
            \end{myEnumerate}
          \end{myEnumerate}
        \end{myEnumerate}
      \end{myEnumerate}
    \end{myEnumerate}
  \end{myEnumerate}
\end{myEnumerate}

### 2020-08-28 B547A1/text/conclusion.tex
moved non-repo papers from overleaf into this repo

\chapter{Conclusion}
\label{ch:conclusion}
%The proposed system is a standalone password manager equipped with the
%USB device. The device is hard to be tampered and is securely to store
%the pair (e.g., key:password) of user's credentials, where the key is
%a unique identity requiring to match the corresponding password
%(e.g. website, hostname, username, \dots). The application of the USB
%device provides an interface for a user to access the encrypted
%partition in the USB device.  The credentials that a user wants to
%store are located in the encrypted partition of the USB device which
%can be decrypted only with the primary password that a user setup.
%The default password is provided for the user to decrypt the encrypted
%partition for the first usage. The application forces a user to update
%the default password with a different password.  The USB device is
%updated with the new release of the application which can be
%downloaded from the designated webserver. The webserver uses the two
%factor authentication to ensure the authentictiy and integrity of the
%applications and the USB device. The user is required to register to
%the website with the public key of the USB device.  The new
%application is digitally signed with the factory's public key. Once
%the new application is verified to be authentic, it replaces the
%existing version in the USB device and restarts the application to
%load up with the new version.

%We have setup a threat model for the proposed system by using the
%standard STRIDE threat.  Our threat model is based on the iterative
%design. We review our proposed system architecture, dataflows, and
%external entities. Any potential threats are identified by using the
%STRIDE methodology and augmented by experimenting with the elevation
%of privilege game cards.  We discuss any identified trheats and assign
%a bug number to each one. Finally we decide whether we accept it as a
%system risk or not and perfom a risk likelihood vs impact for all
%accepted risks using a modified OWASP risk methodology.

Our threat model identified 43 threats and 19 risks (see
Figure~\ref{fig:riskmatrix}) associated with the system threats. Of
the 19 identified risks, 10 were classified as high risks largely
because any compromise of the system password led to complete loss of
confidentiality.  On the other hand, the likelihood of the threats is
4.05 in average within 1 to 10 range and the 10 high risks has 4.56. A
prinicpal attack vector was through the webserver using a malware
infected software/firmware image that is passed to an unaware
device. This attack tree is described in \nameref{ch:a4}. The risk of
this threat is also high category 2.

While the system appears to have significant vulnerabilities, we
believe that the likelihood element of the risk severity is overstated
in part because the OWASP risk methodology used does not actually
consider likelihood in the in light of design and mitigations. A
lesson learned that would apply to an even more comprehensive threat
model would be to iterate against mitigations which would reduce many
of the high risk threats to lower severity. If a compromise were to
occur, the impact is quite severe for most risks.  Finally, we do
transfer one risk to the user, that of properly protecting the USB
device and using it only on machines determined to be safe and free
from compromise.

### 2020-08-28 B547A1/text/introduction.tex
moved non-repo papers from overleaf into this repo

\chapter{Introduction}
\label{ch:introduction}

This report satisfies the requirements of assignment
1~\cite{2017myersa1handout}, which requires us to perform a
\textit{comprehensive} threat model of a password manager.  We describe the
password manager, its functionality and provide state, swimlane and dataflow diagrams
which we use in our threat model. We also perform risk analysis, describe
mitigations in the threat model and provide attack trees for several of our
highest rated threats.  \sidenote[][-0.5in]{Our password manager is hosted on a
specialized USB device which emphasizes hardware functionality as a means of
providing acceptable protection to stored passwords.}

\par We threat modelled using the STRIDE approach, focusing on data
flows and security-relevant elements.
%\sidenote[][]{STRIDE: \emph{Spoofing},\emph{Tampering},\emph{Information Disclosure},\emph{Denial of Service}, and \emph{Elevation of Privilege}}
Similarly, we manage our threats and risks using
Shostack~\cite{shostackbook} and modified OWASP
methodologies~\cite{owasprisk}.

%We are allowed discretion in the choice of architectures but the architecture
%should provide for the remembering of many passwords for websites,
%products, appliances and other uses which require passwords.

%\par The report is limited to no more than 7 pages of diagrams and 40
%pages total length.  The final product is a threat model of the system
%and any other documentation that would be necessary to understand how
%the proposed system would run.  If any sections of the threat model
%are empty, they should still be included in the report and the report
%should indicate that they were intentionally left empty.

### 2020-08-28 B547A1/text/mitigation.tex
moved non-repo papers from overleaf into this repo

\chapter{Mitigations}
\label{ch:mitigations}
In this chapter we describe the mitigations to the threats we found in
the previous chapter.

\section{Mitigation by Threat}
The threat tables detail a proposed mitigation for each identified
threat.  The final implementation is left to the designers who would
address each threat through the bug list. For a product in
development, additional visibility can be afforded by generating
requirements or test cases that identify each threat or risk.

\section{Risk and Prioritization}
\label{sec:risk}
While it is beyond the scope of this assignment to describe any technical or
business processes used to manage risk, we implicitly manage risk through our
assumptions and mitigations.  For example, the risk from the threat of loss of
the USB (resulting in a denial of service) is transferred to the owner of the
device by assuming that the owner will take adequate precautions to prevent loss
or theft of the device.  For the unlikely lucky guess threat, the risk is
accepted because the likelihood of properly guessing a complex password in a
very small, fixed number of tries is unlikely.

\begin{marginfigure}
    \centering
    \includegraphics[width=\linewidth]{riskcats}
    \caption{Risk Categories Used in Threat Modeling}
    \label{fig:riskcats}
\end{marginfigure}

In general, we would expect to pay most attention to priority risks in
increasing order; for example, a category one risk would be addressed before a
category 3 threat risk.

\subsection{Risk Determination}
We used a modified OWASP risk methodology to determine the risks posed by each
threat or group of similar threats. Detailed information concerning the assessed
risks can be found in \nameref{ch:a4}.

### 2020-08-28 B547A1/text/risk.tex
moved non-repo papers from overleaf into this repo

The risk methodology used in this assignment is taken from the OWASP Risk
Methodology~\cite{owasprisk} with modifications to the risk
matrix.\sidenote[][4.5cm]{OWASP uses a 3x3 table whereas we use a 5x5 table for
better granularity.}

Risk categories are defined in table ~\ref{tab:risk_cat}.\\

Each accepted threat is given a likelihood score which is based on threat agent
and vulnerability factors as well as an impact score which is based on technical
impact and business impact.  These two scores are used to enter the risk matrix
in figure~\ref{fig:riskmatrix} which will yield a risk category.

\begin{table}[h]
    \centering
    \begin{tabular}{l l l }
    \hline
    Risk Category & Definition \\
    \hline
    1 & Critical\\
    2 & High\\
    3 & Med-High\\
    4 & Med\\
    5 & Med-Low\\
    6 & Low \\
    \end{tabular}
    \caption{Risk Category Definitions}
    \label{tab:risk_cat}
\end{table}

\begin{table*}[ht]
    \centering
    \includegraphics[width=\linewidth]{risk_sum}
    \caption{Threat Risk Summary Table}
    \label{tab:risksum}
\end{table*}

\begin{figure*}[]
    \centering
    \includegraphics[width=\linewidth]{risk_matrix}
    \caption{Populated Risk Matrix}
    \label{fig:riskmatrix}
\end{figure*}

By convention, category 1 or 2 risks require immediate attention and management
while category 5 and 6 risks can most likely be monitored.The password manager's
risk are summarized using the table in table~\ref{tab:risksum}.  Detailed
information concerning the factors used to asses likelihood and impact for each
risk is found in table~\ref{tab:likelihood} and table~\ref{tab:impact}.

\begin{sidewaystable}[]
    \includegraphics[]{owasp_likelihood}
    \caption{Threat Likelihood Table}
    \label{tab:likelihood}
\end{sidewaystable}

\begin{sidewaystable*}[]
    \includegraphics[]{owasp_impact}
    \caption{Threat Impact Table}
    \label{tab:impact}
\end{sidewaystable*}

### 2020-08-28 B547A1/text/system_overview.tex
moved non-repo papers from overleaf into this repo

\chapter{System Overview}
\label{ch:System Overview}
The system modeled is a web-enabled password manager that resides on a
tamper resistant USB thumb drive, modeled after the
IronKey\texttrademark~secure USB thumb drive. The password manager
allows for a user to enjoy the benefits of a secure offline password
manager with the option of receiving software and firmware updates from the
cloud. The USB device also hosts an embedded crypto system which can be
leveraged to provide powerful cryptographic tools which provide some assurance
that the tools have not been compromised. Because security is the
primary objective, the password file (the file which stores user
names, passwords, and their associated applications, appliances, or
domains) is not written to persistent media outside the USB device.

\begin{marginfigure}%
\centering
  \includegraphics[width=0.25\linewidth]{s1000-vertical}
  \caption{Picture of the IronKey USB drive.  More information can be
found at \url{www.ironkey.com}}
  \label{fig:ik}
\end{marginfigure}

\section{System Description}
\label{sec:sysdesc}

The concept of operation is for the user to, after having inserted the
device into a computer, launch the provided trusted application from
the USB device, which has been mounted as a CD-ROM.  The application's
public key is sent to the computer along with the application
executable, and used to establish a temporary encrypted session with
the USB (in which the corresponding private key is stored).  Within
the temporary session, a random shared key is generated by the USB
device and passed to the application to establish the secure
application session.  Once the application is running, the user is in
a secure but anonymous session with the USB device.  From this point,
the user can choose to (a) submit a master password to the application
in order to access his/her password store, (b) perform a secure
software update on the device, or (c) perform a hard reset of the
device, restoring the device to factory condition and losing any
password store data.
\begin{enumerate}[label=(\alph*)]
    \item When the user submits his/her master password, the USB
device will attempt to decrypt a header file stored on the device.  If
successful, the device will initiate a secure user session with the
application, encrypted using a the symmetric secure application key as
well as a key derived from the user's password.  In this session, the
user (via the application) may perform privileged operations with the
password store file.
    \item When the user performs a secure software update, the
application interrogates the web server as to the latest software
release and downloads the digitally signed file.  Next, the
application sends the file to the USB device for the update.  Within
the USB device, the updated software is checked using the factory's
public key and, if authentic, generates a new public/private key pair
to store on the device, overwriting the current version of the
software.  Finally, the USB device forces a restart of the device and
the application using the new version.
    \item When the user elects to perform a hard reset on the device,
the device overwrites the header file and password store file with a
stored from-factory version, thereby permanently erasing any
previously stored user password information and restoring the device
to a factory state.
\end{enumerate}
\par The USB device firmware and software can be updated via a secure Internet
channel which accesses a webserver to download the device files to the USB
device.  Following download, the device authenticates the downloaded files and
updates the device.
The way to save the password via the password manager is straightforward. The
password is passed along with the corresponding key (e.g. a website, a username,
hostname, \dots) to the password manager when the secure user session is
setup as shown in Figures~\ref{fig:dfd_app} and ~\ref{fig:dfd_usb}. The key
must be unique. Once
the password is stored with the corresponding key to the encrypted partition in
the USB, a user can retrieve the password with the unique given key during the
secure user session. The retrieved password can be blinded out and copied to the
temporary memory of the system (e.g. clipboard) for users' convenience. Of
course, if a user inquires, the password can be displayed in a plain text but
preferably in an image format.

\begin{figure}
    \centering
    \includegraphics{DFD_Main}
    \caption{Main System Data Flow Diagram}
    \label{fig:dfd_main}
\end{figure}
\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{DFD_App}
    \caption{System Data Flow Diagram in the App side}
    \label{fig:dfd_app}
\end{figure*}

\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{DFD_USB}
    \caption{System Data Flow Diagram in the USB side}
    \label{fig:dfd_usb}
\end{figure*}

%\begin{figure*}
%    \centering
%    \includegraphics[width=0.8\linewidth]{webservdfd1}
%    \caption{System Data Flow Diagram for the Webserver}
%    \label{fig:wsdfd1}
%\end{figure*}

\begin{figure*}
    \centering
    \includegraphics[width=0.8\linewidth]{DFD_Webserver}
    \caption{System Data Flow Diagram for the Webserverer}
    \label{fig:wsdfd}
\end{figure*}

The Password manager consists of two major components, the USB device and a User
Application. The USB device when plugged in the system is loaded as a CD
ROM. The Get/Load function from the application gets the public key and begins
the message
encryption. Once the user application is authenticated and loaded there can be 4
possibilities:
\begin{enumerate}
\item The password manager is used for the first time by the user
\item Correct credentials are entered by the user
\item Incorrect credentials are entered by the user
\item User enters incorrect credentials multiple times
\end{enumerate}

When the user uses the iron key for the first time, the header is decrypted with
the default key. Once the header decrypt is successful the user is forced to
reset the default password. When correct credentials are entered by the user,
the password key is sent to the USB which then decrypts the header with that
key.  Once the header is decrypted a user session is started. The counter
which stores the number of invalid attempts is reset.The user has the option to
update the user password. The decrypted file is read into the memory and a new
password key and password file/header is written to the USB. A pair of key and
password is stored in the USB device. The password is queried from the USB
device using the given key.

If the user enters an invalid credentials, the header decrypt fails and the
password try counter is incremented.\marginnote{The USB itself does not possess
any logging capability save for the try counter which also serves as an
initiator to clear the system after a number of incorrect tries.} The user is
then prompted to enter a correct password. When a user enters a wrong password
multiple times which is equal to some threshold value, the reset counter
overwrites the header and password file by default.

The user application can be updated. When the update is successful, the USB is
updated and a force restart happens with the new version. When an update fails
the current version of the app is maintained.

The system data flow diagram is shown in figure \ref{fig:dfd_main}. Swim lane diagrams for several key processes are shown in figures~\ref{fig:dfd_secapp},~\ref{fig:dfd_secureuser}, and~\ref{fig:dfd_unsecure}.

\begin{figure*}
    \centering
    \includegraphics[width=0.8\linewidth]{DFD_SecureApp}
    \caption{Swim Lane Diagrams for Device Reset, Start User Session (Secure) and Software Update}
    \label{fig:dfd_secapp}
\end{figure*}

\begin{figure*}
    \centering
    \includegraphics[width=0.8\linewidth]{DFD_SecureUser}
    \caption{Swim Lane Diagrams for Password and File Store Operations}
    \label{fig:dfd_secureuser}
\end{figure*}

\begin{figure}
    \centering
    \includegraphics[width=0.7\linewidth]{DFD_Unsecure}
    \caption{Swim Lane Diagram for Start User Session (Unsecure)}
    \label{fig:dfd_unsecure}
\end{figure}

\section{Functional Requirements}
\label{sec:funcreq}
The functional requirements of the system are as follows:
\begin{enumerate}
    \item{Physically protect the USB against known tamper attacks including
physical attacks and USB electrical channel attacks.}\sidenote{We do not
enumerate all the possible USB physical and electrical attacks, nor do we
describe in detail the mitigations to these attacks.  It is assumed that
appropriate hardware antitamper features are incorporated into the USB design.}
    \item{Host all cryptography on the USB device.}
    \item{Allow for remote updates of USB software and firmware.}
    \item{Initiate device wipe after a set number of incorrect login attempts.}
    \item{Allow for the secure storage and retrieval of user websites and login
information.}
    \item{Provide for a secure, random password generator capability.}
\end{enumerate}

### 2020-08-28 B547A1/text/threat_list.tex
moved non-repo papers from overleaf into this repo

\cleardoublepage
\chapter{Appendix 2: Threat List}
\label{ch:threatlist}

The final system threat table is shown in figure~\ref{tab:threatlist}.  Each
entry in the threat table has an associated ID, who found the threat, the DFD
diagram element which applies to the threat, the type of threat as given by the
STRIDE model, and a mitigation.  Each entry is also assign a bug ID and finally
a determination of whether the entry is accepted into the larger threat risk
table.  \marginnote{In practice, the use of design rules can be very effective
because they form tangible requirements which are push down to the lowest level
and then reviewed as the software product matures at each design review.  This
keeps key security practices up front and visible throughout the design process.
Finally,incorporating design rules into an evaluation plan ensures the integrety
and security of the final product.}
\par If one threat is similar to another, it may not be entered into the threat
table but rather a threat risk that best decstibes the collective threat is
entered into the table.  This does not mean that the threat is ignored as it
still has a bug id associated with it which must be dispositioned.  In some
cases the threat is not carried to the risk table but instead may be referred to
a design rule which is then checked off during software component design reviews
and testing.

\begin{table*}
\centering
  \includegraphics[width=\linewidth]{threat_list_final}
  \caption{Final Table of Threats}
  \label{tab:threatlist}
\end{table*}

### 2020-08-28 B547A1/text/threat_model.tex
moved non-repo papers from overleaf into this repo

\chapter{Threat Model}
\label{ch:threatmodel}
In this chapter we detail our threat modelling approach, document our iterative
process which iterates the threats, identifies mitigations and manages risk.
Our threat model is an iterative model.  Our approach is as follows:
\begin{enumerate}
    \item{Individually review system architecture and dataflows and identify any
potential threats using the STRIDE methodoly augmented by the elevation of
privilege game cards shown in figure~\ref{fig:spoofcard} through
figure~\ref{fig:eopcard}.}
    \item{As a group, discuss any identified threats and gain concurrence and
understanding of the threat.}
    \item{Assign a bug number to the threat.}
    \item{Decide whether or not to accept this threat as a system risk.  Note
this step is intended to \emph{group} like threats and to allow for a more
detailed assessment of the risk presented by the threat to allow for better
triage of threats given limited time and budgets.}
    \item{Perform a threat risk likelihood vs impact for all accepted risks
using a \emph{modified} OWASP risk methodology.}
    \item{Review mitigations and if necessary iterate on high risk threats.}
\end{enumerate}

\section{Adversary}
\label{sec:adversary}
In order to make secure systems, we must always consider the system design and
threat given a specified adversary.  For our system, our adversary is a moderate
capability Blackhat hacker who seeks to compromise our password system for
financial gain.

\section{Assumptions}
\label{sec:assumptions}
As part of our threat model, we identified several security assumptions which
are documented in \nameref{ch:a1}.  Several assumptions demand deeper discussion.
\begin{enumerate}
    \item{\emph{The OS is secure.} While this may seems trivial to note this
assumption, a fundamental premise of the password manager design is that the OS
functions leveraged by the device such as USB load and clipboard have not been
compromised.  If these key OS functions are compromised then any data in transit
will be vulnerable to disclosure.  We minimize our attack surface by using
volatile memory only and using the USB as our sole source of cryptographic
functionality to the maximum extent possible.}
    \item{\emph{The user will physically protect the device.} In our scenario
analysis, permanent denial of service will result from loss of the device.
Since we do not provide any backup (in order to minimize the attack surface), it
is the user's responsibility to have a disaster plan in case of physical loss.
We are quite confident that passwords will not be compromised in the event of a
loss, but a loss also means that a user will no longer have access to the
password store.}
    \item{\emph{Our adversary is looking for financial gain.} Additionally, our
adversary has the resources to employ moderate BlackHat capabilities but will
not engage in any attacks where the attack cost will exceed the expected
financial gain.}\sidenote[][]{This means that given our measly financial assets,
most attackers will not employ national level attack against our password
manager just to confirm how poor we really are!}  These assumptions are pretty
straight forward.
\end{enumerate}

\section{Threats}
\label{sec:threats}

\subsection{Spoofing Threats}

\begin{marginfigure}%
\centering
  \includegraphics[width=0.4\linewidth]{spoofcard}
  \caption{Spoofing Card from the Elevation of Privilege Game}
  \label{fig:spoofcard}
\end{marginfigure}

In this section we identify the threat to each major element of our system,
usually by using a dataflow diagram.  We conduct a STRIDE analysis on each major
data flow or component using the Microsoft Elevation of Privilege game and
personal inspection.\sidenote{The elevation of privilege game can be found at
\url{https://www.microsoft.com/en-us/SDL/adopt/eop.aspx}} For each threat we
decide whether or not to add it to our risk chart; for similar threats or
threats that depend on a similar attack tree we enter only one threat into the
risk table but track the remaining threats in the bug table to ensure they are
properly managed and mitigated. As we iterated threats, we noted several
limitations of the OWASP Risk Management Methodology:
\begin{enumerate}
    \item{The OWASP model places considerable emphasis on non-repudiation.  For
our analysis, many of the threats can be realized by simple possession of the
USB device which does not require logging of any sort.  Following the risk
methodology model without any modification will yield the highest assessment
(likelihood) for detectability. In the future we would consider amending the
rubric for this particular likelihood factor.}
    \begin{marginfigure}%
    \centering
    \includegraphics[width=0.4\linewidth]{tampercard}
    \caption{Tamper Card from the Elevation of Privilege Game}
    \label{fig:tampercardcard}
    \end{marginfigure}

    \item{The technical impact categories and rubrics were very helpful, however
since many of these attacks were aimed at obtaining the USB password, they rated
artificially high because the cascade of compromise that resulted in obtaining a
password. The rubric for either technical impact or likelihood did not allow for
consideration of existing countermeasures or design features in the device.
Again, this resulted in artificially high risk assessments.  For a second
iteration, we would recommend including an assessment of design features and
countermeasures intended to address the threat or related threat as a better way
to better estimate the likelihood of a risk.}

    \begin{marginfigure}%
    \centering
    \includegraphics[width=0.4\linewidth]{repudcard}
    \caption{Repudiation Card from the Elevation of Privilege Game}
    \label{fig:repudcard}
    \end{marginfigure}

    \item{The business impact categories while useful in a real world exercise
were not applicable to this threat model.  In the future we would look at
substituting more applicable rubrics in place of the existing business impact
factors of the OWASP model.}
\end{enumerate}

In spite of the limitations we discuss, the OWASP model was very useful in
drawing our attention to the highest risk threats.  We benefited from using the
5x5 risk matrix as opposed to the 3x3 OWASP risk matrix because we were able to
break out threats and risk to a finer degree. Spoofing threats deal with an
attacker's ability to masquerade as an individual, process or some other entity
in order to gain access to data and systems.  During our first iteration we
identified several spoofing threats.

\begin{table*}[ht]
    \centering
    \includegraphics[width=\linewidth]{spoof_sum}
    \caption{Summary of Spoofing Threats Found in First Threat Iteration}
    \label{tab:spoofsum}
\end{table*}
We identified 5 threats and adjudicated them. Each threat received a bug id
whether or not they were entered into the risk tables.  Threats that were
similar to other evaluated threats were not accepted into the risk table as
noted. Several spoofing threats were elements in a multi-type threat such as
spoofing resulting in information disclosure. When possible, we identified
potential mitigations for each threat which informs the bug list which is where
the majority of system development is drawn.

\subsection{Tampering Threats}

We identified 10 tampering threats and accepted 5 into the risk tables. Threat ID 004,
``Malware Infested Update'' is a significant threat and represents a category of
several different threats which are intended to compromise the USB device by
delivering malware which can alter the reset counter, keys, and data within the
USB device. To explore the details of this attack vector, we provide
an attack tree as appendix~\nameref{ch:a4}.  Our tampering threats and
potential mitigations are depicted in
table~\ref{tab:tampersum}.

\begin{table*}[]
    \centering
    \includegraphics[width=\linewidth]{tamper_sum}
    \caption{Summary of Tampering Threats Found in First Threat Iteration}
    \label{tab:tampersum}
\end{table*}

\begin{marginfigure}[0.25in]%
\centering
  \includegraphics[width=0.4\linewidth]{infodisclcard}
  \caption{Information Disclosure Card from the Elevation of Privilege Game}
  \label{fig:spoofcard}
\end{marginfigure}

\subsection{Repudiation Threats}

We identified 12 potential repudiation threats.  The webserver used to store and present software and firmware updates
represents the most significant repudiation attack surface.  Compromise of this
element of the password manager can allow an attacker to post tampered software
and firmware files which can result in a full compromise of the password manager
and its contents.
\begin{table*}[ht]
    \centering
    \includegraphics[width=\linewidth]{repud_sum}
    \caption{Summary of Repudiation Threats Found in First Threat Iteration}
    \label{tab:repudsum}
\end{table*}

\subsection{Information Disclosure Threats}
We discovered several information disclosure threats during our analysis.  Many
of these threats where aimed at discovering passwords and keys while one
particular threat was aimed at fuzzing the USB to discover any potential leaks
caused by incorrect input.  A particularly nefarious information disclosure threat
was focused on recovering keys from a wiped device that had been transferred to
another user.

\begin{table*}[ht]
    \centering
    \includegraphics[width=\linewidth]{info_sum}
    \caption{Summary of Information Disclosure Threats Found in First Threat Iteration}
    \label{tab:infosum}
\end{table*}

\subsection{Denial Of Service Threats}
Denial of service for our password manager occurs when either the device is
cleared or no longer in the possession of its owner.  Loss of the USB device
represents a considerable denial of service threat because no backup features
are provisioned.
\begin{table*}[ht]
    \centering
    \includegraphics[width=\linewidth]{dos_sum}
    \caption{Summary of Denial of Service Threats Found in First Threat Iteration}
    \label{tab:dossum}
\end{table*}

Loss of control of the USB which results in an attacker possessing the device
may allow for the opportunity for other more sophisticated attacks, or quite
simply an attack that wipes the device because an attacker entered the wrong
password too many times.

\subsection{Elevation of Privilege Threats}
We identified three elevation of privilege threats, one associated with fuzzing
the USB to discovery entry points, another attack that interrupts the device
reset procedure which retains the old password during a device reset and finally a malformed request to the webserver database.  Elevation
of privilege threat are summarized in table~\ref{tab:eopsum}.

\begin{table*}[ht]
    \centering
    \includegraphics[width=\linewidth]{eop_sum}
    \caption{Summary of Elevation of Privilege Threats Found in First Threat Iteration}
    \label{tab:eopsum}
\end{table*}

\section{Threat Summary by Highest Risk}
Following the OWASP risk mitigation methodology, we accepted 19 risks associated with the threats that we identified.  The full risks, risk tables and methodologies are
summarized in \nameref{ch:a3}. We identified 8 category 2 (\emph{\textbf{high risk}})
threats which represents a significant majority of our risks.  A significant
portion of these threats were associated with compromise of of the password
(impact: 9) despite relatively low likelihood scores.  A summary of our threats
and their risk categories can be found in \nameref{ch:a3}, table~\ref{tab:risksum}.

\begin{marginfigure}%
\centering
  \includegraphics[width=0.4\linewidth]{doscard}
  \caption{Denial of Service Card from the Elevation of Privilege Game}
  \label{fig:doscard}
\end{marginfigure}

\section{Attack Trees}
Several threats were enabled by a similar attack vector.  For example, we
identified multiple threats associated with compromised software and firmware
updates.  Given the strong identification of multiple threats using this vector,
we examined this attack vector using an attack tree.

The attack tree that we have come up with as a good example is compromise the
update of application in the target USB device. The USB device needs to update
its application/software as new features are added and security patches are
required from time to time. Attackers can apply the Man In The Middle (MITM)
attack to the access between the USB device and the web server. The web server
is hosted to provide the new release of the applications. The detailed processes
of this threat is described in the \nameref{ch:a4}.

\begin{marginfigure}%
\centering
  \includegraphics[width=0.4\linewidth]{eopcard}
  \caption{Elevation of Privilege Card from the Elevation of Privilege Game}
  \label{fig:eopcard}
\end{marginfigure}

### 2020-08-28 B547_ICWSM-17B547/language_model.tex
moved non-repo papers from overleaf into this repo

\documentclass{article}

\usepackage{multirow}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{subfigure}

\begin{document}

\section{Association between the act of cross-posting and the drift of language usage of cross-posters}

The raw text of reddit posts (\texttt{body} field) were preprocessed, where HTML and other markup tags were removed. Then each post was tokenized into sentences, and each sentence was represented as a Bigram sequence of tokenized words. We performed two controlled computational experiments on the reddit text data processed as above using statistical language models.

In the first experiment, we investigated if the cross-posters' linguistic pattern in their HOME subreddit has manifested a significant change between the pre- and post-crossposting stage. As a control, we also considered the posts in the same subreddit that are from authors who never cross-posted. These posts were split into those that occurred in the first and second half of an author's lifespan.

For each group, a fixed number (100,000) of sentences were bootstrapped (i.e. sampled with replacement) from the available text data (Table \ref{lm:tab1}), and for each sentence a log-transformed probability score was computed using Bigram language model (with Stupid backoff smoothing [CITE]). The Bigram model was trained on the text data from non-crossposters in the same subreddit (disjoint from those used in the test set), which contains the most posts compared to those from cross-posters, and represents the mainstream linguistic pattern of the subreddit. The total numbers of available sentences in the four groups from which the above test and training set were sampled are briefly summarized in Table \ref{lm:tab1}.

As illustrated in Figure \ref{lm:fig1_1}, the linguistic pattern of both cross-posters and non-crossposters appeared to have drifted away from the mainstream of their HOME subreddit, when comparing the average of the log-probability of sentences in the early stage ("Pre-crossposting" and "First half") and late stage ("Post-crossposting" and "Second half") of an author's lifespan. Notice that the linguistic pattern of Feminism-to-MensRights cross-posters has drifted to a much larger degree compared to the Feminism non-crossposters. To evaluate the statistical significance of the observed difference in Feminism as well as the counterpart in MensRights, the procedure of training and test set sampling and language model building was repeated a hundred times, and for each repetition the delta ($\Delta_{cp}$) between "Post-crossposting" and "Pre-crossposting", and the delta ($\Delta_{Ncp}$) between "Second half" and "First half" was computed. The average of $\Delta_{cp}$ and the average of $\Delta_{Ncp}$ over the hundred repetitions was computed (Figure \ref{lm:fig1_2}). We found that, for Feminism, the average of $\Delta{cp}$ (0.1149) was found to be statistically different (t-test, p-value=$2.5e-184$) from that of $\Delta_{Ncp}$ (0.0182). However, for MensRights the means of two samples of delta (0.0325 and 0.0320) was not found to be statistically different (t-test, p-value=0.5524), which may suggest that Feminism-to-MensRights cross-posters appeared to be more radical or aggressive than the MensRights-to-Feminism cross-posters in terms of the change of linguistic patterns.

\begin{table}
    \centering
    \begin{tabular}{l l r r}\hline
        & & MensRights & Feminism\\
        & &            &         \\  
        \multirow{2}{*}{\textbf{Cross-poster}}      & "Pre-crossposting" & 820,850   & 77,460\\
                                                    & "Post-crossposting"  & 2,494,351 & 149,112\\
        \multirow{2}{*}{\textbf{Non-Crossposter}}   & "Fist half"      & 1,315,380 & 193,058\\
                                                    & "Second half"       & 1,391,906 & 242,324\\\hline
    \end{tabular}\
    \label{lm:tab1}
    \caption{Count of sentences of posts from cross-posters and non-crossposters. Bigram model was trained using 150,000 sentences from "First half" and "Second half", respectively, and 100,000 sentences were bootstrapped from the remaining in each of the four groups as the test set.}
\end{table}

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.5\textwidth}
        \includegraphics{images/LM_figure_1_1.png}
        \label{lm:fig1_1}
    \end{subfigure}
    \begin{subfigure}[b]{0.5\textwidth}
        \includegraphics{images/LM_figure_1_2.png}
        \label{lm:fig1_2}
    \end{subfigure}[b]{0.5\textwidth}
    \caption{Association between cross-posting and drift of linguistic pattern}
\end{figure}

In the second experiment, we investigated whether the linguistic pattern of all authors (cross-posters and non-crossposters alike) has changed over the entire lifecycle of either subreddit. The posts were sorted in increasing order of the timestamp, and sorted list were split into 50 approximately equal-sized chunks. A log-probability was computed for each sentence in one chunk of data, using a Bigram model trained on the sentences from the remaining 49 chunks, and this was repeated for each of the 50 chunks. As a comparison, we considered an irrelevant subreddit "Cooking" on which this process was repeated. We plotted the average of the log-probability of sentences in each chunk as a function of the index (1 to 50). As shown in Figure \ref{lm:fig2}, for MensRights and Feminism the agreement of the local and the global linguistic pattern reached the maximum in the first half the lifecyle, and then decreases over time. In contrast, the linguistic pattern of the subreddit "Cooking" remains stable. It's also interesting to note that the evolution of the linguistic pattern in MensRights and Feminism appear to be correlated with each other ($\rho$ = 0.6636), in contrast to their correlation with the Cooking subreddit ($\rho$ = 0.3487 and 0.0445 for MensRights and Feminism).

\begin{figure}
    \centering
    \includegraphics{images/LM_figure_2.png}
    \label{lm:fig2_1}
    \caption{The evolution of linguistic pattern over the lifecyle of subreddits}
\end{figure}

\end{document}

### 2020-08-28 B547_ICWSM-17B547/paper.tex
moved non-repo papers from overleaf into this repo

\documentclass[letterpaper]{article}
% \textit{Required Packages}
\usepackage{aaai17}
\usepackage{times}
\usepackage{amsmath}
\usepackage{helvet}
\usepackage{courier}
\setlength{\pdfpagewidth}{8.5in}
\setlength{\pdfpageheight}{11in}
%\usepackage{hyperref}

\usepackage{footnote}
\usepackage{dblfloatfix}
\usepackage{booktabs}

\usepackage[flushleft]{threeparttable}

\usepackage{graphicx}
\graphicspath{ {images/} }

\usepackage{float}
\usepackage[draft,inline,nomargin]{fixme}

\usepackage{adjustbox}
\usepackage{multirow}

\usepackage[draft]{changes} 
\newcommand{\mustfix}[1]{\fixme{\hl{#1}}}
\newcommand{\pleasenote}[1]{\fxnote{\hl{#1}}}
\newcommand{\hlfixme}[1]{\fixme{\hl{#1}}}
\newcommand{\fabio}[1]{{\textcolor{blue}{fabio: #1}}}
\newcommand{\yy}[1]{{\textcolor{green}{yy: #1}}}
\newcommand{\apu}[1]{{\textcolor{purple}{apu: #1}}}
\newcommand{\pat}[1]{{\textcolor{red}{Pat: #1}}}
\newcommand{\del}[1]{{\textcolor{gray}{#1}}}

%\%\%\%\%\%\%\%\%\%
% \textit{PDFINFO for PDF\LaTeX{}}
% Uncomment and complete the following for metadata (your paper must compile with PDF\LaTeX{})
\pdfinfo{
/Title (When do ideologues talk to each other? Analyzing factors associated with cross-talk between Feminism and Men's Right Subreddits)
/Author (Patrick Shaffer,Sarvothaman Madhavan,Tyler Dell,Chao Ji,Meng Li,Apu Kapadia,Fabio Rojas)
/Keywords (Reddit, Polarization, Boundary Crossing)
}
%\%\%\%\%\%\%\%\%\%
% \textit{Section Numbers}
% Uncomment if you want to use section numbers
% and change the 0 to a 1 or 2
%\setcounter{secnumdepth}{2}
\setlength\titlebox{2.5in}
%\%\%\%\%\%\%\%\%\%
% \textit{Title, Author, and Address Information}
\begin{document}

\title{When do ideologues talk to each other? Analyzing factors associated with cross-talk between Feminism and Men's Rights Subreddits}

%\author{Patrick Shaffer, Sarvothaman Madhavan, Tyler Dell, Chao Ji, Meng Li, Yong-Yeol Ahn, Apu Kapadia\\
%School of Informatics and Computing\\
%Indiana University Bloomington, IN, USA\\
%patshaff@indiana.edu, madhavas@iu.edu, tdell,jic,li545, yyahn, kapadia@indiana.edu\\
%\AND
%Fabio Rojas\\Department of Sociology \\
%Indiana University Bloomington, IN, USA\\
%frojas@indiana.edu\\}

\author{anonymized}
%\%\%\%\%\%\%\%\%\%
% \textit{Body of Paper Begins}

\maketitle
%\pat{fix overfull box in author names- use first initials?}

\begin{abstract}
Although social media enhances communication across geographical and cultural boundaries, it can also exacerbate the formation of isolated ideological communities. To foster healthy discourse, it is critical to understand how and when opposing groups communicate with each other. Motivated by prior studies linking group participation and the tendency to reach out to other groups, this paper \yy{investigates how multiple aspects of activity measures are associated with the behavior of addressing an ideologically opposed group (``cross-talk'').} \yy{the following sounds something too trivial. higher activity -> more likely to cross talk just by pure chance. } \del{tests the hypothesis that addressing an ideologically opposed group (``cross-talk'') is associated with being highly active in social media.} We test this hypothesis with data from over 133,000 users from two ideologically opposed communities --- the subreddits dedicated to Men's Rights and Feminism. For each user, we collect longitudinal data on their activities to predict when they post a comment on  the opposing subreddit. We apply a proportional hazard model to estimate the importance of various activity variables. We also investigate an important consequence of cross-talk: the linguistic drift between cross-posters and their home subreddit. \yy{per my previous comment, just say we need to make people more active is an empty statement. We did not look at whether more interaction (cross talk) can lead to better discourse. we don't know what's in the cross posts. no content analysis right? It may well exacerbate the polarization.} \yy{we need a better argument} Together, these two results depict one possible path for social change on social media. Individual characteristics, such as user activity and network position, encourage movement across boundaries, which then leads to changes in language and discourse. Thus, polarization might be mitigated by encouraging social media users to interact with others online, especially those who already have a history of crossing boundaries.

\end{abstract}

\section{Introduction}

People form groups based on shared interests, values, and social practices~\cite{gerber2010sociology}. Although group memberships provide individuals with the benefits of developing social identities, they can also lead to detrimental behaviors such as group polarization~\cite{tajfel1979individualsandgroups,tajfel1970experiments,tajfel1979integrative}. Furthermore, the boundaries between communities can often be rigid, and individuals often take it upon themselves to create and enforce social norms about who is to be included or excluded from the group~\cite{sherif1961robbers}. The creation and policing of social boundaries has been heavily investigated by social scientists and researchers in social informatics~\cite{lamont2002study}, and there is now a wealth of knowledge and discussion about social boundaries, whether they emerge from ethnic differences, cultural variance, or professional status ~\cite{abbott1988system,wimmer2008elementary,wimmer2013ethnic} 
%(CITES =  Abbott 1990; Wimmer 2008; 2013). \apu{CITES} %\apu{need citations for each topic in this sentence. In general since we say ``heavily studied'' and ``wealth of knowledge'' we need 6--7 citations at least, and not just one.}.

Traditional studies of groups have identified various behaviors that may result in crossing of boundaries, adoption of innovations, or other brokerage behaviors~\cite{rogers2003diffusion,tilley2005boundsocties}, yet little research has focused on the precursors to this crossing behavior, in particular the emergence of interactions between individuals in one group and an ideologically opposed group.

In this work, we seek to understand the factors affecting when individuals begin to interact with an ideologically opposed group. The question of when ``cross-talk'' occurs is important because the interaction between polarized groups gives us insight into the factors that encourage individuals to engage with people who may disagree with them, which can lead to a richer understanding of when network positions reinforce cross-group interactions.

By studying these precursors to cross-group contact, we add to our knowledge of how groups form ties with one another and how cross-group contacts lead to social change. Sometimes, cross-group contacts may be antagonistic, such as when people intentionally post inflammatory material (``trolling'') and reinforce group boundaries~\cite{bergstrom2011don}. In other cases, cross-group interaction may bring two groups together, e.g., when the members of one group might try to be conciliatory to another group~\cite{cehajic2008forgiveandforget}. While this study does not examine the quality of cross-talk (e.g., antagonistic vs. conciliatory), it provides an important contextualization of cross-talk~--- we offer evidence that individual characteristics of social media users and their network position influence the propensity to address the opposing group (or ``outgroup'').

In this paper, we address this issue by presenting a theory of when individuals in one social group begin to interact with their outgroup and we test our hypotheses with data from Reddit, a popular website where people hold discussions on topics of their choice (e.g., basketball or the 2016 U.S. election). We posit that individuals with a high level of participation on social media and those who have many interactions with others are more likely than others to become cross-posters, and study how various aspects of their activities manifest into the cross-talking behavior. We also investigate the consequence of talk across groups. We show that cross-talking behavior is associated with change in their language, away from the language of the user's original home Subreddit. These two findings indicate a two-stage process: heightened online behavior leads to cross-group talk, which is then associated with changed linguistic patterns. \yy{is it something trivial or more than that? You have different audience, topics, etc., thus it is natural to assume that your language is different in a differnet subreddit. Is this more than that? Maybe it's good to clarify this point early. }

Drawing on prior research on outgroup interaction, we argue that that heightened online participation should be associated with outgroup contacts~\cite{pettigrew1998intergroup}. To test our theory, we examine online participation based on multiple features. First, we examine the frequency and intensity of a user's posts. A user who posts a lot is not only more likely to cross-talk by pure chance, but also to do so because they are more likely to seek out a new audience. Second, we argue that the direction of the communication matters as well. We examine the difference between the comments written by a user and the comments written in response to a user's comments. \yy{in what way? what are the measures (variables) here?} Third, we suggest that contact with individuals who already cross-talk should be associated with an increased likelihood that a user will cross-talk (which may happen as a result of 'contagion' or 'homophily').

We test hypotheses drawn from this theory with data from two large ideologically opposed communities on Reddit, the ``Men's Rights'' and ``Feminism'' subreddits. We view this pair of online communities as a valuable example of two groups separated by a strong social boundary. \yy{the following sentence doesn't sound quite right. revise?} These groups are dedicated to discussing two groups, men and women, who are often seen as having conflicting interests. Furthermore, the proponents of each ideological viewpoint often depict the proponents of the other viewpoint as antagonists. Thus, the boundary between men's rights advocates and feminists should be a particularly strong one that is interesting to examine and provides a test of the strength of the factors that may encourage boundary crossing~\cite{menzies2007backlash,messner2000politicsofmasculinities}. %\fabio{pat, add a cite here} \yy{any literature about Feminism and men's right movements?}\pat{done}

%We test these models with data from the main subreddits dedicated to Men's Rights and Feminism. 
We analyze the data in an event history framework\yy{citatin needed}. For each Reddit user, we collected data for every month after they first appear in either subreddit. Then, we estimated the effect of these possibly time-dependent covariates on the probability that the user will begin to post in the other subreddit (e.g., a ``Feminism'' participant posts in ``Men's Rights''). The correlations are estimated using a proportional hazards, or Cox model, which provides an estimate of the multiplicative effect of the covariates on the transition between states (i.e., switch from not cross-talking to the first instance of cross-talk)~\cite{yamaguchi1991eventhistanal}.  We report bivariate models that link a single variable with cross-talk and multivariate models that include control variables.

Then, we focus on how cross-posters change. We compare the text that cross-posters produce before speaking to the opposite group and after they speak to the opposite group. Literature on outgroup interaction suggests that interactions with out-group members have the potential to change identities and behavior~\cite{brown1999changing}. Consistent with this theory, we find that the language of cross-posters changes after the first cross-post. The comments written \yy{where? the other subreddit or the home subreddit?} after the first cross-post are less similar to the language of the home Reddit than the comments written before cross-posting.

%People form groups based on shared interests, values, and social practices~\cite{gerber2010sociology}. Although group memberships provide individuals with the benefits of developing social identities, they can also lead to detrimental behaviors such as group polarization~\cite{tajfel1979individualsandgroups,tajfel1970experiments,tajfel1979integrative}. Furthermore, the boundaries between communities can often be rigid, and individuals often take it upon themselves to create and enforce social norms about who is to be included or excluded from the group~\cite{sherif1961robbers}. The creation and policing of social boundaries has been heavily investigated by social scientists and researchers in social informatics~\cite{lamont2002study}, and there is now a wealth of knowledge and discussion about social boundaries, whether they emerge from ethnic differences, cultural variance, or professional status \apu{need citations for each topic in this sentence. In general since we say ``heavily studied'' and ``wealth of knowledge'' we need 6--7 citations at least, and not just one.}. 
%Traditional studies of groups have identified various behaviors that may result in crossing of boundaries, adoption of innovations or other brokerage behaviors~\cite{rogers2003diffusion, tilley2005identityboundariessocties}, yet little research has focused on the precursors to this behavior, in particular the emergence of interactions between individuals in one group and an ideologically opposed group. 
%In this work, we seek to understand the factors affecting when individuals begin to interact with another group, especially an ideologically opposed group. The question of when `cross-talk' occurs is important because the interaction between polarized groups gives us insight into the factors that encourage individuals to engage with people who may disagree with them, which can lead to a richer understanding of when network positions reinforce cross-group interactions. 
%By studying these precursors to cross-group contact, we add to our knowledge of how groups form ties with one another and how cross-group contacts lead to social change. Sometimes, cross-group contacts may be antagonistic, such as when people intentionally post inflammatory material (``trolling'' and reinforcing group boundaries). In other cases, cross-group interaction may bring two groups together. The members of one group might try to be conciliatory to another group~\cite{cehajic2008forgiveandforget}. While this study does not examine the quality of cross-talk (e.g., antagonistic vs. conciliatory), it provides an important contextualization of cross-talk. We offer evidence that individual characteristics of social media users \apu{such as...} and their network position influence the propensity to address their `outgroup', i.e., the group that they do not identify with or are ideologically opposed to.
%We present a theory of when individuals in one social group begin to interact with their outgroup and we test our hypotheses with data from Reddit, a popular website where people hold discussions on topics of their choice (e.g., basketball or the 2016 U.S. election). Our theory posits that the movement across social boundaries reflects three important social processes as described next.\footnote{We note that our theory only addresses the presence of cross-group contact, not the type of contact.}
%First, the more information an individual in an online community provides to readers, the more likely that they will cross-talk to an ideologically opposite group. We suggest that these individuals who provide a lot of information seek contact with others who may be different. Second, individuals are embedded in social networks which may be associated with a propensity to address an ideologically opposite group. \apu{Does this sentence belong in the previous category?: People who generate discussion online might be more likely to seek contacts with the ideologically opposite group.} People who attract discussion, via comments on what they write, might be encouraged to post their ideas to an ideologically opposed group. Thus, we suggest that the in- and out-degree within an online community's communication network will be associated with the propensity for cross-talk. Third, we suggest that cross-talk will be clustered in the social network. That is, those who cross-talk will be associated with others who cross-talk. This may happen because individuals share similar characteristics (homophily) or one cross-talker may directly influence others (contagion).
%We test hypotheses drawn from this theory with data from two large ideologically opposed communities on Reddit, the `Men's Rights' and `Feminism' subreddits. We view this pair of online communities as a valuable example of two groups separated by a strong social boundary. These groups are dedicated to discussing two groups, men and women, who are often seen as having conflicting interests. Furthermore, the proponents of each ideological viewpoint often depict the proponents of the other viewpoint as antagonists. Thus, the boundary between Men's Rights advocates and Feminists should be a particularly strong one that is interesting to examine and provides a test of the strength of the factors that may encourage boundary crossing~\cite{menzies2007backlash,messner2000politicsofmasculinities}. %\fabio{pat, add a cite here} \yy{any literature about Feminism and men's right movements?}\pat{done}
%
%Thus, using data from the `Men's Rights' and `Feminism' subreddits, we test the following hypotheses:
%\begin{itemize}
    %\item \textbf{H1}: The probability that a Reddit user will `cross-talk' or interact with an ideologically opposite group is correlated with their `openness', as measured by the information they reveal about themselves in their profile, the number of posts they write, and the length of their posts.
    %\item \textbf{H2}: The probability that a Reddit user will `cross-talk' or interact with an ideologically opposite subreddit is affected by their `network interactions', as measured by the number of people who comment on their posts and number of other people's posts that they comment on. 
    %\item \textbf{H3}: The probability that a Reddit user will `cross-talk' or interact with an ideologically opposite subreddit is affected by their interactions with other people who have cross-talked in earlier time periods. \apu{Pat: add ``as measured by...'' I presume we mean they have commented or been commented on by people who have cross-talked in the past?}
%\end{itemize}

%\yy{I feel, at this point, that the hypotheses are all confounded with each other and that it is quite tricky to test those.} \fabio{we have controls like account age, also we can do a factor analysis to address who items in the regression group together}
%We test these hypotheses with data from the main subreddits dedicated to Men's Rights and Feminism. We analyze the data in an event history framework: For each reddit user, we collected data for every month after they first appeared in either subreddit. Then, we estimated the effect of these possibly time-dependent covariates on the probability that the user will begin to post in the other subreddit (e.g., a ``Feminism'' participant posts in ``Men's Right's''). The correlations are estimated using a proportional hazards, or Cox, model, which provides an estimate of the multiplicative effect of the covariates on the transition between states (i.e., switch from not cross-talking to the first instance of cross-talk)~\cite{yamaguchi1991eventhistanal}. 
%\fabio{PAT - add cite to Cox 78 or the Yamaguchi book}\pat{done}
%People form groups based on shared interests, values, and social practices~\cite{gerber2010sociology}. Although group memberships provide individuals with the benefits of developing social identities, they can also lead to detrimental behaviors such as group polarization~\cite{tajfel1979individualsandgroups,tajfel1970experiments,tajfel1979integrative}. Furthermore, the boundaries between communities can often be rigid, and individuals often take it upon themselves to create and enforce social norms about who is to be included or excluded from the group~\cite{sherif1961robbers}. The creation and policing of social boundaries has been heavily investigated by social scientists and researchers in social informatics~\cite{lamont2002study}, and there is now a wealth of knowledge and discussion about social boundaries, whether they emerge from ethnic differences, cultural variance, or professional status \apu{need citations for each topic in this sentence. In general since we say ``heavily studied'' and ``wealth of knowledge'' we need 6--7 citations at least, and not just one.}. 
%Traditional studies of groups have identified various behaviors that may result in crossing of boundaries, adoption of innovations or other brokerage behaviors~\cite{rogers2003diffusion, tilley2005identityboundariessocties}, yet little research has focused on the precursors to this behavior, in particular the emergence of interactions between individuals in one group and an ideologically opposed group. 
%In this work, we seek to understand the factors affecting when individuals begin to interact with another group, especially an ideologically opposed group. The question of when `cross-talk' occurs is important because the interaction between polarized groups gives us insight into the factors that encourage individuals to engage with people who may disagree with them, which can lead to a richer understanding of when network positions reinforce cross-group interactions. 
%By studying these precursors to cross-group contact, we add to our knowledge of how groups form ties with one another and how cross-group contacts lead to social change. Sometimes, cross-group contacts may be antagonistic, such as when people intentionally post inflammatory material (``trolling'' and reinforcing group boundaries. In other cases, cross-group interaction may bring two groups together. The members of one group might try to be conciliatory to another group~\cite{cehajic2008forgiveandforget}. While this study does not examine the quality of cross-talk (e.g., antagonistic vs. conciliatory), it provides an important contextualization of cross-talk. We offer evidence that individual characteristics of social media users and their network position influence the propensity to address the ``outgroup''.
%In this paper, we address this issue by presenting a theory of when individuals in one social group begin to interact with their outgroup and we test our hypotheses with data from Reddit, a popular website where people hold discussions on topics of their choice (e.g., basketball or the 2016 election). We posit that individuals with a high level of participation on social media and those who have many interactions with others are more likely than others to become cross-posters. We also investigate a separate, but related, issue, the consequence of talk across groups. We provide evidence that cross-talk is associated with drifting away from the user's original home reddit. These two findings indicates a two stage process: heightened online behavior leads to cross-group talk, which is then associated with changed linguistic patterns. 
%Drawing on prior research on outgroup interaction, we examine different features of online participation. First, we focus on the frequency and intensity of a user's posts. We argue that the amount of communication is associated with cross-talk because those who communicate a great deal might be more likely to seek out new audiences. Second, we argue that the direction of the communication matters as well. We examine the difference between the comments written by a user and the comments written in response to a user's comments.  Third, we suggest that contact with individuals who already cross-talk should be associated with an increased probability that a user will cross-talk, which may happen as a result of contagion or homophily. 
%We test  hypotheses drawn from this theory with data from two large ideologically opposed communities on Reddit, the ``Men's Rights'' and ``Feminism'' subreddits. We view this pair of online communities as a valuable example of two groups separated by a strong social boundary. These groups are dedicated to discussing two groups, men and women, who are often seen as having conflicting interests. Furthermore, the proponents of each ideological viewpoint often depict the proponents of the other viewpoint as antagonists. Thus, the boundary between men's rights advocates and feminists should be a particularly strong one that is interesting to examine and provides a test of the strength of the factors that may encourage boundary crossing~\cite{menzies2007backlash,messner2000politicsofmasculinities}. %\fabio{pat, add a cite here} \yy{any literature about Feminism and men's right movements?}\pat{done}
%We test these models with data from the main subreddits dedicated to Men's Rights and Feminism. We analyze the data in an event history framework: For each reddit user, we collected data for every month after they first appear in either subreddit. Then, we estimated the effect of these possibly time-dependent covariates on the probability that the user will begin to post in the other subreddit (e.g., a ``Feminism'' participant posts in ``men's rights). The correlations are estimated using a proportional hazards, or Cox, model, which provides an estimate of the multiplicative effect of the covariates on the transition between states (i.e., switch from not "cross-talking" to the first instance of "cross-talk")~\cite{yamaguchi1991eventhistanal}.  We report bivariate models that link a single variable with cross-talk and multivariate models that include control variables. 
%Then, we focus on how cross-posters change. We compare the text that cross-posters produce before speaking to the opposite group and after they speak to the opposite group. Literature on outgroup interaction suggests that interactions with out-group members have the potential to change identities and behavior. Consistent with this theory, we find that the language of cross-posters changes after the first cross-post. The comments written after the first cross-post are less similar to the language of the home reddit than the comments written before cross-posting.

\section{Social Boundaries and Boundary Crossing}
We view cross-group contact as an example of a broader social process~--- the creation, maintenance, and erosion of group boundaries. A basic observation about human beings is that they create groups where people have shared histories, values, and ideologies~\cite{turner1987rediscovering}. Sometimes, there are strong boundaries between groups, which means that individuals are not likely to interact with people from the ``other'' group or anyone from outside the group~\cite{weisel2015groupconflict}. Examples of strong inter-group boundaries include ethnic divisions, religious denominations, and political parties. In each case, there have been times when there was a strong social boundary between groups. For example, political scientists have documented that there are relatively strong boundaries between Democrats and Republicans in the United States Congress, where bipartisanship is now at a historical low point~\cite{andris2015riseofpartisanship}.

\subsection{The Origin and Evolution of the Term `Boundary Crossing'}

The term `boundary crossing' originated from research concerning the sharing of knowledge between disciplines and professions where factors such as professional affiliations and jargon created occupational boundaries~\cite{wenger1998communitiesofpractice}. Later, researchers approached contact between groups using different disciplinary lenses. Psychologists, for example, have studied outgroup interactions and seek to identify the emotional and cognitive factors that are associated with the willingness to participate in or talk to an outgroup. Sociologists have argued that boundary crossing represents an erosion of social divisions. It is argued that, for example, marriage across ethnic groups is an indicator of weak ethnic boundaries~\cite{qian2007social}. Management scholars argue that linking groups through outgroup contact is bridging a structural hole in a firm, which enhances innovation because it creates new flows of information and provides opportunities for combining ideas \cite{burt2004structuralholes}.

%In our two subreddits, we noted differences in language which may contribute to the boundaries. More importantly, we see a clear delineation of beliefs and ideas about a common topic --- 
%\yy{use three hyphens --- for m-dash. Fix all m-dashes in the paper.}\pat{done}
%men and women. \citeauthor{hara2014frameworkforkowledgesharing} describe this type of boundary as a cognitive boundary that is ``mentally created'' through ``personal beliefs, worldviews and understanding''~\cite{hara2014frameworkforkowledgesharing}.
%\yy{fix all citations; cite like this~\cite{something}. the tilde creates an empty space and attach the citation to the last word.}\pat{done}
\subsection{Relationship to Political Discussions}

The division between Men's Rights and Feminist proponents is an example of political polarization, which is when individuals sort themselves into groups based on ideological or political affiliations. This sort of social boundary has been extensively studied in both social informatics and political science. \citeauthor{adamic2004divided}~(\citeyear{adamic2004divided}) studied the behavior of political bloggers in the context of the 2004 U.S. Presidential Election and found that two groups, Democrats and Republicans, exhibited different behaviors particularly with regards to how they linked to each other's posts~\cite{adamic2004divided}. 
%\fabio{Unless, I'm computing an inner product, I avoid the word orthogonal in print}
Recent work in political science has examined the extent of polarization in different political contexts. For example, multiple studies have found that Congress is polarized in the sense that there is little cooperation between Republican and Democratic legislators~\cite{andris2015riseofpartisanship}. Other studies of public opinion have documented that people with specific opinions are more likely than before to be firmly associated with a single party~\cite{gould2015partyaffilpoliticalbelief}. E.g., pro-life voters are associated with the Republican party at very high rates, while the opposite is true for pro-choice voters who align themselves with the Democratic party~\cite{killian2008abortion}.

% Continuing in the context of politics, ~\citeauthor{abramowitz2008polarization} refuted a previous study's assertion that political polarization resulted in lower voter turnout and found instead that "polarization energizes the electorate and stimulates political participation" ~\cite{abramowitz2008polarization}. \fabio{NOt sure if this passage follows logically}

Strong political boundaries are also present on social media~\cite{connover2011politicalpoltwitter}. Numerous commentators have lamented the ability of people to create ``bubbles'' online where they speak primarily, or exclusively, to those who believe the same things. It is also observed that people in one political group often view the other group in adversarial terms, which is consistent with~\citeauthor{allport1979natureofprejudice}'s~(\citeyear{allport1979natureofprejudice}) notion of an ``ingroup'' that defines itself in reference to an ``outgroup''. This observation is consistent with the current study, as analyses of the Men's Rights and Feminist movements document how each group views the other as an antagonist~\cite{messner2000politicsofmasculinities,menzies2007backlash}.

%\fabio{add cites}\pat{done although this is a repeat of previous cites is that ok?}

%each subreddit is the other's out-group~\cite{allport1979natureofprejudice}. ~\citeauthor{guerra2013measureofpolarization} presented a metric which characterizes the polarization of two groups based on a study of their boundaries~\cite{guerra2013measureofpolarization}.

\subsection{The Importance of Crossing Boundaries}
We also view our work as a contribution to the study of how boundaries are crossed. First, the existence of a contact, or interaction, between two groups is an example of an individual bridging a ``structural hole'' in a social network~\cite{burt2004structuralholes,burt2007burtbrokerage}. Creating a connection between two otherwise disconnected groups is associated with creativity and performance~\cite{burt2004structuralholes,burt2007burtbrokerage}. Second, the linking of two groups facilitates the transmission of information between groups. Third, contact between people may, in some cases, result in the changing of opinions or attitudes. Recent work in political psychology suggests that, in some cases, interpersonal contact between people can change political opinions~\cite{broockman2016reducingtransphobia}. Thus, if cross-talk occurs and it results in positive communication, it may contribute to the reduction of political polarization. Therefore, it is important to understand the individual and contextual factors that contribute to communication between groups. In this paper we use the terms ``crosspost'' and ``cross-talk'' to describe the interactions (using comments) between members of our polarized subreddits. A crossposter is a user who has posted in the opposite subreddit.

%\yy{isn't there something called contact hypothesis or something? (contact between groups -> better mutual understanding) oh yes: https://en.wikipedia.org/wiki/Contact_hypothesis}

\subsection{Theorizing the Factors that Promote Outgroup Interaction and Studying its Consequences}

We rely on prior research to help us develop hypotheses about the cross-talk in social media. First, a number of previous researchers~\cite{bierhoff2002prosocial} have noted that pro-social people tend to engage in outgroup interaction and boundary crossing. That is, those with a large number of interactions tend to be those that cross social boundaries. Second, working across `structural holes' in networks is often associated with individuals who are often high in `status'. For example, demographers have found that high status individuals often `move up' by marrying persons from even higher status ethnic groups (e.g., high socioeconomic status (SES) African Americans are more likely to marry Whites than low SES African Americans~\cite{wimmer2013ethnic}). In social media, this suggests that degree in communication networks might be associated with cross-talk or outgroup interaction.  Third, recent work in cognitive psychology suggests that even the presence of outgroup members disrupts an individual's identification with their own group~\cite{hobson2016mere}). Thus, being confronted with individuals from outgroups and those who cross post should encourage individuals to reach across social divides. These research findings suggest that frequency and intensity of online interactions, network in- and out-degree, and contact with outgroup members and cross-posters should be associated with cross-posting.

There are many possible consequences of interaction across groups. The previous discussion suggests one consequence that we will focus on in this paper~--- the loosening of ingroup identification. If it is true that cross-talk might interrupt ingroup identification, then we might expect language will become systematically different than the language of their original group (we call this ``linguistic drift''). The contact of cross-posters with a new group will create a systematic change in how people express themselves online.

%\citeauthor{Zsolt2011NetworkEffectsandPersInfluences} studied the diffusion process of an online social network at the individual level and identified ``degree'' and ``clustering'' effects which influence the adoption of a behavior~\cite{Zsolt2011NetworkEffectsandPersInfluences}. Given the strong polarization of the two groups under study, we note that brokerage in the traditional sense may not occur but analogies to diffusion of a behavior still exist.  For example, while a majority of the authors \pat{cite number} we studied posted only in their preferred subreddit, a small number of authors posted in both subreddits. If we treat the behavior to be diffused as interacting (posting comments) with the other subreddit, then the we can still apply some of the established diffusion theories such as found in ~\citeauthor{Zsolt2011NetworkEffectsandPersInfluences}.  One difficulty presented by our data is that we are unable to determine when the members of one group begin to read the comments and posts in the opposite group (lurking); we can only establish an interaction when an author has made a comment. \citeauthor{sun2014understandinglurkers} provides a comprehensive literature review on the topics of lurkers.

%\fabio{sorry to comment the next section out, but I really couldn't follow it at all}\pat{no problem it was rough anyway}
%\citeauthor{deutsch2011handbook} identified the cause of intergroup conflict as containing ``incompatibilities, behaviors and sentiments''~\cite{deutsch2011handbook}.
%\yy{there are a lot of social psychology references about intergroup conflicts, social identity, etc. this part is too thin.}
%Men'sRights and Feminism exhibit characteristics of homophily which reinforces the polarization and strong boundaries between the two subreddits. Our data spans a considerable time frame, from the creation of Reddit through July 2016. Any changes to the community are therefore reflected in the data.  ~\citeauthor{singer2014evolutionofreddit} found that Reddit appears to be transitioning from "the front page of the internet" to a platform that is   "increasingly self-referential community that focuses on and reinforces its own user-generated image- and textual content over external sources"~\cite{singer2014evolutionofreddit}. \citeauthor{kim2011socialnetworkspoliticaldifferences} found that the use of social networks can lead to exposure to "dissimilar political views"~\cite{kim2011socialnetworkspoliticaldifferences}.  We also note that the authors of the comments under study are anonymous or at least pseudonymous and may use boundary regulation behaviors such as multiple profiles described in ~\cite{stutzman2012boundaryregulation} or may not even be human (bots).

%\citeauthor{currarini2009economicmodeloffriendship} provides detailed background concerning homophily of social groups and the relationship of homophily to group size. The two subreddits observed in this paper differ in size; Men'sRights having considerably more active authors than Feminism~\cite{currarini2009economicmodeloffriendship}.

%\citeauthor{brundidge2010encounteringdifference} describes the concept of \textit{inadvertency} as it relates to polarization, in particular political polarization in which people who seek a heterogeneous social environment are inadvertently exposed to political difference.  While we have not completed a study of the language and type of interaction of the authors who post in both the Men'sRights and Feminsim subreddits, their existence leave open the possibility of a mechanism for inadvertency.

%\subsection{Diffusion and Brokerage}

%\yy{not entirely convinced that we can use the concept of brokerage in the context of polarized groups as mentioned above. }

\section{Background}

Reddit\footnote{http://www.reddit.com/} is a social sharing and topical discussion forum founded in 2005. It is one of the largest online information sharing and discussion forums, with more than 82 billion page views and more than 73 million posts in 2015~\cite{reddit2015Reddit2015}. Reddit consists of numerous ``subreddits,'' each of which is a topical thread or category containing related content. The users who manage, or edit, a subreddit may impose rules about what may be posted.

Users can anonymously create an account which allows them to participate in existing subreddits or create new subreddits concerning a topic of their choosing. There is a subreddit for almost any topic: movies, gaming, sports, politics, and so on. Users can subscribe to subreddits that interest them.

User subreddit activities include submiting URLs, writing posts, and commenting to posts or other comments. Users may signal their approval, or disapproval, of content by up- or down-voting posts or comments. Upvotes move submissions up the page to make them more viewable to other users, and downvotes move them down the page to make them less viewable. Every submission is coupled with a comment section. Users can comment in reply to the original submission or to another user's comment. Comments can also be voted on with upvotes and downvotes. Upvotes also move comments up the comment section, and downvotes move comments down the comment section. If a comment receives too many downvotes, then it is hidden from view, and users must click to see the comment and in some cases, too many downvotes leads to removal. Users are given link `karma' and comment `karma', which measure how well their content is viewed by other users. A user's link and comment karma denotes their total difference of submission and comment upvotes and downvotes.

%\del{In 2015 Reddit received 82.43 billion page views, 73.15 million submissions, and 725.85 million comments.}

\subsection{/r/Men'sRights (MR) and /r/Feminism (Fem)}

The two subreddits under study can be found at https://www.reddit.com/r/Men'sRights/ and https://www.reddit.com/r/Feminism/. Each subreddit page contains the number of subscribers, intent of the subreddit, rules and generally some links to introductory material to get any reader up to speed on the subreddit topic. The feminist moderators' goal for the subreddit is to serve as ``a space for discussing and promoting awareness of issues related to equality for women.'' The Men's Rights moderators' intentions are to serve as ``place for those who wish to discuss Men's Rights and the ways said rights are infringed upon.'' Of interest, the Men's Rights subreddit homepage links to an article by A Voice for Men, a Men's Rights advocacy group, titled ``What's the Difference Between the Men's Rights Movement and Feminism''~\cite{white2011differencebetweenMRandFem}. The article claims that between Men's Right's and Feminism ``there can be no common ground.'' The author claims that there can be no common ground until feminists ``first acknowledge that their movement has spread anti-male bigotry throughout western governments''. % Figure \ref{fig:net_diag} depicts the polarization between these two subreddits.

%\begin{figure}[htb]
%    \centering
    %\includegraphics[scale=0.15]{network_dig}
    %\caption{Network Diagram Showing Polarization \yy{it's misleading as the blue nodes and edges block the orange ones. it should be tweaked so that it shows the connection structure better. who drew it? where's the raw file?}}\pat{diagram came from our slack discussions - I'll have to see if the repository contains any better representations}\yy{i remember it's from Meng? I think we should invite her as a coauthor if we were to put a network viz. }
%    \label{fig:net_diag}
%\end{figure}

%\subsection{Interactions Between /r/Men'sRights and /r/Feminism}
% Analysis of authors who post in MR and Fem reveals three distinct groups: authors who post only in MR, authors who post only in Fem and authors who post in both subreddits. Figure \ref{fig:bc_ratio} depicts a timeline of normalized author count by author type. As the larger subreddit, /r/Men'sRights possesses the largest amount of authors, followed by /r/Feminism and finally, authors who comment in both subreddits.  Visualization of comment data in this manner can also lead to other interesting observations, for example the spike in authors who posted in both subreddits during 2012 and 2013 which indicates abnormal rates of intergroup interaction.

\subsection{Data}
The data was retrieved from Google BigQuery and aggregated on https://www.pushshift.io/. %BigQuery is a restful web service for big datasets that works in conjunction with Google Storage launched in May of 2010. BigQuery hosts Reddit comment data. The data set begins in 2005 and is updated to the present. 
As of July 2016, there were over 1.4 billion total comments in the reddit comment dataset.
Table~\ref{tab:datasum} summarizes the comment data after cleaning.

\begin{table}[htb]
    \centering
    \caption{Reddit Commment Data Summary                      }
    \resizebox{\columnwidth}{!}{ 
    \begin{tabular}{|c|c|c|c|c|}
    \hline
    Subreddit & Start Date& Stop Date& \# of Authors&\# of Comments  \\
    \hline
    \hline
    /r/Feminism& 1/11/2009&7/31/2016&31,772&213,082\\
    /r/Men'sRights&3/21/2008&7/31/2016&115,073&2,452,442\\
    \hline
    total&&&146,845&2,665,524\\
    \hline
    \end{tabular}}
    \label{tab:datasum}
\end{table}

Because we are focusing on users, we only use comments that are associated with known authors. Comments may not have an associated author if the original author deleted their account. Such comments were removed from the dataset.

%\del{Prior to analysis, the comment data set was cleaned and reduced in size by removing comments where the author had been deleted.  Author deletion in Reddit can occur when the user deletes their account after a comment has been posted. While the comments retain their unique comment id, the linkage to an author is removed.  Often, the comment body is deleted as well. These comments provide minimal value to our longitudinal study and our analysis benefited from their removal. }

%all comments from users of deleted accounts were removed. Since the authors of these comments are unknown, they were removed from the dataset.

We removed comments from two well-known bots: \texttt{AutoModerator} and \texttt{OriginalPostSearcher}. AutoModerator is a feature of Reddit that allows moderators to establish rules that are automatically enforced. This bot will leave messages indicating that a user violated the subreddit's rules and the comment or post was deleted. OriginalPostSearcher automatically cites cross-posts (``x-posts'') across reddit. A cross-post is a post in a subreddit that was originally posted in a different subreddit. OriginalPostSearcher will automatically comment a citation of the cross post with a link to the original subreddit, author's page, and the original post. We excluded the content generated by these bots from the analysis. From a total of 3,301,859 total comments in the Men's Rights and Feminism dataset, 636,339 comments  were removed before analysis. %(no author- 634,797, AutoModerator - 1,480, and OriginalPostSearcher - 62 comments). \yy{maybe discuss as a limitation in the discussion}

\section{Model Estimation Strategy and Results}
\yy{I'd just say ``Results'' as the section title}

\subsection{Descriptive Results}

We begin with a qualitative description on three types of users: those who post in Men's Rights or Feminism exclusively and those who have posted in both. Figure~\ref{fig:bc_ratio} shows the proportion of each group over time. \yy{The figure is unclear. is this cumulative or decided each month? a user never disappears?} A few things are apparent. First, the Men's Rights subreddit is much more popular than the Feminism subreddit. Second, the majority of users will post in one subreddit and remain there and this is stable over time \yy{this statement is a bit confusing if the figure is cumulative}. Only about 8\% of users would ever post in the other subreddit. Third, there is some modest variation in time. Around 2012, there was an increase cross-talk between these two groups. \yy{if it's a unfounded claim (although it's something that can be checked), maybe we should not say anything about it..} \del{This may be due to the heightened atmosphere generated by the 2012 U.S. presidential election.}

\begin{figure}[tbph]
    \centering
    \includegraphics[scale=0.41]{bc_ratios_seabornstyle}
    \caption{Normalized Trajectories of Active Authors by Author Type}
    \label{fig:bc_ratio}
\end{figure}

%\subsection{Statistical Modelling Strategy}
\subsection{Proportional Hazard Model}

Here, we test our hypotheses by applying an event history analysis to a longitudinal data set of user behaviors. For each user, we create time-dependent variables that measure each independent variable over time. \yy{monthly?} The dependent variable is set to zero for all users in the first time period, and switches to one when the user posts to the ``other'' subreddit. That is, if a user first appears in Men's Rights, the dependent variable is zero until they post in Feminism. \yy{one important measurement missing here: to convincingly argue that the home subreddit is really meaningful, we need to show that the first subreddit the users post is what they will post most in the future. among the 8\% of users, in how many users their posting in the first subreddit overwhelmingly dominate their posts? This is critical for arguing that the ``first subreddit'' is meaningful.} Conversely, a user who appears in Feminism has a dependent variable scored as zero until they post in the Men's Rights subreddit. Since we are modelling the one-way transition between states, we do not use data from the period after the first instance of cross-talk. Thus, for most cases, the dependent variable is zero for all time periods because, as Figure \ref{fig:bc_haz} shows, most users never post in the ``other'' subreddit. For approximately 8 percent of cases, the dependent variable will eventually switch to one. Our initial exploration of the data suggested that monthly time periods were appropriate. Most cross-talking starts, if it ever happens, approximately eighteen months after the initial post in the ``first'' subreddit, which suggests that daily or weekly measurements would create zero inflated data and yearly measurements would suppress variation in the time to first cross-talk.

The proportional hazard model is a statistical model of the transition between states. The intuition of the model is that the odds of transitioning between states depends on two things, an underlying hazard rate that varies over time and the effects of the covariates (features). Some variables, like gender, might be fixed while others vary over time, like yearly income. The model estimates a multiplicative effect of the covariates on the rate of transitions. When estimated using maximum partial likelihood estimation, the model produces coefficients that are very similar to a logistic regression model, the coefficients indicate a multiplicative effects on the odds of transition~\cite{blossfeld2014event}. %%% ---- Sarvo added the equation ---- %%%%
The Cox proportional hazard model has the form (at time $t$)\\
\begin{equation}
\begin{split}
\lambda(t | X_{i}) & = \lambda_{0}(t) \exp(\beta_{1}X_{i1}+..+\beta_{p}X_{ip}) \\ 
&= \lambda_{0}(t) \exp (X_{i}^T \beta), 
\end{split}
\end{equation}
where $\lambda$ is a \emph{hazard function}, $X_{i} = (X_{i1},. . ., X_{ip})^T$ is the covariate vector for subject $i$, and $\beta$ is the coefficient vector.

%%% ---- Sarvo added the equation ---- %%%%

%\del{The main purpose of this paper is to test three hypotheses about the tendency for a Reddit user to begin posting in the ``opposite'' subreddit. The modelling strategy that we employ is event history analysis. That is, we created a longitudinal data set of Reddit users that contains information about the reddit user, their behavior, and when they begin to post in the ``other'' subreddit.} \del{The transition from ``not boundary'' to ``boundary crossing'' is modelled as a discrete event. Thus, the data is organized in a longitudinal matter.} The independent variables may change over time and the dependent variable is equal to zero until the event occurs. In that time period, the dependent variable switches to one. \yy{more precise definitions and proper citations please}

Our independent variables are defined in the following way. First, we produce two measures of openness: the number of characters in the user's account name (which is fixed over time) \yy{needs justification} and the average monthly number of words they write in their ``first'' subreddit (e.g., Men's Rights or Feminism) \yy{again, it's conspicuous here that we don't actually know whether the first subreddit align well with ppl's identity.}. Second, we produce measures of network in- and out-degree. We collect the number of monthly comments they write on other redditors posts (out-degree) and the number of monthly comments made on posts written by the user (in-degree). To assess the degree of clustering among cross-talkers, we measure the monthly number of cross-posters who comment on a user's posts and the monthly number of comments the user leaves on threads started by cross-talkers.

We include the following control variables in the analysis. First, we include the account age in months to account for the fact that people who have been on Reddit long might be more willing to rech out to different groups. Second, we include a measure of the user's quality of comments by include the average monthly score (up votes minus down votes). This addresses the possibility that cross-posting is associated with people who are more effective communicators. We also include variables that account for the interaction between the user and those from within their subreddit or the opposing subreddit. We also include dummy variables for each subreddit to account for unmeasured user heterogeneity and yearly dummy variables to account for the possibility that the tendency to cross post varies over time.

\begin{table}[htb]
    \centering
    \caption{Comment Variables and Descriptions}
    \resizebox{\columnwidth}{!}{ 
    \begin{tabular}{|c|p{6cm}|}
    \hline
    Variable &	Definition\\
    \hline
    \hline
    First Cross-Talk& 1 if the user cross-talked for the first time in current month, 0 else\\
    Name Length& Username Length \\
    %account.post.karma& Total Account post Score\\
    %account.comment.karma& Total Account comment Score\\
    %karma\_ratio& Ratio of Account Score to Comment Score\\
   Account Age& Number of days since account was created\\
    Feminism & 1 if Homesubreddit is Feminism, 0 else\\
    wordiness & number of words per/post in a month \\
    Comments Written by User & number of outgoing comments in a month\\
    Comments on User’s Posts & number of incoming comments in a month\\
    Average Score & Mean score in a month\\
    Crossposter Talks to User & 1 if a crossposter talked to me, 0 else\\
    User Talks to Crossposter & 1 if I talked to a crossposter, 0 else\\
    %other\_sr\_to\_me&	1 if someone from the other subreddit talked to me, 0 otherwise\\
    %me\_to\_other\_srs&	1 if I talked to someone from other subreddit, 0 otherwise\\
    \hline
    \end{tabular}}    
    \label{tab:field_def}
\end{table}

We present the results of the event history analysis in two parts. First, we discuss selected hazard curves, which provide intuition about the effects of the independent variables on the decision to post in the ideologically opposite reddit. A hazard curve visualizes the proportion of people over time who adopt a new behavior. The hazard curves start in the month in which the oldest subreddit began - Feminism - which is labeled ``30'' because thirty months passed since the beginning of reddit. Then, we will present the results of the Cox proportional hazards model.

Figure~\ref{fig:bc_haz} shows the hazard curve by original ``home'' subreddit. \yy{define the hazard rate plz} That is, we look at the hazard rate for the entire population and separately for each subreddit. \yy{unclear how it is aggregated} As suggested by the Figure~\ref{fig:bc_haz}, there is a difference between the two ideological subreddits, with Feminist redditors cross-talking at a higher rate than their counterparts. \yy{I don't think we have evidence to claim this. As I wrote before, it can purely come from the size difference. There can be many null models that may explain this without assuming any systematic differences between participants.} \del{This strongly suggests that there are systematic differences between participants in the Men`s Rights and Feminism reddits.} Figure~\ref{fig:ctc_haz} shows that having contact with a cross poster has a substantial effect on cross posting. \yy{activity-controlled? if not, it may be rather trivial (more activity -> more likely to cross post by pure chance). unclear to me how these curves are obtained. precise definitions are needed. } Figure \ref{fig:outmean_haz}. shows that those who leave comments on other posts are much more likely to post.

\begin{figure}[ht]{}
\centering
    \includegraphics[scale=0.75]{Overall_Hazard_thick}
    \caption{Hazard Curve by Home Subreddit}
    \label{fig:bc_haz}
\end{figure}

%We found other differences. Figure \ref{fig:ctc_haz} shows that there are smaller, modest differences between those individuals who have had contact with other \pat{authors who have posted in the opposite subreddit (``boundary crossers'')} \del{boundary crossers} when compared to those who have not had contact.

\begin{figure}[ht]{}
\centering
    \includegraphics[scale=0.75]{contact_Hazard_final}
    \caption{Contact Hazard Curve}
    \label{fig:ctc_haz}
\end{figure}

\begin{figure}[ht]{}
\centering
    \includegraphics[scale=0.75]{outmean_Hazard_final}
    \caption{User Replies to Others Hazard Curve}
    \label{fig:outmean_haz}
\end{figure}
%In event history analysis, there are multiple methods for estimating the effects of the time-dependent co-variates. Here, we employ a Cox proportional hazards model, which estimates the effects of the independent variables on the rate at which cases move to a new state. The coefficients are interpreted as multiplicative effects on the hazard rate.

%\subsection{Results}
We begin with descriptive statistics. Figures~\ref{fig:bc_haz}-\ref{fig:outmean_haz} look at hazard curves \yy{precise definition needed} calculated for the entire sample. Overall, we find that only 8\% of all users in the Feminist and Men's Rights subreddits ever comment in the corresponding opposing group.\yy{this is already mentioned above, and it is not new information from the hazard curve} Figure~\ref{fig:bc_haz} indicates that interaction between Men's Rights and Feminism authors is a relatively uncommon behavior. \yy{same. nothing new here} The hazard curves also show important variation between subreddits. We found that users who begin in /r/Feminism were much more likely to post in /r/Men'sRights than the opposite.  Roughly speaking, /r/Feminist authors are 2.5 times as likely to cross the boundary into Men's Rights as their counterparts. \yy{no. without evidence that shows the community size difference does not matter, the following statement is not supported. there can be many null models that explain this without any difference in behavior but with pure size difference. } \del{This strongly suggests that these groups are highly heterogeneous. This may be due to selection effects or the norms that are adopted in the various groups regarding having discussions with others.}

\begin{table}
\caption{Descriptive Statistics and Correlations}
\centering
\begin{threeparttable}
\begin{adjustbox}{max width=\columnwidth}
\begin{tabular}{|c||r|r|r|r|r|}
\hline
& \multicolumn{5}{|c|}{\textbf{Descriptive Statistics}} \\\hline
\multicolumn{1}{|c||}{\textbf{Variable}}& \multicolumn{1}{l|}{\textbf{N}} & \multicolumn{1}{l|}{\textbf{mean}} & \multicolumn{1}{l|}{\textbf{sd}} & \multicolumn{1}{l|}{\textbf{min}} & \multicolumn{1}{l||}{\textbf{max}} \\\hline
Name Length & 4,605,955 & 10.7558 & 3.7168 & 3  & 20\\
First Cross-Talk & 4,605,955 & 0.0019 & 0.0437 & 0 & 1 \\
Wordiness& 4,605,955 & 3.5409 & 20.7327 & 0     & 1744 \\
Comments Written by User & 4,605,955 & 0.0379 & 0.1759 & 0 & 2 \\
Comments on User's Posts & 4,605,955 & 0.0402 & 0.2370 & 0 & 33 \\
Average Score & 4,605,955 & 0.3843 & 4.0352 & -266  & 1216 \\
Account Age & 4,302,176 & 1770.37 & 624.4882 & 1& 4021  \\
Feminism & 4,605,955 & 0.1512 & 0.3583 & 0  & 1 \\
Crossposter Talks to User & 4,605,955 & 0.0046 & 0.0675 & 0 & 1 \\
User Talks to Crossposter & 4,605,955 & 0.0044 & 0.0663 & 0  & 1\\
Difference between comments out and in & 4,605,955 & -0.0023 & 0.2071 & -33   & 2 \\\hline

\end{tabular}%
\end{adjustbox}
\end{threeparttable}
\label{tab:descrstat}
%\caption{\yy{sorry but the font size is way too small. We can't put a table like this. Same for the other tables. The all text in the table should be in legible size. maybe a couple of points (pt) smaller than the body text is fine. but no smaller than that.}}
\end{table}

\begin{table*}
\caption{Correlation Table}
\centering
\begin{threeparttable}
\begin{adjustbox}{max width=\textwidth}
\begin{tabular}{|p{4cm}||r|r|r|r|r|r|r|r|r|r|r|}
\hline
& \multicolumn{11}{c|}{\textbf{Correlations}} \\\hline
\multicolumn{1}{|c||}{\textbf{Variable}} & \multicolumn{1}{l|}{\textbf{Name Length}} & \multicolumn{1}{p{2cm}|}{\textbf{First Cross-Talk}} & \multicolumn{1}{l|}{\textbf{Wordiness}} & \multicolumn{1}{p{2cm}|}{\textbf{Comments Written by User}} & \multicolumn{1}{p{2cm}|}{\textbf{Comments on User's Posts}} & \multicolumn{1}{p{2cm}|}{\textbf{Average Score}} & \multicolumn{1}{p{2cm}|}{\textbf{Account Age}} & \multicolumn{1}{l|}{\textbf{Feminism}} & \multicolumn{1}{p{2cm}|}{\textbf{Crossposter Talks to User}} & \multicolumn{1}{p{2cm}|}{\textbf{User Talks to Crossposter}} & \multicolumn{1}{p{2cm}|}{\textbf{Difference between comments out and in}} \\\hline
Name Length & 1.0000 &  & &  &  &  & &  &  & & \\
First Cross-Talk & 0.0003 & 1.0000 & & & & & & & & &  \\
Wordiness& -0.0003 & 0.1119 & 1.0000 &  &  & &       &       &       &       &  \\
Comments Written by User & -0.0008 & 0.1207 & 0.4515 & 1.0000 & & & & & & &  \\
Comments on User's Posts & 0.0000 & 0.0909 & 0.4125 & 0.5287 & 1.0000 & & & & & &  \\
Average Score & -0.0010 & 0.0351 & 0.1991 & 0.2512 & 0.4260 & 1.0000 & & & & &  \\
Account Age & -0.2014 & -0.0105 & -0.0465 & -0.0586 & -0.0482 & -0.0273 & 1.0000 &       & & &  \\
Feminism & 0.0331 & 0.0233 & 0.0078 & -0.0215 & -0.0154 & -0.0059 & -0.1583 & 1.0000 & & &  \\
Crossposter Talks to User & -0.0009 & 0.1289 & 0.1859 & 0.2094 & 0.2428 & 0.0831 & -0.0109 & 0.0162 & 1.0000 & &  \\
User Talks to Crossposter & -0.0004 & 0.1294 & 0.1726 & 0.2501 & 0.1624 & 0.0627 & -0.0104 & 0.0158 & 0.6140 & 1.0000 &  \\
Difference between comments out and in & -0.0007 & -0.0018 & -0.0894 & 0.2419 & -0.6957 & -0.2743 & 0.0055 & -0.0006 & -0.1003 & 0.0260 & 1.0000 \\\hline

\end{tabular}%
\end{adjustbox}
\end{threeparttable}
\label{tab:descrstatcorr}
%\caption{\yy{sorry but the font size is way too small. We can't put a table like this. Same for the other tables. The all text in the table should be in legible size. maybe a couple of points (pt) smaller than the body text is fine. but no smaller than that.}}
\end{table*}

%Similarly, Figure \ref{fig:outmean_haz} shows modest differences between \del{users} \pat{authors} who have an above average number of replies to others' posts are much more likely to become \del{cross posters} \pat{boundary crossers}. Together, these hazard models suggest that openness and contact with boundary crossers are associated with boundary crossing.

Table~\ref{tab:regrtab} reports a series of models that independently, then jointly, test the hypotheses about the effects of different aspects of online behavior on boundary crossing. In many cases, our measures of the volume of online participation have positive and significant effects, yet they are small in size. For example, in Models 1 and 2, the effect of the average number of words per post is small ($0.005$). Thus, it would take hundreds of words per post to generate a large effect according to this model. Similar things can be said about the other variables in this table. The effects are small, though statistically significant.

The next table shows a much stronger effect for having an interaction with other users who are already cross-posting. Depending on the direction of the communication, the effect of having an interaction with a cross-poster is roughly 4.6 and reduces to 3.3 with the inclusion of control variables (see Models 9-12). This is larger than nearly every other variable with a significant effect included in the model.

Table 6 shows the full model with all independent variables and control variables included. In this model, many of the variables are still significant but the control variables have reduced the size of the main effects. In this model, contact with cross-posters, both to and from, has a significant effect but it is much attenuated. The variable of interest with the largest effect is the number of comments written by the user ($2.98$).

These tables suggest the following about the relationship between the intensity of online interactions and cross-posting or cross-talk across these two online communities. First, there is a modest effect of the intensity and frequency of interaction on the initiation of outgroup interaction. Second, the largest effects tend to be associated with behavior where the user initiates the behavior. Talk to a cross-poster has a larger effect on cross-talk than being talke to by a cross-poster. A similar result was found for being the source and target of comments. Third, the control variables in Table 6 are either not significant or their estimated is extremely small. This suggests that the differences between reddits and the differences over time that were discussed in the descriptive statistics should be attributed to changes in the number of users who dedicated participants in the reddits.

\yy{maybe good to extract each table to a separate file. easier to edit.}

\begin{table*}
\caption{Modelling User Activity on Cross-talk}
\centering
\begin{threeparttable}
\begin{adjustbox}{max width=\textwidth}
\begin{tabular}{|p{4cm}|c|cccc|cccc|cccc|cccc|cccc|cccc|}
\hline
\multicolumn{1}{|r|}{} &       & \multicolumn{4}{c|}{\textbf{Model 1}} & \multicolumn{4}{c|}{\textbf{Model 2}} & \multicolumn{4}{c|}{\textbf{Model 3}} & \multicolumn{4}{c|}{\textbf{Model 4}} & \multicolumn{4}{c|}{\textbf{Model 5}} & \multicolumn{4}{c|}{\textbf{Model 6}} \\\hline

\multicolumn{1}{|c|}{\textbf{Variable}} & \textbf{Value} & \multicolumn{1}{c|}{\textbf{coef}} & \multicolumn{1}{c|}{\textbf{z}} & \multicolumn{1}{c|}{\textbf{p-value}} & \textbf{sig} & \multicolumn{1}{c|}{\textbf{coef}} & \multicolumn{1}{c|}{\textbf{z}} & \multicolumn{1}{c|}{\textbf{p-value}} & \textbf{sig} & \multicolumn{1}{c|}{\textbf{coef}} & \multicolumn{1}{c|}{\textbf{z}} & \multicolumn{1}{c|}{\textbf{p-value}} & \textbf{sig} & \multicolumn{1}{c|}{\textbf{coef}} & \multicolumn{1}{c|}{\textbf{z}} & \multicolumn{1}{c|}{\textbf{p-value}} & \textbf{sig} & \multicolumn{1}{c|}{\textbf{coef}} & \multicolumn{1}{c|}{\textbf{z}} & \multicolumn{1}{c|}{\textbf{p-value}} & \textbf{sig} & \multicolumn{1}{c|}{\textbf{coef}} & \multicolumn{1}{c|}{\textbf{z}} & \multicolumn{1}{c|}{\textbf{p-value}} & \textbf{sig} \\\hline

Wordiness &       & \multicolumn{1}{r|}{0.0060} & \multicolumn{1}{r|}{148.8} & \multicolumn{1}{r|}{\textless2e-16} & \multicolumn{1}{r|}{***} & \multicolumn{1}{r|}{0.0053} & \multicolumn{1}{r|}{110.506} & \multicolumn{1}{r|}{\textless2e-16} & \multicolumn{1}{r|}{***} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &  \\
Comments on User's Posts &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{0.3307} & \multicolumn{1}{r|}{121.1} & \multicolumn{1}{r|}{\textless2e-16} & \multicolumn{1}{l|}{***} & \multicolumn{1}{r|}{0.3532} & \multicolumn{1}{r|}{90.617} & \multicolumn{1}{r|}{\textless2e-16} & \multicolumn{1}{l|}{***} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &  \\
Comments Written by User &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{3.8596} & \multicolumn{1}{r|}{177.7} & \multicolumn{1}{r|}{\textless2e-16} & \multicolumn{1}{l|}{***} & \multicolumn{1}{r|}{3.559} & \multicolumn{1}{r|}{143.92} & \multicolumn{1}{r|}{\textless2e-16} & \multicolumn{1}{l|}{***} \\\hline

\textbf{Controls} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &  \\\hline

Account age &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{-0.0009} & \multicolumn{1}{r|}{-34.102} & \multicolumn{1}{r|}{\textless2e-16} & \multicolumn{1}{r|}{***} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{-0.0010} & \multicolumn{1}{r|}{-36.247} & \multicolumn{1}{r|}{\textless2e-16} & \multicolumn{1}{r|}{***} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{-0.0004} & \multicolumn{1}{r|}{-17.445} & \multicolumn{1}{r|}{\textless2e-16} & \multicolumn{1}{r|}{***} \\
Name Length &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{-0.0157} & \multicolumn{1}{r|}{-5.115} & \multicolumn{1}{r|}{3.13e-07} & \multicolumn{1}{r|}{***} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{-0.0172} & \multicolumn{1}{r|}{-5.606} & \multicolumn{1}{r|}{2.07e-08} & \multicolumn{1}{r|}{***} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{-0.0106} & \multicolumn{1}{r|}{-3.456} & \multicolumn{1}{r|}{0.0005} & \multicolumn{1}{r|}{***} \\
Feminism &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{1.003} & \multicolumn{1}{r|}{41.491} & \multicolumn{1}{r|}{\textless2e-16} & \multicolumn{1}{r|}{***} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{1.068} & \multicolumn{1}{r|}{44.17} & \multicolumn{1}{r|}{\textless2e-16} & \multicolumn{1}{r|}{***} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{1.194} & \multicolumn{1}{r|}{49.991} & \multicolumn{1}{r|}{\textless2e-16} & \multicolumn{1}{r|}{***} \\
Average Score &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{0.0062} & \multicolumn{1}{r|}{25.161} & \multicolumn{1}{r|}{\textless2e-16} & \multicolumn{1}{r|}{***} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{0.0009} & \multicolumn{1}{r|}{2.481} & \multicolumn{1}{r|}{0.0131} & \multicolumn{1}{r|}{*} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{0.0048} & \multicolumn{1}{r|}{9.34} & \multicolumn{1}{r|}{\textless2e-16} & \multicolumn{1}{r|}{***} \\
2010 & \multicolumn{1}{l|}{\textbf{TRUE}} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{12.73} & \multicolumn{1}{r|}{0.631} & \multicolumn{1}{r|}{0.5281} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{99.12} & \multicolumn{1}{r|}{Inf} & \multicolumn{1}{r|}{\textless2e-16} & \multicolumn{1}{r|}{***} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{31.8} & \multicolumn{1}{r|}{Inf} & \multicolumn{1}{r|}{\textless2e-16} & \multicolumn{1}{r|}{***} \\
2011 & \multicolumn{1}{l|}{\textbf{TRUE}} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{12.51} & \multicolumn{1}{r|}{1.701} & \multicolumn{1}{r|}{0.0890} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{17.98} & \multicolumn{1}{r|}{0.158} & \multicolumn{1}{r|}{0.8746} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{23.8} & \multicolumn{1}{r|}{0.011} & \multicolumn{1}{r|}{0.9909} &  \\
2012 & \multicolumn{1}{l|}{\textbf{TRUE}} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{12.42} & \multicolumn{1}{r|}{2.816} & \multicolumn{1}{r|}{0.0049} & \multicolumn{1}{r|}{**} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{15.43} & \multicolumn{1}{r|}{0.774} & \multicolumn{1}{r|}{0.4391} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{16.09} & \multicolumn{1}{r|}{0.507} & \multicolumn{1}{r|}{0.6125} &  \\
2013 & \multicolumn{1}{l|}{\textbf{TRUE}} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{11.18} & \multicolumn{1}{r|}{3.65} & \multicolumn{1}{r|}{0.0003} & \multicolumn{1}{r|}{***} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{16.76} & \multicolumn{1}{r|}{0.342} & \multicolumn{1}{r|}{0.7326} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{15.29} & \multicolumn{1}{r|}{0.6} & \multicolumn{1}{r|}{0.5484} &  \\
2014 & \multicolumn{1}{l|}{\textbf{TRUE}} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{10.56} & \multicolumn{1}{r|}{3.74} & \multicolumn{1}{r|}{0.0002} & \multicolumn{1}{r|}{***} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{15.24} & \multicolumn{1}{r|}{0.564} & \multicolumn{1}{r|}{0.5725} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{14.88} & \multicolumn{1}{r|}{0.604} & \multicolumn{1}{r|}{0.5462} &  \\
2015 & \multicolumn{1}{l|}{\textbf{TRUE}} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{10.08} & \multicolumn{1}{r|}{2.965} & \multicolumn{1}{r|}{0.0030} & \multicolumn{1}{r|}{**} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{16.22} & \multicolumn{1}{r|}{0.22} & \multicolumn{1}{r|}{0.8257} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{19.8} & \multicolumn{1}{r|}{0.043} & \multicolumn{1}{r|}{0.9655} &  \\\hline

Number of User Months &       & \multicolumn{4}{c|}{\textbf{4,605,955}} & \multicolumn{4}{c|}{\textbf{4,605,955}} & \multicolumn{4}{c|}{\textbf{4,605,955}} & \multicolumn{4}{c|}{\textbf{4,605,955}} & \multicolumn{4}{c|}{\textbf{4,605,955}} & \multicolumn{4}{c|}{\textbf{4,605,955}} \\\hline
Authors who interact with outgroup &       & \multicolumn{4}{c|}{\textbf{8,816}} & \multicolumn{4}{c|}{\textbf{8,021}} & \multicolumn{4}{c|}{\textbf{8,816}} & \multicolumn{4}{c|}{\textbf{8,021}} & \multicolumn{4}{c|}{\textbf{8,816}} & \multicolumn{4}{c|}{\textbf{8,021}} \\\hline
Missing Observations &       & \multicolumn{4}{c|}{\textbf{0}} & \multicolumn{4}{c|}{\textbf{303,779}} & \multicolumn{4}{c|}{\textbf{0}} & \multicolumn{4}{c|}{\textbf{303,779}} & \multicolumn{4}{c|}{\textbf{0}} & \multicolumn{4}{c|}{\textbf{303,779}} \\\hline
$R^2$ &       & \multicolumn{4}{c|}{\textbf{0.001}} & \multicolumn{4}{c|}{\textbf{0.01}} & \multicolumn{4}{c|}{\textbf{0.001}} & \multicolumn{4}{c|}{\textbf{0.001}} & \multicolumn{4}{c|}{\textbf{0.005}} & \multicolumn{4}{c|}{\textbf{0.013}} \\\hline
Chis-squared Value &       & \multicolumn{4}{c|}{\textbf{4890}} & \multicolumn{4}{c|}{\textbf{42570}} & \multicolumn{4}{c|}{\textbf{2613}} & \multicolumn{4}{c|}{\textbf{41183}} & \multicolumn{4}{c|}{\textbf{22339}} & \multicolumn{4}{c|}{\textbf{55725}} \\\hline

\multicolumn{3}{l}{notes: * $p \leq 0.05$ ; ** $p \leq 0.01$ ; *** $p \leq 0.001$}{}&
\multicolumn{5}{l}{}&
\multicolumn{5}{l}{}&
\multicolumn{5}{l}{}\\
\end{tabular}
\end{adjustbox}
\end{threeparttable}
\label{tab:regrtab}
\end{table*}

\begin{table*}
\caption{Modelling Contact with Other Users on Cross-talk}
\centering
\begin{threeparttable}
 \begin{adjustbox}{max width=\textwidth}
                \begin{tabular}{|p{3.75cm}|c|cccc|cccc|cccc|cccc|cccc|cccc|}\hline
    \multicolumn{1}{|r|}{} &       & \multicolumn{4}{c|}{\textbf{Model 7}} & \multicolumn{4}{c|}{\textbf{Model 8}} & \multicolumn{4}{c|}{\textbf{Model 9}} & \multicolumn{4}{c|}{\textbf{Model 10}} & \multicolumn{4}{c|}{\textbf{Model 11}} & \multicolumn{4}{c|}{\textbf{Model 12}} \\\hline
    \textbf{Variable} & \multicolumn{1}{l|}{\textbf{Value}} & \multicolumn{1}{l|}{\textbf{coef}} & \multicolumn{1}{l|}{\textbf{z}} & \multicolumn{1}{l|}{\textbf{p-value}} & \multicolumn{1}{l|}{\textbf{sig}} & \multicolumn{1}{l|}{\textbf{coef}} & \multicolumn{1}{l|}{\textbf{z}} & \multicolumn{1}{l|}{\textbf{p-value}} & \multicolumn{1}{l|}{\textbf{sig}} & \multicolumn{1}{l|}{\textbf{coef}} & \multicolumn{1}{l|}{\textbf{z}} & \multicolumn{1}{l|}{\textbf{p-value}} & \multicolumn{1}{l|}{\textbf{sig}} & \multicolumn{1}{l|}{\textbf{coef}} & \multicolumn{1}{l|}{\textbf{z}} & \multicolumn{1}{l|}{\textbf{p-value}} & \multicolumn{1}{l|}{\textbf{sig}} & \multicolumn{1}{l|}{\textbf{coef}} & \multicolumn{1}{l|}{\textbf{z}} & \multicolumn{1}{l|}{\textbf{p-value}} & \multicolumn{1}{l|}{\textbf{sig}} & \multicolumn{1}{l|}{\textbf{coef}} & \multicolumn{1}{l|}{\textbf{z}} & \multicolumn{1}{l|}{\textbf{p-value}} & \multicolumn{1}{l|}{\textbf{sig}} \\\hline
    Difference between comments out and in &       & \multicolumn{1}{r|}{-0.1624} & \multicolumn{1}{r|}{-5.53} & \multicolumn{1}{r|}{3.3e-08} & \multicolumn{1}{r|}{***} & \multicolumn{1}{r|}{0.1826} & \multicolumn{1}{r|}{4.461} & \multicolumn{1}{r|}{8.2e-06} & \multicolumn{1}{r|}{***} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &  \\
    Crossposter Talks to User & \multicolumn{1}{l|}{TRUE} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{4.6017} & \multicolumn{1}{r|}{172.7} & \multicolumn{1}{r|}{\textless2e-16} & \multicolumn{1}{l|}{***} & \multicolumn{1}{r|}{3.321} & \multicolumn{1}{r|}{109.73} & \multicolumn{1}{r|}{\textless2e-16} & \multicolumn{1}{l|}{***} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &  \\
    User Talks to Crossposter & \multicolumn{1}{l|}{TRUE} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{4.6121} & \multicolumn{1}{r|}{172.2} & \multicolumn{1}{r|}{\textless2e-16} & \multicolumn{1}{l|}{***} & \multicolumn{1}{r|}{3.366} & \multicolumn{1}{r|}{111.77} & \multicolumn{1}{r|}{\textless2e-16} & \multicolumn{1}{l|}{***} \\\hline
    \textbf{Controls} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &  \\\hline
    Account age &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{-0.0011} & \multicolumn{1}{r|}{-39.209} & \multicolumn{1}{r|}{\textless2e-16} & \multicolumn{1}{r|}{***} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{-0.0008} & \multicolumn{1}{r|}{-30.697} & \multicolumn{1}{r|}{\textless2e-16} & \multicolumn{1}{r|}{***} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{-0.0008} & \multicolumn{1}{r|}{-31.128} & \multicolumn{1}{r|}{\textless2e-16} & \multicolumn{1}{r|}{***} \\
    Name Length &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{-0.0197} & \multicolumn{1}{r|}{-6.459} & \multicolumn{1}{r|}{1.06e-10} & \multicolumn{1}{r|}{***} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{-0.0160} & \multicolumn{1}{r|}{-5.226} & \multicolumn{1}{r|}{1.74e-07} & \multicolumn{1}{r|}{***} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{-0.0172} & \multicolumn{1}{r|}{-5.623} & \multicolumn{1}{r|}{1.88e-08} & \multicolumn{1}{r|}{***} \\
    Feminism &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{1.031} & \multicolumn{1}{r|}{42.84} & \multicolumn{1}{r|}{\textless2e-16} & \multicolumn{1}{r|}{***} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{0.9174} & \multicolumn{1}{r|}{38.095} & \multicolumn{1}{r|}{\textless2e-16} & \multicolumn{1}{r|}{***} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{0.9217} & \multicolumn{1}{r|}{38.289} & \multicolumn{1}{r|}{\textless2e-16} & \multicolumn{1}{r|}{***} \\
    Average Score &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{0.0084} & \multicolumn{1}{r|}{31.146} & \multicolumn{1}{r|}{\textless2e-16} & \multicolumn{1}{r|}{***} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{0.0051} & \multicolumn{1}{r|}{21.762} & \multicolumn{1}{r|}{\textless2e-16} & \multicolumn{1}{r|}{***} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{0.0073} & \multicolumn{1}{r|}{32.957} & \multicolumn{1}{r|}{\textless2e-16} & \multicolumn{1}{r|}{***} \\
    2010 & \multicolumn{1}{l|}{\textbf{TRUE}} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{88.4} & \multicolumn{1}{r|}{Inf} & \multicolumn{1}{r|}{\textless2e-16} & \multicolumn{1}{r|}{***} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{17.47} & \multicolumn{1}{r|}{0.086} & \multicolumn{1}{r|}{0.932} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{17.39} & \multicolumn{1}{r|}{0.088} & \multicolumn{1}{r|}{0.93} &  \\
    2011 & \multicolumn{1}{l|}{\textbf{TRUE}} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{19.62} & \multicolumn{1}{r|}{0.076} & \multicolumn{1}{r|}{0.9390} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{18.26} & \multicolumn{1}{r|}{0.143} & \multicolumn{1}{r|}{0.886} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{18.33} & \multicolumn{1}{r|}{0.14} & \multicolumn{1}{r|}{0.8880} &  \\
    2012 & \multicolumn{1}{l|}{\textbf{TRUE}} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{19.5} & \multicolumn{1}{r|}{0.128} & \multicolumn{1}{r|}{0.8980} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{17.49} & \multicolumn{1}{r|}{0.275} & \multicolumn{1}{r|}{0.784} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{17.42} & \multicolumn{1}{r|}{0.287} & \multicolumn{1}{r|}{0.7740} &  \\
    2013 & \multicolumn{1}{l|}{\textbf{TRUE}} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{15.37} & \multicolumn{1}{r|}{0.63} & \multicolumn{1}{r|}{0.5280} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{16.25} & \multicolumn{1}{r|}{0.413} & \multicolumn{1}{r|}{0.68} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{16.18} & \multicolumn{1}{r|}{0.423} & \multicolumn{1}{r|}{0.6720} &  \\
    2014 & \multicolumn{1}{l|}{\textbf{TRUE}} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{14.59} & \multicolumn{1}{r|}{0.7} & \multicolumn{1}{r|}{0.4840} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{15.68} & \multicolumn{1}{r|}{0.422} & \multicolumn{1}{r|}{0.673} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{15.56} & \multicolumn{1}{r|}{0.448} & \multicolumn{1}{r|}{0.6540} &  \\
    2015 & \multicolumn{1}{l|}{\textbf{TRUE}} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{13.98} & \multicolumn{1}{r|}{0.591} & \multicolumn{1}{r|}{0.5540} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{15.15} & \multicolumn{1}{r|}{0.354} & \multicolumn{1}{r|}{0.724} &       & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} & \multicolumn{1}{r|}{} &       & \multicolumn{1}{r|}{15.08} & \multicolumn{1}{r|}{0.365} & \multicolumn{1}{r|}{0.7150} &  \\\hline
    Number of User Months &       & \multicolumn{4}{c|}{\textbf{4,605,955}} & \multicolumn{4}{c|}{\textbf{4,605,955}} & \multicolumn{4}{c|}{\textbf{4,605,955}} & \multicolumn{4}{c|}{\textbf{4,605,955}} & \multicolumn{4}{c|}{\textbf{4,605,955}} & \multicolumn{4}{c|}{\textbf{4,605,955}} \\\hline
    Authors who interact with outgroup &       & \multicolumn{4}{c|}{\textbf{8,816}} & \multicolumn{4}{c|}{\textbf{8,021}} & \multicolumn{4}{c|}{\textbf{8,816}} & \multicolumn{4}{c|}{\textbf{8,021}} & \multicolumn{4}{c|}{\textbf{8,816}} & \multicolumn{4}{c|}{\textbf{8,021}} \\\hline
    Missing Observations &       & \multicolumn{4}{c|}{\textbf{0}} & \multicolumn{4}{c|}{\textbf{303,779}} & \multicolumn{4}{c|}{\textbf{0}} & \multicolumn{4}{c|}{\textbf{303,779}} & \multicolumn{4}{c|}{\textbf{0}} & \multicolumn{4}{c|}{\textbf{303,779}} \\\hline
    $R^2$ &       & \multicolumn{4}{c|}{\textbf{0}} & \multicolumn{4}{c|}{\textbf{0.009}} & \multicolumn{4}{c|}{\textbf{0.003}} & \multicolumn{4}{c|}{\textbf{0.011}} & \multicolumn{4}{c|}{\textbf{0.003}} & \multicolumn{4}{c|}{\textbf{0.011}} \\\hline
    Chis-squared Value &       & \multicolumn{4}{c|}{\textbf{16.76}} & \multicolumn{4}{c|}{\textbf{39136}} & \multicolumn{4}{c|}{\textbf{12641}} & \multicolumn{4}{c|}{\textbf{46110}} & \multicolumn{4}{c|}{\textbf{12498}} & \multicolumn{4}{c|}{\textbf{46227}} \\\hline

\multicolumn{3}{l}{notes: * $p \leq 0.05$ ; ** $p \leq 0.01$ ; *** $p \leq 0.001$}{}&
\multicolumn{5}{l}{}&
\multicolumn{5}{l}{}&
\multicolumn{5}{l}{}\\
    \end{tabular}%

\end{adjustbox}
\end{threeparttable}
\label{tab:regrtab2}
\end{table*}

\begin{table}[htb]
\caption{Saturated Regression Model}
\centering
\begin{threeparttable}
 \begin{adjustbox}{max width=\columnwidth}
 \begin{tabular}{|p{4cm}|c|r|r|r|}
    \hline
          & \multicolumn{4}{c|}{\textbf{Controlled Saturated Model}} \\
    \hline
    \textbf{Variables} & \multicolumn{1}{l|}{\textbf{coef}} & \multicolumn{1}{l|}{\textbf{z}} & \multicolumn{1}{l|}{\textbf{p-value}} & \multicolumn{1}{l|}{\textbf{sig}} \\
    \hline
    Wordiness & \multicolumn{1}{r|}{0.004} & 44.55 & \textless2e-16 & *** \\
    Comments on User's Posts & \multicolumn{1}{r|}{0.208} & 18.82 & \textless2e-16 & *** \\
    Comments Written by User & \multicolumn{1}{r|}{2.98} & 105.11 & \textless2e-16 & *** \\
    Crossposter Talks to User & \multicolumn{1}{r|}{1.1} & 23.62 & \textless2e-16 & *** \\
    User Talks to Crossposter & \multicolumn{1}{r|}{0.746} & 16.11 & \textless2e-16 & *** \\\hline

    \textbf{Controls} &       &       &       &  \\\hline

    Account Age & \multicolumn{1}{r|}{-3e-04} & -14.37 & \textless2e-16 & *** \\
    Average Score & \multicolumn{1}{r|}{-1e-04} & -0.1  & 0.9222 &  \\
    Feminism & \multicolumn{1}{r|}{0.994} & 40.96 & \textless2e-16 & *** \\
    Name Length & \multicolumn{1}{r|}{-0.011} & -3.43 & 0.0006 & *** \\
    2010  & \multicolumn{1}{r|}{15.1} & 0.21  & 0.8372 &  \\
    2011  & \multicolumn{1}{r|}{14.9} & 0.63  & 0.5295 &  \\
    2012  & \multicolumn{1}{r|}{13.9} & 1.22  & 0.2211 &  \\
    2013  & \multicolumn{1}{r|}{13.1} & 1.47  & 0.142 &  \\
    2014  & \multicolumn{1}{r|}{12.8} & 1.47  & 0.1407 &  \\
    2015  & \multicolumn{1}{r|}{12.3} & 1.14  & 0.2527 &  \\\hline

    \textbf{Number of User Months} & \multicolumn{4}{c|}{4,605,955} \\\hline

    \textbf{Authors who interact
 with outgroup} & \multicolumn{4}{c|}{8,021} \\\hline

    \textbf{Missing Observations} & \multicolumn{4}{c|}{303,779} \\\hline

    \textbf{$R^2$} & \multicolumn{4}{c|}{0.013} \\\hline

    \textbf{Chis-squared Value} & \multicolumn{4}{c|}{57858} \\
    \hline
    \multicolumn{5}{l}{notes: * $p \leq 0.05$ ; ** $p \leq 0.01$ ; *** $p \leq 0.001$}{}\\
    \end{tabular}%

\end{adjustbox}
\end{threeparttable}
\label{tab:regrtab3}
\end{table}

\subsection{An Implication of Cross-talk Between Men's Rights and Feminism: Linguistic Drift}

Here, we examine one consequence of cross-talking or cross-posting across these two subreddits. We provide evidence that cross-posting is associated with linguistic drift. Users who cross-post in the opposing subreddit resemble less their ``native'' or home subreddit after they start addressing the opposite reddit. \yy{I think ``which posts?'' is important to clarify. Are we saying that user A's posts on the home subreddit drift after posting to the other subreddit? or user A's posts (across both subreddits)? The latter can be trivial. }  To study the linguistic similarity of Reddit users, we preprocessed the raw text of Reddit posts, where HTML and other markup tags were removed. Then, each post was tokenized into bigrams.

In our first analysis, we investigated if the cross-posters' linguistic pattern in their home subreddit has manifested a significant change between the pre- and post-crossposting stage. As a comparison, we also considered the posts in the same subreddit that are from authors who never cross-posted. These posts were split into those that occurred in the first and second half of an author's lifespan.

\begin{table}
    \caption{Count of Sentences of Posts From Cross-posters and Non-crossposters.}
    \centering
    \begin{adjustbox}{max width=\columnwidth}

    \begin{tabular}{|l l| r r|}\hline
        & & Men'sRights & Feminism\\
        \hline
        \hline
        \multirow{2}{*}{\textbf{Cross-poster}}& ``Pre-crossposting'' & 820,850   & 77,460\\
            & ``Post-crossposting''  & 2,494,351 & 149,112\\
            \hline
        \multirow{2}{*}{\textbf{Non-Crossposter}}   & ``First half''      & 1,315,380 & 193,058\\
            & ``Second half''       & 1,391,906 & 242,324\\\hline
    \end{tabular}
    \label{lm:tab1}
    \end{adjustbox}
\end{table}

For each group, a fixed number (100,000) of sentences were bootstrapped (i.e. sampled with replacement) from the available text data (Table~\ref{lm:tab1}), and for each sentence a log-transformed probability score was computed using bigram language model (with the Stupid backoff smoothing~\cite{Brants07largelanguage}). The bigram model was trained on the text data from non-crossposters in the same subreddit (disjoint from those used in the test set), which contains the most posts compared to those from cross-posters, and represents the mainstream linguistic pattern of the subreddit. The total number of available sentences in the four groups from which the above test and training set were sampled are briefly summarized in Table~\ref{lm:tab1}.

%Bigram model was trained using 150,000 sentences from ``First half'' and ``Second half'', respectively, and 100,000 sentences were bootstrapped from the remaining in each of the four groups as the test set.

As illustrated in Figure~\ref{lm:fig1_1}, the linguistic pattern of both cross-posters and non-crossposters appeared to have drifted away from the mainstream of their home subreddit, when comparing the average of the log-probability of sentences in the early stage (``Pre-crossposting'' and ``First half'') and late stage (``Post-crossposting'' and ``Second half'') of an author's lifespan. Notice that the linguistic pattern of Feminism-to-Men'sRights cross-posters has drifted to a much larger degree compared to the Feminism non-crossposters. To evaluate the statistical significance of the observed difference in Feminism as well as the counterpart in Men's Rights, the procedure of training and test set sampling and language model building was repeated a hundred times, and for each repetition the delta ($\Delta_{cp}$) between ``Post-crossposting'' and ``Pre-crossposting'', and the delta ($\Delta_{Ncp}$) between ``Second half'' and ``First half'' was computed. The average of $\Delta_{cp}$ and the average of $\Delta_{Ncp}$ over the hundred repetitions was computed (Figure ~\ref{lm:fig1_2}). We found that, for Feminism, the average of $\Delta{cp}$ (0.1149) was found to be statistically different (t-test, p-value$\simeq 0$) from that of $\Delta_{Ncp}$ (0.0182). However, for Men'sRights the means of two samples of delta (0.0325 and 0.0320) was not found to be statistically different (t-test, p-value=0.5524), which may suggest that Feminism-to-Men'sRights cross-posters \del{appeared to be more radical or aggressive}\yy{no evidence on the nature of the content} than the Men'sRights-to-Feminism cross-posters in terms of the change of linguistic patterns.

\begin{figure}[ht]
    \centering
    \includegraphics[scale = 0.4]{LM_figure_1_1}
    \caption{Change in Linguistic Pattern for Cross-posters and Non-crossposters.}
    \label{lm:fig1_1}
\end{figure}

\begin{figure}[ht]
    \centering
    \includegraphics[scale = 0.4]{LM_figure_1_2}
    \caption{\yy{not sure what's being measured here} Feminism-to-Men'sRights Cross-posters appear to be different in Linguistic Pattern.}
    \label{lm:fig1_2}
\end{figure}

In the second analysis, we investigated whether the linguistic pattern of all authors (cross-posters and non-crossposters alike) has changed over the entire lifecycle of either subreddit. The posts were sorted in increasing order of the timestamp, and sorted list were split into 50 approximately equal-sized sub-samples. A log-probability was computed for each sentence in one portion of data, using a bigram model trained on the sentences from the remaining 49 samples, and this was repeated for each of the 50 data samples. As a comparison, we considered an irrelevant subreddit ``Cooking'' on which this process was repeated. We plotted the average of the log-probability of sentences in each sample as a function of the index (1 to 50). As shown in Figure XXX~\ref{lm:fig2}, the agreement of the local and the global linguistic pattern for both subreddits reached the maximum in the first half the lifecyle, and then decreases over time. In contrast, the linguistic pattern of the subreddit ``Cooking'' remains stable. It's also interesting to note that the evolution of the linguistic pattern in Men'sRights and Feminism appear to be correlated with each other ($\rho$ = 0.6636), in contrast to their correlation with the Cooking subreddit ($\rho$ = 0.3487 and 0.0445 for Men'sRights and Feminism). Taken together, these analyses of the language patterns of reddit users show that there are important consequence of cross-group contact.

\begin{figure}[ht]
    \centering
    \includegraphics[scale=0.4]{images/LM_figure_2.png}
    \label{lm:fig2}
    \caption{The Evolution of Linguistic Pattern over the Lifecyle of Subreddits}
\end{figure}

\subsection{Discussion}

We hypothesized that the tendency to address an ideologically opposite group was associated with wordiness, network degree and exposure to others who cross-post. Our data indicate that there is support most of these hypotheses. However, in many cases, the size of the effect is  small. The exception is the variable for having contact with others who cross-post. That has an  large effect. We noted earlier that there could be multiple explanations for this effect. There may be direct contagion or it may reflect homophily. If there is contagion, then further research can investigate the reasons why. For example, one possibility is that the user may learn about the ideologically opposite reddit through contact with a cross-poster. Alternatively, cross-talk may tend to be antagonistic. The appearance of cross-talkers may incite a user to visit the ideologically opposite reddit in an attempt to continue a dispute.

It may also be the case that the spread of cross-talking or cross-posting behavior may be due to unobserved variables and not the contact with a cross-talker. For example, perhaps people with certain preferences for topics that overlap with Men's Rights and Feminism would naturally talk with each other. One issue might be the discussion of books that critique either Feminist or Men's Rights ideas from the other perspective. People who would be interested in such a debate might cross-post and comment on each other's text, even if they did not influence each other. Further research can assess whether the effects reported in this study are attributable to such unobserved variables.

The analysis of linguist drift shows that cross-talk between ideologically opposite groups might be an important component of cultural change. As individuals gain exposure to others from different ideological orientations, they may change the content of their speech. Although this does not mean that they adopt the opposite side's opinions, they may be changing in ways that shift the conversation within their own group.

\subsection{Conclusion}

The main purpose of this study is to investigate the conditions under which individuals might choose to engage with a group that holds conflicting beliefs. Our evidence shows that a mixture of individual and structural factors are associated with this behavior. Our study of linguistic drift shows that cross-talking, in this case, leads to users who are substantively different in their discourse than their colleagues in their ``home'' subreddit. This drift might be one source of cultural and social change more generally. The exposure to outside groups leads to drift that, in some cases, may lead to a broader change in beliefs. By studying when cross-talk happens and some of the consequences, our research alludes to a larger account of how social boundaries, crossing, and linguistic change play a role in social change.

%\section{Discussion}
%[Fabio and Pat]
%We use the term \textit{crossposter} to describe those authors who interact with their ideological outgroup by posting in that subreddit, in this case,  /r/Men'sRights and /r/Feminism. Our hazard curves show that interactions between these two polarized subreddits continues throughout the nearly seven years of comment data for these subreddits, although there appear to be external factors of a temporal nature which affect the number of authors who emerge to interact as shown in figure \ref{fig:bc_yoe}.

%\begin{figure}[htb]
    %\centering
    %\includegraphics[scale=0.4]{bc_yr_of_emergence}
    %\caption{Crossposter Year of Emergence}
    %\label{fig:bc_yoe}
%\end{figure}

%\fxfatal{\pat{consistent use and spelling of Crossposter / Boundary Crosser}}
%Table \ref{tab:datasum} summarizes the comment data set and shows how the two subreddits differ by size with /r/Men'sRights being significantly larger in terms of numbers of authors and comments when compared to /r/Feminsim. Figure \ref{fig:bc_ratio} depicts the normalized count of comment authors by date of emergence and author type (/r/Men'sRights only, /r/Feminism only, and crossposter).  /r/Men'sRights authors are by far the largest count, followed by /r/Feminism authors and finally crossposters.  
%While we have not studied the cause of these (it is left for future work), we have been able to attribute several features relating to the author and how they post in their home subreddit as potential indicators. One particularly interesting finding is that the first subreddit in which an author posts is a good indicator of their potential to cross boundaries.  Consider the data in figure \ref{fig:bc_tot} where /r/Feminism contributes an appreciable number of boundary crossing authors given its relatively small size when compared with /r/Men'sRights.

%\begin{figure}[ht]
%    \centering
    %\includegraphics[scale=0.4]{tot_bc_count}
    %\caption{Boundary Crossers by Subreddit}
%    \label{fig:bc_tot}
%\end{figure}

%\subsection{Association between the act of cross-posting and the drift of linguistic pattern of cross-posters}
%Also of interest is our finding that an author's language changes after interacting with the opposite subreddit.Figure~\ref{fig:LMchange} depicts the language changes before and after crossposting for authors by their principal subreddit using a bigram language model which analyzes author comments.  Notice that the \pat{need some words from Chao Ji here...}

%The raw text of reddit posts (\texttt{body} field) were preprocessed, where HTML and other markup tags were removed. Then each post was tokenized into sentences, and each sentence was represented as a Bigram sequence of tokenized words. We performed two controlled computational experiments on the reddit text data processed as above using statistical language models.

%In the first experiment, we investigated if the cross-posters' linguistic pattern in their HOME subreddit has manifested a significant change between the pre- and post-crossposting stage. As a comparison, we also considered the posts in the same subreddit that are from authors who never cross-posted. These posts were split into those that occurred in the first and second half of an author's lifespan.

%For each group, a fixed number (100,000) of sentences were bootstrapped (i.e. sampled with replacement) from the available text data (Table ~\ref{lm:tab1}), and for each sentence a log-transformed probability score was computed using Bigram language model (with Stupid backoff smoothing ~\cite{Brants07largelanguage}). The Bigram model was trained on the text data from non-crossposters in the same subreddit (disjoint from those used in the test set), which contains the most posts compared to those from cross-posters, and represents the mainstream linguistic pattern of the subreddit. The total numbers of available sentences in the four groups from which the above test and training set were sampled are briefly summarized in Table ~\ref{lm:tab1}.

%As illustrated in Figure ~\ref{lm:fig1_1}, the linguistic pattern of both cross-posters and non-crossposters appeared to have drifted away from the mainstream of their HOME subreddit, when comparing the average of the log-probability of sentences in the early stage ("Pre-crossposting" and "First half") and late stage ("Post-crossposting" and "Second half") of an author's lifespan. Notice that the linguistic pattern of Feminism-to-Men'sRights cross-posters has drifted to a much larger degree compared to the Feminism non-crossposters. To evaluate the statistical significance of the observed difference in Feminism as well as the counterpart in Men'sRights, the procedure of training and test set sampling and language model building was repeated a hundred times, and for each repetition the delta ($\Delta_{cp}$) between "Post-crossposting" and "Pre-crossposting", and the delta ($\Delta_{Ncp}$) between "Second half" and "First half" was computed. The average of $\Delta_{cp}$ and the average of $\Delta_{Ncp}$ over the hundred repetitions was computed (Figure ~\ref{lm:fig1_2}). We found that, for Feminism, the average of $\Delta{cp}$ (0.1149) was found to be statistically different (t-test, p-value=$2.5e-184$) from that of $\Delta_{Ncp}$ (0.0182). However, for Men'sRights the means of two samples of delta (0.0325 and 0.0320) was not found to be statistically different (t-test, p-value=0.5524), which may suggest that Feminism-to-Men'sRights cross-posters appeared to be more radical or aggressive than the Men'sRights-to-Feminism cross-posters in terms of the change of linguistic patterns.

%We also analyzed the change in language used in /r/Men'sRights and /f/Feminism over time when compared to a less controversial subreddit, /r/Cooking.  \pat{write findings here}

%\begin{figure}[ht]
%    \centering
%    \includegraphics[scale=0.4]{LM_figure_2}
%    \caption{Changes in SR Language Over Time}
%    \label{fig:SRlangchanges}
%\end{figure}

%In the second experiment, we investigated whether the linguistic pattern of all authors (cross-posters and non-crossposters alike) has changed over the entire lifecycle of either subreddit. The posts were sorted in increasing order of the timestamp, and sorted list were split into 50 approximately equal-sized chunks. A log-probability was computed for each sentence in one chunk of data, using a Bigram model trained on the sentences from the remaining 49 chunks, and this was repeated for each of the 50 chunks. As a comparison, we considered an irrelevant subreddit "Cooking" on which this process was repeated. We plotted the average of the log-probability of sentences in each chunk as a function of the index (1 to 50). As shown in Figure ~\ref{lm:fig2}, for Men'sRights and Feminism the agreement of the local and the global linguistic pattern reached the maximum in the first half the lifecyle, and then decreases over time. In contrast, the linguistic pattern of the subreddit "Cooking" remains stable. It's also interesting to note that the evolution of the linguistic pattern in Men'sRights and Feminism appear to be correlated with each other ($\rho$ = 0.6636), in contrast to their correlation with the Cooking subreddit ($\rho$ = 0.3487 and 0.0445 for Men'sRights and Feminism).

%\section{Conclusion}
%[Fabio]
%\%\%\%\%\%\%\%\%\%
% \textit{References and End of Paper}
\bibliographystyle{aaai}
\bibliography{refs.bib}
\end{document}

### 2020-08-28 acmart-master/samplebody-conf.tex
moved non-repo papers from overleaf into this repo

\section{Introduction}

The \textit{proceedings} are the records of a conference.\footnote{This
  is a footnote}  ACM seeks
to give these conference by-products a uniform, high-quality
appearance.  To do this, ACM has some rigid requirements for the
format of the proceedings documents: there is a specified format
(balanced double columns), a specified set of fonts (Arial or
Helvetica and Times Roman) in certain specified sizes, a specified
live area, centered on the page, specified size of margins, specified
column width and gutter size.

\section{The Body of The Paper}
Typically, the body of a paper is organized into a hierarchical
structure, with numbered or unnumbered headings for sections,
subsections, sub-subsections, and even smaller sections.  The command
\texttt{{\char'134}section} that precedes this paragraph is part of
such a hierarchy.\footnote{This is a footnote.} \LaTeX\ handles the
numbering and placement of these headings for you, when you use the
appropriate heading commands around the titles of the headings.  If
you want a sub-subsection or smaller part to be unnumbered in your
output, simply append an asterisk to the command name.  Examples of
both numbered and unnumbered headings will appear throughout the
balance of this sample document.

Because the entire article is contained in the \textbf{document}
environment, you can indicate the start of a new paragraph with a
blank line in your input file; that is why this sentence forms a
separate paragraph. \cite{steinj.2018}

\subsection{Type Changes and {\itshape Special} Characters}

We have already seen several typeface changes in this sample.  You can
indicate italicized words or phrases in your text with the command
\texttt{{\char'134}textit}; emboldening with the command
\texttt{{\char'134}textbf} and typewriter-style (for instance, for
computer code) with \texttt{{\char'134}texttt}.  But remember, you do
not have to indicate typestyle changes when such changes are part of
the \textit{structural} elements of your article; for instance, the
heading of this subsection will be in a sans serif\footnote{Another
  footnote here.  Let's make this a rather long one to see how it
  looks.} typeface, but that is handled by the document class file.
Take care with the use of\footnote{Another footnote.}  the
curly braces in typeface changes; they mark the beginning and end of
the text that is to be in the different typeface.

You can use whatever symbols, accented characters, or non-English
characters you need anywhere in your document; you can find a complete
list of what is available in the \textit{\LaTeX\ User's Guide}
\cite{Lamport:LaTeX}.

\subsection{Math Equations}
You may want to display math equations in three distinct styles:
inline, numbered or non-numbered display.  Each of
the three are discussed in the next sections.

\subsubsection{Inline (In-text) Equations}
A formula that appears in the running text is called an
inline or in-text formula.  It is produced by the
\textbf{math} environment, which can be
invoked with the usual \texttt{{\char'134}begin\,\ldots{\char'134}end}
construction or with the short form \texttt{\$\,\ldots\$}. You
can use any of the symbols and structures,
from $\alpha$ to $\omega$, available in
\LaTeX~\cite{Lamport:LaTeX}; this section will simply show a
few examples of in-text equations in context. Notice how
this equation:
\begin{math}
  \lim_{n\rightarrow \infty}x=0
\end{math},
set here in in-line math style, looks slightly different when
set in display style.  (See next section).

\subsubsection{Display Equations}
A numbered display equation---one set off by vertical space from the
text and centered horizontally---is produced by the \textbf{equation}
environment. An unnumbered display equation is produced by the
\textbf{displaymath} environment.

Again, in either environment, you can use any of the symbols
and structures available in \LaTeX\@; this section will just
give a couple of examples of display equations in context.
First, consider the equation, shown as an inline equation above:
\begin{equation}
  \lim_{n\rightarrow \infty}x=0
\end{equation}
Notice how it is formatted somewhat differently in
the \textbf{displaymath}
environment.  Now, we'll enter an unnumbered equation:
\begin{displaymath}
  \sum_{i=0}^{\infty} x + 1
\end{displaymath}
and follow it with another numbered equation:
\begin{equation}
  \sum_{i=0}^{\infty}x_i=\int_{0}^{\pi+2} f
\end{equation}
just to demonstrate \LaTeX's able handling of numbering.

\subsection{Citations}
Citations to articles~\cite{bowman:reasoning,
clark:pct, braams:babel, herlihy:methodology},
conference proceedings~\cite{clark:pct} or maybe
books \cite{Lamport:LaTeX, salas:calculus} listed
in the Bibliography section of your
article will occur throughout the text of your article.
You should use BibTeX to automatically produce this bibliography;
you simply need to insert one of several citation commands with
a key of the item cited in the proper location in
the \texttt{.tex} file~\cite{Lamport:LaTeX}.
The key is a short reference you invent to uniquely
identify each work; in this sample document, the key is
the first author's surname and a
word from the title.  This identifying key is included
with each item in the \texttt{.bib} file for your article.

The details of the construction of the \texttt{.bib} file
are beyond the scope of this sample document, but more
information can be found in the \textit{Author's Guide},
and exhaustive details in the \textit{\LaTeX\ User's
Guide} by Lamport~\shortcite{Lamport:LaTeX}.

This article shows only the plainest form
of the citation command, using \texttt{{\char'134}cite}.

Some examples.  A paginated journal article \cite{Abril07}, an enumerated
journal article \cite{Cohen07}, a reference to an entire issue \cite{JCohen96},
a monograph (whole book) \cite{Kosiur01}, a monograph/whole book in a series (see 2a in spec. document)
\cite{Harel79}, a divisible-book such as an anthology or compilation \cite{Editor00}
followed by the same example, however we only output the series if the volume number is given
\cite{Editor00a} (so Editor00a's series should NOT be present since it has no vol. no.),
a chapter in a divisible book \cite{Spector90}, a chapter in a divisible book
in a series \cite{Douglass98}, a multi-volume work as book \cite{Knuth97},
an article in a proceedings (of a conference, symposium, workshop for example)
(paginated proceedings article) \cite{Andler79}, a proceedings article
with all possible elements \cite{Smith10}, an example of an enumerated
proceedings article \cite{VanGundy07},
an informally published work \cite{Harel78}, a doctoral dissertation \cite{Clarkson85},
a master's thesis: \cite{anisi03}, an online document / world wide web
resource \cite{Thornburg01, Ablamowicz07, Poker06}, a video game (Case 1) \cite{Obama08} and (Case 2) \cite{Novak03}
and \cite{Lee05} and (Case 3) a patent \cite{JoeScientist001},
work accepted for publication \cite{rous08}, 'YYYYb'-test for prolific author
\cite{SaeediMEJ10} and \cite{SaeediJETC10}. Other cites might contain
'duplicate' DOI and URLs (some SIAM articles) \cite{Kirschmer:2010:AEI:1958016.1958018}.
Boris / Barbara Beeton: multi-volume works as books
\cite{MR781536} and \cite{MR781537}.

A couple of citations with DOIs: \cite{2004:ITE:1009386.1010128,
  Kirschmer:2010:AEI:1958016.1958018}.

Online citations: \cite{TUGInstmem, Thornburg01, CTANacmart}.

\subsection{Tables}
Because tables cannot be split across pages, the best
placement for them is typically the top of the page
nearest their initial cite.  To
ensure this proper ``floating'' placement of tables, use the
environment \textbf{table} to enclose the table's contents and
the table caption.  The contents of the table itself must go
in the \textbf{tabular} environment, to
be aligned properly in rows and columns, with the desired
horizontal and vertical rules.  Again, detailed instructions
on \textbf{tabular} material
are found in the \textit{\LaTeX\ User's Guide}.

Immediately following this sentence is the point at which
Table~\ref{tab:freq} is included in the input file; compare the
placement of the table here with the table in the printed
output of this document.

\begin{table}
  \caption{Frequency of Special Characters}
  \label{tab:freq}
  \begin{tabular}{ccl}
    \toprule
    Non-English or Math&Frequency&Comments\\
    \midrule
    \O & 1 in 1,000& For Swedish names\\
    $\pi$ & 1 in 5& Common in math\\
    \$ & 4 in 5 & Used in business\\
    $\Psi^2_1$ & 1 in 40,000& Unexplained usage\\
  \bottomrule
\end{tabular}
\end{table}

To set a wider table, which takes up the whole width of the page's
live area, use the environment \textbf{table*} to enclose the table's
contents and the table caption.  As with a single-column table, this
wide table will ``float'' to a location deemed more desirable.
Immediately following this sentence is the point at which
Table~\ref{tab:commands} is included in the input file; again, it is
instructive to compare the placement of the table here with the table
in the printed output of this document.

\begin{table*}
  \caption{Some Typical Commands}
  \label{tab:commands}
  \begin{tabular}{ccl}
    \toprule
    Command &A Number & Comments\\
    \midrule
    \texttt{{\char'134}author} & 100& Author \\
    \texttt{{\char'134}table}& 300 & For tables\\
    \texttt{{\char'134}table*}& 400& For wider tables\\
    \bottomrule
  \end{tabular}
\end{table*}
% end the environment with {table*}, NOTE not {table}!

It is strongly recommended to use the package booktabs~\cite{Fear05}
and follow its main principles of typography with respect to tables:
\begin{enumerate}
\item Never, ever use vertical rules.
\item Never use double rules.
\end{enumerate}
It is also a good idea not to overuse horizontal rules.

\subsection{Figures}

Like tables, figures cannot be split across pages; the best placement
for them is typically the top or the bottom of the page nearest their
initial cite.  To ensure this proper ``floating'' placement of
figures, use the environment \textbf{figure} to enclose the figure and
its caption.

This sample document contains examples of \texttt{.eps} files to be
displayable with \LaTeX.  If you work with pdf\LaTeX, use files in the
\texttt{.pdf} format.  Note that most modern \TeX\ systems will convert
\texttt{.eps} to \texttt{.pdf} for you on the fly.  More details on
each of these are found in the \textit{Author's Guide}.

\begin{figure}
\includegraphics{fly}
\caption{A sample black and white graphic.}
\end{figure}

\begin{figure}
\includegraphics[height=1in, width=1in]{fly}
\caption{A sample black and white graphic
that has been resized with the \texttt{includegraphics} command.}
\end{figure}

As was the case with tables, you may want a figure that spans two
columns.  To do this, and still to ensure proper ``floating''
placement of tables, use the environment \textbf{figure*} to enclose
the figure and its caption.  And don't forget to end the environment
with \textbf{figure*}, not \textbf{figure}!

\begin{figure*}
\includegraphics{flies}
\caption{A sample black and white graphic
that needs to span two columns of text.}
\end{figure*}

\begin{figure}
\includegraphics[height=1in, width=1in]{rosette}
\caption{A sample black and white graphic that has
been resized with the \texttt{includegraphics} command.}
\end{figure}

\subsection{Theorem-like Constructs}

Other common constructs that may occur in your article are the forms
for logical constructs like theorems, axioms, corollaries and proofs.
ACM uses two types of these constructs:  theorem-like and
definition-like.

Here is a theorem:
\begin{theorem}
  Let $f$ be continuous on $[a,b]$.  If $G$ is
  an antiderivative for $f$ on $[a,b]$, then
  \begin{displaymath}
    \int^b_af(t)\,dt = G(b) - G(a).
  \end{displaymath}
\end{theorem}

Here is a definition:
\begin{definition}
  If $z$ is irrational, then by $e^z$ we mean the
  unique number that has
  logarithm $z$:
  \begin{displaymath}
    \log e^z = z.
  \end{displaymath}
\end{definition}

The pre-defined theorem-like constructs are \textbf{theorem},
\textbf{conjecture}, \textbf{proposition}, \textbf{lemma} and
\textbf{corollary}.  The pre-defined de\-fi\-ni\-ti\-on-like constructs are
\textbf{example} and \textbf{definition}.  You can add your own
constructs using the \textsl{amsthm} interface~\cite{Amsthm15}.  The
styles used in the \verb|\theoremstyle| command are \textbf{acmplain}
and \textbf{acmdefinition}.

Another construct is \textbf{proof}, for example,

\begin{proof}
  Suppose on the contrary there exists a real number $L$ such that
  \begin{displaymath}
    \lim_{x\rightarrow\infty} \frac{f(x)}{g(x)} = L.
  \end{displaymath}
  Then
  \begin{displaymath}
    l=\lim_{x\rightarrow c} f(x)
    = \lim_{x\rightarrow c}
    \left[ g{x} \cdot \frac{f(x)}{g(x)} \right ]
    = \lim_{x\rightarrow c} g(x) \cdot \lim_{x\rightarrow c}
    \frac{f(x)}{g(x)} = 0\cdot L = 0,
  \end{displaymath}
  which contradicts our assumption that $l\neq 0$.
\end{proof}

\section{Conclusions}
This paragraph will end the body of this sample document.
Remember that you might still have Acknowledgments or
Appendices; brief samples of these
follow.  There is still the Bibliography to deal with; and
we will make a disclaimer about that here: with the exception
of the reference to the \LaTeX\ book, the citations in
this paper are to articles which have nothing to
do with the present subject and are used as
examples only.
%\end{document}  % This is where a 'short' article might terminate

\appendix
%Appendix A
\section{Headings in Appendices}
The rules about hierarchical headings discussed above for
the body of the article are different in the appendices.
In the \textbf{appendix} environment, the command
\textbf{section} is used to
indicate the start of each Appendix, with alphabetic order
designation (i.e., the first is A, the second B, etc.) and
a title (if you include one).  So, if you need
hierarchical structure
\textit{within} an Appendix, start with \textbf{subsection} as the
highest level. Here is an outline of the body of this
document in Appendix-appropriate form:
\subsection{Introduction}
\subsection{The Body of the Paper}
\subsubsection{Type Changes and  Special Characters}
\subsubsection{Math Equations}
\paragraph{Inline (In-text) Equations}
\paragraph{Display Equations}
\subsubsection{Citations}
\subsubsection{Tables}
\subsubsection{Figures}
\subsubsection{Theorem-like Constructs}
\subsubsection*{A Caveat for the \TeX\ Expert}
\subsection{Conclusions}
\subsection{References}
Generated by bibtex from your \texttt{.bib} file.  Run latex,
then bibtex, then latex twice (to resolve references)
to create the \texttt{.bbl} file.  Insert that \texttt{.bbl}
file into the \texttt{.tex} source file and comment out
the command \texttt{{\char'134}thebibliography}.
% This next section command marks the start of
% Appendix B, and does not continue the present hierarchy
\section{More Help for the Hardy}

Of course, reading the source code is always useful.  The file
\path{acmart.pdf} contains both the user guide and the commented
code.

\begin{acks}
  The authors would like to thank Dr. Yuhua Li for providing the
  MATLAB code of the \textit{BEPS} method.

  The authors would also like to thank the anonymous referees for
  their valuable comments and helpful suggestions. The work is
  supported by the \grantsponsor{GS501100001809}{National Natural
    Science Foundation of
    China}{http://dx.doi.org/10.13039/501100001809} under Grant
  No.:~\grantnum{GS501100001809}{61273304}
  and~\grantnum[http://www.nnsf.cn/youngscientists]{GS501100001809}{Young
    Scientists' Support Program}.

\end{acks}

### 2020-08-28 sandbox/main.tex
moved non-repo papers from overleaf into this repo

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath}

\title{sandbox}
\author{johndstein }
\date{September 2016}

\usepackage{natbib}
\usepackage{graphicx}

\begin{document}

\maketitle

\section{Introduction}
There is a theory which states that if ever anyone discovers exactly what the Universe is for and why it is here, it will instantly disappear and be replaced by something even more bizarre and inexplicable.
There is another theory which states that this has already happened.

\begin{equation} A = B \cap C \end{equation}

\begin{figure}[h!]
\centering
\includegraphics[scale=1.7]{universe.jpg}
\caption{The Universe}
\label{fig:univerise}
\end{figure}

\section{Conclusion}
``I always thought something was fundamentally wrong with the universe'' \citep{adams1995hitchhiker}

\bibliographystyle{plain}
\bibliography{references}
\end{document}

### 2020-08-28 secMod/main.tex
moved files into subdir

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}

\title{TBD Title}
\author{John Stein (jodstein@iu.edu)}
\date{TBD}

\usepackage{natbib}
\usepackage{graphicx}
\usepackage{setspace}\singlespacing
\usepackage[left=1in, right=1in, top=1in, bottom=1in]{geometry}
\usepackage{appendix}
\usepackage{amsmath}
\DeclareMathOperator*{\argmax}{arg\,max}
\usepackage{tikz}
\usetikzlibrary{positioning, shapes, arrows}
\usepackage[ruled]{algorithm2e}
\usepackage{amssymb}

\begin{document}

\begin{titlepage}
    % https://stackoverflow.com/questions/3141702/vertically-centering-a-title-page
    \null
    \nointerlineskip
    \vfill
    \let\snewpage \newpage
    \let\newpage \relax
    \maketitle
    \let \newpage \snewpage
    \vfill
    \thispagestyle{empty}
\end{titlepage}

\tableofcontents
\thispagestyle{empty}

%%%%%%%%%%%%%%%%% Paper %%%%%%%%%%%%%%%%%%%%%%
\newpage
\setcounter{page}{1}

\begin{abstract}
    As system complexity increases to meet growing expectations of almost every market in our networked world, traditional methodologies of System Security Engineering (SSE, inclusive of Cybersecurity, Information Assurance, and Anti-Tamper) are becoming untenable.  SSE, as a discipline, is not unique in this regard.  Indeed, Model Based System Engineering represents a rapidly growing capability within the field of engineering to address precisely this need.  However, SSE can be regarded as an extreme exemplar of this phenomena.  This is because with increasing complexity comes a larger attack surface (i.e. a greater number of potentially vulnerable element instantiations within the system which an attacker may use as entry or pivot points to achieve his/her objective(s)).  Since the attacker need only find one or few such vulnerable elements, growing system complexity very much favors the attacker over the developer.  As such, there is a need for capabilities which leverage computational analysis techniques to address complexity (something humans do not perform well) in order to generate a comprehensive-but-minimal data representation that is informative, relevant, and actionable to a human SSE practitioner to enable a holistic risk management approach (something that machines do not perform well).
\end{abstract}

\section{TBD}

\section{Problem Formulation}

\subsection{Knowledge Representation}

Ontologies:

\begin{itemize}

    \item Triplets (subject, predicate, object) \cite{jenkins2010ontologies}. Extensibility, Top-down and bottom-up.

    \item Basic Formal Ontology \cite{iso/iec_dis_21838-2}

    \item Reference Ontologies - Reference Ontology for SE (Top-Level Ontology (TLO) $\rightarrow$ Reference Ontology $\rightarrow$ Application Ontology).  Building out an ontology for Systems Engineering based on ISO/IEC/IEEE 15288. \cite{orellana2019ontology}

    \item Application Ontologies

\end{itemize}

Explainability. I think the thought here is to discover relationships between attacks, targets, goals, and context.  \textcolor{red}{I need to ping Apu and make sure I understand his thoughts about how explainability applies to this area.}:

\begin{itemize}

    \item Leino\/Sen\/Datta propose approach to explain CNNs based on a particular layer's $s$ influence over a particular measure $f$ with respect to a given distribution of interest $P$, which can be an instance, a class, the difference between two classes, or an entire dataset.  This approach is interesting because it appears to be very powerful and highly tunable to answer granular and specific questions. \cite{leino2018influence}

\end{itemize}

\subsection{Mapping Data to Ontologies}

\subsection{Reasoning over Models}

\begin{itemize}

    \item Berger - introduces Extended DFDs, maps these onto graphs, performs rule checking against ‘knowledge base’ built manually from CWE and CAPEC; uses Microsoft Threat Modeling Tool \cite{berger2016automatically}

    \item Audinot - Defines regular grammar for attack trees, operations on that grammar, and automata for those operations.  Trace semantics are implemented to determine (atk) goal reachability and, if quantitative info is available, optimal reachability (priced, timed, weighted, etc.).  With the optimal paths, certain positions (atks) are considered ‘useful’ and require further refinement. \cite{audinot2018guided}

\end{itemize}

\subsection{Security Analysis Augmentation}

Natural Language Processing:

\begin{itemize}

    \item learning to recognize program dependency graphs (PDG) for secure and insecure use case patterns, and then implements nudge factors to help the (Stack Overflow) user choose secure examples \cite{fischer2019stack}

    \item edge2vec can learn heterogeneous node embeddings (nodes/edges can be typed, and graphs very similar to text) \cite{gao2018edge2vec}.  This could be very useful in prediction tasks (augmentation, text completion, pattern matching).

\end{itemize}

\section{Related Work}

\begin{itemize}

    \item Jurjens - Security preservation semantics for Class/structure diagrams, statechart diagrams, and interaction/swimlane diagrams using UML.  Security is co-mingled with design. \cite{jurjens2001towards}

    \item Apvrille - further formalizes the application of SysML-sec to embedded hardware, uses EVITA vehicle architectural example, and demonstrates formal verification ability for a key distribution protocol \cite{apvrille2013sysml}

    \item Apvrille - focus on integration of reqts and attack analysis against a system design effort using the Y-chart methodology (mapping of app models onto arch models) to achieve formal V\&V using SysML (TTool).  Security is on separate diagrams from design/structure. \cite{roudier2015sysml}

\end{itemize}

%%%%%%%%%%%%%%%%% BIB %%%%%%%%%%%%%%%%%%%%
\newpage
\addcontentsline{toc}{section}{References}
\bibliographystyle{plain}
\bibliography{references}

%%%%%%%%%%%%%%%%% Appendices %%%%%%%%%%%%%%%%%%%%

\appendixtitleon
\begin{appendices}

%%%%%%%%%% Appendix A %%%%%%%%%%

\clearpage
\begin{center}
    \section{Title}
\end{center}

\end{appendices}

\end{document}

### 2020-08-28 iufa17dmproject/article_3.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Stylish Article
% LaTeX Template
% Version 2.1 (1/10/15)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% Original author:
% Mathias Legrand (legrand.mathias@gmail.com) 
% With extensive modifications by:
% Vel (vel@latextemplates.com)
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[fleqn,10pt]{SelfArx} % Document font size and equations flushed left

\usepackage[english]{babel} % Specify a different language here - english by default
\usepackage{lipsum} % Required to insert dummy text. To be removed otherwise
\setlength{\columnsep}{0.55cm} % Distance between the two columns of text
\setlength{\fboxrule}{0.75pt} % Width of the border around the abstract
\definecolor{color1}{RGB}{0,0,90} % Color of the article title and sections
\definecolor{color2}{RGB}{0,20,20} % Color of the boxes behind the abstract and headings
\usepackage{hyperref} % Required for hyperlinks
\hypersetup{hidelinks,colorlinks,breaklinks=true,urlcolor=color2,citecolor=color1,linkcolor=color1,bookmarksopen=false,pdftitle={Title},pdfauthor={Author}}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{multirow}
\usepackage{amsmath}
\usepackage{amssymb}
\def\Plus{\texttt{+}}	
\def\Minus{\texttt{-}} 
\usepackage{amsfonts}
\usepackage{alltt}
\usepackage{url}
\usepackage{relsize}
\usepackage{adjustbox}
\usepackage{graphicx}

%----------------------------------------------------------------------------------------
%	ARTICLE INFORMATION
%----------------------------------------------------------------------------------------

\JournalInfo{Computer Science\\School of Informatics, Computing, and Engineering\\Indiana University, Bloomington, IN, USA} % Journal information
\Archive{Datamining B565 Fall 2017}  % Additional notes (e.g. copyright, DOI, review/research article)

\PaperTitle{Kaggle Competitions:\\ \ Spooky Author Identification\\ \ Statoil/C-CORE Iceberg Classifier Challenge} % Article title

\Authors{Ashay Sawant\textsuperscript{1}*, John Stein\textsuperscript{2}*, Samanvitha Pradhan\textsuperscript{3}*} % Authors
\affiliation{\textsuperscript{1}\textit{Data Science, School of Informatics, Computing, and Engineering, Indiana University, Bloomington, IN, USA}} % Author affiliation
\affiliation{*\textbf{Corresponding author}: ahsawant@iu.edu} % Corresponding author

\affiliation{\textsuperscript{2}\textit{Computer Science, School of Informatics, Computing, and Engineering, Indiana University, Bloomington, IN, USA}} % Author affiliation
\affiliation{*\textbf{Corresponding author}: jodstein@iu.edu} % Corresponding author

\affiliation{\textsuperscript{3}\textit{Data  Science, School of Informatics, Computing, and Engineering, Indiana University, Bloomington, IN, USA}} % Author affiliation
\affiliation{*\textbf{Corresponding author}: sspradha@iu.edu} % Corresponding author

\Keywords{NLP --- Natural Language Processing --- SAR --- Image Recognition --- HOG --- Histogram of Oriented Gradients --- Logistic Regression --- Random Forest --- XGBoost} % Keywords - if you don't want any simply remove all the text between the curly brackets
\newcommand{\keywordname}{Keywords} % Defines the keywords heading name
\usepackage{abstract}

%----------------------------------------------------------------------------------------
%	Executive Summary
%----------------------------------------------------------------------------------------

\Abstract{\textbf{Executive Summary} 
	This report summarizes two datamining projects undertaken by the authors. The first project attempts to classify `scattered pages' from works of fiction written by Edgar Allan Poe, HP Lovecraft, and Mary Wollstonecraft Shelley according to their author.  Various approaches were explored, including Naive Bayes, Decision Trees, Maximun Entropy. The best-performing approach was achieved by Maximum Entropy. The second project attempts to classify Synthetic Aperture Radar images according to whether the image has captured an iceberg or not.  We have used two approaches including Convolution Neural Networks and Histogram of Oriented Gradients to process the image in a way to obtain custom features. Finally, we performed regression on custom features using Logistic Regression, Random Forest and XGBoost Algorithms. The best-performing approach was achieved by Logistic Regression using custom features, with a Log Loss of 0.3682.
}

\begin{document}

\renewcommand{\abstractname}{}  
\renewcommand{\absnamepos}{}

\flushbottom % Makes all text pages the same height
\maketitle % Print the title and abstract box
\tableofcontents % Print the contents section
\thispagestyle{empty} % Removes page numbering from the first page

%----------------------------------------------------------------------------------------
%	ARTICLE CONTENTS
%----------------------------------------------------------------------------------------

\section{Introduction}
The two projects presented herein (Spooky Author Identification and Iceberg Classifier Challenge) are the results of the combined efforts of the authors to meet the corresponding challenges published by Kaggle. \cite{kaggle-author} \cite{kaggle-ice}  Kaggle.com is a site dedicated to the Data Science community where entities can post problems and challenges to be taken on by Data Science academics and enthusiasts.  Readers can find more details on each of these challenges at Kaggle's website \footnote{\url{https://www.kaggle.com/}} in the Competitions section.

Data Mining (DM) has three broad areas of application: answering questions, exploring and discovering relationships, and annotation.  Kaggle competitions, including those described in this paper, are a convenient way for an entity with a requirement in one of these application spaces to draw from a large community of willing problem solvers to find effective and efficient solutions to satisfy their requirement.  Spooky Author Identification is considered a `Playground Competition', meaning the underlying motivation is to encourage increased participation in the Kaggle community and potentially spur innovative advances in to standing challenges in Natural Language Processing.  In the case of Iceberg Classification, the underlying requirement is for Statoil to accurately and efficiently identify icebergs that could threaten their sea-based natural resources operations.  In both cases, DM presents an opportunity to discover useful relationships between written word and their author or between SAR imagery and the target being imaged, to enable classification accuracy and efficiency well beyond human performance.

\subsection{Author Identification}
The Author Identification problem is to accurately classify whether a given statement of text (an excerpt) is from one of the Horror(Spooky) Authors namely, Edgar Allen Poe, H P Lovecraft, and Mary Wollstonecraft Craft. The data has three classes as “EAP”, “HPL”, and “MWS”. The training set has three columns, one identifier, one column of text, and one column as the class label with one of the three classes. The excerpts are anywhere from a quote to a description. The classifier must calculate the probability that and map the text to one of the authors. The test data set has an identifier and a text per row and the trained classifier must classify this set. Kaggle evaluates the goodness of the classifier using the below stated Log Loss function. A good classifier must have a low log loss.
$$LL = - \frac{1}{N} \sum_{i} \sum_{j} y_ij log(p_ij)$$

\subsection{Statoi/C-CORE Iceberg Classifier Challenge}

The Iceberg Classifier Challenge is to accurately classify whether a Synthetic Aperture Radar (SAR) image contains an iceberg or something that is not an iceberg, which makes it a binary classification problem.  The data provided includes an $75 \times 75$ image that represents horizontally polarized backscatter intensity (\texttt{band\_1}), a $75 \times 75$ image that represents vertically polarized backscatter intensity (\texttt{band\_2}), and the angle of incidence between Earth-normal and the transmit beam (\texttt{inc\_angle}).  Training data also contains the class label (\texttt{is\_iceberg}).  The classifier is to calculate the probability that the image contains an iceberg (1) or not (0).  Kaggle evaluates the goodness of the classifier using Log Loss, shown below.

$$LL = - \frac{1}{N} \sum_{i=1}^N t_i log(y(x_i)) + (1-t_i) log(1-y(x_i))$$

The best classifiers will have Log Loss scores near zero. Receiver Operator Characteristics (ROC) was also used to evaluate the True Positive vs False Positive behavior of the classifier.  High-performing classifier will have ROC curves that resemble a step function and with an area under the curve (AUC) near 1.0.  \footnote{ROC finds its origins in World War 2 as radar operators were evaluated on their ability to distinguish real targets from false targets and clutter. \cite{Lusted1217}}

\section{Datamining}

\begin{itemize}[noitemsep]

\item What is DM?\\
DM seeks to inform decision making by answering questions about, exploring or discovering relationships in, or providing annotation of data.

\item What does it yield?\\
  DM yields information that is not possible or practical to directly observe, based on data that is practical to observe.  This information may be predicted trends in human behavior, perceived relationships or clusters within high-dimension data, or a prediction of future events.  In all cases, the yielded information is deemed valuable to someone who is then enabled to make decisions based on that information.

\item What are the general steps?

\begin{itemize}

	\item The first (and most difficult) step in DM is developing the problem statement: What decision or action is the DM effort required to inform or enable, and what must the DM effort yield in order to satisfy that requirement?

	\item The second step is data acquisition.  Acquisition of data is non-trivial.  It can be difficult, expensive, and in some cases even prohibited or restricted by law.  Therefore, the miner must be intentional about pursuing the acquisition of data that is both needed (from the problem statement) and available, whether by manual observation, survey, or automated sensing/recording.

	\item The third (and most time consuming) step in data-mining is data pre-processing.  The data must be analyzed, including identifying unknown, missing, or outlier records, examining attributes and their domains, and visualizing the data for first-order behaviors, trends, and distributions.  The data must then be integrated and/or cleaned, including dealing with unknown, missing, or bad data, enriching the data from other sources, and/or transforming the data to work better in the algorithm.

	\item The fourth step is to actually mine the data (i.e. answer questions, explore or discover relationships, or annotate).  The methodology, assumptions, models, and parameters of how the mining is performed can depend on the objective, the data itself, and the miner themselves.

	\item The fifth and final step in DM is to interpret the output of the mining step and either validate that it meets the stated objective/requirement of the problem statement, or determine that changes to the pre-processing or mining steps are warranted and start again with a modified approach.

\end{itemize}

\item What is clustering \textit{vs.} classification?\\
  Classification refers to the task of predicting the label of some unobserved target attribute given a set of observed attributes for some record.  The label is generally a single member of a finite list of possible labels.  Clustering refers to the task of identifying or discovering relationships among data.  In other words, are there properties or measures for which subsets of data can be interpreted as similar or dissimilar in a meaningful way?  Clustering aims to discover these `clumps' or similar data as well as the properties or measurements from which the similarity can be determined.  It is useful to think of Classification as a supervised learning task because it is notionally possible to produce a training set with labeled records, and Clustering as an unsupervised learning task because the number and meaning of the discovered clusters is not known apriori (if they were, they could be labeled and it would be called Classification).

\item What is a loss function?\\
 A loss function is a function which attempts to assign a real value to some undesirable property of an intermediate outcome during the mining step.  This value (known as cost or loss) may represent error, variance, or bias of a model, clustering, or classifier.  The actual value of loss is usually not meaningful.  Instead, the goal is generally to minimize the loss over some parameter or variable during an optimization step.

\end{itemize}

\subsection{Data Preprossing}

\begin{itemize}[noitemsep]

\item What are the steps, and what challenges does each present?\\
  The first step in data preprocessing is data analysis.  Data analysis can include examining histograms, bar/pair-plots, frequency tables, etc.  The goal is to get familiar with the data attributes and their domains, frequency of values, trends and distributions, and first-order behaviors.  From the analysis, the two primary remaining tasks are selection and modification.  Selection (i.e. sampling and attribute selection) refers to deciding which subset or records and/or which subset of attributes for each record to use for the mining task.  Modification (i.e. aggregation, dimensionality reduction, discretization, and transformation) refers to changing existing, creating new, or transforming, attributes and their values.

\begin{itemize}

	\item Sampling refers to choosing a subset of the total data from which to develop the analysis technique - often because working on the entire dataset would result in more resource investment than it would improve the performance of the technique.  The challenge with sampling is that, especially for high-dimensionality data, the data data can quickly become sparse, resulting in a greater potential for over-fitting due to the small data size and high dimensionality (degrees of freedom).

	\item Attribute Selection refers to selecting which attributes to include in the analysis and which to exclude, based on some preliminary analysis.  Manual attribute selection (or de-selection) is challenging because humans can only perceive date in two, three, or perhaps four dimensions.  One can only reasonably expect to act on the most obvious of attribute properties without impacting the analysis in unexpected ways.

	\item Aggregation refers to manually grouping several attributes or attribute values together, preferably while preserving some higher-level relationship among those being grouped.  Aggregation can be relatively straightforward in some cases, but in other cases it can be very difficult to group attributes or values in a way that recognizes the essential common property of its members.

	\item Dimensionality Reduction refers to the creation of fewer new attributes which are combinations of the many old attributes.  Two common methods include Principal Component Analysis (PCA) and Neural Networks.  The primary challenge with dimensionality reduction is that the meaning of the original attributes are all but lost as the data are mapped to the new, reduced feature space.

	\item Discretization refers to the assignment of values from a continous-valued attributes into one of finite-many bins.

	\item Transformation refers to the mapping of old attribute values to new attribute values - generally reversible.
\end{itemize}
Finally, data with unknown, erroneous, or missing values must be handled.  Some of these handling techniques may be implemented in the mining algorithm(s) themselves, but some may be performed prior to the mining step.  Unknown/missing data handling is more challenging for some algorithms (i.e. regression) than it is for others (i.e. decision trees).  Some algorithms need each record to have good data, or they will simply not work.  Other algorithms can be made to handle unknown/missing data in a reliable way and achieve good performance.  Still other algorithms can operate on records having missing data with no extra effort.  The primary challenge, especially for high-dimensionality data, is that the impact of the handling technique can be different for every attribute, and must be understood and chosen carefully.

\item What is the general load (time, space, \$) for preprocessing?\\
  Computationally, preprocessing may consume $O(n)$ or $O(n log m)$ time where $n$ is the number of records to process and $m$ may be search steps if matching, inclusion, or sorting operations are required.  Preprocessing may consume anywhere from constant space to $O(n)$ space, if the records are processed one at a time and stored back into their original location, or copies are made.  Practically speaking, preprocessing can consume upwards of 80\% of the labor hours, and therefore ~80\% of the funding, of the overall DM effort.  That is because it is the most human-intensive step of the process.  Automation of the preprocessing step, although possible, is of limited benefit if the DM problems vary significantly in size, scope, data representation, purpose, etc.  The ability to predict the preprocessing impacts of data across problem spaces is indeed an analytics problem unto itself.

\end{itemize}

\subsection{Mining, Interpretation, and Action}

\paragraph{Top Ten Algorithms in DM}
Below is a brief overview of the top 10 algorithms in DM. \cite{Wu:2007:TAD:1327434.1327436}

\begin{itemize}
	\item{\textbf{C4.5:}}\\ C4.5 is an algorithm that can be used for decisions trees or rulesets. For trees, tt essentially builds a decision tree by recursively splitting the data on attributes until the data is fully explained by the tree.  It then prunes the tree by calculating a binomial error estimate at each of the leaves, and comparing that with the error that would be introduced by collapsing to the parent node, and repeating until complete.  For rulesets, it similarly generates rules representing every complete path through the notional unpruned tree (i.e. the rules generated fully explain the data).  Then, C4.5 discards conditions in the generated rules to minimize the binomial error estimate using hill-climing techniques.  Finally, in both cases, a default rule is created in case data do not satisfy any of the remaining rules.

	\item{\textbf{K-Means:}}\\ K-Means naively assigns K centroids to the space shared by the data, and then assigns each data point to the centroid that is closest (according to some distance function).  Next, the algorithm computes new locations for the existing centroids based on that data points that are assigned to them.  Then the data points are reassigned, if they are now closer to a different centroid.  The process continues until convergence, which is guaranteed, and the centroids are taken to be the resulting clusters.
	\item{\textbf{Support Vector Machines:}} SVMs attempt to find some hyperplan that fully seperates the data into their proper classes, so that the hyperplane can be used to estimate classes for future data.  It chooses the hyperplane with the maximum distance from the nearest data.  Many times the data is not linearly separable.  Two primary approaches have been adopted in such cases: (a) the introduction of a slack variable can be used to assign a penalty to any hyperplane that incorrectly classifies some of the training data, and the task becomes to minimize that penalty, and (b) the data can be kernelized, or mapped onto a new space, which is linearly separable using traditional SVM. 
	\item{\textbf{Apriori:}} The Apriori algorithm was developed to efficiently find frequent sets.  It relies on the truth that if an item is not frequent in some set, then any superset containing that item is also not frequent in that set.  Starting with singletons, candidate frequent subsets are generated from frequent subsets of lower cardinality, and then eliminated, thereby eliminating the need to check for frequency on subsets that could not be frequent.
	\item{\textbf{Expectation Maximization:}}\\ EM is a more general case of K-Means.  For any probabilistic model, the parameters for that model are initialized to some starting value, and then the model is then used to cluster the training data.  In many cases, the clustering is performed via soft-assignment, meaning each data point is assigned a probability of belonging to each cluster.  Based on the new clustering, parameters for the model are recalculated to best fit the data assigned to each.  Then the data are reassigned, and the process repeats until convergence (to some threshold) is achieved.
	\item{\textbf{Page Rank:}}\\ Page Rank is intended to find relative importance of hub nodes in a complex network, commonly applied to web pages on the Internet.  Essentially each page is assigned a rank value that is the sum of all contributions of pages that link to it.  Each incoming contribution is the contributing page's rank value divided by the number of outgoing links it has.  In this way, the highest page ranks are achieved when a page has many links to it that come from pages with high rank and few outgoing links.
	\item{\textbf{Adaboost:}}\\ Adaboost is a type of ensemble learning algorithm that uses voting to create strong learners from multiple weak learners.  Essentially, the algorithm picks a single (best) learner by which to classify training (labled) data.  It then assigns a weight to each training sample, lower for those that were correctly classified and higher for those that were not.  It then selects a new (best) feature and similarly redistributes the weight to misclassified examples.  After the last weak learner is used, each learner is then weighted according to the weighted error they would have using the final weights on the misclassified examples in each learner.  The final weighted weak learners represents the new stronger ensemble learner.
	\item{\textbf{K Nearest Neighbors:}}\\ KNN is a relatively simple algorithm that finds the K nearest neighbors to each data points and either clusters that data point according to whichever cluster was most frequent in the neighbors, or clusters that data point according to a weighted (by distance) vote from the nearest neighbors.
	\item{\textbf{Naive Bayes:}}\\ Naive Bayes classifiers essentially assume that all the features belonging to a dataset are independent of each other, given the class.  Essentially, joint probability distributions are created (rather efficiently using the chain rule, due to the independence assumption) to provide a purely probabilistic estimate of class, based on the training data.
	\item{\textbf{Classification and Regression Trees (CART):}}\\ CART is an umbrella term encompassing several variations on building trees for use with categorical or real-valued data.  In general, they work by building binary classification trees using intelligent selection criteria for choosing attributes to split on (such as information gain or the Gini impurity), and then perform pruning of the resulting tree (or stopping criteria) by minimizing a loss which contains a complexity penalty element and a misclassification (or error, in the real-valued case) element.
\end{itemize}

Despite DM's many benefits, the algorithms do not solve problems or tell users what to do.  It is up to the human Dataminer to clearly understand the problem space, define what they need from the DM solution, exercise the DM process, interpret and validate the results and limitations of the results, and then use the results to inform their action.

New types of problems in DM include finding more stable decision-tree algorithms, decomposition of complex models into ensembles of simpler models, and improvement in feature selection techniques. \cite{Wu:2007:TAD:1327434.1327436} Additional (albeit tangential) topics related to DM include security/privacy implications of DM, the convergence of DM and Internet of Things (IoT), and novel applications of DM.

%------------------------------------------------
%
%  Author Identication
%   
%------------------------------------------------
\section{Author Identification: Full Problem Description}
Edgar Allen Poe, H P Lovecraft, and Mary Wollstonecraft Craft are three popular Horror Fiction authors of their times and whose works have stood the test of time. Poe’s Raven or Shelly’s Frankenstein are works which have defined the genre for centuries. 
In the Spooky Author Challenge, the aim is build a classifier using a training set which contains excerpts from works by the three authors. The training dataset consists of three variables, an identifier, a string of text which is the excerpt and a class label column.  There are 19,560 records in it. The unlabeled test set consists of 8,394 records and the Classifier must label it as “EAP”, “HPL” or “MWS” as its solution

\subsection{Data Analysis}
Any data which is found has to be cleaned and made machine readable. The text data is the only component here. The strings are excerpts from various works. It can be noticed that some are descriptive sentences such as narration while some have dialogs in the text. They are all strings varying lengths. 
One of the text’s features which stands out is that most of the data from MWS are predominantly in first person it is primarily sourced from Frankenstein. 
There are number of such instances of the data can be seen here. The following steps were used:

\begin{itemize}[noitemsep]
\item CSV: As the training and test files were CSVs, when the data was loaded onto MySQL, the comma failed as a delimiter, and had to be processed to make it ‘~’ as this not part of regular sentence structure.
\item Removing of all the punctuations: For the purposes of this assignment, the punctuations were ignored and hence removed as part of pre-processing.
\item Stemming and Stopwords: The words were stemmed and stopwords were eliminated before the text was used for training and testing the data as part of pre-processing. 
\end{itemize}

\subsection{Methods}
There are multiple models available for text processing. The models which were considered for Naïve Bayes’, Decision Trees, SVMs and Maximum Entropy Models.
\begin{itemize}[noitemsep]
\item \textit{Naïve Bayes}: Though Naïve Bayes’ is a relatively simple approach, like Occum’s Razor, a simple approach and fruitful approach. Naïve Bayes’ approach assumes that each feature is independent of one another. They are independently considered, and a probability is assigned to it. This is highly scalable in nature. It follows the following conditional probability method to classify
$$ p(C_k | x_1,x_2,…x_n) $$
$$P(C_k|x) = \frac{p(C_k)p(x|C_x)}{p(x)}$$
Naïve Bayes works particularly well for textual data and hence gives highly reliable output.
Python’s NLTK Naïve Bayes’ classifier was utilized here. This was trained on all the records of the training set and was made to classify the entire test set.
\item \textit{Decision Trees}: Decision Trees or Classification Trees uses various features which is learnt by the classifier to form nodes. Generally, a binary decision is made such as True or False in order to traverse the tree and there are class labels assigned as the root nodes. The path taken from root node to the leaf node serves are the decision made to classify the input as particular class. Decision Trees are easily interpretable by humans owing to their visual nature and hence are highly useful.\\
Python’s NLTK Decision Tree classifier was utilized here. Cross Validation was utilized here, and was trained on sample of the records of the training set and was made to classify the entire test set. The dataset was too large and there were memory errors which occurs when the training set was larger than 5000 records.

\item \textit{SVM}: Support Vector Machines are classifiers used for supervised learning for when the given set of data is linearly separable using a hyperplane of required dimension. SVMs aim at obtaining the best possible solution for the linear separation by attempting to obtain a partition which yields the largest margin. SVMs are useful in classifying larger data sets as they rely only the support vectors to gauge the hyperplane. This is a well renowned tool for text classification.\\
Python’s NLTK and sklearn’s SVC classifier was utilized here. It was trained on sample of the records of the training set and was made to classify the entire test set.

\item \textit{Maximum Entropy}: Maximum Entropy implies that classifiers utilize the most uniform class which meets all the constraints. This has the advantage being applicable to any probability distribution. This classifier unlike Naïve Bayes doesn’t assume independence of features.  Hence, it uses contextual information to obtain information for classification. But this method faces an optimization issue which are improved using an iterative scaling algorithm such as the GIS (Generalized Iterative Scaling) or the IIS (Improved Iterative Scaling) for lambda parameters to be eliminated in the algorithm.\\
Python’s NLTK Maxent classifier was utilized here. Cross Validation was utilized here, and was trained on sample of the records of the training set and was made to classify the entire test set. The dataset was too large and there were memory errors which occurs when the training set was larger than 5000 records.

\end{itemize}
\begin{figure}[ht]
	\begin{minipage}{0.24\textwidth}
		\includegraphics[width=2.0\linewidth]{author/decision_tree.PNG}
	\end{minipage}%
	\caption{Partial Decision Tree}
\end{figure}
%------------------------------------------------

\subsection{Results}
The classifiers each had certain drawbacks as stated above and the results are as follows.
\begin{itemize}[noitemsep]
\item Naïve Bayes performed well with an accuracy rate of 0.910465294448 from the classifier. As stated above this is simple yet a powerful solution for classification.
\item Decision Trees under cross validation gave extremely high accuracy rate of 0.978. This may seem like a perfect classifier, but it most certainly is skewed. There is a high chance of overfitting which has occurred in this classifier implying that the test data hasn’t varied much from the training data. Hence, it is not an optimal solution.
\item SVMs has produced a really low accuracy rate of 0.406 and this was cross validated, but the score failed to improve.
\item Maxent classifier produced accuracy 0.944 with cross validation of 5000 records and 0.789 with 2000 records per iteration.
\end{itemize}
Remember, we're interested in the journey, so simply because an approach failed doesn't mean failure if you discuss the failure!
\subsection{Summary and Future Work}
The Spooky Author Detection project was intended for the classifier to identify Spooky Authors namely, Edgar Allen Poe, H P Lovecraft, and Mary Wollstonecraft Craft. The data has three classes as “EAP”, “HPL”, and “MWS”.  Excerpts from the authors literary work used to train the classifiers to learn the features of the authors. The most reliable results were obtained from the Naïve Bayes’ classifier despite being a simple approach. Maxent classifier obtained better results by improving the cross validation. This project can be deemed a success, despite the obvious gaps in the results mainly because the data was real world data which was worked on. This yielded acceptable results. The handling and management of data was highlighted through this effort.
The potential work can be done using Neural Networks. As stated there are nuances in the data set which were left unexplored for the scope of the current work done, like Shelly’s work being primarily in First person, Shelly inspiring Poe and Lovecraft’s work which may be reflected in the data as homage etc. Using NN such prospects can be explored, and better results might be obtained.

\section{Iceberg: Full Problem Description}

Statoil, an energy company with oil and natural gas operations based largely in the Norwegian Continental Shelf region, has an operational need to track the presence of icebergs that may present a threat to the safety and efficiency of those operations. \cite{statoil} The company has therefore posed the following iceberg classification problem to the data science community.  Statoil partners with C-CORE (a Canadian R\&D corporation) to obtain access to Synthetic Aperture Radar (SAR) imagery data from the Sentinel-1 satellite constellation, and the challenge is to classify the object in each image as an iceberg or non-iceberg (presumably a ship, which is not likely to pose a threat to operations). A labeled training set (1604 records) and an unlabeled test set (8424 records) is provided, and the task is to classify each image as containing an iceberg (\texttt{is\_iceberg}=1) or not (\texttt{is\_iceberg}=0). \cite{kaggle-ice}

\subsection{Data Analysis}
Each record contains three components (besides an identifier):

\begin{itemize}
	\item \texttt{band\_1} - 75$\times$75 pixel image representing horizontally-polarized backscatter intensity from the horizontally-polarized C-band radar transmission
	\item \texttt{band\_2} - 75$\times$75 pixel image representing vertically-polarized backscatter intensity from the horizontally-polarized C-band radar transmission
	\item \texttt{inc\_angle} - angle of incidence (between the transmission path and Earth-normal at the target object)
\end{itemize}

\paragraph{Initial Analysis} Some simple statistical analysis on the image data reveals that maximum and mean values in both bands are meaningful contributors to the classification task, as shown in Figure \ref{max-mean-separable}.

\begin{figure}[ht]
	\centering
	\begin{minipage}{0.24\textwidth}
		\includegraphics[scale=0.2]{iceberg/analysis/band_1_max.png}
	\end{minipage}%
	\begin{minipage}{0.24\textwidth}
		\includegraphics[scale=0.2]{iceberg/analysis/band_2_max.png}
	\end{minipage}
	\begin{minipage}{0.24\textwidth}
		\includegraphics[scale=0.2]{iceberg/analysis/band_1_mean.png}
	\end{minipage}%
	\begin{minipage}{0.24\textwidth}
		\includegraphics[scale=0.2]{iceberg/analysis/band_2_mean.png}
	\end{minipage}
	\caption{Band 1 and 2 Max and Mean}\label{max-mean-separable}
\end{figure}

\paragraph{Further Analysis} The overlap areas in the histograms represent those records that will cause difficulty for a classification model. Additional features are required for good performance.  The following functionality was developed to enable visual analysis as well as experimentation with new useful features:

\begin{itemize}
	\item Incidence angle has a perceived impact on backscatter intensity; see Figure \ref{inc_angle1}.  In this approach, simple linear regression was used to fit a line to background intensity as a function of incidence angle.  The intercepts for this line were used to normalize all the images to the mean incidence angle.  Those images whose incidence angle were not defined (NA) were not altered (i.e. assumed to have mean incidence angle).  The adjusted data is shown in Figure \ref{inc_angle2}.

	\item Simple convolution matrix filtering was implemented to achieve some rudimentary image processing.  Specifically, 3$\times$3 kernels were used for smoothing, sharpening, edge detection, and gradient. \cite{polarimetric} These filters were used to automatically produce images during visual analysis to aid in new feature development.  Examples are shown in Figure \ref{filtering}.  Notably, the plots for band 2 indicate potential higher-mode dependency, but that was not explored further.

	\item A simple center-finding function was implemented to calculate the sum of a sliding window over the image, and the position which yields the largest sum is taken to be centered on the object of interest (OoI).  Several of the features are calculated solely from a fixed-size mini-matrix centered on that position.  That mini-matrix is herein referred to as the `target'.  The intent was yield feature values that are more influenced by target pixels than by non-target pixels.

	\item Histogram of Oriented Gradient (HOG) descriptors provide excellent performance compared to other feature sets for human detection\cite{hog}. Hence, we decided to observe the performance of algorithm on Iceburg detection. The technique counts the number of occurrences of gradient orientation in localized portions of an image. The descriptors are computed on a dense grid of uniformly spaced cells and uses overlapping local contrast normalization for improved accuracy. We have used 'HOG' function from the package \textbf{OpenImageR} to compute HOG descriptors for both band\_1 and band\_2 images. We got the 9-point descriptor for each band\_1 and band\_2 image.    
\end{itemize}

\begin{figure}\centering
	\begin{minipage}{0.45\linewidth}
		\includegraphics[width=.9\linewidth]{iceberg/analysis/b1_bg_intensity-ing_angle_prenorm.png}
	\end{minipage}%
	\begin{minipage}{0.45\linewidth}
		\includegraphics[width=.9\linewidth]{iceberg/analysis/b2_bg_intensity-ing_angle_prenorm.png}
	\end{minipage}
	\caption{Intensity Dependence on Incidence Angle}\label{inc_angle1}
\end{figure}

\begin{figure}\centering
	\begin{minipage}{0.45\linewidth}
		\includegraphics[width=.9\linewidth]{iceberg/analysis/b1_bg_intensity-ing_angle.png}
	\end{minipage}%
	\begin{minipage}{0.45\linewidth}
		\includegraphics[width=.9\linewidth]{iceberg/analysis/b2_bg_intensity-ing_angle.png}
	\end{minipage}
	\caption{Normalized Intensity}\label{inc_angle2}
\end{figure}

\begin{figure}\centering
	\includegraphics[width=.9\linewidth]{iceberg/analysis/525ab75c.png}
	\caption{Image Filtering}\label{filtering}
\end{figure}

\paragraph{Hypotheses} From visual analysis and research in the domain area \cite{wesche} \cite{Ship-Iceberg_CNN} \cite{howell}, several hypotheses were developed which drove new feature experimentation.

\begin{itemize}
	\item Ships are relatively sharp and angular, whereas icebergs are smooth and irregular, so differences in surface backscatter intensities (mean, max, variance) may yield class information.
	\item Volume backscatter properties of ice and metal \cite{howell} are different, so comparisons between band 1 and band 2 may yield some class information.  
	\item Icebergs (being made of frozen water) may have some observable characteristics that are similar to the surrounding water, whereas ships may not.  Therefore, comparison of the OoI to the background seawater may yield class information.
	\item The vast majority of iceberg volume is below the ocean's surface, and the angle of the ice as it protrudes from the water's surface is likely to be off-normal, whereas that of ships is likely to be near-normal.  Therefore, gradient and edge-detection approaches near the object-to-ocean transition area may yield class information.
	\item Incident angle is likely to have an effect on both the intensity and polarity of the return signal. \cite{makynen}
	\item We have computed Histogram of Oriented Gradients(HOG) descriptor for both positive and negative observations. By Positive observation, we mean the observation for which response 'is\_iceburg' is $1$ and Negative observation is the one for which reponse 'is\_iceburg' is $0$. Intuition behind using HOG is to convert $75\times75$ image into a set of numeric values, where each value can be used as separate feature. By this phenomenon, we are able to extract 9 features per image, i.e. for $band\_1 = 9$ and for $band\_2 = 9$. This processed data is used to train regression model and predict test results. 
\end{itemize}

\subsection{Methods}

\paragraph{Model Selection} Several classification methodologies were considered for this classification task: Convolution Neural Networks (CNN), Logistic Regression (LgR), Random forest, XGBoost.  CNN methods have been shown to exhibit good performance on image recognition tasks. \cite{NIPS2012_4824} \cite{behnke}  Two challenges with Neural Networks are that (a) the more naive examples tend towards high-complexity and so require large data sets to achieve high accuracy and low variance, and (b) the hidden nature of the internal neurons pose challenges to applying useful domain knowledge to the classification problem.  They work particularly well in applications where they learn telling features of a visible object within the hidden layers. In the context of this task, the size, shape, and orientation of the OoI is not believed to be as consistent and would be difficult to normalize, which would be highly desirable for a CNN approach.  CNNs, therefore, were not implemented. \cite{lecun}

LgR can be thought of as a special case of a Neural Network which has a single neuron that uses a logit function as its output.  The Since the single neuron is at the outer layer, its inputs and outputs are known.  This makes it straightforward to apply some domain knowledge to the raw data in order to provide inputs (features) that are thought to contain class information.  Additionally, the low complexity of the model makes it ideal for smaller training sets, such as the one provided by Kaggle. \cite{mccullagh} \cite{lim} The Logistic Regression Algorithm is shown in Algorithm \ref{alg:1}.  For this project, the authors used R's built-in \texttt{glm} function.

\paragraph{Logistic Regression} Essentially, LgR uses an initial guess of feature weights to transform the data feature vector into a value between 0 and 1, representing the probabilities that each record belongs to class 1.  A likelihood function and its derivative (with respect to the weights) are calculated, representing the curve of the likelihood that the weights are optimal over the domain of W.  The derivative use used to influence the magnitude and direction for which the weight vector should be adjusted.  The user-defined value $\alpha$ is also used to influence the step size of these adjustments.  This processes is repeated until the gradient is sufficiently small, representing the peak (or valley) of the likelihood function, indicating an optimal weight vector has been reached.

\begin{algorithm}
	\caption{Logistic Regression}
	\label{alg:1}
	\begin{algorithmic}[1]
		\STATE{\ {\bf INPUT:} data $X_{n \times (d-1)}$, class labels $C_{n \times 1}$ }
		\STATE{\ {\bf OUTPUT:} weights $W_{1 \times d}$ }
		\STATE \%\% Define gradient descent parameters
		\STATE{\ $\epsilon \leftarrow$ user threshold }
		\STATE{\ $\alpha \leftarrow$ user step size }
		\STATE \%\% Add ones column for offset
		\STATE{\ $X \leftarrow \left[ X_{n \times (d-1)} \vec{1}_{n \times 1} \right] $ }
		\STATE \%\% Initialize weights
		\STATE{\ $W \leftarrow \vec{0}_{d \times 1}$ } 
		\WHILE { $\left| \nabla_W L(W) \right| > \epsilon$ }
		\STATE \%\% Class probability as function of weights
		\STATE{\ $y_i(W) \leftarrow \frac{1}{1 + e^{-x_iW}} \forall x_i \in X$ }
		\STATE \%\% Likelihood as function of weights
		\STATE{\ $L(W) \leftarrow -\sum_{i=1}^{n} c_i log(y_i) + (1-c_i) log(1-y_i)$ }
		\STATE \%\% Gradient of likelihood wrt weight vectors
		\STATE{\ $\nabla_W L(W) \leftarrow \sum_{i=1}^n x_i (y_i - c_i)$ }
		\STATE \%\% Update weights
		\STATE{\ $W \leftarrow W + \alpha \nabla_W L(W)$ }
		\ENDWHILE
		\RETURN $W$
	\end{algorithmic}
\end{algorithm}

\paragraph{Random Forest}
Random Forests algorithm, creates multiple random decision trees using only important set of the features. It take samples from the training dataset with replacement, and trees are constructed by reducing the correlation between individual classifiers. Instead of choosing best split to construct trees, it uses random subset of features for each split. Advantage of using Random Forest Algorithm is that we don't have to prune trees to prevent over-fitting of the model on the training dataset.\\
We have used \textbf{randomForest} package for the Random Forest Algorithm. We have tuned following parameters of the function to get better prediction.
\begin{enumerate}
	\item[\textbf{1.}] ntree = Number of trees to grow. This should not be set to too small a number, to ensure that every input row gets predicted at least a few times.\\
	\item[\textbf{2.}] nodesize = Minimum size of terminal nodes. Setting this number larger causes smaller trees to be grown.\\ 
	\item[\textbf{3.}] mtry = Number of variables randomly sampled as candidates at each split.\\
\end{enumerate}

\paragraph{XGBoost}
XGBoost(eXtreme Gradient Boosting) is an advanced implementation of gradient boosting algorithm. Boosting is sequential processing technique which works on the principle of ensemble. It gives improved prediction by combining multiple weak learners. At $n^{th}$ instant, the model outcomes are weighed based on previous instant $n-1^{th}$. The outcomes which are predicted correctly gets a lower weight whereas miss-classified predictions get higher weights.\\
Highlights of XGBoost:
\begin{itemize}
	\item[1] We have used \textbf{xgboost} package of R. Computational part of xgboost method is implementated in C++ and it can be multithreaded on single machine.
	\item[2] Customized Objective: XGBoost can be used to train model with custom objective function. Hence, user can define custom objective to improve the model.
	\item[3] Early Stopping: While building the decision trees, if the number of trees to be build is large then waiting time increases. Early stopping parameter when set, XGBoost will terminate the training process if performance is getting worst.  
	\item[4] Handle Missing Values: XGBoost uses soft-coding method to handle missing values. When it encounters a missing value while splitting, it assigns direction instead of numerical value to the missing value.
	\item[5] Feature Importance: XGBoost tries to find best feature and splitting point to optemize the objective function. It calculates the information gain over the feature to decide feature importance.
\end{itemize}

\paragraph{Features} The LgR modeling approach enabled experimentation of new potential features, which became the next focus.  Table \ref{more-features} summarizes, in some detail, a few of the more interesting features that were explored for LgR based on the previously mentioned visual analysis and domain research.  Additional features are briefly described below.

\begin{table*}[ht!]
	\caption{Additional Features for LgR}\label{more-features}
	\begin{tabular}{l p{.45\linewidth} l}
		\toprule
		Feature & Description & Histogram \\
		\hline \vspace{10pt} \\
		\rotatebox[origin=r]{90}{\texttt{tar.sum.gs.sum.mean}} &
			The target mini-matrix for each band is added together.  The result is filtered by vertical gradient and horizontal gradient kernel in parallel. Those two results are then squared and added together, and the mean of the result is taken to be the new feature.  The intent of this feature is to let both bands contribute to squared gradient result.  Two theories are at work here: (a) The edges of ships are vertical and so one might expect a sharper ship-to-water transition, resulting in a higher gradient, and (b) since icebergs are more similar in texture and composition to water than are ships, that may also contibute to a sharper contrast between ship and water, again resulting in a higher gradient. &
			\begin{minipage}[t]{0.35\linewidth}
				\adjustimage{width=1\linewidth,valign=t}{iceberg/analysis/tar_sum_gs_sum_mean.png}
			\end{minipage}\\%
		\rotatebox[origin=r]{90}{\texttt{tb.mean.dif.dif}} & 
			Visual analysis revealed most images to be rather centered on the OoI.  A fixed-size border mask (think picture frame) was created to provide the subset of pixels presumed to reside in the background.  For each band, the background mean was subtracted from the target mean.  The restult for band 2 was then subtracted from the result for band 1, and the final result was taken to be the feature.  The intent was essentially to see if the difference in target-to-background contrast between bands may contribute to classification.  Put another way, the depolarizing properties of ice may be similar to that of the surrounding water, creating a consistently lower contrast image in band 2 as compared with band 1 for icerbergs. &
			\begin{minipage}[t]{0.35\linewidth}%\centering
				\adjustimage{width=1\linewidth,valign=t}{iceberg/analysis/tb_mean_dif_dif.png}
			\end{minipage}\\%
		\rotatebox[origin=r]{90}{\texttt{tar.cor}} & 
			The band 1 and band 2 images are first smoothed to reduce extreme values caused by noise or speckle and zero-adjusted.  Next, all of the pixel values that are greater than 0.7 of the maximum value are kept, the others discarded.  These pixels are then used as a mask to extract all of the pixels in the raw image which are presumed to belong to the OoI.  Finally, a correlation is calculated between these pixels between band 1 and band 2, and the resulting value is taken to be the feature.  The intent is to capture how well band 2 correlates to band 1 over the OoI.  The theory is that depolarization effect would affect correlation and be different for icebergs than for ships. &
			\begin{minipage}[t]{0.35\linewidth}%\centering
				\adjustimage{width=1\linewidth,valign=t}{iceberg/analysis/tar_cor.png}
			\end{minipage}%
	\end{tabular}
\end{table*}

\begin{itemize}
	\item{\textbf{\texttt{band.1/2.tar.var}}} Variation of image pixels in the target for each band.  The seed for this idea originated from Bentes \cite{Ship-Iceberg_CNN} who recognized variance among other signal characteristics as harmful in a Neural Network approach and therefore something to be reduced.  The intent is to see whether there is class information found within the variance of the return signal from the target.
	\item{\textbf{\texttt{band.1/2.tb.mean.dif}}} The difference of means of the target and border regions of the image pixels for each band.  The intent is to see if there is class information derivable from return signal intensity from the target as compared with the background.
	\item{\textbf{\texttt{band.1/2.tar.gvs.mean}}} The target in each band was filtered using a vertical gradient filter and then squared, and the mean of the result was taken for each band.  The intent is to see if there is class information in the object-to-ocean interface area that would be captured by the gradient filter.
	\item{\textbf{\texttt{band.1/2.tar.ghs.mean}}} The target in each band was put through a horizontal gradient filter and then squared, and the mean of the result was taken for each band.
	\item{\textbf{\texttt{band.1/2.tar.mean}}} A mean was taken of the target pixels for each band.
	\item{\textbf{\texttt{tar.gvs.dif}}} The target in each band was filtered using a vertical gradient filter, squared, and meaned.  The result from band 2 was subtracted from that of band 1.
	\item{\textbf{\texttt{tar.ghs.dif}}} The target in each band was filtered using a horizontal gradient filter, squared, and meaned.  The results from band 2 was subtracted from that of band 1.
\end{itemize}

\paragraph{Challenges} Some challenges and drawbacks to the chosen modeling approaches include the following.

\paragraph{}Logistic Regression:
\begin{itemize}
	\item{Structure Simplicity:} LgR is capable of creating arbitarily complex models based on the number input features.  However, as a statistical tool, it lacks some of the structural features of other models such as decision trees and neural nets.  Because of this, the LgR model alone is not capable of branching on specific features.
	\item{Feature Selection:} During experimentation with new and existing feature, simple histogram analysis revealed many of the candidate features as containing some class information (i.e. good separability of classes over the feature).  However, pairs of features are highly correlated and others were not (See figure \ref{correlation}).  It is difficult to choose which features to include manually.
\end{itemize}

\begin{figure}
	\includegraphics[width=.9\linewidth]{iceberg/analysis/corr_matrix.png}
	\caption{LgR Feature Correlation}\label{correlation}
\end{figure}

\paragraph{}Histogram of Oriented Gradients:
\begin{itemize}
	\item{Parameter selection:} As we have mentioned earlier, we have used \textbf{OpenImageR} package for HOG implementation. The function takes following set of parameters:
	\begin{itemize}
		\item[1] \textbf{image:} Matrix or 3-dimensional array. 
		\item[2] \textbf{cells:} Number of divisions(cells).
		\item[3] \textbf{orientations:} Number of orientation bins.
	\end{itemize}
We tried various combinations of these parameters to obtain best results in terms of log loss. But higher values of cells and orientations resulted into higher order of HOG descriptors and lots of features in processed dataset which affected the performace. Hence, we tuned these parameters as 'cells' = 1 and 'orientation' = 9 to get 9-point descriptor.   
\end{itemize}
\paragraph{Implementation} 
\paragraph{}Lgr - Given the manual nature of the feature development and experimentation approach chosen, it was important to automate updates to the analysis, training, and validation approach to the greatest extent possible.  Automation of these aspects enables rapid identification and incorporation of improvements to the feature set.  To this end, all database interaction, image processing, and feature calculation was centralized to a single code source.  An analysis script was run after each change to the feature set to quickly visualize some characteristics of new features.  In addition to automation, V-fold cross validation was utilized as the primary means of model performance prediction.  It should be noted that test data used in cross-validation was present during all data analysis tasks, which may introduce bias as a result in the model selection process. \cite{cawley}  Finally, a simple greedy algorithm was chosen to select which features to use during the training process.  Essentially, starting from an empty feature set, the feature which (when added to the set) results in the lowest Log Loss on the test data during V-fold cross validation is added.  A subset of the (now ordered) feature list which results in the lowest total Log Loss on the test data is then selected for use.  The author notes this algorithm is not guaranteed to find the optimal set of features on which to train.  In fact it was observed that some of the later Kaggle submissions (which represented a superset of features as compared with past submissions) achieved worse performance!  This is because the optimal subset of features was never encountered during the greedy selection process.

\begin{algorithm}
	\caption{LgR Feature Selection}
	\label{alg:3}
	\begin{algorithmic}[1]
		\STATE{\ \textbf{Def:} $ vFoldXVal(\Delta \left[ features \right] ) \rightarrow LogLoss_{test}$ }
		\STATE{\ {\bf INPUT:} feature list $F$, labeled data $\Delta$ }
		\STATE{\ {\bf OUTPUT:} selected feature list $F_S$ }
		\STATE{\ $F_S \leftarrow \emptyset$ }
		\STATE{\ $F_S.last \leftarrow $ NA }
		\STATE{\ $F_S.lossmin \leftarrow \infty$ }
		\WHILE{\ $F \neq \emptyset$ }
		\STATE{\ $f_{best} \leftarrow$ NA }
		\STATE{\ $f_{best}.loss \leftarrow \infty$}
		\FOR{\ $f$ in $F$}
		\STATE{\ $loss \leftarrow vFoldXVal(\Delta\left[\{F_S\} \cup \{f\} \right])$ }
		\IF{\ $loss < f_{best}.loss$ }
		\STATE{\ $f_{best} \leftarrow f$ }
		\STATE{\ $f_{best}.loss \leftarrow loss$ }
		\IF{\ $loss < F_S.lossmin$ }
		\STATE{\ $F_S.last \leftarrow f$ }
		\STATE{\ $F_S.lossmin \leftarrow loss$ }
		\ENDIF
		\ENDIF
		\ENDFOR
		\STATE{\ $F_S \leftarrow \{F_S\} \cup \{f_{best}\}$ }
		\STATE{\ $F \leftarrow F \setminus \{f_{best}\}$ }
		\ENDWHILE
		\STATE{\ $F_S \leftarrow F_S\left[1:F_S.last\right]$ }
		\RETURN $F_S$
	\end{algorithmic}
\end{algorithm}

\paragraph{}HOG with LgR - As we have observed in initial analysis, the training data has 133 $NA$ values particularly in 'inc\_angle' feature which is around 9\% of the training data. Removing these $NA$ values would have resulted in less training examples compared to test data. Hence, we tried to impute these values using \textbf{mice} package. The package is robust in terms of imputing missing values. We have generated 5-possible set of imputed data.\\
Next, we calculated HOG descriptors for both band\_1 and band\_2 images for both training and test data. We divided training data set into training and validation set to check the performance of LgR. Finally, we performed regression on the test dataset and processed the prediction in a way to make submission on Kaggle.

\paragraph{}HOG with XGBoost - As mentioned earlier, we pre-processed the data in similar way by imputing missing values and then computing HOG descriptors for both band\_1 and band\_2 images.
We divided training data set into training and validation set to check the performance of XGBoost.
We have tuned parameters of XGBoost to following values to obtain best results:
\begin{itemize}
	\item \textbf{nrounds:} 100
	\item \textbf{max.depth:} 2
	\item \textbf{eta:} 1
	\item \textbf{nthread:} 2
\end{itemize}

\paragraph{}HOG with Random Forest - 
We divided training data set into training and validation set to check the performance of Random Forest Algorithm.
We have tuned parameters of \textbf{randomForest} to following values to obtain best results:
\begin{itemize}
	\item \textbf{ntree:} 50
	\item \textbf{nodesize:} 10
	\item \textbf{mtry:} 19
\end{itemize}

\subsection{Results}

\paragraph{Processing Performance}
Feature calculation, model training and feature selection, and classification were performed on a personal laptop, the specifications of which are found in Table \ref{machine-specs-lgr} for LgR and \ref{machine-specs} for HOG LgR.  Processing times for Lgr are shown in Table \ref{proc-time-lgr}.

\begin{table}[H]
	\centering
	\caption{Machine Specifications for LgR}\label{machine-specs-lgr}
	\begin{tabular}{r l}
		\toprule
		\textbf{Processor:} 		& Intel Core i5-5200 @ 2.2 GHz x 4 \\
		\textbf{Memory:}			& 8 GiB \\
		\textbf{Operating System:}	& Ubuntu 16.04 LTS \\
		\textbf{Applications:}		& R and MySQL \\
		\textbf{R Packages:}		& RMySQL, rjson, grid, ggplot2, \\
									& corrplot, pROC
	\end{tabular}
\end{table}

\begin{table}[H]
	\centering
	\caption{Machine Specifications for HOG with LgR, XGBoost and Random Forest}\label{machine-specs}
	\begin{tabular}{r l}
		\toprule
		\textbf{Processor:} 		& Intel Core i7-8550 @ 1.9 GHz \\
		\textbf{Memory:}			& 16 GiB \\
		\textbf{Operating System:}	& Windows 10 \\
		\textbf{Applications:}		& R and MySQL \\
		\textbf{R Packages:}		& ROSE, OpenImageR, rjson, grid,\\
		& RMySQL, mice
	\end{tabular}
\end{table}

\begin{table}[H]
	\centering
	\caption{Processing Time (mm:ss) for LgR}\label{proc-time-lgr}
	\begin{tabular}{r l}
		\toprule
		\textbf{Feature Calculation:} 		& 02:22 \\
		\textbf{Feature Selection:}			& 01:26 \\
		\textbf{Model Training:}			& 00:00  \\
		\textbf{Test Data Classification:}	& 13:04 \\
		\textbf{Total:}						& 16:52
	\end{tabular}
\end{table}

\paragraph{LgR Feature Selection Performance}
One might assume that the feature selection ordering may well correlate well with those chosen by visual examination of the histograms, but that is not the case.  Since some of the features inform on some of the others, once a feature is selected and incorporated into the model, the addition of a feature that is highly correlated to one already in the set would result in little performance improvement.  The performance of the LgR model as a function of features added is shown in Figure \ref{feat_selection}.  Overall, the performance on test data tracks relatively well to performance on training data, suggesting that overfitting is not occuring to a great extent for the first dozen or so features.

\begin{figure}
	\includegraphics[width=.9\linewidth]{iceberg/feat_sel.png}
	\caption{Performance During Feature Selection for LgR}\label{feat_selection}
\end{figure}

\paragraph{Predictive Performance}
LgR - The best performance achieved as of the writing of this paper is \textbf{0.3682}, which is about the 18th percentile as of the submission of this paper. \footnote{Kaggle's scoring script uses Log Loss, which is consistent with the evaluation methods used herein}.  The Receiver Operator Characteristics (ROC) plot is shown in Figure \ref{roc-lgr}.\\

\begin{figure}
	\centering
	\includegraphics[width=0.9\linewidth]{iceberg/ROC_Results.png}
	\caption{Logistic Regression ROC Results}\label{roc-lgr}
	\small
	Sensitivity represents icebergs correctly classified as icebergs, and specificity represents non-icebergs incorrectly classified as icebergs.
\end{figure}

HOG with LgR - When we used Histogram of Oriented Gradients with Logistic Regression, we are able to achieve \textbf{0.5505} score after submitting the predictions on test data to Kaggle. Figure \ref{roc-hoglgr} shows the ROC plot for the prediction made on validation data. Area Under Curve (AUC) for the model is \textbf{0.7133157}. Log-loss for the trained model is \textbf{0.6262356}.\\

\begin{figure}
	\centering
	\includegraphics[width=0.9\linewidth]{iceberg/hog_logit_roc.png}
	\caption{HOG with Logistic Regression ROC plot}\label{roc-hoglgr}
	\small
	True-Positive rate represents icebergs correctly classified as icebergs, and False-Positive rate represents non-icebergs incorrectly classified as icebergs.
\end{figure}

HOG with XGBoost - When we used Histogram of Oriented Gradients with XGBoost, we are able to achieve \textbf{0.5986} score after submitting the predictions on test data to Kaggle. Figure \ref{roc-hogxg} shows the ROC plot for the prediction made on validation data. Area Under Curve (AUC) for the model is \textbf{0.8170044}. Log-loss for the trained model is \textbf{0.6610249}.\\

\begin{figure}
	\centering
	\includegraphics[width=0.9\linewidth]{iceberg/xgboost.png}
	\caption{HOG with XGBoost ROC plot}\label{roc-hogxg}
	\small
	True-Positive rate represents icebergs correctly classified as icebergs, and False-Positive rate represents non-icebergs incorrectly classified as icebergs.
\end{figure}

HOG with Random Forest - When we used Histogram of Oriented Gradients with Random Forest, we are able to achieve \textbf{0.5345} score after submitting the predictions on test data to Kaggle. Figure \ref{roc-hogrf} shows the ROC plot for the prediction made on validation data. Area Under Curve (AUC) for the model is \textbf{0.7778883}. Log-loss for the trained model is \textbf{0.5769474}.\\

\begin{figure}
	\centering
	\includegraphics[width=0.9\linewidth]{iceberg/random_forest.png}
	\caption{HOG with Random Forest ROC plot}\label{roc-hogrf}
	\small
	True-Positive rate represents icebergs correctly classified as icebergs, and False-Positive rate represents non-icebergs incorrectly classified as icebergs.
\end{figure}

\subsection{Summary and Future Work}
\paragraph{Summary}
The performance of the LgR classifier resulting from this effort was not exemplary.  The methodology chosen for feature experimentation and calculation was manual and labor intensive, and frankly the problem is not within the author's domain of experience.  The author hypothesizes that the high level of mutual information shared by many of the custom features ultimately limited learning performance of the model.  This hypothesis is supported by both Figure \ref{correlation} and Figure \ref{feat_selection}.  The author considers the project a success, despite the average results achieved.  The lessons learned about how to plan and structure a DM project as well as some of the practical aspects of implementing algorithms and dealing with real data will serve the author well.

\paragraph{Future Work}
Some potential future improvements to the LgR approach would be to adopt an entropy-based or correlation-based feature selection algorithm rather than the simple greedy algorithm described herein - that would enable the selection of features according to the information they contain about class, derived from their probability density functions or correlation to class label vs other existing features, respectively.

Further, one could revisit the Neural Network approach which would allow for nonlinearity and higher complexity.  Neural Networks could incorporate both custom features (as described in the LgR approach) as well as image-oriented features such as the image processing filters discussed herein or the Histogram of Oriented Gradients.  Additionally, some additional domain expertise applied to this problem would surely result in more relevant features which capture more of the known physical phenomena related to SAR imagery that could impact classification performance.  The higher-mode behavior observed in the plots of background intensity as a function of incidence angle is just one example.

%------------------------------------------------
\phantomsection
\section*{Acknowledgments} % The \section*{} command stops section numbering
The authors would like to thank our professor, Dr. Mehmet Dalkilic, for his superior guidance and quick wit in the classroom; our parents for their continued recognition that we have not yet achieved our full potential; and Indiana weather, which provided ceaseless encouragement to stay indoors and pursue Datamining greatness.
\addcontentsline{toc}{section}{Acknowledgments} % Adds this section to the table of contents

%----------------------------------------------------------------------------------------
%	REFERENCE LIST
%----------------------------------------------------------------------------------------
\phantomsection
\bibliographystyle{unsrt}
\bibliography{sample}

%----------------------------------------------------------------------------------------

\end{document}

### 2020-08-31 CyberKGX/body.tex
cleaned old paper info from CyberKGX files

\section{Introduction}
TBD

\paragraph{Our contributions.} TBD

\section{Methodology}

\subsection{Computing Environment}

TBD

\subsection{Data}

TBD

\subsection{Models}

TBD

\subsection{Task}
TBD

\section{Results}

TBD

\section{Conclusions}

TBD

\section{Future Work}

TBD

\begin{acks}

This research was supported in part by Lilly Endowment, Inc., through its support for the Indiana University Pervasive Technology Institute. \cite{PTI} % for Carbonate use

TBD - remember compute resources and other boilerplate acks

\end{acks}

\newpage
\appendix
%\newpage
%Appendix A
\section{Appendix A Title}

TBD

%Appendix B
\section{Appendix B Title}

TBD

### 2020-08-31 CyberKGX/main.tex
cleaned old paper info from CyberKGX files

--- 2018-09-03 sample-sigconf.tex
@@
 \usepackage{booktabs} % For formal tables
+
+\usepackage{subcaption} % for multiple subfigures
+\usepackage{footnote}
+\makesavenoteenv{tabular}
+\makesavenoteenv{table}
 
@@
 \begin{document}
-\title{A Covert Channel Detector for Machine Learning Models}
+\title{Learning to extend a Cybersecurity Knowledge Graph}
 %\titlenote{Produces the permission block, and
@@
 \begin{abstract}
-It has been shown that malicious model training code can augment sensitive training data with specially-crafted samples to enable later exfiltration of the sensitive data by an attacker who has black-box access to the completed model, but no direct access to the sensitive training data.
-\cite{DBLP:journals/corr/abs-1709-07886}  This is known as a Covert Channel for machine learning models.
-It has also been shown an attacker can infer whether a specific sample was included in the training set for a black-box model if the attacker has knowledge of the model's structure or can generate models with similar structure.
-\cite{DBLP:journals/corr/ShokriSS16}  This is known as Membership Inference on machine learning models.
-We demonstrate that a variation of the Membership Inference `attack' can be used to detect whether a given machine learning model may have a covert channel.
-We review some performance factors of the detector, and discuss how performance is directly related to model capacity and sparsity of the sensitive and maliciously-crafted data.
+The landscape of cybersecurity research and intelligence is vast and complex.
+Engineers, developers, and administrators of information technology (IT) systems are faced with the monumental task of knowing about, accessing, making sense of, and making decisions from, cybersecurity knowledge.
+\textcolor{red}{For example, …}
+
+Toward that goal, various repositories of cybersecurity knowledge, including some that are machine-readable \textcolor{red}{(refs)}, have been developed and curated by the community.
+These efforts have improved the coherence and alignment of interrelated lines of cybersecurity research, and created many opportunities for more powerful machine-aided security analysis.
+
+Many projects have endeavored to perform meaningful learning tasks aimed at discovery of new cybersecurity knowledge \textcolor{red}{(refs)} or using existing knowledge to automate or augment design and analysis of systems and software \textcolor{red}{(refs)}.
+With a few exceptions (SemFuzz, others), semantic aspects of most of these learning tasks are limited, or at least not accessible, from the perspective of external semantic reasoning.
+More recently, some of this cybersecurity knowledge has been assembled in the form of knowledge graphs (SEPSES ref) and linked data.
+We believe that representation of cybersecurity knowledge as linked entities and relations is a critical step toward enabling higher semantic learning tasks in the cybersecurity domain.
+
+This project seeks to develop a model to expand a cybersecurity knowledge graph generated from structured, manually entered information to include new knowledge from unstructured documents.
+When encountering a new, unstructured document, the model (1) identifies cybersecurity-related entities contained in the document, (2) classifies each entity type, and (3) identifies and classifies relations between entities.
+The resulting model, therefore, has the capability to propose additions to structured cybersecurity repositories (when it encounters entities or relations that are not currently in the repository) as well as provide graph-based reasoning and context capabilities to new, unstructured information.
+
 \end{abstract}
@@
 
-\input{samplebody-conf}
+\input{body}
 
 \bibliographystyle{ACM-Reference-Format}
-\bibliography{sample-bibliography}
+\bibliography{bibliography}
 

### 2021-03-09 CyberKGX/abstract.tex
CyberKG prior to significant rewrite.

Security analysis is an inherently complex problem.  Vulnerabilities can manifest at any level of a system, ranging from flaws in logic, primitives, or algorithms, to flaws in the configuration or inappropriate use of otherwise correct components/constituents, to unpredictable and undesirable sytem- or system-of-systems-level behavior not attributable to any of the components/constituents in isolation (emergent).  For ease of reference, we label these manifestations as Class 1, 2, and 3, respectively.

Traditional cybersecurity research and cyberintelligence approaches endeavor to produce security data (weaknesses, product vulnerabilities, attack patterns, etc.) to enable the broader cybersecurity community to perform informed assessment and decision making within the cyber threat landscape.  These traditional approaches have proven demonstrably successful in producing security data (dozens of new vulnerabilities are reported every day \cite{noauthor_browse_nodate}).

While these approaches have enabled the development of automation capability to track and manage known vulnerabilities \cite{noauthor_cve_nodate}, they fall short of enabling informed assessment and decision making with respect to Class 2 and Class 3 vulnerabilities \textcolor{red}{(need support for this claim)}.

% Toward that goal, various repositories of structured cybersecurity knowledge \textcolor{red}{(refs)}, have been developed and curated by the community.  These efforts have improved the coherence and alignment of interrelated lines of cybersecurity research, and created many opportunities for more powerful machine-aided security analysis.

% Many projects have endeavored to perform meaningful learning tasks aimed at discovery of new cybersecurity knowledge \textcolor{red}{(refs)} or using existing knowledge to automate or augment design and analysis of systems and software \textcolor{red}{(refs)}.  With a few exceptions (SemFuzz, others), semantic aspects of most of these learning tasks are limited, or at least not accessible, from the perspective of external semantic reasoning.  More recently, some of this cybersecurity knowledge has been assembled in the form of knowledge graphs (SEPSES ref) and linked data.  We believe that representation of cybersecurity knowledge as linked entities and relations is a critical step toward enabling higher semantic learning tasks in the cybersecurity domain.

% This project seeks to develop a model to expand a cybersecurity knowledge graph generated from structured, manually entered information to include new knowledge from unstructured documents.  When encountering a new, unstructured document, the model (1) identifies cybersecurity-related entities contained in the document, (2) classifies each entity type, and (3) identifies and classifies relations between entities.  The resulting model, therefore, has the capability to propose additions to structured cybersecurity repositories (when it encounters entities or relations that are not currently in the repository) as well as provide graph-based reasoning and context capabilities to new, unstructured information.

### 2021-03-09 CyberKGX/main.tex
CyberKG prior to significant rewrite.

--- 2018-09-03 sample-sigconf.tex
@@
 \begin{abstract}
-It has been shown that malicious model training code can augment sensitive training data with specially-crafted samples to enable later exfiltration of the sensitive data by an attacker who has black-box access to the completed model, but no direct access to the sensitive training data.
-\cite{DBLP:journals/corr/abs-1709-07886}  This is known as a Covert Channel for machine learning models.
-It has also been shown an attacker can infer whether a specific sample was included in the training set for a black-box model if the attacker has knowledge of the model's structure or can generate models with similar structure.
-\cite{DBLP:journals/corr/ShokriSS16}  This is known as Membership Inference on machine learning models.
-We demonstrate that a variation of the Membership Inference `attack' can be used to detect whether a given machine learning model may have a covert channel.
-We review some performance factors of the detector, and discuss how performance is directly related to model capacity and sparsity of the sensitive and maliciously-crafted data.
+\input{abstract}
 \end{abstract}

### 2021-03-09 CyberKGX/uc1.tex
CyberKG prior to significant rewrite.

A software engineer is hired to develop a back-end application for a new Internet of Things (IoT) device.  She is given a software requirements specification, a hardware interface specification, a small budget, and tight schedule.  The developer appreciates the importance of security and privacy, especially in the IoT market, and understands that her design implementation decisions will have significant impact on security and privacy, but she is not a cybersecurity expert and cannot afford to dedicate a major portion of the schedule or budget to security.

\paragraph{Status Quo} Near the beginning of the project, she performs an internet survey of existing popular IoT backend frameworks, narrows down to 2 candidates, and engages in some internet research on pros and cons of each, including security.  The number of security hits returned on each is overwhelming.  Some vulnerabilities relate only to older versions and have since been patched.  Some vulnerabilities manifest from poor implmentation.  Some vulnerabilities seem to be instances of weaknesses found in all candidate frameworks.  Hence, most of the search results are not actionable and the developer simply chooses the framework that best supports other project requirements.  In the best case, she may track documented vulnerabilities associated with the chosen framework and implement mitigations into the system.

\paragraph{With CyberKGX}
